### 2017-02-07 abstract.tex
first

In this work we address the problem of finding unsafe behaviors with
given a safety property for a hybrid dynamical system. We treat the
system as a black box and approach the problem in two distinct steps.
First, we model the behavior of the system using a piecewise affine
discrete time model. Such a model is incomplete in nature and is
computed with respect to the given property. Next, we encode the
search for falsification as a bounded model checking query and use an
SMT solver to find a counter example. If found, we check the existance
of the violation in the original system to ensure reproducibility.

%Time bounded LTL property in 
%Models from verificaiton
%model without data
%black box model/data -> structured model

### 2017-02-07 intro.tex
first

%\chapter{Relational Modeling for Falsification}
%\label{chap:rel}

In \chapref{abs}, we analyzed systems by restricting ourselves to
their black box semantics. This enabled us to reason about the
state-space reachability of the system without a direct analysis of
its structure. Using a coarse abstraction which we searched
on-the-fly, we could observe the local dynamics as required. This gave
us an efficient procedure to find abstract counter-examples.  However,
to concretize the counter-examples, a `closer look' or refinement of
the abstraction is required. The grid based state-space abstraction
was refined by splitting the relevant cells (abstract states) into
smaller ones. As noted previously, due to the curse of dimensionality,
such a uniform splitting is an expensive operation, and can not scale to
higher dimensions.

In this chapter, we further our exploration of trajectory segment
based methods; and explore an alternative approach to overcome the
explosion in abstract states. Instead of selectively refining the
abstraction, we compute a model of the black box system and use
bounded model checking to find a concrete counter-example in the
model~\footnote{Due to modeling errors, this might not be reproducible in
the original black-box system.}. We then use this counter-example to
guide the search towards a counter-example in the original system.

More specifically, we use regression to quantitatively estimate the
discovered relations, which were witnessed by trajectory segments, by
affine maps. These maps approximate locally observed behaviors, which
are incorporated into the sampled reachability graph as edge
annotations.  The resulting graph or the Piece-Wise Affine (PWA)
relational model can be interpreted as an infinite state discrete
transition system and model checked for time bounded safety
properties. A counter-example if found, can indicate the presence of a
violation in the system.
%Furthermore, using linear programming, we can over-approximate the
%neighbhourhood of the counter-example in the model.

Towards the end of the chapter, we discuss extensions to data driven
approaches (instead of simulator driven), where instead of a
simulator, a fixed set of data is provided as the behavioral
description of the system. Using a combination of relational modeling,
and program analysis we outline the future work for falsifying
properties of SDCS.

\section{Overview}
\label{sec:ovw}

We now briefly describe the approach. Given a numerical simulator
$\simulate$ and an initial abstraction defined by a quantization
function $\quant_\epsilon$ and a time step $\Delta$, we use
scatter-and-simulate (parameterized by number of samples $n$ as
described in \chapref{abs}) to explore the abstract graph
$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
\areach{\Delta} C'$.

Instead of using a CEGAR like loop (as in \chapref{abs}), we use the
generated trajectory segments to learn quantitative models describing
the local behavior of the system. These models are defined by a set of
relations $R \subseteq \HybridStates \times \HybridStates$ for each
edge of the reachability graph.

\input{figs/segtraj_rel_enriched.tex}

Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
trajectory segment between the respective cells. The graph abstraction
$G$ only states that there exists a state $\x \in C$ from which the
system can evolve to a future state $\x' \in C'$. To increase the
precision we iteratively refined the abstraction by state
splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
which non-deterministically describe relations between $\x \in C$ and $\x'
\in C'$. This is illustrated in \figref{enriched-edge} (compare with
\figref{segtraj}).

The enriched graph $G^R$ captures the underlying local forward
dynamics describing the evolution of the system in each abstract
state. We represent the dynamics using an affine model with an
interval error. Such a model can either be approximated using learning
methods or computed as a sound (over) approximation using reachability
set computation methods. Because we assume black box semantics, we
only present the former. Using regression analysis on the $start$ and
$end$ states of the witnessed trajectory segments between two cells,
we compute an approximate discrete map along with an error estimate.
Moreover, using the simulation function $\simulate$, additional
trajectory segments (or data) can be generated if required. The data
can be separated into a training set and testing set to compute the
map and the error respectively.  The latter case can be explored if
the symbolic dynamics of the system are known. A tool like
\flowstar~\cite{chen2013flow} can be used to find the reachable set
map.

Observe that $G^R$, a directed reachability graph, is rich enough to
search for concrete behaviors in the system. We call it a time
parameterized PWA relational abstraction. It can be interpreted as an
infinite state discrete transition system, and we can use
off-the-shelf bounded model checkers to find concrete violations of a
given safety property, and even other temporal properties.  We now
discuss the background required to present our ideas.

\section{Background}

In this section, we present the basics of relational abstractions and
fundamentals of regression analysis.

\subsection{Relational Abstraction}

For general hybrid dynamical systems, algorithmic ways for computing
analytical solutions do not exist. Discretization (along with suitable
abstractions) is often employed to transform the systems into a
discrete transition system. Relational abstractions is one such idea,
which abstracts continuous dynamics by discrete relations using
appropriate \textit{reachability invariants}. Both time independent
and time dependent relations have been proposed; the former captures
all reachable states over all time, whereas, the latter explicitly
includes time by relating reachable states to time. Thus it
can prove timing properties whereas, time independent relational
abstractions can not. In both cases, the resulting abstractions can be
interpreted as discrete transition systems and can be analyzed using
model checkers. We briefly discuss these two types of abstractions.

\mypara{Timeless Relational Abstraction}

Relational abstractions for hybrid systems were proposed in
\cite{Sankaranarayanan+Tiwari/2011/Relational}. For a hybrid automaton
model, they can summarize the continuous dynamics of each mode using a
binary relation over continuous states. The resulting relations are
timeless (independent) and hence valid for all time as long as
the mode invariant is satisfied. The relations take the general form
of $R(\x,\x') \bowtie 0$, where $\bowtie$ represents one of the
relational operators $=, \ge, \le, <, >$. For example, an abstraction
that captures the monotonicity with respect to time for the
differential equation $\dot{x} = 2$ is $x' > x$. The abstraction
capturing the relation between the set of ODEs: $\dot{x} = 2$,
$\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.

Such relation summaries, discretize the continuous evolution of a
given system into an infinite state transition system, in which safety
properties can be verified using $k$-induction, or falsified using
bounded model checkers. The relationalization procedure involves
finding suitable invariants in a chosen abstract domain, such as
affine abstractions, eigen abstractions or box abstractions
\cite{Sankaranarayanan+Tiwari/2011/Relational}.  For this, different
techniques like template based invariant generation
\cite{Gulwani+Tiwari/2008/Constraint,
Colon+Sankaranarayanan+Sipma/03/Linear} can be used.

\mypara{Timed and Time-Aware Relational Abstractions}

As the above discussed relations are timeless, we cannot reason over
the timing properties of the original system. Moreover, they are not
suitable for time-triggered systems. Timed relational
abstractions~\cite{zutshi2012timed} and time-aware relational
abstractions~\cite{mover2013time} were proposed to overcome these
shortcomings by explicitly including time in the relations. In
addition, both can be more precise than their timeless counterpart,
but unlike it, assume continuous affine dynamics.

To analyze time-triggered systems like an SDCS modeled by an affine
hybrid automata~\footnote{An affine hybrid automata is restricted to
the ODEs, resets and guards being affine.}, timed relational
abstractions can be computed directly from the solutions of the
underlying affine ODEs.  Recall that the solution of an affine ODE
$\dot{\x} = A\x$ can be represented using the matrix exponential as
$\x(t) = e^{tA}\x(0)$. This gives us the timed relation $\x' =
e^{tA}\x$.  Clearly, the relation is non-linear with respect to time.
However, for the case of SDCS with a fixed sampling time period, we
obtain linear relations in $\x$.  We demonstrated the usefulness of
timed relational abstractions for the case of linear systems
in \cite{zutshi2012timed}, and now incorporate the idea to discretize
black box dynamical systems.

Similar to timeless relational abstractions, time-aware relational
abstractions \cite{mover2013time} construct binary relations between
the current state $\x$ of the system and any future reachable state
$\x'$. The difference lies with the latter also constructing relations
between the current time $t$ and any future time $t'$. This is
achieved by a case by case analysis of the eigen structure of the
matrix $A$ (of an affine ODE). Separate abstractions are used for the
case of linear systems with constant rate, real eigen values and complex eigenvalues.

\subsection{Learning Dynamics using Simple Linear Regression}

\mypara{Regression Analysis}

In statistics, regression is the problem of finding a
\textit{predictor}, which can suitably predict the relationship
between the given set of observed input $\x$ and output $\y$ vectors.
In other words, assuming that $\y$ depends on $\x$, regression
strategies find either a parameterized or a non-parameterized
prediction function to explain the dependence. We now discuss simple
linear regression, which is parametric in nature and searches for an
affine predictor. It is also called \textit{ordinary least squares}.

\mypara{Ordinary Least Squares (OLS)}

Let the data set be comprised of $N$ input and output pairs $(\x,
y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
the case in the current context of \textit{finding the best fit}, the
problem is over-determined; there are more equations than
unknowns. Hence, a single affine function cannot be found which
satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
need to find the `best' choice for $\vec{a}$ and $b$. This is formally
defined using a loss function. For the case of simple linear
regression or OLS, the loss function is the sum of squares of the
errors in prediction.  The task is then to determine the vector of
coefficients $\vec{a}$ and an offset constant $b$, such that the least
square error of the affine predictor is minimized for the given data
set.
\begin{equation}
    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
\label{reg}
\end{equation}
To ease the presentation, we can rewrite the above as a homogeneous
expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
\[
    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
\]
The solution of OLS can be analytically computed as
\[ Ab = (X^TX)^{-1}X^T\y\]
where $X$ is the matrix representing the horizontal stacking of all
$\hat{\x}$.
The details can be found in several texts on learning and statistics
\cite{friedman2001elements}.

Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
where $\x\in\reals^n$ we can use OLS to approximate its trajectories
at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
$A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
\begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
    global relational model for the system. The data set for the
    regression is a set of pairs $\setof{(\x,\x') | \x' =
    \simulate(\x, \Delta)}$ and can be generated on demand. Another
    set can be generated to compute the error $\delta$ as an interval
    of vectors, where each element is an interval $\delta_i \in
    [\delta_{min}, \delta_{max}]$.

    Affine maps are poor
    approximations for arbitrary non-linear functions. Hence, we use a
    collection of affine maps to approximate the local behaviors (in
    state-space) of the system $\simulate$. This results in a piece
    wise approximation, as described in the next section.

\subsection{Piecewise Affine (PWA) Transition System}

We define the PWA transition system as a transition system
(ref.~\secref{DiscSys}) $\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
where each transition $\tau \in \scrT$, is associated with a
transition relation of the form
\[
    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
\]
where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
The assertion on the initial states $\Theta$ is also affine.

We can now use a PWA transition system $\rho$ to approximate the
behavior of a hybrid dynamical system $\System$ defined by a simulator
$\simulate_{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
form, we can only approximate general hybrid systems.}. Each
transition of $\rho$ represents a discrete time step $\Delta$. Let the
state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
predicates $g_\tau$ are given by a conjunction of hyper-planes and
hence can be represented in matrix form as a polyhedron (defined by
$m$ constraints) in the state-space
\[
g:C\x - \vd \le 0
\]
where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
$m$. The affine maps $f_\tau$ can be represented in matrix form as
$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
an error term $\delta$ which is defined by a vector of intervals
$[\delta_{\min},\delta_{\max}]$. Hence,
\[
f:\x \mapsto A\x + \vb + \delta
\]
defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
\in f_\tau(\x)}$.

% Given a state vector $\x\in\reals^n$, a guarded affine map
% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
% guard is satisfied.

We now formalize the PWA affine transition system for a dynamical system.
\begin{definition}[PWA Transition System]

Given a hybrid dynamical system over a state-space
$\HybridStates^n$, a PWA transition system $\rho$ is given by the
tuple $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in
\scrT$ are affine transition relations and $\Theta$ is an affine
predicate over $\x$ and the states $\x\in\HybridStates^n$. The
transition relation is then defined by $\scrT$ with $n$ transition
relations as follows

\begin{equation}
    \scrT = \left\{
        \begin{array}{ll}
            g_1(\x) \land g'_1(\x') \implies \x' \in f_1(\x) \\
            \ldots\\
            g_n(\x) \land g'_n(\x') \implies \x' \in f_n(\x) \\
        \end{array}
    \right.
\end{equation}

\end{definition}

where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
denotes the next state of the system.

A PWA model is \textit{deterministic}, iff for every state
$\x\in\HybridStates^n$, a unique guarded affine map is satisfied and
all errors $\delta$ are singletons. However, in practice such a case
rarely exists. A PWA model will be usually non-deterministic, both due
to multiple choice of transition relations and the reachable states at
the $k^{th}$ time step (or after $k\Delta$ units of time) being a set.
Abusing the notation, we denote the set of states reachable in a
single step in $\rho$ by $\x' = \rho(\x)$.

A PWA transition system is \textit{complete}, iff for every state
$\x\in\HybridStates^n$, there exists at least one satisfied
transition relation $\scrT$. If this is not the case, the system can
deadlock, with no further executions. This usually results from
incorrect modeling of SDCS, and from here on, we do not consider such
cases.

% Given a dynamical system over a continuous state-space
% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
% dynamical system to a finite set of guarded affine maps
% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
% rule by the affine map of the satisfied $\gm$.

% \begin{equation}
%     \pwa = \left\{
%         \begin{array}{ll}
%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
%             \ldots & \ldots\\
%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
%         \end{array}
%     \right.
% \end{equation}

% \end{definition}

% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
% Abusing the notation, we denote the next state computed by the PWA
% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
% for every state $\x\in\ContStates$, a unique guarded affine map is
% satisfied. A PWA model is \textit{complete}, iff for every state
% $\x\in\ContStates$, there exists at least one satisfied guarded affine
% map $\gm$.

\section{Relational Modeling}

We now describe how relational models can be computed for a given
black box system by enriching the abstract reachability graph obtained
by scatter-and-simulate. We first formalize the notion of an enriched
graph, and then show how they can be interpreted as a PWA transition
system using $k$-relational modeling.

\subsection{Abstract Enriched Graph}

The existential abstraction relation in \chapref{abs} was defined as
follows:
\[
%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
\]
The abstract relation $\areach{t}$ can be enriched by incorporating
the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
system, such a relation can be rarely represented using an exact affine map.
This is due to the presence of non-linear and hybrid behaviors. But,
an affine map can always be estimated with an error.

If the system dynamics are completely specified in the form of a white
box model, we can use first order approximations to find the affine
expressions for the relations. Using a tool like \flowstar, we can
obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
that, every element $b_i$ of the vector $\vb$ is contained in the
respective scalar interval $b_i\in[b^l_i,b^h_i]$.

For the case of black box systems, such a sound approximation is not
possible. Instead, we rely on a statistical method like simple linear
regression to estimate the affine map $\amap: A\x + \vb$. We then
estimate the error $\delta$ and generalize $\amap$ to an interval
affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
abstraction with the below relation
\[
    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
\]

However, in the presence of complex non-linear behavior, using a
single affine map can lead to a poor approximation. To increase the
precision, one can use more than one affine map. Let us denote this set as
$R_{(C,C')}$ (ref.~\secref{ovw}).
\[
    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
\]

\begin{figure}[!htbp]
\begin{center}
\tikzstyle{line} = [thick]
\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
\begin{tikzpicture}

    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
    \draw[arw] (-3,0) -- (-1.2,0);
    \draw[arw] (1.2,0) -- (3,0);

\node at (-3.4,0) {$G$};
\node at (3.4,0) {$G^R$};
\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};

\end{tikzpicture}
\end{center}
\vspace*{-.3cm}
\caption{Using OLS, $G^R$ is computed by determining the appropriate
$R_{(C,C')}$ for each edge of $G$.}
\label{fig:enrichment}
\vspace*{-.3cm}
\end{figure}

As shown in \figref{enrichment}, we use OLS to compute an enrichment
of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
of the enriched edge are clearly non-deterministic. Form a state $\x
\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
is satisfied. This interpretation results in an infinite state
transition system. We now detail the construction of the set
$R_{(C,C')}$ using $k$-relational modeling.

% We now compute a PWA model for a given black box system by estimating
% the dynamics using OLS. We can estimate the affine relations between
% every cell, but as discussed in ~\chapref{abs}, it is a futile
% approach.  Instead, we use the same heuristic as before:
% scatter-and-simulate, to select the cell relations for estimation. We
% build the reachability graph as before, but in addition, annotate each
% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
% respective relation. The reachability graph thus computed, is a
% transition system. Using off-the-shelf bounded model checkers, we
% reason about the system's safety properties.

\subsection{$k$ - Relational Modeling}

\begin{figure}[!htbp]
\vspace*{-.2cm}
\begin{center}
\tikzstyle{line} = [thick]
\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
\begin{tikzpicture}

\begin{scope}[scale=1]
\draw [line] (-1.0,0) circle (0.5);
\draw [line] (-4.0,0) circle (0.5);
\draw [line] (2.0,1) circle (0.5);
\draw [line] (2.0,-1) circle (0.5);
\draw[arw] (-0.5,0) -- (1.5,1.0);
\draw[arw] (-0.5,0) -- (1.5,-1.0);
\draw[arw] (-3.5,0) -- (-1.5,0.0);
\node at (-4.0,0) {$C$};
\node at (-1.0,0) {$C'$};
\node at (2.0,1.0) {$C''_1$};
\node at (2.0,-1.0) {$C''_2$};
\draw [line] (2.0,-1) circle (0.5);
\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
\node at (-4.0,-2.5) {$k=0$};
\node at (-1.0,-2.5) {$k=1$};
\node at (+2.0,-2.5) {$k=2$};
\draw [fill=black] (-4.0,-2.0) circle (0.05);
\draw [fill=black] (-1.0,-2.0) circle (0.05);
\draw [fill=black] (2.0,-2.0) circle (0.05);

\end{scope}
\end{tikzpicture}
\end{center}
\vspace*{-.3cm}
\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
\label{fig:k-rel}
\vspace*{-.3cm}
\end{figure}

Given a $G$, when computing the set of relations for an edge $R$, we
consider only the set of abstract states reachable at a specific time
step.  Intuitively, one might consider the states reachable in one
time step $\Delta$.  However, such a process can be generalized by
looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
k\Delta}$ time steps.

Using \figref{k-rel}, we illustrate this notion. To compute
$R_{(C,C')}$, we
split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
$k\Delta$) as follows.  We observe the local behavior of the system by
noting the evolution of the system $\System$ from a state $\x(t)
\in C$ for a time length dependant on $k$. For
\begin{itemize}
    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
        and split them in two sets (a) and (b) on the basis of the
        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
        \begin{enumerate}[(a)]
            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
        \end{enumerate}
    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
        and split them in to multiple sets on the basis of the cells
        reached at time $(t+\Delta),\; (t+2\Delta),\;
        \ldots,\;(t+n\Delta)$.
\end{itemize}

$k$-relational modeling can be understood as a heuristic, which uses
the underlying abstraction to differentiate behaviors of the system
which `diverge' (or are revealed to be `distinct' at a future time).
Such a heuristic can be useful in increasing the precision of the
learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
$n$ and illustrate using examples.

\mypara{$0$-Relational Model}

\begin{figure}[!htbp]
\begin{center}
\tikzstyle{line} = [thick]
\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
\begin{tikzpicture}
\begin{scope}[scale=1]

    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);

\node at (0.0, 0.0) {$C$};
\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
\node at (2,1.5) {$\pi_1$};
\node at (2,0.5) {$\pi_2$};
\node at (1.7,-1) {$\pi_3$};
\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
\node at (0,-2.5) {$t=0$};
\node at (4,-2.5) {$t=\Delta$};
\draw [fill=black] (0.0,-2.0) circle (0.05);
\draw [fill=black] (4.0,-2.0) circle (0.05);

\end{scope}
\begin{scope}[xshift=0,yshift=3.0cm,scale=1]
    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
    \implies A_1\vx + \vb_1 + \delta_1$};
\end{scope}

\end{tikzpicture}
\end{center}
\vspace*{-.3cm}
\caption{All the trajectory segments will be used to construct the
model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
\label{fig:k0}
\vspace*{-.3cm}
\end{figure}

When $k=0$, the $0$-relational model is a PWA transition system which only
defines the evolution of states $\x$, and does not specify the
reachable cell. The guard predicates of the relations are defined over
cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.

We use regression to estimate the dynamics for the outgoing trajectory
segments form a cell $C$. Hence, the data set $\ds$ for the regression
includes all trajectory segments beginning from the same cell
\[
    \ds = \setof{\pi_t | start(\pi_t) \in C}.
\]
This includes trajectory segments ending in different cells, as shown
in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
\begin{equation*}
    \scr{T} = \left\{
        \begin{array}{ll}
            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
            \ldots \\
            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
        \end{array}
    \right.
\end{equation*}

Note that this can be quite imprecise when the cells are big,
containing regions of state-space with complex dynamics. This is true
for both non-linear systems and hybrid dynamical system, where a cell
can contain two or more modes with differing continuous dynamics.

\mypara{$1$-Relational Model}

\begin{figure}[!htbp]
\begin{center}
\tikzstyle{line} = [thick]
\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
\begin{tikzpicture}
\begin{scope}[scale=1]

    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);

\node at (0.0, 0.0) {$C$};
\node at (4.0, 2.0) {$C'_1$};
\node at (4.0, -2.0) {$C'_2$};
\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
\node at (2,1.5) {$\pi_1$};
\node at (2,0.5) {$\pi_2$};
\node at (1.7,-1) {$\pi_3$};
\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
\node at (0,-4.0) {$t=0$};
\node at (4,-4.0) {$t=\Delta$};
\draw [fill=black] (0.0,-3.5) circle (0.05);
\draw [fill=black] (4.0,-3.5) circle (0.05);

\end{scope}
\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
    \node at (2,0) {\begin{minipage}{\textwidth}
    \footnotesize \begin{equation*}
        \scr{T} = \left\{
        \begin{array}{ll}
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
        \end{array}
    \right.
\end{equation*}
\end{minipage}};
\end{scope}

\end{tikzpicture}
\end{center}
\vspace*{-.3cm}
\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
%Cardinality($R_{(C,C')}$) = $2$.}
\label{fig:k1}
\vspace*{-.3cm}
\end{figure}

% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
%     \node at (2,0) {\footnotesize $\begin{equation*}
%     \scr{T} = \left\{
%         \begin{array}{ll}
%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
%         \end{array}
%     \right.
% \end{equation*}$};
% \end{scope}

To improve the preciseness of learnt dynamics, we include the
reachability relation in the
regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
comprised only of trajectory segments $\pi_t$ which start and end in
the same set of cells.
\[
    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
\]
For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
$N$ transition relations.

\figref{k1} illustrates the $k=1$ refinement of the case shown in
\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
relations are constructed. The resulting $1$-relational model
is at least as precise as the corresponding $0$-relational model.

\mypara{$k$-Relational Model}

\begin{figure}[!htbp]
\begin{center}
\tikzstyle{line} = [thick]
\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
\begin{tikzpicture}
\begin{scope}

    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;

%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);

    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);

\node at (0.0, 0.0) {$C$};
\node at (4.0, 2.0) {$C'_1$};
\node at (8.0, 2.0) {$C''_1$};
\node at (8.0, -2.0) {$C''_2$};
%\node at (4.0, -2.0) {$C'_2$};
\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
\node at (2,1.5) {$\pi_1$};
\node at (2,0.5) {$\pi_2$};
\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
\node at (0,-4.0) {$t=0$};
\node at (4,-4.0) {$t=\Delta$};
\node at (8,-4.0) {$t=2\Delta$};
\draw [fill=black] (0.0,-3.5) circle (0.05);
\draw [fill=black] (4.0,-3.5) circle (0.05);
\draw [fill=black] (8.0,-3.5) circle (0.05);
\end{scope}

\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
    \node at (4,0) {\begin{minipage}{\textwidth}
        \footnotesize \begin{equation*}
    \scr{T} = \left\{
        \begin{array}{ll}
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
        \end{array}
    \right.
\end{equation*}
\end{minipage}};
\end{scope}

\end{tikzpicture}
\end{center}
\vspace*{-.3cm}
\caption{The $k=2$ refinement further splits the data set $\ds_1$
    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
    non-deterministically defining the system behavior.}
\label{fig:k2}
\vspace*{-.3cm}
\end{figure}

Finally, we generalize the relational models to $k$-relational PWA
models. A $k$-relational model is constructed by using $k$ length
\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
\textit{connected} iff its cost $\cost(\SegTraj) = 0$.

Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
\textit{similar} if their trajectory segments have the same sequence
of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
\[
    \forall i\in\setof{1 \ldots k}.
    start(\pi_{t_i}) \in C_i
    \land start(\pi'_{t_i}) \in C_i
    \land end(\pi_{t_i}) \in C_{i+1}
    \land end(\pi'_{t_i}) \in C_{i+1}
\]

For every cell $C$ in the reachability graph, we construct a data set
$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
length \textit{connected} and \textit{similar} segmented trajectories.
%\[
%    \ds = \setof{\pi_{\tau_i} |
%        \pi_{\tau_i}
%        start(\pi_{\tau_{i}}) \in C_i
%        \land
%        end(\pi_{\tau_{i}}) \in C_i}
%\]

\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.

\section{Bounded Model Checking Black Box Systems}
% include an algorithm
We now describe the entire procedure in three steps.
\begin{enumerate}
    \item Given a black box system $\System$ specified by
        $\simulate_\System$, we
        use scatter-and-simulate with ($\Delta$-length) trajectory
        segments to sample a finite reachability graph
        $\scr{H}(\Delta)$.
    \item Using regression, we enrich the graph relations and annotate
        the edges with the estimated affine maps.
    \item Interpreting the directed reachability graph as a transition
        system, we use a bounded model checker to find fixed length
        violations to safety properties.
\end{enumerate}

\section{An Example: Van der Pol Oscillator}

We now describe the above using the Van der Pol oscillator benchmark
presented in \secref{math-dyn-sys}. Simulations generated using
uniformly random samples are shown in \figref{vdp-cont}. We want to
check the property $P_3$, indicated by the red box, given the initial
set indicated by the green box. The abstraction is defined by
$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
where each cell is of size $0.2 \times 0.2$ units.
Scatter-and-simulate is then used to construct the abstract graph $G$,
using $2$ samples per cell and the time step $\Delta = 0.1s$. The
complete process follows.

\begin{enumerate}
    \item{\emph{Abstract}} Consider an implicit abstraction induced by the quantization
    function $\quant_\epsilon(\x)$. The corresponding reachability
    graph $\scr{H}(\Delta)$ (with time step $\Delta$) is given by
    $\tupleof{\scr{C},\areach{\Delta}, \scr{C}_0, \scr{C}_u}$.
\item{\emph{Discover}}: Using scatter-and-simulate, enumerate a finite number
    of cells (vertices) and the relations (edges) of the graph
    $\scr{H}(\Delta)$. Associate the set of generated trajectory
    segments with their respective originating cells using a map $D:C
    \mapsto \setof{\pi | start(\pi) \in C}$. The discovered
    abstraction is show in \figref{vdp-abs}. As mentioned, red cells
    are unsafe and green cells are initial cells.
\item{\emph{Relationalize}} For each cell $C$, perform regression analysis on
    the respective set of trajectory segments $D(C)$, and compute a
    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(\Delta)$. Annotate
    each edge in the graph $\scr{H}(\Delta)$ with the respective
    relation.

    \figref{vdp-map} shows the cells and the trajectory segments which
    are part of the data sets constructed using the $1$-relational
    modeling. Against each cell, its unique identifier (integer
    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
    relations. Note that self loops result when an observed trajectory
    segment has its $start$ and $end$ states in the same cell.

\item{\emph{Model Check}} The graph $\scr{H}(\Delta)$ can now be viewed as a
    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
    where $C \in vertices(\scr{H}(\Delta))$ and $\x \in
    \HybridStates$. A transition $\tau \in \scrT$ is of the form
    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
    edge(\scr{H}(\Delta))}$.
\item{\emph{Check Counter-example}} The infinite state (but finite location) transition system can
    be model checked to find a concrete counter-example, which if
    exists, can indicate the existence of a similar trace in the
    original black-box system $\System$. The latter check is carried
    out as before, using the numerical simulation function
    $\simulate$. For the given example, the model checker is unable to
    find a counter-example.
\end{enumerate}

\inclfig[width=0.75\textwidth]{vdp_cont_traces.png}{Van der
Pol: continuous trajectories. Red and green boxes indicate unsafe and
initial sets.}
\label{fig:vdp-cont}

\inclfig[width=0.75\textwidth]{vdp_abs.png}{The discovered
abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
initial cells.}
\label{fig:vdp-abs}

\inclfig[width=\textwidth]{vdp_disc_map.png}{Cells and
trajectory segments used by $1$-relational modeling.}
\label{fig:vdp-map}

\inclfig[width=0.40\textwidth]{vdp_graph.png}{Enriched
graph $G^R$. The affine maps $f$ for the transition relations are show
in \tabref{vdp-pwa}.}
\label{fig:vdp-graph}

\begin{table*}[!htbp]
\centering
\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
\label{tab:vdp-pwa}
\begin{tabular}{@{}ccccc@{}}
\toprule
$C$ & $C'$ & $A$ & $b$ & $\delta$\\
\midrule
(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
\midrule
(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
\midrule
(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
\midrule
(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
\midrule
(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
\bottomrule
\end{tabular}
\end{table*}

% (0, -1)   & (0, -1)   &  \begin{bmatrix}0.9944498 & 0.12911959 \\-0.12013468&1.63250531\end{bmatrix} &\begin{bmatrix}-0.00003683\\-0.00085411\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
% (0, -1)   & (0, -2)   &  \begin{bmatrix}0.99523928& 0.1290447  \\-0.10236736&1.6314103 \end{bmatrix} &\begin{bmatrix}-0.00013019\\-0.00289192\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
% \midrule
% (0, -2)   & (0, -2)   &  \begin{bmatrix}0.99565348& 0.12916002 \\-0.09352462&1.6341424 \end{bmatrix} &\begin{bmatrix}-0.00015574\\-0.00339664\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
% (0, -2)   & (0, -4)   &  \begin{bmatrix}0.99668599& 0.12892455 \\-0.0725035 &1.62937613\end{bmatrix} &\begin{bmatrix}-0.00037391\\-0.00796189\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
% \midrule
% (0, -9)   & (-1, -15) &  \begin{bmatrix}0.99985362& 0.12858788 \\-0.09447884&1.61476417\end{bmatrix} &\begin{bmatrix}-0.0018997 \\-0.04200363\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
% \midrule
% (-1, -15) & (-3, -24) &  \begin{bmatrix}0.95473217& 0.12473489 \\-1.29198505&1.47155614\end{bmatrix} &\begin{bmatrix}-0.01248412\\-0.41100129\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
% \midrule
% (-3, -24) & (-4, -29) &  \begin{bmatrix}1.11789749&-0.17261206 \\-1.71017779&0.66006597\end{bmatrix} &\begin{bmatrix}-1.0745344 \\-3.31246863\end{bmatrix}&\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}\\
% \midrule
% (-4, -29) & (-4, -29) &  \begin{bmatrix}0.99796588& 0.01012817 \\-0.41529747&1.02432211\end{bmatrix} &\begin{bmatrix}-0.00135199\\-0.28016983\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
% \midrule
% (-1, -10) & (-2, -16) &  \begin{bmatrix}0.97381953& 0.12768367 \\-0.71242741&1.56764585\end{bmatrix} &\begin{bmatrix}-0.00320265\\-0.11964102\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
% (-1, -10) & (-2, -15) &  \begin{bmatrix}0.97373058& 0.12859163 \\-0.70173855&1.58540872\end{bmatrix} &\begin{bmatrix}-0.0015415 \\-0.08616376\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
% \midrule
% (-2, -16) & (-3, -23) &  \begin{bmatrix}0.92850123& 0.12242812 \\-1.84720528&1.39159826\end{bmatrix} &\begin{bmatrix}-0.02334236\\-0.73348865\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
% \midrule
% (-3, -23) & (-5, -29) &  \begin{bmatrix}3.14947585&-0.81150646 \\-0.96159406&0.23543254\end{bmatrix} &\begin{bmatrix}-3.13860326\\-4.99696623\end{bmatrix}&\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}\\
% \midrule
% (-2, -15) & (-3, -22) &  \begin{bmatrix}0.93156623& 0.12260264 \\-1.75117219&1.40104348\end{bmatrix} &\begin{bmatrix}-0.02214125\\-0.68423569\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
% \midrule
% (-3, -22) & (-5, -29) &  \begin{bmatrix}3.08806462&-0.72332776 \\-1.69271079&0.40308293\end{bmatrix} &\begin{bmatrix}-2.77908467\\-4.57411339\end{bmatrix}&\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}\\

\subsection{Search Parameters}

The search parameters for S3CAM-R include the parameters of S3CAM:
$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
maximum error budget $\delta_{max}$ for OLS and the maximum length of
segmented trajectory for building $k$-relational models.

We have already discussed the effects of $N$, $\epsilon$ and $\Delta$
on S3CAM's performance. However, they also have an effect on
relational modeling. A finer grid with small cells produces more
accurate models than a coarser grid with large cells. Similarly,
small time length trajectory segments result in more accurate
modeling.
%An increase in the number of samples can increases the
%accuracy of regression.

\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
Given a $\delta_{max}$, we keep increasing $k$ during the
$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
satisfied. A $k_{max}$ is introduced to bound the longest
segmented trajectory which can be considered.

% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }

\subsection{Reasons for Failure}
The approach can fail to find a counter-example, even when it exists, in
one of three ways.

\begin{itemize}

    \item \emph{S3CAM Fails} No abstract counter-example is found by
        S3CAM. We can remedy it by increasing search budgets and/or
        restarting.

    \item \emph{BMC Fails} An abstract counter-example is found, but
        the BMC fails to find a concrete counter-example in the PWA
        relational model. The failure can be attributed to either (a)
        a spurious abstract counter-examples or (b) a poorly estimated
        model. The former can be addressed by restarting but the
        latter requires that the maximum model error $\delta_{max}$ be
        decreased.

    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
        is found, and it is successfully concretized in the PWA
        relational model by the BMC. However, it is not reproducible
        in the black box system. This happens when the PWA relational
        model is not precise enough.

\end{itemize}

\section{Implementation and Evaluation}

The implementation was prototyped as S3CAM-R, an extension to our
previously mentioned tool S3CAM (\chapref{case}). OLS regression
routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
Python module for machine learning. SAL~\cite{SAL-SRI}
with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
solver.

We used S3CAM to discover the abstraction graph $G$, which was then
trimmed of the nodes which did not contribute to the error search
(nodes from which error nodes were not reachable). In our experiments we
used $1$-relational modeling to create the transition system from $G$.
Using SAL, we checked for the given safety property. If a concrete
trace was obtained from the BMC, it was further checked for validity
by simulating using $\simulate$ to see if it was indeed an error
trace.  If not, $100$ samples from its associated abstract state was
sampled to check the presence of a counter-example in it
neighborhood. This can be further extended by obtaining different
discrete sequences or discrete trace sequences from the model checker.

We tabulate our preliminary evaluation in \tabref{res-rel}. We used
few of the benchmarks described in \chapref{case}, including the Van
der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
benchmark. As before, we ran S3CAM-R $10$ times with different seeds
and averaged the results. We tabulate both the total time taken and
the time taken by SAL to compute the counter-example and compare
against a newer implementation of S3CAM. Note that different
abstraction parameters are used for S3CAM and S3CAM-R, due to the
differences in which they operate.

\begin{table*}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in seconds (rounded off
to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
Benchmark & BMC & S3CAM-R & S3CAM\\
\midrule
Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
Brusselator               &$<$1  & $4$    & $2$\\
Lorenz                    &$<$1 & $122$  & $35$\\
%B.Ball                    &  &   & \\
Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
Bouncing Ball             &-  &$TO$   & $450$\\

\bottomrule
\end{tabular}
\end{table*}

% 1m8s, 2m56.184s, 1m31.059s[f], 1m32.665s,  1m45.312s, 1m22.397s, 2m1.451s[f], 1m12.118s, 1m26.596s
% 0.12, 0.06,         0.33       0.13,       0.22,     FAIL,       ,X           0.08        0.15

The results show promise, but clearly S3CAM performs better. Also,
S3CAM-R timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
and the bouncing ball. However, we need to explore more benchmarks to
be conclusive. S3CAM-R's performance can be explained by the
difficulty in finding a good abstraction ($\quant_\epsilon$), which
needs to be fine enough, to obtain good prediction models but coarse
enough, to be manageable by the current SMT solvers. Specifically, to
obtain a good quality counter-example from the model checker (and
avoid false positives), the transition system should be created from
models with high accuracy, for which a finer abstraction is required.
However, the exploration of a finer abstraction is exponentially more
complex and results in a much bigger transition system, which the
current state of the art bounded model checkers (which use SMT
solvers) can not handle under reasonable resources.

\section{Conclusion and Future Work}

We have presented another methodology to find falsifications in black
box dynamical systems. Combining the ideas from abstraction based
search (\chapref{abs}), with our previous relational
abstractions~\cite{zutshi2012timed}, we proposed $k$-relational
modeling to compute richer abstractions. Simple linear regression can
be used to estimate the local dynamics of the trajectory segments and
compute PWA relational models which can be interpreted as a transition
system.  These can then be checked for safety violations using an
off-the-shelf bounded model checker.  Finally, we implemented the
ideas as a tool S3CAM-R and demonstrated its performance on a few examples.

\subsection{Improvements}

The approach seems promising as it avoids an expensive refinement
step. But, before it can be applied to more complex systems, we need
to overcome its shortcomings.  As a future extension, we are
investigating ideas to improve the performance of the underlying three
main sub-processes, namely, abstraction, modeling, and BMC. We detail
these below.

\begin{enumerate}

\item \emph{Abstraction} The abstract domain we have used, has been
    confined to the interval or rectangular domain. Such a domain is
    very coarse when compared to more general polyhedral domains, using
    which, we can gain better precision. Due to polyhedral domains
    being computational expensive, their integration in S3CAM needs to
    be explored thoroughly.

\item \emph{Modeling} We have used $k$-relational modeling to increase
    the precision of the enriched abstractions, however, there are
    several other refinement strategies that remain to be explored.
    This includes refining the abstract states by using guard
    predicates (similar to predicate abstraction), using non-linear
    templates in parametric regression to compute more precise
    quadratic relations, and using an adaptive time step instead of a
    fixed one to address the non-linearities due to a long time
    horizon.

\item \emph{Model Checking} As of now, we use SMT solvers to search
    for concrete counter-examples. Although they are very efficient,
    owing to extensive engineering effort, reachability of transition
    systems resulting from dynamical systems remains a difficult
    problem. As the time horizon of the safety property increases, the
    possible combinations of discrete transitions increase
    exponentially. Hence, to find a counterexample which is a sequence
    of discrete transitions over a `long' time horizon is not
    tractable for most, but the simplest of dynamical systems.
    Instead, we can use linear programming solvers, by enumerating
    each path in the graph $G^R$. Note that the constraints are all
    linear (conjunctions) along an abstract path. However, in the
    worst case, the number of paths in a graph can be of the order
    $n!$ where $n$ is the number of vertices of a graph. Hence, this
    will not be feasible, unless, we can prioritize paths by using a
    triage process similar to one used by S3CAM and using a budget on
    the maximum number of paths.

    Another approach to address the issue would be to use an adaptive
    time discretization technique, where relations over both shorter
    and longer time steps are computed, and the SMT solver can
    `select' the time step precise enough to find a counter-example.

\item \emph{SMT Solvers} One major impediment to our approach is the
    fact that SMT solvers use the theory of reals with \emph{exact
    precision}. This is important for verification approaches, but can
    be relaxed for the problem of falsification. An SMT solver which
    uses approximate reasoning but returns robust counter-examples
    will be as useful, and perhaps more efficient.

\end{enumerate}

\subsection{Data Driven Analysis}

Finally, we would like to mention that our approach is `simulation'
driven. It can be easily modified to be `data driven', by working with
a fixed set of data. Given a data set, we then need to automatically
find a good abstraction and a high fidelity PWA transition system
using which, we can summarize the black box hybrid dynamical system.
Apart from model checking the transition system, one can extend it to
the analysis of SDCS. More specifically, by combining the transition
system model of a plant with the control software, a model checker
like CBMC \cite{kroening2014cbmc} can be used to do a closed loop
symbolic analysis of the SDCS. This can be an alternative to S3CAM-X.

%which will be
%very useful to for analyzing data for potential property violations.

### 2017-02-07 relational.tex
first

--- 2017-02-07 intro.tex
@@
-%\chapter{Relational Modeling for Falsification}
-%\label{chap:rel}
+\newcommand{\pwa}{\rho}
+\newcommand{\map}{f}
+\newcommand{\amap}{f}
+\newcommand{\gm}{\scr{T}}
+\newcommand{\guard}{g}
+\newcommand{\ds}{\scr{D}}
+
+\chapter{Relational Modeling for Falsification}
+\label{chap:rel}
 
@@
 
-\input{figs/segtraj_rel_enriched.tex}
+\input{figs/relational/segtraj_rel_enriched.tex}
 
@@
 
-\inclfig[width=0.75\textwidth]{vdp_cont_traces.png}{Van der
+\inclfig[width=0.75\textwidth]{relational/vdp_cont_traces.png}{Van der
 Pol: continuous trajectories. Red and green boxes indicate unsafe and
@@
 
-\inclfig[width=0.75\textwidth]{vdp_abs.png}{The discovered
+\inclfig[width=0.75\textwidth]{relational/vdp_abs.png}{The discovered
 abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
@@
 
-\inclfig[width=\textwidth]{vdp_disc_map.png}{Cells and
+\inclfig[width=\textwidth]{relational/vdp_disc_map.png}{Cells and
 trajectory segments used by $1$-relational modeling.}
@@
 
-\inclfig[width=0.40\textwidth]{vdp_graph.png}{Enriched
+\inclfig[width=0.40\textwidth]{relational/vdp_graph.png}{Enriched
 graph $G^R$. The affine maps $f$ for the transition relations are show

### 2017-02-07 sample.tex
first

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%2345678901234567890123456789012345678901234567890123456789012345678901234567890
%        1         2         3         4         5         6         7         8

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed

\title{\LARGE \bf
Preparation of Papers for IEEE Sponsored Conferences \& Symposia*
}

\author{Albert Author$^{1}$ and Bernard D. Researcher$^{2}$% <-this % stops a space
\thanks{*This work was not supported by any organization}% <-this % stops a space
\thanks{$^{1}$Albert Author is with Faculty of Electrical Engineering, Mathematics and Computer Science,
        University of Twente, 7500 AE Enschede, The Netherlands
        {\tt\small albert.author@papercept.net}}%
\thanks{$^{2}$Bernard D. Researcheris with the Department of Electrical Engineering, Wright State University,
        Dayton, OH 45435, USA
        {\tt\small b.d.researcher@ieee.org}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

\input{abstract}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

This template provides authors with most of the formatting specifications needed for preparing electronic versions of their papers. All standard paper components have been specified for three reasons: (1) ease of use when formatting individual papers, (2) automatic compliance to electronic requirements that facilitate the concurrent or later production of electronic products, and (3) conformity of style throughout a conference proceedings. Margins, column widths, line spacing, and type styles are built-in; examples of the type styles are provided throughout this document and are identified in italic type, within parentheses, following the example. Some components, such as multi-leveled equations, graphics, and tables are not prescribed, although the various table text styles are provided. The formatter will need to create these components, incorporating the applicable criteria that follow.

\section{PROCEDURE FOR PAPER SUBMISSION}

\subsection{Selecting a Template (Heading 2)}

First, confirm that you have the correct template for your paper size. This template has been tailored for output on the US-letter paper size. 
It may be used for A4 paper size if the paper size setting is suitably modified.

\subsection{Maintaining the Integrity of the Specifications}

The template is used to format your paper and style the text. All margins, column widths, line spaces, and text fonts are prescribed; please do not alter them. You may note peculiarities. For example, the head margin in this template measures proportionately more than is customary. This measurement and others are deliberate, using specifications that anticipate your paper as one part of the entire proceedings, and not as an independent document. Please do not revise any of the current designations

\section{MATH}

Before you begin to format your paper, first write and save the content as a separate text file. Keep your text and graphic files separate until after the text has been formatted and styled. Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph. Do not add any kind of pagination anywhere in the paper. Do not number text heads-the template will do that for you.

Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:

\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract. Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined. Do not use abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}

\begin{itemize}

\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as Ò3.5-inch disk driveÓ.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ÒWb/m2Ó or Òwebers per square meterÓ, not Òwebers/m2Ó.  Spell out units when they appear in text: Ò. . . a few henriesÓ, not Ò. . . a few HÓ.
\item Use a zero before decimal points: Ò0.25Ó, not Ò.25Ó. Use Òcm3Ó, not ÒccÓ. (bullet list)

\end{itemize}

\subsection{Equations}

The equations are an exception to the prescribed specifications of this template. You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font). To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled. Number equations consecutively. Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop. To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents. Italicize Roman symbols for quantities and variables, but not Greek symbols. Use a long dash rather than a hyphen for a minus sign. Punctuate equations with commas or periods when they are part of a sentence, as in

$$
\alpha + \beta = \chi \eqno{(1)}
$$

Note that the equation is centered using a center tab stop. Be sure that the symbols in your equation have been defined before or immediately following the equation. Use Ò(1)Ó, not ÒEq. (1)Ó or Òequation (1)Ó, except at the beginning of a sentence: ÒEquation (1) is . . .Ó

\subsection{Some Common Mistakes}
\begin{itemize}

\item The word ÒdataÓ is plural, not singular.
\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ÒoÓ.
\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ÒinsetÓ, not an ÒinsertÓ. The word alternatively is preferred to the word ÒalternatelyÓ (unless you really mean something that alternates).
\item Do not use the word ÒessentiallyÓ to mean ÒapproximatelyÓ or ÒeffectivelyÓ.
\item In your paper title, if the words Òthat usesÓ can accurately replace the word ÒusingÓ, capitalize the ÒuÓ; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ÒaffectÓ and ÒeffectÓ, ÒcomplementÓ and ÒcomplimentÓ, ÒdiscreetÓ and ÒdiscreteÓ, ÒprincipalÓ and ÒprincipleÓ.
\item Do not confuse ÒimplyÓ and ÒinferÓ.
\item The prefix ÒnonÓ is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ÒetÓ in the Latin abbreviation Òet al.Ó.
\item The abbreviation Òi.e.Ó means Òthat isÓ, and the abbreviation Òe.g.Ó means Òfor exampleÓ.

\end{itemize}

\section{USING THE TEMPLATE}

Use this sample document as your LaTeX source file to create your document. Save this file as {\bf root.tex}. You have to make sure to use the cls file that came with this distribution. If you use a different style file, you cannot expect to get required margins. Note also that when you are creating your out PDF file, the source file is only part of the equation. {\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size. Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output. }

It is impossible to account for all possible situation, one would encounter using \TeX. If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.

\subsection{Headings, etc}

Text heads organize the topics on a relational, hierarchical basis. For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic. If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced. Styles named ÒHeading 1Ó, ÒHeading 2Ó, ÒHeading 3Ó, and ÒHeading 4Ó are prescribed.

\subsection{Figures and Tables}

Positioning Figures and Tables: Place figures and tables at the top and bottom of columns. Avoid placing them in the middle of columns. Large figures and tables may span across both columns. Figure captions should be below the figures; table heads should appear above the tables. Insert figures and tables after they are cited in the text. Use the abbreviation ÒFig. 1Ó, even at the beginning of a sentence.

\begin{table}[h]
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ. If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ. Do not label axes with a ratio of quantities and units. For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó

\section{CONCLUSIONS}

A conclusion section is not required. Although a conclusion may review the main points of the paper, do not replicate the abstract as the conclusion. A conclusion might elaborate on the importance of the work or suggest applications and extensions.

\addtolength{\textheight}{-12cm}   % This command serves to balance the column lengths
                                  % on the last page of the document manually. It shortens
                                  % the textheight of the last page by a suitable amount.
                                  % This command does not take effect until the next page
                                  % so it should come on the page before the last. Make
                                  % sure that you do not shorten the textheight too much.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{APPENDIX}

Appendixes should appear before the acknowledgment.

\section*{ACKNOWLEDGMENT}

The preferred spelling of the word ÒacknowledgmentÓ in America is without an ÒeÓ after the ÒgÓ. Avoid the stilted expression, ÒOne of us (R. B. G.) thanks . . .Ó  Instead, try ÒR. B. G. thanksÓ. Put sponsor acknowledgments in the unnumbered footnote on the first page.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

References are important to the reader; therefore, each citation must be complete and correct. If at all possible, references should be commonly available publications.

\begin{thebibliography}{99}

\bibitem{c1} G. O. Young, ÒSynthetic structure of industrial plastics (Book style with paper title and editor),Ó 	in Plastics, 2nd ed. vol. 3, J. Peters, Ed.  New York: McGraw-Hill, 1964, pp. 15Ð64.
\bibitem{c2} W.-K. Chen, Linear Networks and Systems (Book style).	Belmont, CA: Wadsworth, 1993, pp. 123Ð135.
\bibitem{c3} H. Poor, An Introduction to Signal Detection and Estimation.   New York: Springer-Verlag, 1985, ch. 4.
\bibitem{c4} B. Smith, ÒAn approach to graphs of linear forms (Unpublished work style),Ó unpublished.
\bibitem{c5} E. H. Miller, ÒA note on reflector arrays (Periodical styleÑAccepted for publication),Ó IEEE Trans. Antennas Propagat., to be publised.
\bibitem{c6} J. Wang, ÒFundamentals of erbium-doped fiber amplifiers arrays (Periodical styleÑSubmitted for publication),Ó IEEE J. Quantum Electron., submitted for publication.
\bibitem{c7} C. J. Kaufman, Rocky Mountain Research Lab., Boulder, CO, private communication, May 1995.
\bibitem{c8} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ÒElectron spectroscopy studies on magneto-optical media and plastic substrate interfaces(Translation Journals style),Ó IEEE Transl. J. Magn.Jpn., vol. 2, Aug. 1987, pp. 740Ð741 [Dig. 9th Annu. Conf. Magnetics Japan, 1982, p. 301].
\bibitem{c9} M. Young, The Techincal Writers Handbook.  Mill Valley, CA: University Science, 1989.
\bibitem{c10} J. U. Duncombe, ÒInfrared navigationÑPart I: An assessment of feasibility (Periodical style),Ó IEEE Trans. Electron Devices, vol. ED-11, pp. 34Ð39, Jan. 1959.
\bibitem{c11} S. Chen, B. Mulgrew, and P. M. Grant, ÒA clustering technique for digital communications channel equalization using radial basis function networks,Ó IEEE Trans. Neural Networks, vol. 4, pp. 570Ð578, July 1993.
\bibitem{c12} R. W. Lucky, ÒAutomatic equalization for digital communication,Ó Bell Syst. Tech. J., vol. 44, no. 4, pp. 547Ð588, Apr. 1965.
\bibitem{c13} S. P. Bingulac, ÒOn the compatibility of adaptive controllers (Published Conference Proceedings style),Ó in Proc. 4th Annu. Allerton Conf. Circuits and Systems Theory, New York, 1994, pp. 8Ð16.
\bibitem{c14} G. R. Faulhaber, ÒDesign of service systems with priority reservation,Ó in Conf. Rec. 1995 IEEE Int. Conf. Communications, pp. 3Ð8.
\bibitem{c15} W. D. Doyle, ÒMagnetization reversal in films with biaxial anisotropy,Ó in 1987 Proc. INTERMAG Conf., pp. 2.2-1Ð2.2-6.
\bibitem{c16} G. W. Juette and L. E. Zeffanella, ÒRadio noise currents n short sections on bundle conductors (Presented Conference Paper style),Ó presented at the IEEE Summer power Meeting, Dallas, TX, June 22Ð27, 1990, Paper 90 SM 690-0 PWRS.
\bibitem{c17} J. G. Kreifeldt, ÒAn analysis of surface-detected EMG as an amplitude-modulated noise,Ó presented at the 1989 Int. Conf. Medicine and Biological Engineering, Chicago, IL.
\bibitem{c18} J. Williams, ÒNarrow-band analyzer (Thesis or Dissertation style),Ó Ph.D. dissertation, Dept. Elect. Eng., Harvard Univ., Cambridge, MA, 1993. 
\bibitem{c19} N. Kawasaki, ÒParametric study of thermal and chemical nonequilibrium nozzle flow,Ó M.S. thesis, Dept. Electron. Eng., Osaka Univ., Osaka, Japan, 1993.
\bibitem{c20} J. P. Wilkinson, ÒNonlinear resonant circuit devices (Patent style),Ó U.S. Patent 3 624 12, July 16, 1990.

\end{thebibliography}

\end{document}

### 2017-02-07 intro.tex
adjusted sizes

@@
 
-\begin{scope}[scale=1]
+\begin{scope}[scale=0.8]
 \draw [line] (-1.0,0) circle (0.5);
@@
 \begin{tikzpicture}
-\begin{scope}[scale=1]
+\begin{scope}[scale=0.5]
 
@@
 \end{scope}
-\begin{scope}[xshift=0,yshift=3.0cm,scale=1]
+\begin{scope}[xshift=0,yshift=3.0cm,scale=0.5]
     \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
@@
 \end{scope}
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (2,0) {\begin{minipage}{\textwidth}
+\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
+    \node at (2,0) {\begin{minipage}{\linewidth}
     \footnotesize \begin{equation*}
@@
 \begin{tikzpicture}
-\begin{scope}
+\begin{scope}[scale=0.5]
 
@@
 
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (4,0) {\begin{minipage}{\textwidth}
+\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
+    \node at (4,0) {\begin{minipage}{\linewidth}
         \footnotesize \begin{equation*}
@@
 
-\inclfig[width=0.75\textwidth]{vdp_cont_traces.png}{Van der
+\inclfig[width=0.75\linewidth]{vdp_cont_traces.png}{Van der
 Pol: continuous trajectories. Red and green boxes indicate unsafe and
@@
 
-\inclfig[width=0.75\textwidth]{vdp_abs.png}{The discovered
+\inclfig[width=0.75\linewidth]{vdp_abs.png}{The discovered
 abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
@@
 
-\inclfig[width=\textwidth]{vdp_disc_map.png}{Cells and
+\inclfig[width=\linewidth]{vdp_disc_map.png}{Cells and
 trajectory segments used by $1$-relational modeling.}
@@
 
-\inclfig[width=0.40\textwidth]{vdp_graph.png}{Enriched
+\inclfig[width=0.40\linewidth]{vdp_graph.png}{Enriched
 graph $G^R$. The affine maps $f$ for the transition relations are show

### 2017-02-07 intro.tex
updated

@@
-%\chapter{Relational Modeling for Falsification}
-%\label{chap:rel}
-
-In \chapref{abs}, we analyzed systems by restricting ourselves to
-their black box semantics. This enabled us to reason about the
-state-space reachability of the system without a direct analysis of
-its structure. Using a coarse abstraction which we searched
-on-the-fly, we could observe the local dynamics as required. This gave
-us an efficient procedure to find abstract counter-examples.  However,
-to concretize the counter-examples, a `closer look' or refinement of
-the abstraction is required. The grid based state-space abstraction
-was refined by splitting the relevant cells (abstract states) into
-smaller ones. As noted previously, due to the curse of dimensionality,
-such a uniform splitting is an expensive operation, and can not scale to
-higher dimensions.
-
-In this chapter, we further our exploration of trajectory segment
-based methods; and explore an alternative approach to overcome the
-explosion in abstract states. Instead of selectively refining the
-abstraction, we compute a model of the black box system and use
-bounded model checking to find a concrete counter-example in the
-model~\footnote{Due to modeling errors, this might not be reproducible in
-the original black-box system.}. We then use this counter-example to
-guide the search towards a counter-example in the original system.
+Model based design has been quite successful in the industry. 
+Tools like \SIMULINK are often used to create models describing hybrid
+dynamical systems. Testing such systems for safety properties remains
+a hard problem, and the current approaches consist of either guided
+testing using numerical simulations or translating to hybrid automata.
+We now suggest an approach which lies in the middle.
+
+In this work we propose a two step process for searching violations of
+the properties. This involves (a) learning a piece-wise affine model
+and (b) exhaustively searching it for violations. It must be noted
+that the violation found in the model might not be valid in the
+original system, depending on the accuracy of the model.
+% concretizing the violation in the original system.
+
+The piece-wise affine model is the aggregation of affine maps, each
+approximating observed local behavior of the system.  The model is
+discrete time and `relational'. It relates the states reachable in a
+fixed time step $\Delta$ to a given state.
+
+In our previous work~\cite{zutshi2014multiple}, we analyzed systems by
+restricting ourselves to their black box semantics. This enabled us to
+reason about the state-space reachability of the system without a
+direct analysis of its structure. Using a coarse abstraction which we
+searched on-the-fly, we could observe the local dynamics as required.
+This gave us an efficient procedure to find abstract counter-examples.
+However, to concretize the counter-examples, a `closer look' or
+refinement of the abstraction is required. The grid based state-space
+abstraction was refined by splitting the relevant cells (abstract
+states) into smaller ones. As noted previously, due to the curse of
+dimensionality, such a uniform splitting is an expensive operation,
+and can not scale to higher dimensions.
+
+We further our exploration of trajectory segment based methods; and
+explore an alternative approach to overcome the explosion in abstract
+states. Instead of selectively refining the abstraction, we compute a
+model of the black box system and use bounded model checking to find a
+concrete counter-example in the model~\footnote{Due to modeling
+errors, this might not be reproducible in the original black-box
+system.}. We then use this counter-example to guide the search towards
+a counter-example in the original system.
 

### 2017-02-08 intro.tex
.

@@
 model of the black box system and use bounded model checking to find a
-concrete counter-example in the model~\footnote{Due to modeling
-errors, this might not be reproducible in the original black-box
-system.}. We then use this counter-example to guide the search towards
-a counter-example in the original system.
+concrete counter-example in the model. Due to modeling errors, this
+might not be reproducible in the original black-box system. We then
+use this counter-example to guide the search towards a counter-example
+in the original system.
 
@@
 
-Towards the end of the chapter, we discuss extensions to data driven
-approaches (instead of simulator driven), where instead of a
-simulator, a fixed set of data is provided as the behavioral
-description of the system. Using a combination of relational modeling,
-and program analysis we outline the future work for falsifying
-properties of SDCS.
-
-\section{Overview}
-\label{sec:ovw}
-
-We now briefly describe the approach. Given a numerical simulator
-$\simulate$ and an initial abstraction defined by a quantization
-function $\quant_\epsilon$ and a time step $\Delta$, we use
-scatter-and-simulate (parameterized by number of samples $n$ as
-described in \chapref{abs}) to explore the abstract graph
-$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
-number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
-\areach{\Delta} C'$.
-
-Instead of using a CEGAR like loop (as in \chapref{abs}), we use the
-generated trajectory segments to learn quantitative models describing
-the local behavior of the system. These models are defined by a set of
-relations $R \subseteq \HybridStates \times \HybridStates$ for each
-edge of the reachability graph.
-
-\input{figs/segtraj_rel_enriched.tex}
-
-Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
-trajectory segment between the respective cells. The graph abstraction
-$G$ only states that there exists a state $\x \in C$ from which the
-system can evolve to a future state $\x' \in C'$. To increase the
-precision we iteratively refined the abstraction by state
-splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
-set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
-which non-deterministically describe relations between $\x \in C$ and $\x'
-\in C'$. This is illustrated in \figref{enriched-edge} (compare with
-\figref{segtraj}).
-
-The enriched graph $G^R$ captures the underlying local forward
-dynamics describing the evolution of the system in each abstract
-state. We represent the dynamics using an affine model with an
-interval error. Such a model can either be approximated using learning
-methods or computed as a sound (over) approximation using reachability
-set computation methods. Because we assume black box semantics, we
-only present the former. Using regression analysis on the $start$ and
-$end$ states of the witnessed trajectory segments between two cells,
-we compute an approximate discrete map along with an error estimate.
-Moreover, using the simulation function $\simulate$, additional
-trajectory segments (or data) can be generated if required. The data
-can be separated into a training set and testing set to compute the
-map and the error respectively.  The latter case can be explored if
-the symbolic dynamics of the system are known. A tool like
-\flowstar~\cite{chen2013flow} can be used to find the reachable set
-map.
-
-Observe that $G^R$, a directed reachability graph, is rich enough to
-search for concrete behaviors in the system. We call it a time
-parameterized PWA relational abstraction. It can be interpreted as an
-infinite state discrete transition system, and we can use
-off-the-shelf bounded model checkers to find concrete violations of a
-given safety property, and even other temporal properties.  We now
-discuss the background required to present our ideas.
-
-\section{Background}
-
-In this section, we present the basics of relational abstractions and
-fundamentals of regression analysis.
-
-\subsection{Relational Abstraction}
-
-For general hybrid dynamical systems, algorithmic ways for computing
-analytical solutions do not exist. Discretization (along with suitable
-abstractions) is often employed to transform the systems into a
-discrete transition system. Relational abstractions is one such idea,
-which abstracts continuous dynamics by discrete relations using
-appropriate \textit{reachability invariants}. Both time independent
-and time dependent relations have been proposed; the former captures
-all reachable states over all time, whereas, the latter explicitly
-includes time by relating reachable states to time. Thus it
-can prove timing properties whereas, time independent relational
-abstractions can not. In both cases, the resulting abstractions can be
-interpreted as discrete transition systems and can be analyzed using
-model checkers. We briefly discuss these two types of abstractions.
-
-\mypara{Timeless Relational Abstraction}
-
-Relational abstractions for hybrid systems were proposed in
-\cite{Sankaranarayanan+Tiwari/2011/Relational}. For a hybrid automaton
-model, they can summarize the continuous dynamics of each mode using a
-binary relation over continuous states. The resulting relations are
-timeless (independent) and hence valid for all time as long as
-the mode invariant is satisfied. The relations take the general form
-of $R(\x,\x') \bowtie 0$, where $\bowtie$ represents one of the
-relational operators $=, \ge, \le, <, >$. For example, an abstraction
-that captures the monotonicity with respect to time for the
-differential equation $\dot{x} = 2$ is $x' > x$. The abstraction
-capturing the relation between the set of ODEs: $\dot{x} = 2$,
-$\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
-
-Such relation summaries, discretize the continuous evolution of a
-given system into an infinite state transition system, in which safety
-properties can be verified using $k$-induction, or falsified using
-bounded model checkers. The relationalization procedure involves
-finding suitable invariants in a chosen abstract domain, such as
-affine abstractions, eigen abstractions or box abstractions
-\cite{Sankaranarayanan+Tiwari/2011/Relational}.  For this, different
-techniques like template based invariant generation
-\cite{Gulwani+Tiwari/2008/Constraint,
-Colon+Sankaranarayanan+Sipma/03/Linear} can be used.
-
-\mypara{Timed and Time-Aware Relational Abstractions}
-
-As the above discussed relations are timeless, we cannot reason over
-the timing properties of the original system. Moreover, they are not
-suitable for time-triggered systems. Timed relational
-abstractions~\cite{zutshi2012timed} and time-aware relational
-abstractions~\cite{mover2013time} were proposed to overcome these
-shortcomings by explicitly including time in the relations. In
-addition, both can be more precise than their timeless counterpart,
-but unlike it, assume continuous affine dynamics.
-
-To analyze time-triggered systems like an SDCS modeled by an affine
-hybrid automata~\footnote{An affine hybrid automata is restricted to
-the ODEs, resets and guards being affine.}, timed relational
-abstractions can be computed directly from the solutions of the
-underlying affine ODEs.  Recall that the solution of an affine ODE
-$\dot{\x} = A\x$ can be represented using the matrix exponential as
-$\x(t) = e^{tA}\x(0)$. This gives us the timed relation $\x' =
-e^{tA}\x$.  Clearly, the relation is non-linear with respect to time.
-However, for the case of SDCS with a fixed sampling time period, we
-obtain linear relations in $\x$.  We demonstrated the usefulness of
-timed relational abstractions for the case of linear systems
-in \cite{zutshi2012timed}, and now incorporate the idea to discretize
-black box dynamical systems.
-
-Similar to timeless relational abstractions, time-aware relational
-abstractions \cite{mover2013time} construct binary relations between
-the current state $\x$ of the system and any future reachable state
-$\x'$. The difference lies with the latter also constructing relations
-between the current time $t$ and any future time $t'$. This is
-achieved by a case by case analysis of the eigen structure of the
-matrix $A$ (of an affine ODE). Separate abstractions are used for the
-case of linear systems with constant rate, real eigen values and complex eigenvalues.
-
-\subsection{Learning Dynamics using Simple Linear Regression}
-
-\mypara{Regression Analysis}
-
-In statistics, regression is the problem of finding a
-\textit{predictor}, which can suitably predict the relationship
-between the given set of observed input $\x$ and output $\y$ vectors.
-In other words, assuming that $\y$ depends on $\x$, regression
-strategies find either a parameterized or a non-parameterized
-prediction function to explain the dependence. We now discuss simple
-linear regression, which is parametric in nature and searches for an
-affine predictor. It is also called \textit{ordinary least squares}.
-
-\mypara{Ordinary Least Squares (OLS)}
-
-Let the data set be comprised of $N$ input and output pairs $(\x,
-y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
-the case in the current context of \textit{finding the best fit}, the
-problem is over-determined; there are more equations than
-unknowns. Hence, a single affine function cannot be found which
-satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This is formally
-defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine the vector of
-coefficients $\vec{a}$ and an offset constant $b$, such that the least
-square error of the affine predictor is minimized for the given data
-set.
-\begin{equation}
-    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
-\label{reg}
-\end{equation}
-To ease the presentation, we can rewrite the above as a homogeneous
-expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
-$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
-\[
-    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
-\]
-The solution of OLS can be analytically computed as
-\[ Ab = (X^TX)^{-1}X^T\y\]
-where $X$ is the matrix representing the horizontal stacking of all
-$\hat{\x}$.
-The details can be found in several texts on learning and statistics
-\cite{friedman2001elements}.
-
-Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
-where $\x\in\reals^n$ we can use OLS to approximate its trajectories
-at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
-$A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
-\begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
-    global relational model for the system. The data set for the
-    regression is a set of pairs $\setof{(\x,\x') | \x' =
-    \simulate(\x, \Delta)}$ and can be generated on demand. Another
-    set can be generated to compute the error $\delta$ as an interval
-    of vectors, where each element is an interval $\delta_i \in
-    [\delta_{min}, \delta_{max}]$.
-
-    Affine maps are poor
-    approximations for arbitrary non-linear functions. Hence, we use a
-    collection of affine maps to approximate the local behaviors (in
-    state-space) of the system $\simulate$. This results in a piece
-    wise approximation, as described in the next section.
-
-\subsection{Piecewise Affine (PWA) Transition System}
-
-We define the PWA transition system as a transition system
-(ref.~\secref{DiscSys}) $\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
-where each transition $\tau \in \scrT$, is associated with a
-transition relation of the form
-\[
-    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
-\]
-where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
-$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
-The assertion on the initial states $\Theta$ is also affine.
-
-We can now use a PWA transition system $\rho$ to approximate the
-behavior of a hybrid dynamical system $\System$ defined by a simulator
-$\simulate_{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
-form, we can only approximate general hybrid systems.}. Each
-transition of $\rho$ represents a discrete time step $\Delta$. Let the
-state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
-states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
-predicates $g_\tau$ are given by a conjunction of hyper-planes and
-hence can be represented in matrix form as a polyhedron (defined by
-$m$ constraints) in the state-space
-\[
-g:C\x - \vd \le 0
-\]
-where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
-$m$. The affine maps $f_\tau$ can be represented in matrix form as
-$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
-an error term $\delta$ which is defined by a vector of intervals
-$[\delta_{\min},\delta_{\max}]$. Hence,
-\[
-f:\x \mapsto A\x + \vb + \delta
-\]
-defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
-\in f_\tau(\x)}$.
-
-% Given a state vector $\x\in\reals^n$, a guarded affine map
-% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
-% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
-% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
-% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
-% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
-% guard is satisfied.
-
-We now formalize the PWA affine transition system for a dynamical system.
-\begin{definition}[PWA Transition System]
-
-Given a hybrid dynamical system over a state-space
-$\HybridStates^n$, a PWA transition system $\rho$ is given by the
-tuple $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in
-\scrT$ are affine transition relations and $\Theta$ is an affine
-predicate over $\x$ and the states $\x\in\HybridStates^n$. The
-transition relation is then defined by $\scrT$ with $n$ transition
-relations as follows
-
-\begin{equation}
-    \scrT = \left\{
-        \begin{array}{ll}
-            g_1(\x) \land g'_1(\x') \implies \x' \in f_1(\x) \\
-            \ldots\\
-            g_n(\x) \land g'_n(\x') \implies \x' \in f_n(\x) \\
-        \end{array}
-    \right.
-\end{equation}
-
-\end{definition}
-
-where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
-$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
-denotes the next state of the system.
-
-A PWA model is \textit{deterministic}, iff for every state
-$\x\in\HybridStates^n$, a unique guarded affine map is satisfied and
-all errors $\delta$ are singletons. However, in practice such a case
-rarely exists. A PWA model will be usually non-deterministic, both due
-to multiple choice of transition relations and the reachable states at
-the $k^{th}$ time step (or after $k\Delta$ units of time) being a set.
-Abusing the notation, we denote the set of states reachable in a
-single step in $\rho$ by $\x' = \rho(\x)$.
-
-A PWA transition system is \textit{complete}, iff for every state
-$\x\in\HybridStates^n$, there exists at least one satisfied
-transition relation $\scrT$. If this is not the case, the system can
-deadlock, with no further executions. This usually results from
-incorrect modeling of SDCS, and from here on, we do not consider such
-cases.
-
-% Given a dynamical system over a continuous state-space
-% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
-% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
-% dynamical system to a finite set of guarded affine maps
-% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
-% rule by the affine map of the satisfied $\gm$.
-
-% \begin{equation}
-%     \pwa = \left\{
-%         \begin{array}{ll}
-%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
-%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
-%             \ldots & \ldots\\
-%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
-%         \end{array}
-%     \right.
-% \end{equation}
-
-% \end{definition}
-
-% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
-% Abusing the notation, we denote the next state computed by the PWA
-% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
-% for every state $\x\in\ContStates$, a unique guarded affine map is
-% satisfied. A PWA model is \textit{complete}, iff for every state
-% $\x\in\ContStates$, there exists at least one satisfied guarded affine
-% map $\gm$.
+% Towards the end of the chapter, we discuss extensions to data driven
+% approaches (instead of simulator driven), where instead of a
+% simulator, a fixed set of data is provided as the behavioral
+% description of the system. Using a combination of relational modeling,
+% and program analysis we outline the future work for falsifying
+% properties of SDCS.
 

### 2017-02-08 prelims.tex
.

--- 2017-02-07 intro.tex
@@
-%\chapter{Relational Modeling for Falsification}
-%\label{chap:rel}
-
-In \chapref{abs}, we analyzed systems by restricting ourselves to
-their black box semantics. This enabled us to reason about the
-state-space reachability of the system without a direct analysis of
-its structure. Using a coarse abstraction which we searched
-on-the-fly, we could observe the local dynamics as required. This gave
-us an efficient procedure to find abstract counter-examples.  However,
-to concretize the counter-examples, a `closer look' or refinement of
-the abstraction is required. The grid based state-space abstraction
-was refined by splitting the relevant cells (abstract states) into
-smaller ones. As noted previously, due to the curse of dimensionality,
-such a uniform splitting is an expensive operation, and can not scale to
-higher dimensions.
-
-In this chapter, we further our exploration of trajectory segment
-based methods; and explore an alternative approach to overcome the
-explosion in abstract states. Instead of selectively refining the
-abstraction, we compute a model of the black box system and use
-bounded model checking to find a concrete counter-example in the
-model~\footnote{Due to modeling errors, this might not be reproducible in
-the original black-box system.}. We then use this counter-example to
-guide the search towards a counter-example in the original system.
-
-More specifically, we use regression to quantitatively estimate the
-discovered relations, which were witnessed by trajectory segments, by
-affine maps. These maps approximate locally observed behaviors, which
-are incorporated into the sampled reachability graph as edge
-annotations.  The resulting graph or the Piece-Wise Affine (PWA)
-relational model can be interpreted as an infinite state discrete
-transition system and model checked for time bounded safety
-properties. A counter-example if found, can indicate the presence of a
-violation in the system.
-%Furthermore, using linear programming, we can over-approximate the
-%neighbhourhood of the counter-example in the model.
-
-Towards the end of the chapter, we discuss extensions to data driven
-approaches (instead of simulator driven), where instead of a
-simulator, a fixed set of data is provided as the behavioral
-description of the system. Using a combination of relational modeling,
-and program analysis we outline the future work for falsifying
-properties of SDCS.
-
-\section{Overview}
-\label{sec:ovw}
-
-We now briefly describe the approach. Given a numerical simulator
-$\simulate$ and an initial abstraction defined by a quantization
-function $\quant_\epsilon$ and a time step $\Delta$, we use
-scatter-and-simulate (parameterized by number of samples $n$ as
-described in \chapref{abs}) to explore the abstract graph
-$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
-number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
-\areach{\Delta} C'$.
-
-Instead of using a CEGAR like loop (as in \chapref{abs}), we use the
-generated trajectory segments to learn quantitative models describing
-the local behavior of the system. These models are defined by a set of
-relations $R \subseteq \HybridStates \times \HybridStates$ for each
-edge of the reachability graph.
-
-\input{figs/segtraj_rel_enriched.tex}
-
-Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
-trajectory segment between the respective cells. The graph abstraction
-$G$ only states that there exists a state $\x \in C$ from which the
-system can evolve to a future state $\x' \in C'$. To increase the
-precision we iteratively refined the abstraction by state
-splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
-set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
-which non-deterministically describe relations between $\x \in C$ and $\x'
-\in C'$. This is illustrated in \figref{enriched-edge} (compare with
-\figref{segtraj}).
-
-The enriched graph $G^R$ captures the underlying local forward
-dynamics describing the evolution of the system in each abstract
-state. We represent the dynamics using an affine model with an
-interval error. Such a model can either be approximated using learning
-methods or computed as a sound (over) approximation using reachability
-set computation methods. Because we assume black box semantics, we
-only present the former. Using regression analysis on the $start$ and
-$end$ states of the witnessed trajectory segments between two cells,
-we compute an approximate discrete map along with an error estimate.
-Moreover, using the simulation function $\simulate$, additional
-trajectory segments (or data) can be generated if required. The data
-can be separated into a training set and testing set to compute the
-map and the error respectively.  The latter case can be explored if
-the symbolic dynamics of the system are known. A tool like
-\flowstar~\cite{chen2013flow} can be used to find the reachable set
-map.
-
-Observe that $G^R$, a directed reachability graph, is rich enough to
-search for concrete behaviors in the system. We call it a time
-parameterized PWA relational abstraction. It can be interpreted as an
-infinite state discrete transition system, and we can use
-off-the-shelf bounded model checkers to find concrete violations of a
-given safety property, and even other temporal properties.  We now
-discuss the background required to present our ideas.
-
-\section{Background}
-
 In this section, we present the basics of relational abstractions and
@@
 % map $\gm$.
-
-\section{Relational Modeling}
-
-We now describe how relational models can be computed for a given
-black box system by enriching the abstract reachability graph obtained
-by scatter-and-simulate. We first formalize the notion of an enriched
-graph, and then show how they can be interpreted as a PWA transition
-system using $k$-relational modeling.
-
-\subsection{Abstract Enriched Graph}
-
-The existential abstraction relation in \chapref{abs} was defined as
-follows:
-\[
-%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
-    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
-\]
-The abstract relation $\areach{t}$ can be enriched by incorporating
-the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
-system, such a relation can be rarely represented using an exact affine map.
-This is due to the presence of non-linear and hybrid behaviors. But,
-an affine map can always be estimated with an error.
-
-If the system dynamics are completely specified in the form of a white
-box model, we can use first order approximations to find the affine
-expressions for the relations. Using a tool like \flowstar, we can
-obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
-\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
-that, every element $b_i$ of the vector $\vb$ is contained in the
-respective scalar interval $b_i\in[b^l_i,b^h_i]$.
-
-For the case of black box systems, such a sound approximation is not
-possible. Instead, we rely on a statistical method like simple linear
-regression to estimate the affine map $\amap: A\x + \vb$. We then
-estimate the error $\delta$ and generalize $\amap$ to an interval
-affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
-abstraction with the below relation
-\[
-    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
-    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
-\]
-
-However, in the presence of complex non-linear behavior, using a
-single affine map can lead to a poor approximation. To increase the
-precision, one can use more than one affine map. Let us denote this set as
-$R_{(C,C')}$ (ref.~\secref{ovw}).
-\[
-    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
-\]
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
-    \draw[arw] (-3,0) -- (-1.2,0);
-    \draw[arw] (1.2,0) -- (3,0);
-
-\node at (-3.4,0) {$G$};
-\node at (3.4,0) {$G^R$};
-\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Using OLS, $G^R$ is computed by determining the appropriate
-$R_{(C,C')}$ for each edge of $G$.}
-\label{fig:enrichment}
-\vspace*{-.3cm}
-\end{figure}
-
-As shown in \figref{enrichment}, we use OLS to compute an enrichment
-of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
-the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
-of the enriched edge are clearly non-deterministic. Form a state $\x
-\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
-is satisfied. This interpretation results in an infinite state
-transition system. We now detail the construction of the set
-$R_{(C,C')}$ using $k$-relational modeling.
-
-% We now compute a PWA model for a given black box system by estimating
-% the dynamics using OLS. We can estimate the affine relations between
-% every cell, but as discussed in ~\chapref{abs}, it is a futile
-% approach.  Instead, we use the same heuristic as before:
-% scatter-and-simulate, to select the cell relations for estimation. We
-% build the reachability graph as before, but in addition, annotate each
-% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
-% respective relation. The reachability graph thus computed, is a
-% transition system. Using off-the-shelf bounded model checkers, we
-% reason about the system's safety properties.
-
-\subsection{$k$ - Relational Modeling}
-
-\begin{figure}[!htbp]
-\vspace*{-.2cm}
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-\begin{scope}[scale=1]
-\draw [line] (-1.0,0) circle (0.5);
-\draw [line] (-4.0,0) circle (0.5);
-\draw [line] (2.0,1) circle (0.5);
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-0.5,0) -- (1.5,1.0);
-\draw[arw] (-0.5,0) -- (1.5,-1.0);
-\draw[arw] (-3.5,0) -- (-1.5,0.0);
-\node at (-4.0,0) {$C$};
-\node at (-1.0,0) {$C'$};
-\node at (2.0,1.0) {$C''_1$};
-\node at (2.0,-1.0) {$C''_2$};
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
-\node at (-4.0,-2.5) {$k=0$};
-\node at (-1.0,-2.5) {$k=1$};
-\node at (+2.0,-2.5) {$k=2$};
-\draw [fill=black] (-4.0,-2.0) circle (0.05);
-\draw [fill=black] (-1.0,-2.0) circle (0.05);
-\draw [fill=black] (2.0,-2.0) circle (0.05);
-
-\end{scope}
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
-\label{fig:k-rel}
-\vspace*{-.3cm}
-\end{figure}
-
-Given a $G$, when computing the set of relations for an edge $R$, we
-consider only the set of abstract states reachable at a specific time
-step.  Intuitively, one might consider the states reachable in one
-time step $\Delta$.  However, such a process can be generalized by
-looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
-k\Delta}$ time steps.
-
-Using \figref{k-rel}, we illustrate this notion. To compute
-$R_{(C,C')}$, we
-split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
-$k\Delta$) as follows.  We observe the local behavior of the system by
-noting the evolution of the system $\System$ from a state $\x(t)
-\in C$ for a time length dependant on $k$. For
-\begin{itemize}
-    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
-    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
-    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
-        and split them in two sets (a) and (b) on the basis of the
-        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
-        \begin{enumerate}[(a)]
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
-        \end{enumerate}
-    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
-        and split them in to multiple sets on the basis of the cells
-        reached at time $(t+\Delta),\; (t+2\Delta),\;
-        \ldots,\;(t+n\Delta)$.
-\end{itemize}
-
-$k$-relational modeling can be understood as a heuristic, which uses
-the underlying abstraction to differentiate behaviors of the system
-which `diverge' (or are revealed to be `distinct' at a future time).
-Such a heuristic can be useful in increasing the precision of the
-learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
-$n$ and illustrate using examples.
-
-\mypara{$0$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=1]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
-\node at (0,-2.5) {$t=0$};
-\node at (4,-2.5) {$t=\Delta$};
-\draw [fill=black] (0.0,-2.0) circle (0.05);
-\draw [fill=black] (4.0,-2.0) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=3.0cm,scale=1]
-    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
-    \implies A_1\vx + \vb_1 + \delta_1$};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{All the trajectory segments will be used to construct the
-model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
-\label{fig:k0}
-\vspace*{-.3cm}
-\end{figure}
-
-When $k=0$, the $0$-relational model is a PWA transition system which only
-defines the evolution of states $\x$, and does not specify the
-reachable cell. The guard predicates of the relations are defined over
-cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.
-
-We use regression to estimate the dynamics for the outgoing trajectory
-segments form a cell $C$. Hence, the data set $\ds$ for the regression
-includes all trajectory segments beginning from the same cell
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C}.
-\]
-This includes trajectory segments ending in different cells, as shown
-in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
-\begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            \ldots \\
-            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
-        \end{array}
-    \right.
-\end{equation*}
-
-Note that this can be quite imprecise when the cells are big,
-containing regions of state-space with complex dynamics. This is true
-for both non-linear systems and hybrid dynamical system, where a cell
-can contain two or more modes with differing continuous dynamics.
-
-\mypara{$1$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=1]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (2,0) {\begin{minipage}{\textwidth}
-    \footnotesize \begin{equation*}
-        \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
-    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
-    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
-%Cardinality($R_{(C,C')}$) = $2$.}
-\label{fig:k1}
-\vspace*{-.3cm}
-\end{figure}
-
-% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-%     \node at (2,0) {\footnotesize $\begin{equation*}
-%     \scr{T} = \left\{
-%         \begin{array}{ll}
-%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-%         \end{array}
-%     \right.
-% \end{equation*}$};
-% \end{scope}
-
-To improve the preciseness of learnt dynamics, we include the
-reachability relation in the
-regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
-comprised only of trajectory segments $\pi_t$ which start and end in
-the same set of cells.
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
-\]
-For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
-$N$ transition relations.
-
-\figref{k1} illustrates the $k=1$ refinement of the case shown in
-\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
-= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
-relations are constructed. The resulting $1$-relational model
-is at least as precise as the corresponding $0$-relational model.
-
-\mypara{$k$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
-    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;
-
-%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-
-    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (8.0, 2.0) {$C''_1$};
-\node at (8.0, -2.0) {$C''_2$};
-%\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
-\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\node at (8,-4.0) {$t=2\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-\draw [fill=black] (8.0,-3.5) circle (0.05);
-\end{scope}
-
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (4,0) {\begin{minipage}{\textwidth}
-        \footnotesize \begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The $k=2$ refinement further splits the data set $\ds_1$
-    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
-    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
-    non-deterministically defining the system behavior.}
-\label{fig:k2}
-\vspace*{-.3cm}
-\end{figure}
-
-Finally, we generalize the relational models to $k$-relational PWA
-models. A $k$-relational model is constructed by using $k$ length
-\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
-\textit{connected} iff its cost $\cost(\SegTraj) = 0$.
-
-Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
-$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
-\textit{similar} if their trajectory segments have the same sequence
-of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
-\[
-    \forall i\in\setof{1 \ldots k}.
-    start(\pi_{t_i}) \in C_i
-    \land start(\pi'_{t_i}) \in C_i
-    \land end(\pi_{t_i}) \in C_{i+1}
-    \land end(\pi'_{t_i}) \in C_{i+1}
-\]
-
-For every cell $C$ in the reachability graph, we construct a data set
-$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
-length \textit{connected} and \textit{similar} segmented trajectories.
-%\[
-%    \ds = \setof{\pi_{\tau_i} |
-%        \pi_{\tau_i}
-%        start(\pi_{\tau_{i}}) \in C_i
-%        \land
-%        end(\pi_{\tau_{i}}) \in C_i}
-%\]
-
-\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.
-
-\section{Bounded Model Checking Black Box Systems}
-% include an algorithm
-We now describe the entire procedure in three steps.
-\begin{enumerate}
-    \item Given a black box system $\System$ specified by
-        $\simulate_\System$, we
-        use scatter-and-simulate with ($\Delta$-length) trajectory
-        segments to sample a finite reachability graph
-        $\scr{H}(\Delta)$.
-    \item Using regression, we enrich the graph relations and annotate
-        the edges with the estimated affine maps.
-    \item Interpreting the directed reachability graph as a transition
-        system, we use a bounded model checker to find fixed length
-        violations to safety properties.
-\end{enumerate}
-
-\section{An Example: Van der Pol Oscillator}
-
-We now describe the above using the Van der Pol oscillator benchmark
-presented in \secref{math-dyn-sys}. Simulations generated using
-uniformly random samples are shown in \figref{vdp-cont}. We want to
-check the property $P_3$, indicated by the red box, given the initial
-set indicated by the green box. The abstraction is defined by
-$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
-where each cell is of size $0.2 \times 0.2$ units.
-Scatter-and-simulate is then used to construct the abstract graph $G$,
-using $2$ samples per cell and the time step $\Delta = 0.1s$. The
-complete process follows.
-
-\begin{enumerate}
-    \item{\emph{Abstract}} Consider an implicit abstraction induced by the quantization
-    function $\quant_\epsilon(\x)$. The corresponding reachability
-    graph $\scr{H}(\Delta)$ (with time step $\Delta$) is given by
-    $\tupleof{\scr{C},\areach{\Delta}, \scr{C}_0, \scr{C}_u}$.
-\item{\emph{Discover}}: Using scatter-and-simulate, enumerate a finite number
-    of cells (vertices) and the relations (edges) of the graph
-    $\scr{H}(\Delta)$. Associate the set of generated trajectory
-    segments with their respective originating cells using a map $D:C
-    \mapsto \setof{\pi | start(\pi) \in C}$. The discovered
-    abstraction is show in \figref{vdp-abs}. As mentioned, red cells
-    are unsafe and green cells are initial cells.
-\item{\emph{Relationalize}} For each cell $C$, perform regression analysis on
-    the respective set of trajectory segments $D(C)$, and compute a
-    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
-    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(\Delta)$. Annotate
-    each edge in the graph $\scr{H}(\Delta)$ with the respective
-    relation.
-
-    \figref{vdp-map} shows the cells and the trajectory segments which
-    are part of the data sets constructed using the $1$-relational
-    modeling. Against each cell, its unique identifier (integer
-    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
-    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
-    relations. Note that self loops result when an observed trajectory
-    segment has its $start$ and $end$ states in the same cell.
-
-\item{\emph{Model Check}} The graph $\scr{H}(\Delta)$ can now be viewed as a
-    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
-    where $C \in vertices(\scr{H}(\Delta))$ and $\x \in
-    \HybridStates$. A transition $\tau \in \scrT$ is of the form
-    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
-    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
-    edge(\scr{H}(\Delta))}$.
-\item{\emph{Check Counter-example}} The infinite state (but finite location) transition system can
-    be model checked to find a concrete counter-example, which if
-    exists, can indicate the existence of a similar trace in the
-    original black-box system $\System$. The latter check is carried
-    out as before, using the numerical simulation function
-    $\simulate$. For the given example, the model checker is unable to
-    find a counter-example.
-\end{enumerate}
-
-\inclfig[width=0.75\textwidth]{vdp_cont_traces.png}{Van der
-Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}
-\label{fig:vdp-cont}
-
-\inclfig[width=0.75\textwidth]{vdp_abs.png}{The discovered
-abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
-initial cells.}
-\label{fig:vdp-abs}
-
-\inclfig[width=\textwidth]{vdp_disc_map.png}{Cells and
-trajectory segments used by $1$-relational modeling.}
-\label{fig:vdp-map}
-
-\inclfig[width=0.40\textwidth]{vdp_graph.png}{Enriched
-graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}
-\label{fig:vdp-graph}
-
-\begin{table*}[!htbp]
-\centering
-\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
-\label{tab:vdp-pwa}
-\begin{tabular}{@{}ccccc@{}}
-\toprule
-$C$ & $C'$ & $A$ & $b$ & $\delta$\\
-\midrule
-(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
-\midrule
-(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
-\midrule
-(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% (0, -1)   & (0, -1)   &  \begin{bmatrix}0.9944498 & 0.12911959 \\-0.12013468&1.63250531\end{bmatrix} &\begin{bmatrix}-0.00003683\\-0.00085411\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -1)   & (0, -2)   &  \begin{bmatrix}0.99523928& 0.1290447  \\-0.10236736&1.6314103 \end{bmatrix} &\begin{bmatrix}-0.00013019\\-0.00289192\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -2)   & (0, -2)   &  \begin{bmatrix}0.99565348& 0.12916002 \\-0.09352462&1.6341424 \end{bmatrix} &\begin{bmatrix}-0.00015574\\-0.00339664\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -2)   & (0, -4)   &  \begin{bmatrix}0.99668599& 0.12892455 \\-0.0725035 &1.62937613\end{bmatrix} &\begin{bmatrix}-0.00037391\\-0.00796189\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -9)   & (-1, -15) &  \begin{bmatrix}0.99985362& 0.12858788 \\-0.09447884&1.61476417\end{bmatrix} &\begin{bmatrix}-0.0018997 \\-0.04200363\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-1, -15) & (-3, -24) &  \begin{bmatrix}0.95473217& 0.12473489 \\-1.29198505&1.47155614\end{bmatrix} &\begin{bmatrix}-0.01248412\\-0.41100129\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-3, -24) & (-4, -29) &  \begin{bmatrix}1.11789749&-0.17261206 \\-1.71017779&0.66006597\end{bmatrix} &\begin{bmatrix}-1.0745344 \\-3.31246863\end{bmatrix}&\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}\\
-% \midrule
-% (-4, -29) & (-4, -29) &  \begin{bmatrix}0.99796588& 0.01012817 \\-0.41529747&1.02432211\end{bmatrix} &\begin{bmatrix}-0.00135199\\-0.28016983\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-1, -10) & (-2, -16) &  \begin{bmatrix}0.97381953& 0.12768367 \\-0.71242741&1.56764585\end{bmatrix} &\begin{bmatrix}-0.00320265\\-0.11964102\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% (-1, -10) & (-2, -15) &  \begin{bmatrix}0.97373058& 0.12859163 \\-0.70173855&1.58540872\end{bmatrix} &\begin{bmatrix}-0.0015415 \\-0.08616376\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-2, -16) & (-3, -23) &  \begin{bmatrix}0.92850123& 0.12242812 \\-1.84720528&1.39159826\end{bmatrix} &\begin{bmatrix}-0.02334236\\-0.73348865\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -23) & (-5, -29) &  \begin{bmatrix}3.14947585&-0.81150646 \\-0.96159406&0.23543254\end{bmatrix} &\begin{bmatrix}-3.13860326\\-4.99696623\end{bmatrix}&\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}\\
-% \midrule
-% (-2, -15) & (-3, -22) &  \begin{bmatrix}0.93156623& 0.12260264 \\-1.75117219&1.40104348\end{bmatrix} &\begin{bmatrix}-0.02214125\\-0.68423569\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -22) & (-5, -29) &  \begin{bmatrix}3.08806462&-0.72332776 \\-1.69271079&0.40308293\end{bmatrix} &\begin{bmatrix}-2.77908467\\-4.57411339\end{bmatrix}&\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}\\
-
-\subsection{Search Parameters}
-
-The search parameters for S3CAM-R include the parameters of S3CAM:
-$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
-maximum error budget $\delta_{max}$ for OLS and the maximum length of
-segmented trajectory for building $k$-relational models.
-
-We have already discussed the effects of $N$, $\epsilon$ and $\Delta$
-on S3CAM's performance. However, they also have an effect on
-relational modeling. A finer grid with small cells produces more
-accurate models than a coarser grid with large cells. Similarly,
-small time length trajectory segments result in more accurate
-modeling.
-%An increase in the number of samples can increases the
-%accuracy of regression.
-
-\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
-Given a $\delta_{max}$, we keep increasing $k$ during the
-$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
-satisfied. A $k_{max}$ is introduced to bound the longest
-segmented trajectory which can be considered.
-
-% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
-
-\subsection{Reasons for Failure}
-The approach can fail to find a counter-example, even when it exists, in
-one of three ways.
-
-\begin{itemize}
-
-    \item \emph{S3CAM Fails} No abstract counter-example is found by
-        S3CAM. We can remedy it by increasing search budgets and/or
-        restarting.
-
-    \item \emph{BMC Fails} An abstract counter-example is found, but
-        the BMC fails to find a concrete counter-example in the PWA
-        relational model. The failure can be attributed to either (a)
-        a spurious abstract counter-examples or (b) a poorly estimated
-        model. The former can be addressed by restarting but the
-        latter requires that the maximum model error $\delta_{max}$ be
-        decreased.
-
-    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
-        is found, and it is successfully concretized in the PWA
-        relational model by the BMC. However, it is not reproducible
-        in the black box system. This happens when the PWA relational
-        model is not precise enough.
-
-\end{itemize}
-
-\section{Implementation and Evaluation}
-
-The implementation was prototyped as S3CAM-R, an extension to our
-previously mentioned tool S3CAM (\chapref{case}). OLS regression
-routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
-Python module for machine learning. SAL~\cite{SAL-SRI}
-with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
-solver.
-
-We used S3CAM to discover the abstraction graph $G$, which was then
-trimmed of the nodes which did not contribute to the error search
-(nodes from which error nodes were not reachable). In our experiments we
-used $1$-relational modeling to create the transition system from $G$.
-Using SAL, we checked for the given safety property. If a concrete
-trace was obtained from the BMC, it was further checked for validity
-by simulating using $\simulate$ to see if it was indeed an error
-trace.  If not, $100$ samples from its associated abstract state was
-sampled to check the presence of a counter-example in it
-neighborhood. This can be further extended by obtaining different
-discrete sequences or discrete trace sequences from the model checker.
-
-We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-few of the benchmarks described in \chapref{case}, including the Van
-der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
-benchmark. As before, we ran S3CAM-R $10$ times with different seeds
-and averaged the results. We tabulate both the total time taken and
-the time taken by SAL to compute the counter-example and compare
-against a newer implementation of S3CAM. Note that different
-abstraction parameters are used for S3CAM and S3CAM-R, due to the
-differences in which they operate.
-
-\begin{table*}[!htbp]
-\centering
-\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
-time taken by the BMC engine. The total time in seconds (rounded off
-to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
-$TO$ signifies time $>5hr$, after which the search was killed.}
-\label{tab:res-rel}
-\begin{tabular}{@{}llll@{}}
-\toprule
-Benchmark & BMC & S3CAM-R & S3CAM\\
-\midrule
-Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
-Brusselator               &$<$1  & $4$    & $2$\\
-Lorenz                    &$<$1 & $122$  & $35$\\
-%B.Ball                    &  &   & \\
-Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
-Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
-Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
-Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
-Bouncing Ball             &-  &$TO$   & $450$\\
-
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% 1m8s, 2m56.184s, 1m31.059s[f], 1m32.665s,  1m45.312s, 1m22.397s, 2m1.451s[f], 1m12.118s, 1m26.596s
-% 0.12, 0.06,         0.33       0.13,       0.22,     FAIL,       ,X           0.08        0.15
-
-The results show promise, but clearly S3CAM performs better. Also,
-S3CAM-R timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
-and the bouncing ball. However, we need to explore more benchmarks to
-be conclusive. S3CAM-R's performance can be explained by the
-difficulty in finding a good abstraction ($\quant_\epsilon$), which
-needs to be fine enough, to obtain good prediction models but coarse
-enough, to be manageable by the current SMT solvers. Specifically, to
-obtain a good quality counter-example from the model checker (and
-avoid false positives), the transition system should be created from
-models with high accuracy, for which a finer abstraction is required.
-However, the exploration of a finer abstraction is exponentially more
-complex and results in a much bigger transition system, which the
-current state of the art bounded model checkers (which use SMT
-solvers) can not handle under reasonable resources.
-
-\section{Conclusion and Future Work}
-
-We have presented another methodology to find falsifications in black
-box dynamical systems. Combining the ideas from abstraction based
-search (\chapref{abs}), with our previous relational
-abstractions~\cite{zutshi2012timed}, we proposed $k$-relational
-modeling to compute richer abstractions. Simple linear regression can
-be used to estimate the local dynamics of the trajectory segments and
-compute PWA relational models which can be interpreted as a transition
-system.  These can then be checked for safety violations using an
-off-the-shelf bounded model checker.  Finally, we implemented the
-ideas as a tool S3CAM-R and demonstrated its performance on a few examples.
-
-\subsection{Improvements}
-
-The approach seems promising as it avoids an expensive refinement
-step. But, before it can be applied to more complex systems, we need
-to overcome its shortcomings.  As a future extension, we are
-investigating ideas to improve the performance of the underlying three
-main sub-processes, namely, abstraction, modeling, and BMC. We detail
-these below.
-
-\begin{enumerate}
-
-\item \emph{Abstraction} The abstract domain we have used, has been
-    confined to the interval or rectangular domain. Such a domain is
-    very coarse when compared to more general polyhedral domains, using
-    which, we can gain better precision. Due to polyhedral domains
-    being computational expensive, their integration in S3CAM needs to
-    be explored thoroughly.
-
-\item \emph{Modeling} We have used $k$-relational modeling to increase
-    the precision of the enriched abstractions, however, there are
-    several other refinement strategies that remain to be explored.
-    This includes refining the abstract states by using guard
-    predicates (similar to predicate abstraction), using non-linear
-    templates in parametric regression to compute more precise
-    quadratic relations, and using an adaptive time step instead of a
-    fixed one to address the non-linearities due to a long time
-    horizon.
-
-\item \emph{Model Checking} As of now, we use SMT solvers to search
-    for concrete counter-examples. Although they are very efficient,
-    owing to extensive engineering effort, reachability of transition
-    systems resulting from dynamical systems remains a difficult
-    problem. As the time horizon of the safety property increases, the
-    possible combinations of discrete transitions increase
-    exponentially. Hence, to find a counterexample which is a sequence
-    of discrete transitions over a `long' time horizon is not
-    tractable for most, but the simplest of dynamical systems.
-    Instead, we can use linear programming solvers, by enumerating
-    each path in the graph $G^R$. Note that the constraints are all
-    linear (conjunctions) along an abstract path. However, in the
-    worst case, the number of paths in a graph can be of the order
-    $n!$ where $n$ is the number of vertices of a graph. Hence, this
-    will not be feasible, unless, we can prioritize paths by using a
-    triage process similar to one used by S3CAM and using a budget on
-    the maximum number of paths.
-
-    Another approach to address the issue would be to use an adaptive
-    time discretization technique, where relations over both shorter
-    and longer time steps are computed, and the SMT solver can
-    `select' the time step precise enough to find a counter-example.
-
-\item \emph{SMT Solvers} One major impediment to our approach is the
-    fact that SMT solvers use the theory of reals with \emph{exact
-    precision}. This is important for verification approaches, but can
-    be relaxed for the problem of falsification. An SMT solver which
-    uses approximate reasoning but returns robust counter-examples
-    will be as useful, and perhaps more efficient.
-
-\end{enumerate}
-
-\subsection{Data Driven Analysis}
-
-Finally, we would like to mention that our approach is `simulation'
-driven. It can be easily modified to be `data driven', by working with
-a fixed set of data. Given a data set, we then need to automatically
-find a good abstraction and a high fidelity PWA transition system
-using which, we can summarize the black box hybrid dynamical system.
-Apart from model checking the transition system, one can extend it to
-the analysis of SDCS. More specifically, by combining the transition
-system model of a plant with the control software, a model checker
-like CBMC \cite{kroening2014cbmc} can be used to do a closed loop
-symbolic analysis of the SDCS. This can be an alternative to S3CAM-X.
-
-%which will be
-%very useful to for analyzing data for potential property violations.

### 2017-02-08 pwarel.tex
.

--- 2017-02-07 intro.tex
@@
-%\chapter{Relational Modeling for Falsification}
-%\label{chap:rel}
-
-In \chapref{abs}, we analyzed systems by restricting ourselves to
-their black box semantics. This enabled us to reason about the
-state-space reachability of the system without a direct analysis of
-its structure. Using a coarse abstraction which we searched
-on-the-fly, we could observe the local dynamics as required. This gave
-us an efficient procedure to find abstract counter-examples.  However,
-to concretize the counter-examples, a `closer look' or refinement of
-the abstraction is required. The grid based state-space abstraction
-was refined by splitting the relevant cells (abstract states) into
-smaller ones. As noted previously, due to the curse of dimensionality,
-such a uniform splitting is an expensive operation, and can not scale to
-higher dimensions.
-
-In this chapter, we further our exploration of trajectory segment
-based methods; and explore an alternative approach to overcome the
-explosion in abstract states. Instead of selectively refining the
-abstraction, we compute a model of the black box system and use
-bounded model checking to find a concrete counter-example in the
-model~\footnote{Due to modeling errors, this might not be reproducible in
-the original black-box system.}. We then use this counter-example to
-guide the search towards a counter-example in the original system.
-
-More specifically, we use regression to quantitatively estimate the
-discovered relations, which were witnessed by trajectory segments, by
-affine maps. These maps approximate locally observed behaviors, which
-are incorporated into the sampled reachability graph as edge
-annotations.  The resulting graph or the Piece-Wise Affine (PWA)
-relational model can be interpreted as an infinite state discrete
-transition system and model checked for time bounded safety
-properties. A counter-example if found, can indicate the presence of a
-violation in the system.
-%Furthermore, using linear programming, we can over-approximate the
-%neighbhourhood of the counter-example in the model.
-
-Towards the end of the chapter, we discuss extensions to data driven
-approaches (instead of simulator driven), where instead of a
-simulator, a fixed set of data is provided as the behavioral
-description of the system. Using a combination of relational modeling,
-and program analysis we outline the future work for falsifying
-properties of SDCS.
-
-\section{Overview}
+\subsection{Overview}
 \label{sec:ovw}
@@
 discuss the background required to present our ideas.
-
-\section{Background}
-
-In this section, we present the basics of relational abstractions and
-fundamentals of regression analysis.
-
-\subsection{Relational Abstraction}
-
-For general hybrid dynamical systems, algorithmic ways for computing
-analytical solutions do not exist. Discretization (along with suitable
-abstractions) is often employed to transform the systems into a
-discrete transition system. Relational abstractions is one such idea,
-which abstracts continuous dynamics by discrete relations using
-appropriate \textit{reachability invariants}. Both time independent
-and time dependent relations have been proposed; the former captures
-all reachable states over all time, whereas, the latter explicitly
-includes time by relating reachable states to time. Thus it
-can prove timing properties whereas, time independent relational
-abstractions can not. In both cases, the resulting abstractions can be
-interpreted as discrete transition systems and can be analyzed using
-model checkers. We briefly discuss these two types of abstractions.
-
-\mypara{Timeless Relational Abstraction}
-
-Relational abstractions for hybrid systems were proposed in
-\cite{Sankaranarayanan+Tiwari/2011/Relational}. For a hybrid automaton
-model, they can summarize the continuous dynamics of each mode using a
-binary relation over continuous states. The resulting relations are
-timeless (independent) and hence valid for all time as long as
-the mode invariant is satisfied. The relations take the general form
-of $R(\x,\x') \bowtie 0$, where $\bowtie$ represents one of the
-relational operators $=, \ge, \le, <, >$. For example, an abstraction
-that captures the monotonicity with respect to time for the
-differential equation $\dot{x} = 2$ is $x' > x$. The abstraction
-capturing the relation between the set of ODEs: $\dot{x} = 2$,
-$\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
-
-Such relation summaries, discretize the continuous evolution of a
-given system into an infinite state transition system, in which safety
-properties can be verified using $k$-induction, or falsified using
-bounded model checkers. The relationalization procedure involves
-finding suitable invariants in a chosen abstract domain, such as
-affine abstractions, eigen abstractions or box abstractions
-\cite{Sankaranarayanan+Tiwari/2011/Relational}.  For this, different
-techniques like template based invariant generation
-\cite{Gulwani+Tiwari/2008/Constraint,
-Colon+Sankaranarayanan+Sipma/03/Linear} can be used.
-
-\mypara{Timed and Time-Aware Relational Abstractions}
-
-As the above discussed relations are timeless, we cannot reason over
-the timing properties of the original system. Moreover, they are not
-suitable for time-triggered systems. Timed relational
-abstractions~\cite{zutshi2012timed} and time-aware relational
-abstractions~\cite{mover2013time} were proposed to overcome these
-shortcomings by explicitly including time in the relations. In
-addition, both can be more precise than their timeless counterpart,
-but unlike it, assume continuous affine dynamics.
-
-To analyze time-triggered systems like an SDCS modeled by an affine
-hybrid automata~\footnote{An affine hybrid automata is restricted to
-the ODEs, resets and guards being affine.}, timed relational
-abstractions can be computed directly from the solutions of the
-underlying affine ODEs.  Recall that the solution of an affine ODE
-$\dot{\x} = A\x$ can be represented using the matrix exponential as
-$\x(t) = e^{tA}\x(0)$. This gives us the timed relation $\x' =
-e^{tA}\x$.  Clearly, the relation is non-linear with respect to time.
-However, for the case of SDCS with a fixed sampling time period, we
-obtain linear relations in $\x$.  We demonstrated the usefulness of
-timed relational abstractions for the case of linear systems
-in \cite{zutshi2012timed}, and now incorporate the idea to discretize
-black box dynamical systems.
-
-Similar to timeless relational abstractions, time-aware relational
-abstractions \cite{mover2013time} construct binary relations between
-the current state $\x$ of the system and any future reachable state
-$\x'$. The difference lies with the latter also constructing relations
-between the current time $t$ and any future time $t'$. This is
-achieved by a case by case analysis of the eigen structure of the
-matrix $A$ (of an affine ODE). Separate abstractions are used for the
-case of linear systems with constant rate, real eigen values and complex eigenvalues.
-
-\subsection{Learning Dynamics using Simple Linear Regression}
-
-\mypara{Regression Analysis}
-
-In statistics, regression is the problem of finding a
-\textit{predictor}, which can suitably predict the relationship
-between the given set of observed input $\x$ and output $\y$ vectors.
-In other words, assuming that $\y$ depends on $\x$, regression
-strategies find either a parameterized or a non-parameterized
-prediction function to explain the dependence. We now discuss simple
-linear regression, which is parametric in nature and searches for an
-affine predictor. It is also called \textit{ordinary least squares}.
-
-\mypara{Ordinary Least Squares (OLS)}
-
-Let the data set be comprised of $N$ input and output pairs $(\x,
-y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
-the case in the current context of \textit{finding the best fit}, the
-problem is over-determined; there are more equations than
-unknowns. Hence, a single affine function cannot be found which
-satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This is formally
-defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine the vector of
-coefficients $\vec{a}$ and an offset constant $b$, such that the least
-square error of the affine predictor is minimized for the given data
-set.
-\begin{equation}
-    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
-\label{reg}
-\end{equation}
-To ease the presentation, we can rewrite the above as a homogeneous
-expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
-$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
-\[
-    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
-\]
-The solution of OLS can be analytically computed as
-\[ Ab = (X^TX)^{-1}X^T\y\]
-where $X$ is the matrix representing the horizontal stacking of all
-$\hat{\x}$.
-The details can be found in several texts on learning and statistics
-\cite{friedman2001elements}.
-
-Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
-where $\x\in\reals^n$ we can use OLS to approximate its trajectories
-at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
-$A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
-\begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
-    global relational model for the system. The data set for the
-    regression is a set of pairs $\setof{(\x,\x') | \x' =
-    \simulate(\x, \Delta)}$ and can be generated on demand. Another
-    set can be generated to compute the error $\delta$ as an interval
-    of vectors, where each element is an interval $\delta_i \in
-    [\delta_{min}, \delta_{max}]$.
-
-    Affine maps are poor
-    approximations for arbitrary non-linear functions. Hence, we use a
-    collection of affine maps to approximate the local behaviors (in
-    state-space) of the system $\simulate$. This results in a piece
-    wise approximation, as described in the next section.
-
-\subsection{Piecewise Affine (PWA) Transition System}
-
-We define the PWA transition system as a transition system
-(ref.~\secref{DiscSys}) $\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
-where each transition $\tau \in \scrT$, is associated with a
-transition relation of the form
-\[
-    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
-\]
-where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
-$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
-The assertion on the initial states $\Theta$ is also affine.
-
-We can now use a PWA transition system $\rho$ to approximate the
-behavior of a hybrid dynamical system $\System$ defined by a simulator
-$\simulate_{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
-form, we can only approximate general hybrid systems.}. Each
-transition of $\rho$ represents a discrete time step $\Delta$. Let the
-state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
-states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
-predicates $g_\tau$ are given by a conjunction of hyper-planes and
-hence can be represented in matrix form as a polyhedron (defined by
-$m$ constraints) in the state-space
-\[
-g:C\x - \vd \le 0
-\]
-where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
-$m$. The affine maps $f_\tau$ can be represented in matrix form as
-$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
-an error term $\delta$ which is defined by a vector of intervals
-$[\delta_{\min},\delta_{\max}]$. Hence,
-\[
-f:\x \mapsto A\x + \vb + \delta
-\]
-defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
-\in f_\tau(\x)}$.
-
-% Given a state vector $\x\in\reals^n$, a guarded affine map
-% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
-% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
-% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
-% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
-% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
-% guard is satisfied.
-
-We now formalize the PWA affine transition system for a dynamical system.
-\begin{definition}[PWA Transition System]
-
-Given a hybrid dynamical system over a state-space
-$\HybridStates^n$, a PWA transition system $\rho$ is given by the
-tuple $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in
-\scrT$ are affine transition relations and $\Theta$ is an affine
-predicate over $\x$ and the states $\x\in\HybridStates^n$. The
-transition relation is then defined by $\scrT$ with $n$ transition
-relations as follows
-
-\begin{equation}
-    \scrT = \left\{
-        \begin{array}{ll}
-            g_1(\x) \land g'_1(\x') \implies \x' \in f_1(\x) \\
-            \ldots\\
-            g_n(\x) \land g'_n(\x') \implies \x' \in f_n(\x) \\
-        \end{array}
-    \right.
-\end{equation}
-
-\end{definition}
-
-where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
-$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
-denotes the next state of the system.
-
-A PWA model is \textit{deterministic}, iff for every state
-$\x\in\HybridStates^n$, a unique guarded affine map is satisfied and
-all errors $\delta$ are singletons. However, in practice such a case
-rarely exists. A PWA model will be usually non-deterministic, both due
-to multiple choice of transition relations and the reachable states at
-the $k^{th}$ time step (or after $k\Delta$ units of time) being a set.
-Abusing the notation, we denote the set of states reachable in a
-single step in $\rho$ by $\x' = \rho(\x)$.
-
-A PWA transition system is \textit{complete}, iff for every state
-$\x\in\HybridStates^n$, there exists at least one satisfied
-transition relation $\scrT$. If this is not the case, the system can
-deadlock, with no further executions. This usually results from
-incorrect modeling of SDCS, and from here on, we do not consider such
-cases.
-
-% Given a dynamical system over a continuous state-space
-% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
-% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
-% dynamical system to a finite set of guarded affine maps
-% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
-% rule by the affine map of the satisfied $\gm$.
-
-% \begin{equation}
-%     \pwa = \left\{
-%         \begin{array}{ll}
-%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
-%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
-%             \ldots & \ldots\\
-%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
-%         \end{array}
-%     \right.
-% \end{equation}
-
-% \end{definition}
-
-% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
-% Abusing the notation, we denote the next state computed by the PWA
-% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
-% for every state $\x\in\ContStates$, a unique guarded affine map is
-% satisfied. A PWA model is \textit{complete}, iff for every state
-% $\x\in\ContStates$, there exists at least one satisfied guarded affine
-% map $\gm$.
-
-\section{Relational Modeling}
-
-We now describe how relational models can be computed for a given
-black box system by enriching the abstract reachability graph obtained
-by scatter-and-simulate. We first formalize the notion of an enriched
-graph, and then show how they can be interpreted as a PWA transition
-system using $k$-relational modeling.
-
-\subsection{Abstract Enriched Graph}
-
-The existential abstraction relation in \chapref{abs} was defined as
-follows:
-\[
-%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
-    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
-\]
-The abstract relation $\areach{t}$ can be enriched by incorporating
-the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
-system, such a relation can be rarely represented using an exact affine map.
-This is due to the presence of non-linear and hybrid behaviors. But,
-an affine map can always be estimated with an error.
-
-If the system dynamics are completely specified in the form of a white
-box model, we can use first order approximations to find the affine
-expressions for the relations. Using a tool like \flowstar, we can
-obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
-\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
-that, every element $b_i$ of the vector $\vb$ is contained in the
-respective scalar interval $b_i\in[b^l_i,b^h_i]$.
-
-For the case of black box systems, such a sound approximation is not
-possible. Instead, we rely on a statistical method like simple linear
-regression to estimate the affine map $\amap: A\x + \vb$. We then
-estimate the error $\delta$ and generalize $\amap$ to an interval
-affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
-abstraction with the below relation
-\[
-    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
-    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
-\]
-
-However, in the presence of complex non-linear behavior, using a
-single affine map can lead to a poor approximation. To increase the
-precision, one can use more than one affine map. Let us denote this set as
-$R_{(C,C')}$ (ref.~\secref{ovw}).
-\[
-    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
-\]
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
-    \draw[arw] (-3,0) -- (-1.2,0);
-    \draw[arw] (1.2,0) -- (3,0);
-
-\node at (-3.4,0) {$G$};
-\node at (3.4,0) {$G^R$};
-\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Using OLS, $G^R$ is computed by determining the appropriate
-$R_{(C,C')}$ for each edge of $G$.}
-\label{fig:enrichment}
-\vspace*{-.3cm}
-\end{figure}
-
-As shown in \figref{enrichment}, we use OLS to compute an enrichment
-of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
-the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
-of the enriched edge are clearly non-deterministic. Form a state $\x
-\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
-is satisfied. This interpretation results in an infinite state
-transition system. We now detail the construction of the set
-$R_{(C,C')}$ using $k$-relational modeling.
-
-% We now compute a PWA model for a given black box system by estimating
-% the dynamics using OLS. We can estimate the affine relations between
-% every cell, but as discussed in ~\chapref{abs}, it is a futile
-% approach.  Instead, we use the same heuristic as before:
-% scatter-and-simulate, to select the cell relations for estimation. We
-% build the reachability graph as before, but in addition, annotate each
-% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
-% respective relation. The reachability graph thus computed, is a
-% transition system. Using off-the-shelf bounded model checkers, we
-% reason about the system's safety properties.
-
-\subsection{$k$ - Relational Modeling}
-
-\begin{figure}[!htbp]
-\vspace*{-.2cm}
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-\begin{scope}[scale=1]
-\draw [line] (-1.0,0) circle (0.5);
-\draw [line] (-4.0,0) circle (0.5);
-\draw [line] (2.0,1) circle (0.5);
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-0.5,0) -- (1.5,1.0);
-\draw[arw] (-0.5,0) -- (1.5,-1.0);
-\draw[arw] (-3.5,0) -- (-1.5,0.0);
-\node at (-4.0,0) {$C$};
-\node at (-1.0,0) {$C'$};
-\node at (2.0,1.0) {$C''_1$};
-\node at (2.0,-1.0) {$C''_2$};
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
-\node at (-4.0,-2.5) {$k=0$};
-\node at (-1.0,-2.5) {$k=1$};
-\node at (+2.0,-2.5) {$k=2$};
-\draw [fill=black] (-4.0,-2.0) circle (0.05);
-\draw [fill=black] (-1.0,-2.0) circle (0.05);
-\draw [fill=black] (2.0,-2.0) circle (0.05);
-
-\end{scope}
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
-\label{fig:k-rel}
-\vspace*{-.3cm}
-\end{figure}
-
-Given a $G$, when computing the set of relations for an edge $R$, we
-consider only the set of abstract states reachable at a specific time
-step.  Intuitively, one might consider the states reachable in one
-time step $\Delta$.  However, such a process can be generalized by
-looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
-k\Delta}$ time steps.
-
-Using \figref{k-rel}, we illustrate this notion. To compute
-$R_{(C,C')}$, we
-split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
-$k\Delta$) as follows.  We observe the local behavior of the system by
-noting the evolution of the system $\System$ from a state $\x(t)
-\in C$ for a time length dependant on $k$. For
-\begin{itemize}
-    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
-    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
-    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
-        and split them in two sets (a) and (b) on the basis of the
-        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
-        \begin{enumerate}[(a)]
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
-        \end{enumerate}
-    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
-        and split them in to multiple sets on the basis of the cells
-        reached at time $(t+\Delta),\; (t+2\Delta),\;
-        \ldots,\;(t+n\Delta)$.
-\end{itemize}
-
-$k$-relational modeling can be understood as a heuristic, which uses
-the underlying abstraction to differentiate behaviors of the system
-which `diverge' (or are revealed to be `distinct' at a future time).
-Such a heuristic can be useful in increasing the precision of the
-learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
-$n$ and illustrate using examples.
-
-\mypara{$0$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=1]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
-\node at (0,-2.5) {$t=0$};
-\node at (4,-2.5) {$t=\Delta$};
-\draw [fill=black] (0.0,-2.0) circle (0.05);
-\draw [fill=black] (4.0,-2.0) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=3.0cm,scale=1]
-    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
-    \implies A_1\vx + \vb_1 + \delta_1$};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{All the trajectory segments will be used to construct the
-model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
-\label{fig:k0}
-\vspace*{-.3cm}
-\end{figure}
-
-When $k=0$, the $0$-relational model is a PWA transition system which only
-defines the evolution of states $\x$, and does not specify the
-reachable cell. The guard predicates of the relations are defined over
-cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.
-
-We use regression to estimate the dynamics for the outgoing trajectory
-segments form a cell $C$. Hence, the data set $\ds$ for the regression
-includes all trajectory segments beginning from the same cell
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C}.
-\]
-This includes trajectory segments ending in different cells, as shown
-in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
-\begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            \ldots \\
-            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
-        \end{array}
-    \right.
-\end{equation*}
-
-Note that this can be quite imprecise when the cells are big,
-containing regions of state-space with complex dynamics. This is true
-for both non-linear systems and hybrid dynamical system, where a cell
-can contain two or more modes with differing continuous dynamics.
-
-\mypara{$1$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=1]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (2,0) {\begin{minipage}{\textwidth}
-    \footnotesize \begin{equation*}
-        \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
-    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
-    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
-%Cardinality($R_{(C,C')}$) = $2$.}
-\label{fig:k1}
-\vspace*{-.3cm}
-\end{figure}
-
-% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-%     \node at (2,0) {\footnotesize $\begin{equation*}
-%     \scr{T} = \left\{
-%         \begin{array}{ll}
-%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-%         \end{array}
-%     \right.
-% \end{equation*}$};
-% \end{scope}
-
-To improve the preciseness of learnt dynamics, we include the
-reachability relation in the
-regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
-comprised only of trajectory segments $\pi_t$ which start and end in
-the same set of cells.
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
-\]
-For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
-$N$ transition relations.
-
-\figref{k1} illustrates the $k=1$ refinement of the case shown in
-\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
-= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
-relations are constructed. The resulting $1$-relational model
-is at least as precise as the corresponding $0$-relational model.
-
-\mypara{$k$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
-    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;
-
-%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-
-    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (8.0, 2.0) {$C''_1$};
-\node at (8.0, -2.0) {$C''_2$};
-%\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
-\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\node at (8,-4.0) {$t=2\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-\draw [fill=black] (8.0,-3.5) circle (0.05);
-\end{scope}
-
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (4,0) {\begin{minipage}{\textwidth}
-        \footnotesize \begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The $k=2$ refinement further splits the data set $\ds_1$
-    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
-    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
-    non-deterministically defining the system behavior.}
-\label{fig:k2}
-\vspace*{-.3cm}
-\end{figure}
-
-Finally, we generalize the relational models to $k$-relational PWA
-models. A $k$-relational model is constructed by using $k$ length
-\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
-\textit{connected} iff its cost $\cost(\SegTraj) = 0$.
-
-Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
-$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
-\textit{similar} if their trajectory segments have the same sequence
-of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
-\[
-    \forall i\in\setof{1 \ldots k}.
-    start(\pi_{t_i}) \in C_i
-    \land start(\pi'_{t_i}) \in C_i
-    \land end(\pi_{t_i}) \in C_{i+1}
-    \land end(\pi'_{t_i}) \in C_{i+1}
-\]
-
-For every cell $C$ in the reachability graph, we construct a data set
-$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
-length \textit{connected} and \textit{similar} segmented trajectories.
-%\[
-%    \ds = \setof{\pi_{\tau_i} |
-%        \pi_{\tau_i}
-%        start(\pi_{\tau_{i}}) \in C_i
-%        \land
-%        end(\pi_{\tau_{i}}) \in C_i}
-%\]
-
-\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.
-
-\section{Bounded Model Checking Black Box Systems}
-% include an algorithm
-We now describe the entire procedure in three steps.
-\begin{enumerate}
-    \item Given a black box system $\System$ specified by
-        $\simulate_\System$, we
-        use scatter-and-simulate with ($\Delta$-length) trajectory
-        segments to sample a finite reachability graph
-        $\scr{H}(\Delta)$.
-    \item Using regression, we enrich the graph relations and annotate
-        the edges with the estimated affine maps.
-    \item Interpreting the directed reachability graph as a transition
-        system, we use a bounded model checker to find fixed length
-        violations to safety properties.
-\end{enumerate}
-
-\section{An Example: Van der Pol Oscillator}
-
-We now describe the above using the Van der Pol oscillator benchmark
-presented in \secref{math-dyn-sys}. Simulations generated using
-uniformly random samples are shown in \figref{vdp-cont}. We want to
-check the property $P_3$, indicated by the red box, given the initial
-set indicated by the green box. The abstraction is defined by
-$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
-where each cell is of size $0.2 \times 0.2$ units.
-Scatter-and-simulate is then used to construct the abstract graph $G$,
-using $2$ samples per cell and the time step $\Delta = 0.1s$. The
-complete process follows.
-
-\begin{enumerate}
-    \item{\emph{Abstract}} Consider an implicit abstraction induced by the quantization
-    function $\quant_\epsilon(\x)$. The corresponding reachability
-    graph $\scr{H}(\Delta)$ (with time step $\Delta$) is given by
-    $\tupleof{\scr{C},\areach{\Delta}, \scr{C}_0, \scr{C}_u}$.
-\item{\emph{Discover}}: Using scatter-and-simulate, enumerate a finite number
-    of cells (vertices) and the relations (edges) of the graph
-    $\scr{H}(\Delta)$. Associate the set of generated trajectory
-    segments with their respective originating cells using a map $D:C
-    \mapsto \setof{\pi | start(\pi) \in C}$. The discovered
-    abstraction is show in \figref{vdp-abs}. As mentioned, red cells
-    are unsafe and green cells are initial cells.
-\item{\emph{Relationalize}} For each cell $C$, perform regression analysis on
-    the respective set of trajectory segments $D(C)$, and compute a
-    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
-    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(\Delta)$. Annotate
-    each edge in the graph $\scr{H}(\Delta)$ with the respective
-    relation.
-
-    \figref{vdp-map} shows the cells and the trajectory segments which
-    are part of the data sets constructed using the $1$-relational
-    modeling. Against each cell, its unique identifier (integer
-    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
-    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
-    relations. Note that self loops result when an observed trajectory
-    segment has its $start$ and $end$ states in the same cell.
-
-\item{\emph{Model Check}} The graph $\scr{H}(\Delta)$ can now be viewed as a
-    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
-    where $C \in vertices(\scr{H}(\Delta))$ and $\x \in
-    \HybridStates$. A transition $\tau \in \scrT$ is of the form
-    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
-    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
-    edge(\scr{H}(\Delta))}$.
-\item{\emph{Check Counter-example}} The infinite state (but finite location) transition system can
-    be model checked to find a concrete counter-example, which if
-    exists, can indicate the existence of a similar trace in the
-    original black-box system $\System$. The latter check is carried
-    out as before, using the numerical simulation function
-    $\simulate$. For the given example, the model checker is unable to
-    find a counter-example.
-\end{enumerate}
-
-\inclfig[width=0.75\textwidth]{vdp_cont_traces.png}{Van der
-Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}
-\label{fig:vdp-cont}
-
-\inclfig[width=0.75\textwidth]{vdp_abs.png}{The discovered
-abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
-initial cells.}
-\label{fig:vdp-abs}
-
-\inclfig[width=\textwidth]{vdp_disc_map.png}{Cells and
-trajectory segments used by $1$-relational modeling.}
-\label{fig:vdp-map}
-
-\inclfig[width=0.40\textwidth]{vdp_graph.png}{Enriched
-graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}
-\label{fig:vdp-graph}
-
-\begin{table*}[!htbp]
-\centering
-\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
-\label{tab:vdp-pwa}
-\begin{tabular}{@{}ccccc@{}}
-\toprule
-$C$ & $C'$ & $A$ & $b$ & $\delta$\\
-\midrule
-(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
-\midrule
-(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
-\midrule
-(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% (0, -1)   & (0, -1)   &  \begin{bmatrix}0.9944498 & 0.12911959 \\-0.12013468&1.63250531\end{bmatrix} &\begin{bmatrix}-0.00003683\\-0.00085411\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -1)   & (0, -2)   &  \begin{bmatrix}0.99523928& 0.1290447  \\-0.10236736&1.6314103 \end{bmatrix} &\begin{bmatrix}-0.00013019\\-0.00289192\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -2)   & (0, -2)   &  \begin{bmatrix}0.99565348& 0.12916002 \\-0.09352462&1.6341424 \end{bmatrix} &\begin{bmatrix}-0.00015574\\-0.00339664\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -2)   & (0, -4)   &  \begin{bmatrix}0.99668599& 0.12892455 \\-0.0725035 &1.62937613\end{bmatrix} &\begin{bmatrix}-0.00037391\\-0.00796189\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -9)   & (-1, -15) &  \begin{bmatrix}0.99985362& 0.12858788 \\-0.09447884&1.61476417\end{bmatrix} &\begin{bmatrix}-0.0018997 \\-0.04200363\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-1, -15) & (-3, -24) &  \begin{bmatrix}0.95473217& 0.12473489 \\-1.29198505&1.47155614\end{bmatrix} &\begin{bmatrix}-0.01248412\\-0.41100129\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-3, -24) & (-4, -29) &  \begin{bmatrix}1.11789749&-0.17261206 \\-1.71017779&0.66006597\end{bmatrix} &\begin{bmatrix}-1.0745344 \\-3.31246863\end{bmatrix}&\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}\\
-% \midrule
-% (-4, -29) & (-4, -29) &  \begin{bmatrix}0.99796588& 0.01012817 \\-0.41529747&1.02432211\end{bmatrix} &\begin{bmatrix}-0.00135199\\-0.28016983\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-1, -10) & (-2, -16) &  \begin{bmatrix}0.97381953& 0.12768367 \\-0.71242741&1.56764585\end{bmatrix} &\begin{bmatrix}-0.00320265\\-0.11964102\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% (-1, -10) & (-2, -15) &  \begin{bmatrix}0.97373058& 0.12859163 \\-0.70173855&1.58540872\end{bmatrix} &\begin{bmatrix}-0.0015415 \\-0.08616376\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-2, -16) & (-3, -23) &  \begin{bmatrix}0.92850123& 0.12242812 \\-1.84720528&1.39159826\end{bmatrix} &\begin{bmatrix}-0.02334236\\-0.73348865\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -23) & (-5, -29) &  \begin{bmatrix}3.14947585&-0.81150646 \\-0.96159406&0.23543254\end{bmatrix} &\begin{bmatrix}-3.13860326\\-4.99696623\end{bmatrix}&\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}\\
-% \midrule
-% (-2, -15) & (-3, -22) &  \begin{bmatrix}0.93156623& 0.12260264 \\-1.75117219&1.40104348\end{bmatrix} &\begin{bmatrix}-0.02214125\\-0.68423569\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -22) & (-5, -29) &  \begin{bmatrix}3.08806462&-0.72332776 \\-1.69271079&0.40308293\end{bmatrix} &\begin{bmatrix}-2.77908467\\-4.57411339\end{bmatrix}&\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}\\
-
-\subsection{Search Parameters}
-
-The search parameters for S3CAM-R include the parameters of S3CAM:
-$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
-maximum error budget $\delta_{max}$ for OLS and the maximum length of
-segmented trajectory for building $k$-relational models.
-
-We have already discussed the effects of $N$, $\epsilon$ and $\Delta$
-on S3CAM's performance. However, they also have an effect on
-relational modeling. A finer grid with small cells produces more
-accurate models than a coarser grid with large cells. Similarly,
-small time length trajectory segments result in more accurate
-modeling.
-%An increase in the number of samples can increases the
-%accuracy of regression.
-
-\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
-Given a $\delta_{max}$, we keep increasing $k$ during the
-$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
-satisfied. A $k_{max}$ is introduced to bound the longest
-segmented trajectory which can be considered.
-
-% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
-
-\subsection{Reasons for Failure}
-The approach can fail to find a counter-example, even when it exists, in
-one of three ways.
-
-\begin{itemize}
-
-    \item \emph{S3CAM Fails} No abstract counter-example is found by
-        S3CAM. We can remedy it by increasing search budgets and/or
-        restarting.
-
-    \item \emph{BMC Fails} An abstract counter-example is found, but
-        the BMC fails to find a concrete counter-example in the PWA
-        relational model. The failure can be attributed to either (a)
-        a spurious abstract counter-examples or (b) a poorly estimated
-        model. The former can be addressed by restarting but the
-        latter requires that the maximum model error $\delta_{max}$ be
-        decreased.
-
-    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
-        is found, and it is successfully concretized in the PWA
-        relational model by the BMC. However, it is not reproducible
-        in the black box system. This happens when the PWA relational
-        model is not precise enough.
-
-\end{itemize}
-
-\section{Implementation and Evaluation}
-
-The implementation was prototyped as S3CAM-R, an extension to our
-previously mentioned tool S3CAM (\chapref{case}). OLS regression
-routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
-Python module for machine learning. SAL~\cite{SAL-SRI}
-with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
-solver.
-
-We used S3CAM to discover the abstraction graph $G$, which was then
-trimmed of the nodes which did not contribute to the error search
-(nodes from which error nodes were not reachable). In our experiments we
-used $1$-relational modeling to create the transition system from $G$.
-Using SAL, we checked for the given safety property. If a concrete
-trace was obtained from the BMC, it was further checked for validity
-by simulating using $\simulate$ to see if it was indeed an error
-trace.  If not, $100$ samples from its associated abstract state was
-sampled to check the presence of a counter-example in it
-neighborhood. This can be further extended by obtaining different
-discrete sequences or discrete trace sequences from the model checker.
-
-We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-few of the benchmarks described in \chapref{case}, including the Van
-der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
-benchmark. As before, we ran S3CAM-R $10$ times with different seeds
-and averaged the results. We tabulate both the total time taken and
-the time taken by SAL to compute the counter-example and compare
-against a newer implementation of S3CAM. Note that different
-abstraction parameters are used for S3CAM and S3CAM-R, due to the
-differences in which they operate.
-
-\begin{table*}[!htbp]
-\centering
-\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
-time taken by the BMC engine. The total time in seconds (rounded off
-to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
-$TO$ signifies time $>5hr$, after which the search was killed.}
-\label{tab:res-rel}
-\begin{tabular}{@{}llll@{}}
-\toprule
-Benchmark & BMC & S3CAM-R & S3CAM\\
-\midrule
-Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
-Brusselator               &$<$1  & $4$    & $2$\\
-Lorenz                    &$<$1 & $122$  & $35$\\
-%B.Ball                    &  &   & \\
-Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
-Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
-Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
-Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
-Bouncing Ball             &-  &$TO$   & $450$\\
-
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% 1m8s, 2m56.184s, 1m31.059s[f], 1m32.665s,  1m45.312s, 1m22.397s, 2m1.451s[f], 1m12.118s, 1m26.596s
-% 0.12, 0.06,         0.33       0.13,       0.22,     FAIL,       ,X           0.08        0.15
-
-The results show promise, but clearly S3CAM performs better. Also,
-S3CAM-R timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
-and the bouncing ball. However, we need to explore more benchmarks to
-be conclusive. S3CAM-R's performance can be explained by the
-difficulty in finding a good abstraction ($\quant_\epsilon$), which
-needs to be fine enough, to obtain good prediction models but coarse
-enough, to be manageable by the current SMT solvers. Specifically, to
-obtain a good quality counter-example from the model checker (and
-avoid false positives), the transition system should be created from
-models with high accuracy, for which a finer abstraction is required.
-However, the exploration of a finer abstraction is exponentially more
-complex and results in a much bigger transition system, which the
-current state of the art bounded model checkers (which use SMT
-solvers) can not handle under reasonable resources.
-
-\section{Conclusion and Future Work}
-
-We have presented another methodology to find falsifications in black
-box dynamical systems. Combining the ideas from abstraction based
-search (\chapref{abs}), with our previous relational
-abstractions~\cite{zutshi2012timed}, we proposed $k$-relational
-modeling to compute richer abstractions. Simple linear regression can
-be used to estimate the local dynamics of the trajectory segments and
-compute PWA relational models which can be interpreted as a transition
-system.  These can then be checked for safety violations using an
-off-the-shelf bounded model checker.  Finally, we implemented the
-ideas as a tool S3CAM-R and demonstrated its performance on a few examples.
-
-\subsection{Improvements}
-
-The approach seems promising as it avoids an expensive refinement
-step. But, before it can be applied to more complex systems, we need
-to overcome its shortcomings.  As a future extension, we are
-investigating ideas to improve the performance of the underlying three
-main sub-processes, namely, abstraction, modeling, and BMC. We detail
-these below.
-
-\begin{enumerate}
-
-\item \emph{Abstraction} The abstract domain we have used, has been
-    confined to the interval or rectangular domain. Such a domain is
-    very coarse when compared to more general polyhedral domains, using
-    which, we can gain better precision. Due to polyhedral domains
-    being computational expensive, their integration in S3CAM needs to
-    be explored thoroughly.
-
-\item \emph{Modeling} We have used $k$-relational modeling to increase
-    the precision of the enriched abstractions, however, there are
-    several other refinement strategies that remain to be explored.
-    This includes refining the abstract states by using guard
-    predicates (similar to predicate abstraction), using non-linear
-    templates in parametric regression to compute more precise
-    quadratic relations, and using an adaptive time step instead of a
-    fixed one to address the non-linearities due to a long time
-    horizon.
-
-\item \emph{Model Checking} As of now, we use SMT solvers to search
-    for concrete counter-examples. Although they are very efficient,
-    owing to extensive engineering effort, reachability of transition
-    systems resulting from dynamical systems remains a difficult
-    problem. As the time horizon of the safety property increases, the
-    possible combinations of discrete transitions increase
-    exponentially. Hence, to find a counterexample which is a sequence
-    of discrete transitions over a `long' time horizon is not
-    tractable for most, but the simplest of dynamical systems.
-    Instead, we can use linear programming solvers, by enumerating
-    each path in the graph $G^R$. Note that the constraints are all
-    linear (conjunctions) along an abstract path. However, in the
-    worst case, the number of paths in a graph can be of the order
-    $n!$ where $n$ is the number of vertices of a graph. Hence, this
-    will not be feasible, unless, we can prioritize paths by using a
-    triage process similar to one used by S3CAM and using a budget on
-    the maximum number of paths.
-
-    Another approach to address the issue would be to use an adaptive
-    time discretization technique, where relations over both shorter
-    and longer time steps are computed, and the SMT solver can
-    `select' the time step precise enough to find a counter-example.
-
-\item \emph{SMT Solvers} One major impediment to our approach is the
-    fact that SMT solvers use the theory of reals with \emph{exact
-    precision}. This is important for verification approaches, but can
-    be relaxed for the problem of falsification. An SMT solver which
-    uses approximate reasoning but returns robust counter-examples
-    will be as useful, and perhaps more efficient.
-
-\end{enumerate}
-
-\subsection{Data Driven Analysis}
-
-Finally, we would like to mention that our approach is `simulation'
-driven. It can be easily modified to be `data driven', by working with
-a fixed set of data. Given a data set, we then need to automatically
-find a good abstraction and a high fidelity PWA transition system
-using which, we can summarize the black box hybrid dynamical system.
-Apart from model checking the transition system, one can extend it to
-the analysis of SDCS. More specifically, by combining the transition
-system model of a plant with the control software, a model checker
-like CBMC \cite{kroening2014cbmc} can be used to do a closed loop
-symbolic analysis of the SDCS. This can be an alternative to S3CAM-X.
-
-%which will be
-%very useful to for analyzing data for potential property violations.

### 2017-02-08 bmc.tex
.

--- 2017-02-08 intro.tex
@@
-Model based design has been quite successful in the industry. 
-Tools like \SIMULINK are often used to create models describing hybrid
-dynamical systems. Testing such systems for safety properties remains
-a hard problem, and the current approaches consist of either guided
-testing using numerical simulations or translating to hybrid automata.
-We now suggest an approach which lies in the middle.
-
-In this work we propose a two step process for searching violations of
-the properties. This involves (a) learning a piece-wise affine model
-and (b) exhaustively searching it for violations. It must be noted
-that the violation found in the model might not be valid in the
-original system, depending on the accuracy of the model.
-% concretizing the violation in the original system.
-
-The piece-wise affine model is the aggregation of affine maps, each
-approximating observed local behavior of the system.  The model is
-discrete time and `relational'. It relates the states reachable in a
-fixed time step $\Delta$ to a given state.
-
-In our previous work~\cite{zutshi2014multiple}, we analyzed systems by
-restricting ourselves to their black box semantics. This enabled us to
-reason about the state-space reachability of the system without a
-direct analysis of its structure. Using a coarse abstraction which we
-searched on-the-fly, we could observe the local dynamics as required.
-This gave us an efficient procedure to find abstract counter-examples.
-However, to concretize the counter-examples, a `closer look' or
-refinement of the abstraction is required. The grid based state-space
-abstraction was refined by splitting the relevant cells (abstract
-states) into smaller ones. As noted previously, due to the curse of
-dimensionality, such a uniform splitting is an expensive operation,
-and can not scale to higher dimensions.
-
-We further our exploration of trajectory segment based methods; and
-explore an alternative approach to overcome the explosion in abstract
-states. Instead of selectively refining the abstraction, we compute a
-model of the black box system and use bounded model checking to find a
-concrete counter-example in the model. Due to modeling errors, this
-might not be reproducible in the original black-box system. We then
-use this counter-example to guide the search towards a counter-example
-in the original system.
-
-More specifically, we use regression to quantitatively estimate the
-discovered relations, which were witnessed by trajectory segments, by
-affine maps. These maps approximate locally observed behaviors, which
-are incorporated into the sampled reachability graph as edge
-annotations.  The resulting graph or the Piece-Wise Affine (PWA)
-relational model can be interpreted as an infinite state discrete
-transition system and model checked for time bounded safety
-properties. A counter-example if found, can indicate the presence of a
-violation in the system.
-%Furthermore, using linear programming, we can over-approximate the
-%neighbhourhood of the counter-example in the model.
-
-% Towards the end of the chapter, we discuss extensions to data driven
-% approaches (instead of simulator driven), where instead of a
-% simulator, a fixed set of data is provided as the behavioral
-% description of the system. Using a combination of relational modeling,
-% and program analysis we outline the future work for falsifying
-% properties of SDCS.
-
-\section{Relational Modeling}
-
-We now describe how relational models can be computed for a given
-black box system by enriching the abstract reachability graph obtained
-by scatter-and-simulate. We first formalize the notion of an enriched
-graph, and then show how they can be interpreted as a PWA transition
-system using $k$-relational modeling.
-
-\subsection{Abstract Enriched Graph}
-
-The existential abstraction relation in \chapref{abs} was defined as
-follows:
-\[
-%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
-    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
-\]
-The abstract relation $\areach{t}$ can be enriched by incorporating
-the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
-system, such a relation can be rarely represented using an exact affine map.
-This is due to the presence of non-linear and hybrid behaviors. But,
-an affine map can always be estimated with an error.
-
-If the system dynamics are completely specified in the form of a white
-box model, we can use first order approximations to find the affine
-expressions for the relations. Using a tool like \flowstar, we can
-obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
-\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
-that, every element $b_i$ of the vector $\vb$ is contained in the
-respective scalar interval $b_i\in[b^l_i,b^h_i]$.
-
-For the case of black box systems, such a sound approximation is not
-possible. Instead, we rely on a statistical method like simple linear
-regression to estimate the affine map $\amap: A\x + \vb$. We then
-estimate the error $\delta$ and generalize $\amap$ to an interval
-affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
-abstraction with the below relation
-\[
-    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
-    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
-\]
-
-However, in the presence of complex non-linear behavior, using a
-single affine map can lead to a poor approximation. To increase the
-precision, one can use more than one affine map. Let us denote this set as
-$R_{(C,C')}$ (ref.~\secref{ovw}).
-\[
-    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
-\]
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
-    \draw[arw] (-3,0) -- (-1.2,0);
-    \draw[arw] (1.2,0) -- (3,0);
-
-\node at (-3.4,0) {$G$};
-\node at (3.4,0) {$G^R$};
-\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Using OLS, $G^R$ is computed by determining the appropriate
-$R_{(C,C')}$ for each edge of $G$.}
-\label{fig:enrichment}
-\vspace*{-.3cm}
-\end{figure}
-
-As shown in \figref{enrichment}, we use OLS to compute an enrichment
-of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
-the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
-of the enriched edge are clearly non-deterministic. Form a state $\x
-\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
-is satisfied. This interpretation results in an infinite state
-transition system. We now detail the construction of the set
-$R_{(C,C')}$ using $k$-relational modeling.
-
-% We now compute a PWA model for a given black box system by estimating
-% the dynamics using OLS. We can estimate the affine relations between
-% every cell, but as discussed in ~\chapref{abs}, it is a futile
-% approach.  Instead, we use the same heuristic as before:
-% scatter-and-simulate, to select the cell relations for estimation. We
-% build the reachability graph as before, but in addition, annotate each
-% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
-% respective relation. The reachability graph thus computed, is a
-% transition system. Using off-the-shelf bounded model checkers, we
-% reason about the system's safety properties.
-
-\subsection{$k$ - Relational Modeling}
-
-\begin{figure}[!htbp]
-\vspace*{-.2cm}
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-\begin{scope}[scale=0.8]
-\draw [line] (-1.0,0) circle (0.5);
-\draw [line] (-4.0,0) circle (0.5);
-\draw [line] (2.0,1) circle (0.5);
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-0.5,0) -- (1.5,1.0);
-\draw[arw] (-0.5,0) -- (1.5,-1.0);
-\draw[arw] (-3.5,0) -- (-1.5,0.0);
-\node at (-4.0,0) {$C$};
-\node at (-1.0,0) {$C'$};
-\node at (2.0,1.0) {$C''_1$};
-\node at (2.0,-1.0) {$C''_2$};
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
-\node at (-4.0,-2.5) {$k=0$};
-\node at (-1.0,-2.5) {$k=1$};
-\node at (+2.0,-2.5) {$k=2$};
-\draw [fill=black] (-4.0,-2.0) circle (0.05);
-\draw [fill=black] (-1.0,-2.0) circle (0.05);
-\draw [fill=black] (2.0,-2.0) circle (0.05);
-
-\end{scope}
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
-\label{fig:k-rel}
-\vspace*{-.3cm}
-\end{figure}
-
-Given a $G$, when computing the set of relations for an edge $R$, we
-consider only the set of abstract states reachable at a specific time
-step.  Intuitively, one might consider the states reachable in one
-time step $\Delta$.  However, such a process can be generalized by
-looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
-k\Delta}$ time steps.
-
-Using \figref{k-rel}, we illustrate this notion. To compute
-$R_{(C,C')}$, we
-split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
-$k\Delta$) as follows.  We observe the local behavior of the system by
-noting the evolution of the system $\System$ from a state $\x(t)
-\in C$ for a time length dependant on $k$. For
-\begin{itemize}
-    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
-    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
-    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
-        and split them in two sets (a) and (b) on the basis of the
-        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
-        \begin{enumerate}[(a)]
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
-        \end{enumerate}
-    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
-        and split them in to multiple sets on the basis of the cells
-        reached at time $(t+\Delta),\; (t+2\Delta),\;
-        \ldots,\;(t+n\Delta)$.
-\end{itemize}
-
-$k$-relational modeling can be understood as a heuristic, which uses
-the underlying abstraction to differentiate behaviors of the system
-which `diverge' (or are revealed to be `distinct' at a future time).
-Such a heuristic can be useful in increasing the precision of the
-learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
-$n$ and illustrate using examples.
-
-\mypara{$0$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
-\node at (0,-2.5) {$t=0$};
-\node at (4,-2.5) {$t=\Delta$};
-\draw [fill=black] (0.0,-2.0) circle (0.05);
-\draw [fill=black] (4.0,-2.0) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=3.0cm,scale=0.5]
-    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
-    \implies A_1\vx + \vb_1 + \delta_1$};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{All the trajectory segments will be used to construct the
-model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
-\label{fig:k0}
-\vspace*{-.3cm}
-\end{figure}
-
-When $k=0$, the $0$-relational model is a PWA transition system which only
-defines the evolution of states $\x$, and does not specify the
-reachable cell. The guard predicates of the relations are defined over
-cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.
-
-We use regression to estimate the dynamics for the outgoing trajectory
-segments form a cell $C$. Hence, the data set $\ds$ for the regression
-includes all trajectory segments beginning from the same cell
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C}.
-\]
-This includes trajectory segments ending in different cells, as shown
-in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
-\begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            \ldots \\
-            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
-        \end{array}
-    \right.
-\end{equation*}
-
-Note that this can be quite imprecise when the cells are big,
-containing regions of state-space with complex dynamics. This is true
-for both non-linear systems and hybrid dynamical system, where a cell
-can contain two or more modes with differing continuous dynamics.
-
-\mypara{$1$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
-    \node at (2,0) {\begin{minipage}{\linewidth}
-    \footnotesize \begin{equation*}
-        \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
-    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
-    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
-%Cardinality($R_{(C,C')}$) = $2$.}
-\label{fig:k1}
-\vspace*{-.3cm}
-\end{figure}
-
-% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-%     \node at (2,0) {\footnotesize $\begin{equation*}
-%     \scr{T} = \left\{
-%         \begin{array}{ll}
-%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-%         \end{array}
-%     \right.
-% \end{equation*}$};
-% \end{scope}
-
-To improve the preciseness of learnt dynamics, we include the
-reachability relation in the
-regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
-comprised only of trajectory segments $\pi_t$ which start and end in
-the same set of cells.
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
-\]
-For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
-$N$ transition relations.
-
-\figref{k1} illustrates the $k=1$ refinement of the case shown in
-\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
-= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
-relations are constructed. The resulting $1$-relational model
-is at least as precise as the corresponding $0$-relational model.
-
-\mypara{$k$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
-    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;
-
-%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-
-    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (8.0, 2.0) {$C''_1$};
-\node at (8.0, -2.0) {$C''_2$};
-%\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
-\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\node at (8,-4.0) {$t=2\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-\draw [fill=black] (8.0,-3.5) circle (0.05);
-\end{scope}
-
-\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
-    \node at (4,0) {\begin{minipage}{\linewidth}
-        \footnotesize \begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The $k=2$ refinement further splits the data set $\ds_1$
-    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
-    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
-    non-deterministically defining the system behavior.}
-\label{fig:k2}
-\vspace*{-.3cm}
-\end{figure}
-
-Finally, we generalize the relational models to $k$-relational PWA
-models. A $k$-relational model is constructed by using $k$ length
-\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
-\textit{connected} iff its cost $\cost(\SegTraj) = 0$.
-
-Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
-$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
-\textit{similar} if their trajectory segments have the same sequence
-of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
-\[
-    \forall i\in\setof{1 \ldots k}.
-    start(\pi_{t_i}) \in C_i
-    \land start(\pi'_{t_i}) \in C_i
-    \land end(\pi_{t_i}) \in C_{i+1}
-    \land end(\pi'_{t_i}) \in C_{i+1}
-\]
-
-For every cell $C$ in the reachability graph, we construct a data set
-$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
-length \textit{connected} and \textit{similar} segmented trajectories.
-%\[
-%    \ds = \setof{\pi_{\tau_i} |
-%        \pi_{\tau_i}
-%        start(\pi_{\tau_{i}}) \in C_i
-%        \land
-%        end(\pi_{\tau_{i}}) \in C_i}
-%\]
-
-\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.
-
-\section{Bounded Model Checking Black Box Systems}
 % include an algorithm
@@
 \end{itemize}
-
-\section{Implementation and Evaluation}
-
-The implementation was prototyped as S3CAM-R, an extension to our
-previously mentioned tool S3CAM (\chapref{case}). OLS regression
-routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
-Python module for machine learning. SAL~\cite{SAL-SRI}
-with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
-solver.
-
-We used S3CAM to discover the abstraction graph $G$, which was then
-trimmed of the nodes which did not contribute to the error search
-(nodes from which error nodes were not reachable). In our experiments we
-used $1$-relational modeling to create the transition system from $G$.
-Using SAL, we checked for the given safety property. If a concrete
-trace was obtained from the BMC, it was further checked for validity
-by simulating using $\simulate$ to see if it was indeed an error
-trace.  If not, $100$ samples from its associated abstract state was
-sampled to check the presence of a counter-example in it
-neighborhood. This can be further extended by obtaining different
-discrete sequences or discrete trace sequences from the model checker.
-
-We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-few of the benchmarks described in \chapref{case}, including the Van
-der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
-benchmark. As before, we ran S3CAM-R $10$ times with different seeds
-and averaged the results. We tabulate both the total time taken and
-the time taken by SAL to compute the counter-example and compare
-against a newer implementation of S3CAM. Note that different
-abstraction parameters are used for S3CAM and S3CAM-R, due to the
-differences in which they operate.
-
-\begin{table*}[!htbp]
-\centering
-\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
-time taken by the BMC engine. The total time in seconds (rounded off
-to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
-$TO$ signifies time $>5hr$, after which the search was killed.}
-\label{tab:res-rel}
-\begin{tabular}{@{}llll@{}}
-\toprule
-Benchmark & BMC & S3CAM-R & S3CAM\\
-\midrule
-Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
-Brusselator               &$<$1  & $4$    & $2$\\
-Lorenz                    &$<$1 & $122$  & $35$\\
-%B.Ball                    &  &   & \\
-Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
-Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
-Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
-Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
-Bouncing Ball             &-  &$TO$   & $450$\\
-
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% 1m8s, 2m56.184s, 1m31.059s[f], 1m32.665s,  1m45.312s, 1m22.397s, 2m1.451s[f], 1m12.118s, 1m26.596s
-% 0.12, 0.06,         0.33       0.13,       0.22,     FAIL,       ,X           0.08        0.15
-
-The results show promise, but clearly S3CAM performs better. Also,
-S3CAM-R timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
-and the bouncing ball. However, we need to explore more benchmarks to
-be conclusive. S3CAM-R's performance can be explained by the
-difficulty in finding a good abstraction ($\quant_\epsilon$), which
-needs to be fine enough, to obtain good prediction models but coarse
-enough, to be manageable by the current SMT solvers. Specifically, to
-obtain a good quality counter-example from the model checker (and
-avoid false positives), the transition system should be created from
-models with high accuracy, for which a finer abstraction is required.
-However, the exploration of a finer abstraction is exponentially more
-complex and results in a much bigger transition system, which the
-current state of the art bounded model checkers (which use SMT
-solvers) can not handle under reasonable resources.
-
-\section{Conclusion and Future Work}
-
-We have presented another methodology to find falsifications in black
-box dynamical systems. Combining the ideas from abstraction based
-search (\chapref{abs}), with our previous relational
-abstractions~\cite{zutshi2012timed}, we proposed $k$-relational
-modeling to compute richer abstractions. Simple linear regression can
-be used to estimate the local dynamics of the trajectory segments and
-compute PWA relational models which can be interpreted as a transition
-system.  These can then be checked for safety violations using an
-off-the-shelf bounded model checker.  Finally, we implemented the
-ideas as a tool S3CAM-R and demonstrated its performance on a few examples.
-
-\subsection{Improvements}
-
-The approach seems promising as it avoids an expensive refinement
-step. But, before it can be applied to more complex systems, we need
-to overcome its shortcomings.  As a future extension, we are
-investigating ideas to improve the performance of the underlying three
-main sub-processes, namely, abstraction, modeling, and BMC. We detail
-these below.
-
-\begin{enumerate}
-
-\item \emph{Abstraction} The abstract domain we have used, has been
-    confined to the interval or rectangular domain. Such a domain is
-    very coarse when compared to more general polyhedral domains, using
-    which, we can gain better precision. Due to polyhedral domains
-    being computational expensive, their integration in S3CAM needs to
-    be explored thoroughly.
-
-\item \emph{Modeling} We have used $k$-relational modeling to increase
-    the precision of the enriched abstractions, however, there are
-    several other refinement strategies that remain to be explored.
-    This includes refining the abstract states by using guard
-    predicates (similar to predicate abstraction), using non-linear
-    templates in parametric regression to compute more precise
-    quadratic relations, and using an adaptive time step instead of a
-    fixed one to address the non-linearities due to a long time
-    horizon.
-
-\item \emph{Model Checking} As of now, we use SMT solvers to search
-    for concrete counter-examples. Although they are very efficient,
-    owing to extensive engineering effort, reachability of transition
-    systems resulting from dynamical systems remains a difficult
-    problem. As the time horizon of the safety property increases, the
-    possible combinations of discrete transitions increase
-    exponentially. Hence, to find a counterexample which is a sequence
-    of discrete transitions over a `long' time horizon is not
-    tractable for most, but the simplest of dynamical systems.
-    Instead, we can use linear programming solvers, by enumerating
-    each path in the graph $G^R$. Note that the constraints are all
-    linear (conjunctions) along an abstract path. However, in the
-    worst case, the number of paths in a graph can be of the order
-    $n!$ where $n$ is the number of vertices of a graph. Hence, this
-    will not be feasible, unless, we can prioritize paths by using a
-    triage process similar to one used by S3CAM and using a budget on
-    the maximum number of paths.
-
-    Another approach to address the issue would be to use an adaptive
-    time discretization technique, where relations over both shorter
-    and longer time steps are computed, and the SMT solver can
-    `select' the time step precise enough to find a counter-example.
-
-\item \emph{SMT Solvers} One major impediment to our approach is the
-    fact that SMT solvers use the theory of reals with \emph{exact
-    precision}. This is important for verification approaches, but can
-    be relaxed for the problem of falsification. An SMT solver which
-    uses approximate reasoning but returns robust counter-examples
-    will be as useful, and perhaps more efficient.
-
-\end{enumerate}
-
-\subsection{Data Driven Analysis}
-
-Finally, we would like to mention that our approach is `simulation'
-driven. It can be easily modified to be `data driven', by working with
-a fixed set of data. Given a data set, we then need to automatically
-find a good abstraction and a high fidelity PWA transition system
-using which, we can summarize the black box hybrid dynamical system.
-Apart from model checking the transition system, one can extend it to
-the analysis of SDCS. More specifically, by combining the transition
-system model of a plant with the control software, a model checker
-like CBMC \cite{kroening2014cbmc} can be used to do a closed loop
-symbolic analysis of the SDCS. This can be an alternative to S3CAM-X.
-
-%which will be
-%very useful to for analyzing data for potential property violations.

### 2017-02-08 implementation.tex
.

--- 2017-02-07 intro.tex
@@
-%\chapter{Relational Modeling for Falsification}
-%\label{chap:rel}
-
-In \chapref{abs}, we analyzed systems by restricting ourselves to
-their black box semantics. This enabled us to reason about the
-state-space reachability of the system without a direct analysis of
-its structure. Using a coarse abstraction which we searched
-on-the-fly, we could observe the local dynamics as required. This gave
-us an efficient procedure to find abstract counter-examples.  However,
-to concretize the counter-examples, a `closer look' or refinement of
-the abstraction is required. The grid based state-space abstraction
-was refined by splitting the relevant cells (abstract states) into
-smaller ones. As noted previously, due to the curse of dimensionality,
-such a uniform splitting is an expensive operation, and can not scale to
-higher dimensions.
-
-In this chapter, we further our exploration of trajectory segment
-based methods; and explore an alternative approach to overcome the
-explosion in abstract states. Instead of selectively refining the
-abstraction, we compute a model of the black box system and use
-bounded model checking to find a concrete counter-example in the
-model~\footnote{Due to modeling errors, this might not be reproducible in
-the original black-box system.}. We then use this counter-example to
-guide the search towards a counter-example in the original system.
-
-More specifically, we use regression to quantitatively estimate the
-discovered relations, which were witnessed by trajectory segments, by
-affine maps. These maps approximate locally observed behaviors, which
-are incorporated into the sampled reachability graph as edge
-annotations.  The resulting graph or the Piece-Wise Affine (PWA)
-relational model can be interpreted as an infinite state discrete
-transition system and model checked for time bounded safety
-properties. A counter-example if found, can indicate the presence of a
-violation in the system.
-%Furthermore, using linear programming, we can over-approximate the
-%neighbhourhood of the counter-example in the model.
-
-Towards the end of the chapter, we discuss extensions to data driven
-approaches (instead of simulator driven), where instead of a
-simulator, a fixed set of data is provided as the behavioral
-description of the system. Using a combination of relational modeling,
-and program analysis we outline the future work for falsifying
-properties of SDCS.
-
-\section{Overview}
-\label{sec:ovw}
-
-We now briefly describe the approach. Given a numerical simulator
-$\simulate$ and an initial abstraction defined by a quantization
-function $\quant_\epsilon$ and a time step $\Delta$, we use
-scatter-and-simulate (parameterized by number of samples $n$ as
-described in \chapref{abs}) to explore the abstract graph
-$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
-number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
-\areach{\Delta} C'$.
-
-Instead of using a CEGAR like loop (as in \chapref{abs}), we use the
-generated trajectory segments to learn quantitative models describing
-the local behavior of the system. These models are defined by a set of
-relations $R \subseteq \HybridStates \times \HybridStates$ for each
-edge of the reachability graph.
-
-\input{figs/segtraj_rel_enriched.tex}
-
-Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
-trajectory segment between the respective cells. The graph abstraction
-$G$ only states that there exists a state $\x \in C$ from which the
-system can evolve to a future state $\x' \in C'$. To increase the
-precision we iteratively refined the abstraction by state
-splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
-set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
-which non-deterministically describe relations between $\x \in C$ and $\x'
-\in C'$. This is illustrated in \figref{enriched-edge} (compare with
-\figref{segtraj}).
-
-The enriched graph $G^R$ captures the underlying local forward
-dynamics describing the evolution of the system in each abstract
-state. We represent the dynamics using an affine model with an
-interval error. Such a model can either be approximated using learning
-methods or computed as a sound (over) approximation using reachability
-set computation methods. Because we assume black box semantics, we
-only present the former. Using regression analysis on the $start$ and
-$end$ states of the witnessed trajectory segments between two cells,
-we compute an approximate discrete map along with an error estimate.
-Moreover, using the simulation function $\simulate$, additional
-trajectory segments (or data) can be generated if required. The data
-can be separated into a training set and testing set to compute the
-map and the error respectively.  The latter case can be explored if
-the symbolic dynamics of the system are known. A tool like
-\flowstar~\cite{chen2013flow} can be used to find the reachable set
-map.
-
-Observe that $G^R$, a directed reachability graph, is rich enough to
-search for concrete behaviors in the system. We call it a time
-parameterized PWA relational abstraction. It can be interpreted as an
-infinite state discrete transition system, and we can use
-off-the-shelf bounded model checkers to find concrete violations of a
-given safety property, and even other temporal properties.  We now
-discuss the background required to present our ideas.
-
-\section{Background}
-
-In this section, we present the basics of relational abstractions and
-fundamentals of regression analysis.
-
-\subsection{Relational Abstraction}
-
-For general hybrid dynamical systems, algorithmic ways for computing
-analytical solutions do not exist. Discretization (along with suitable
-abstractions) is often employed to transform the systems into a
-discrete transition system. Relational abstractions is one such idea,
-which abstracts continuous dynamics by discrete relations using
-appropriate \textit{reachability invariants}. Both time independent
-and time dependent relations have been proposed; the former captures
-all reachable states over all time, whereas, the latter explicitly
-includes time by relating reachable states to time. Thus it
-can prove timing properties whereas, time independent relational
-abstractions can not. In both cases, the resulting abstractions can be
-interpreted as discrete transition systems and can be analyzed using
-model checkers. We briefly discuss these two types of abstractions.
-
-\mypara{Timeless Relational Abstraction}
-
-Relational abstractions for hybrid systems were proposed in
-\cite{Sankaranarayanan+Tiwari/2011/Relational}. For a hybrid automaton
-model, they can summarize the continuous dynamics of each mode using a
-binary relation over continuous states. The resulting relations are
-timeless (independent) and hence valid for all time as long as
-the mode invariant is satisfied. The relations take the general form
-of $R(\x,\x') \bowtie 0$, where $\bowtie$ represents one of the
-relational operators $=, \ge, \le, <, >$. For example, an abstraction
-that captures the monotonicity with respect to time for the
-differential equation $\dot{x} = 2$ is $x' > x$. The abstraction
-capturing the relation between the set of ODEs: $\dot{x} = 2$,
-$\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
-
-Such relation summaries, discretize the continuous evolution of a
-given system into an infinite state transition system, in which safety
-properties can be verified using $k$-induction, or falsified using
-bounded model checkers. The relationalization procedure involves
-finding suitable invariants in a chosen abstract domain, such as
-affine abstractions, eigen abstractions or box abstractions
-\cite{Sankaranarayanan+Tiwari/2011/Relational}.  For this, different
-techniques like template based invariant generation
-\cite{Gulwani+Tiwari/2008/Constraint,
-Colon+Sankaranarayanan+Sipma/03/Linear} can be used.
-
-\mypara{Timed and Time-Aware Relational Abstractions}
-
-As the above discussed relations are timeless, we cannot reason over
-the timing properties of the original system. Moreover, they are not
-suitable for time-triggered systems. Timed relational
-abstractions~\cite{zutshi2012timed} and time-aware relational
-abstractions~\cite{mover2013time} were proposed to overcome these
-shortcomings by explicitly including time in the relations. In
-addition, both can be more precise than their timeless counterpart,
-but unlike it, assume continuous affine dynamics.
-
-To analyze time-triggered systems like an SDCS modeled by an affine
-hybrid automata~\footnote{An affine hybrid automata is restricted to
-the ODEs, resets and guards being affine.}, timed relational
-abstractions can be computed directly from the solutions of the
-underlying affine ODEs.  Recall that the solution of an affine ODE
-$\dot{\x} = A\x$ can be represented using the matrix exponential as
-$\x(t) = e^{tA}\x(0)$. This gives us the timed relation $\x' =
-e^{tA}\x$.  Clearly, the relation is non-linear with respect to time.
-However, for the case of SDCS with a fixed sampling time period, we
-obtain linear relations in $\x$.  We demonstrated the usefulness of
-timed relational abstractions for the case of linear systems
-in \cite{zutshi2012timed}, and now incorporate the idea to discretize
-black box dynamical systems.
-
-Similar to timeless relational abstractions, time-aware relational
-abstractions \cite{mover2013time} construct binary relations between
-the current state $\x$ of the system and any future reachable state
-$\x'$. The difference lies with the latter also constructing relations
-between the current time $t$ and any future time $t'$. This is
-achieved by a case by case analysis of the eigen structure of the
-matrix $A$ (of an affine ODE). Separate abstractions are used for the
-case of linear systems with constant rate, real eigen values and complex eigenvalues.
-
-\subsection{Learning Dynamics using Simple Linear Regression}
-
-\mypara{Regression Analysis}
-
-In statistics, regression is the problem of finding a
-\textit{predictor}, which can suitably predict the relationship
-between the given set of observed input $\x$ and output $\y$ vectors.
-In other words, assuming that $\y$ depends on $\x$, regression
-strategies find either a parameterized or a non-parameterized
-prediction function to explain the dependence. We now discuss simple
-linear regression, which is parametric in nature and searches for an
-affine predictor. It is also called \textit{ordinary least squares}.
-
-\mypara{Ordinary Least Squares (OLS)}
-
-Let the data set be comprised of $N$ input and output pairs $(\x,
-y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
-the case in the current context of \textit{finding the best fit}, the
-problem is over-determined; there are more equations than
-unknowns. Hence, a single affine function cannot be found which
-satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This is formally
-defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine the vector of
-coefficients $\vec{a}$ and an offset constant $b$, such that the least
-square error of the affine predictor is minimized for the given data
-set.
-\begin{equation}
-    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
-\label{reg}
-\end{equation}
-To ease the presentation, we can rewrite the above as a homogeneous
-expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
-$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
-\[
-    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
-\]
-The solution of OLS can be analytically computed as
-\[ Ab = (X^TX)^{-1}X^T\y\]
-where $X$ is the matrix representing the horizontal stacking of all
-$\hat{\x}$.
-The details can be found in several texts on learning and statistics
-\cite{friedman2001elements}.
-
-Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
-where $\x\in\reals^n$ we can use OLS to approximate its trajectories
-at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
-$A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
-\begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
-    global relational model for the system. The data set for the
-    regression is a set of pairs $\setof{(\x,\x') | \x' =
-    \simulate(\x, \Delta)}$ and can be generated on demand. Another
-    set can be generated to compute the error $\delta$ as an interval
-    of vectors, where each element is an interval $\delta_i \in
-    [\delta_{min}, \delta_{max}]$.
-
-    Affine maps are poor
-    approximations for arbitrary non-linear functions. Hence, we use a
-    collection of affine maps to approximate the local behaviors (in
-    state-space) of the system $\simulate$. This results in a piece
-    wise approximation, as described in the next section.
-
-\subsection{Piecewise Affine (PWA) Transition System}
-
-We define the PWA transition system as a transition system
-(ref.~\secref{DiscSys}) $\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
-where each transition $\tau \in \scrT$, is associated with a
-transition relation of the form
-\[
-    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
-\]
-where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
-$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
-The assertion on the initial states $\Theta$ is also affine.
-
-We can now use a PWA transition system $\rho$ to approximate the
-behavior of a hybrid dynamical system $\System$ defined by a simulator
-$\simulate_{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
-form, we can only approximate general hybrid systems.}. Each
-transition of $\rho$ represents a discrete time step $\Delta$. Let the
-state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
-states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
-predicates $g_\tau$ are given by a conjunction of hyper-planes and
-hence can be represented in matrix form as a polyhedron (defined by
-$m$ constraints) in the state-space
-\[
-g:C\x - \vd \le 0
-\]
-where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
-$m$. The affine maps $f_\tau$ can be represented in matrix form as
-$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
-an error term $\delta$ which is defined by a vector of intervals
-$[\delta_{\min},\delta_{\max}]$. Hence,
-\[
-f:\x \mapsto A\x + \vb + \delta
-\]
-defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
-\in f_\tau(\x)}$.
-
-% Given a state vector $\x\in\reals^n$, a guarded affine map
-% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
-% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
-% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
-% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
-% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
-% guard is satisfied.
-
-We now formalize the PWA affine transition system for a dynamical system.
-\begin{definition}[PWA Transition System]
-
-Given a hybrid dynamical system over a state-space
-$\HybridStates^n$, a PWA transition system $\rho$ is given by the
-tuple $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in
-\scrT$ are affine transition relations and $\Theta$ is an affine
-predicate over $\x$ and the states $\x\in\HybridStates^n$. The
-transition relation is then defined by $\scrT$ with $n$ transition
-relations as follows
-
-\begin{equation}
-    \scrT = \left\{
-        \begin{array}{ll}
-            g_1(\x) \land g'_1(\x') \implies \x' \in f_1(\x) \\
-            \ldots\\
-            g_n(\x) \land g'_n(\x') \implies \x' \in f_n(\x) \\
-        \end{array}
-    \right.
-\end{equation}
-
-\end{definition}
-
-where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
-$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
-denotes the next state of the system.
-
-A PWA model is \textit{deterministic}, iff for every state
-$\x\in\HybridStates^n$, a unique guarded affine map is satisfied and
-all errors $\delta$ are singletons. However, in practice such a case
-rarely exists. A PWA model will be usually non-deterministic, both due
-to multiple choice of transition relations and the reachable states at
-the $k^{th}$ time step (or after $k\Delta$ units of time) being a set.
-Abusing the notation, we denote the set of states reachable in a
-single step in $\rho$ by $\x' = \rho(\x)$.
-
-A PWA transition system is \textit{complete}, iff for every state
-$\x\in\HybridStates^n$, there exists at least one satisfied
-transition relation $\scrT$. If this is not the case, the system can
-deadlock, with no further executions. This usually results from
-incorrect modeling of SDCS, and from here on, we do not consider such
-cases.
-
-% Given a dynamical system over a continuous state-space
-% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
-% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
-% dynamical system to a finite set of guarded affine maps
-% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
-% rule by the affine map of the satisfied $\gm$.
-
-% \begin{equation}
-%     \pwa = \left\{
-%         \begin{array}{ll}
-%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
-%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
-%             \ldots & \ldots\\
-%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
-%         \end{array}
-%     \right.
-% \end{equation}
-
-% \end{definition}
-
-% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
-% Abusing the notation, we denote the next state computed by the PWA
-% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
-% for every state $\x\in\ContStates$, a unique guarded affine map is
-% satisfied. A PWA model is \textit{complete}, iff for every state
-% $\x\in\ContStates$, there exists at least one satisfied guarded affine
-% map $\gm$.
-
-\section{Relational Modeling}
-
-We now describe how relational models can be computed for a given
-black box system by enriching the abstract reachability graph obtained
-by scatter-and-simulate. We first formalize the notion of an enriched
-graph, and then show how they can be interpreted as a PWA transition
-system using $k$-relational modeling.
-
-\subsection{Abstract Enriched Graph}
-
-The existential abstraction relation in \chapref{abs} was defined as
-follows:
-\[
-%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
-    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
-\]
-The abstract relation $\areach{t}$ can be enriched by incorporating
-the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
-system, such a relation can be rarely represented using an exact affine map.
-This is due to the presence of non-linear and hybrid behaviors. But,
-an affine map can always be estimated with an error.
-
-If the system dynamics are completely specified in the form of a white
-box model, we can use first order approximations to find the affine
-expressions for the relations. Using a tool like \flowstar, we can
-obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
-\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
-that, every element $b_i$ of the vector $\vb$ is contained in the
-respective scalar interval $b_i\in[b^l_i,b^h_i]$.
-
-For the case of black box systems, such a sound approximation is not
-possible. Instead, we rely on a statistical method like simple linear
-regression to estimate the affine map $\amap: A\x + \vb$. We then
-estimate the error $\delta$ and generalize $\amap$ to an interval
-affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
-abstraction with the below relation
-\[
-    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
-    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
-\]
-
-However, in the presence of complex non-linear behavior, using a
-single affine map can lead to a poor approximation. To increase the
-precision, one can use more than one affine map. Let us denote this set as
-$R_{(C,C')}$ (ref.~\secref{ovw}).
-\[
-    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
-\]
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
-    \draw[arw] (-3,0) -- (-1.2,0);
-    \draw[arw] (1.2,0) -- (3,0);
-
-\node at (-3.4,0) {$G$};
-\node at (3.4,0) {$G^R$};
-\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Using OLS, $G^R$ is computed by determining the appropriate
-$R_{(C,C')}$ for each edge of $G$.}
-\label{fig:enrichment}
-\vspace*{-.3cm}
-\end{figure}
-
-As shown in \figref{enrichment}, we use OLS to compute an enrichment
-of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
-the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
-of the enriched edge are clearly non-deterministic. Form a state $\x
-\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
-is satisfied. This interpretation results in an infinite state
-transition system. We now detail the construction of the set
-$R_{(C,C')}$ using $k$-relational modeling.
-
-% We now compute a PWA model for a given black box system by estimating
-% the dynamics using OLS. We can estimate the affine relations between
-% every cell, but as discussed in ~\chapref{abs}, it is a futile
-% approach.  Instead, we use the same heuristic as before:
-% scatter-and-simulate, to select the cell relations for estimation. We
-% build the reachability graph as before, but in addition, annotate each
-% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
-% respective relation. The reachability graph thus computed, is a
-% transition system. Using off-the-shelf bounded model checkers, we
-% reason about the system's safety properties.
-
-\subsection{$k$ - Relational Modeling}
-
-\begin{figure}[!htbp]
-\vspace*{-.2cm}
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-\begin{scope}[scale=1]
-\draw [line] (-1.0,0) circle (0.5);
-\draw [line] (-4.0,0) circle (0.5);
-\draw [line] (2.0,1) circle (0.5);
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-0.5,0) -- (1.5,1.0);
-\draw[arw] (-0.5,0) -- (1.5,-1.0);
-\draw[arw] (-3.5,0) -- (-1.5,0.0);
-\node at (-4.0,0) {$C$};
-\node at (-1.0,0) {$C'$};
-\node at (2.0,1.0) {$C''_1$};
-\node at (2.0,-1.0) {$C''_2$};
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
-\node at (-4.0,-2.5) {$k=0$};
-\node at (-1.0,-2.5) {$k=1$};
-\node at (+2.0,-2.5) {$k=2$};
-\draw [fill=black] (-4.0,-2.0) circle (0.05);
-\draw [fill=black] (-1.0,-2.0) circle (0.05);
-\draw [fill=black] (2.0,-2.0) circle (0.05);
-
-\end{scope}
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
-\label{fig:k-rel}
-\vspace*{-.3cm}
-\end{figure}
-
-Given a $G$, when computing the set of relations for an edge $R$, we
-consider only the set of abstract states reachable at a specific time
-step.  Intuitively, one might consider the states reachable in one
-time step $\Delta$.  However, such a process can be generalized by
-looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
-k\Delta}$ time steps.
-
-Using \figref{k-rel}, we illustrate this notion. To compute
-$R_{(C,C')}$, we
-split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
-$k\Delta$) as follows.  We observe the local behavior of the system by
-noting the evolution of the system $\System$ from a state $\x(t)
-\in C$ for a time length dependant on $k$. For
-\begin{itemize}
-    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
-    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
-    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
-        and split them in two sets (a) and (b) on the basis of the
-        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
-        \begin{enumerate}[(a)]
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
-        \end{enumerate}
-    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
-        and split them in to multiple sets on the basis of the cells
-        reached at time $(t+\Delta),\; (t+2\Delta),\;
-        \ldots,\;(t+n\Delta)$.
-\end{itemize}
-
-$k$-relational modeling can be understood as a heuristic, which uses
-the underlying abstraction to differentiate behaviors of the system
-which `diverge' (or are revealed to be `distinct' at a future time).
-Such a heuristic can be useful in increasing the precision of the
-learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
-$n$ and illustrate using examples.
-
-\mypara{$0$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=1]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
-\node at (0,-2.5) {$t=0$};
-\node at (4,-2.5) {$t=\Delta$};
-\draw [fill=black] (0.0,-2.0) circle (0.05);
-\draw [fill=black] (4.0,-2.0) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=3.0cm,scale=1]
-    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
-    \implies A_1\vx + \vb_1 + \delta_1$};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{All the trajectory segments will be used to construct the
-model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
-\label{fig:k0}
-\vspace*{-.3cm}
-\end{figure}
-
-When $k=0$, the $0$-relational model is a PWA transition system which only
-defines the evolution of states $\x$, and does not specify the
-reachable cell. The guard predicates of the relations are defined over
-cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.
-
-We use regression to estimate the dynamics for the outgoing trajectory
-segments form a cell $C$. Hence, the data set $\ds$ for the regression
-includes all trajectory segments beginning from the same cell
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C}.
-\]
-This includes trajectory segments ending in different cells, as shown
-in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
-\begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            \ldots \\
-            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
-        \end{array}
-    \right.
-\end{equation*}
-
-Note that this can be quite imprecise when the cells are big,
-containing regions of state-space with complex dynamics. This is true
-for both non-linear systems and hybrid dynamical system, where a cell
-can contain two or more modes with differing continuous dynamics.
-
-\mypara{$1$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=1]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (2,0) {\begin{minipage}{\textwidth}
-    \footnotesize \begin{equation*}
-        \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
-    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
-    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
-%Cardinality($R_{(C,C')}$) = $2$.}
-\label{fig:k1}
-\vspace*{-.3cm}
-\end{figure}
-
-% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-%     \node at (2,0) {\footnotesize $\begin{equation*}
-%     \scr{T} = \left\{
-%         \begin{array}{ll}
-%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-%         \end{array}
-%     \right.
-% \end{equation*}$};
-% \end{scope}
-
-To improve the preciseness of learnt dynamics, we include the
-reachability relation in the
-regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
-comprised only of trajectory segments $\pi_t$ which start and end in
-the same set of cells.
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
-\]
-For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
-$N$ transition relations.
-
-\figref{k1} illustrates the $k=1$ refinement of the case shown in
-\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
-= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
-relations are constructed. The resulting $1$-relational model
-is at least as precise as the corresponding $0$-relational model.
-
-\mypara{$k$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
-    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;
-
-%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-
-    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (8.0, 2.0) {$C''_1$};
-\node at (8.0, -2.0) {$C''_2$};
-%\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
-\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\node at (8,-4.0) {$t=2\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-\draw [fill=black] (8.0,-3.5) circle (0.05);
-\end{scope}
-
-\begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-    \node at (4,0) {\begin{minipage}{\textwidth}
-        \footnotesize \begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The $k=2$ refinement further splits the data set $\ds_1$
-    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
-    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
-    non-deterministically defining the system behavior.}
-\label{fig:k2}
-\vspace*{-.3cm}
-\end{figure}
-
-Finally, we generalize the relational models to $k$-relational PWA
-models. A $k$-relational model is constructed by using $k$ length
-\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
-\textit{connected} iff its cost $\cost(\SegTraj) = 0$.
-
-Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
-$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
-\textit{similar} if their trajectory segments have the same sequence
-of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
-\[
-    \forall i\in\setof{1 \ldots k}.
-    start(\pi_{t_i}) \in C_i
-    \land start(\pi'_{t_i}) \in C_i
-    \land end(\pi_{t_i}) \in C_{i+1}
-    \land end(\pi'_{t_i}) \in C_{i+1}
-\]
-
-For every cell $C$ in the reachability graph, we construct a data set
-$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
-length \textit{connected} and \textit{similar} segmented trajectories.
-%\[
-%    \ds = \setof{\pi_{\tau_i} |
-%        \pi_{\tau_i}
-%        start(\pi_{\tau_{i}}) \in C_i
-%        \land
-%        end(\pi_{\tau_{i}}) \in C_i}
-%\]
-
-\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.
-
-\section{Bounded Model Checking Black Box Systems}
-% include an algorithm
-We now describe the entire procedure in three steps.
-\begin{enumerate}
-    \item Given a black box system $\System$ specified by
-        $\simulate_\System$, we
-        use scatter-and-simulate with ($\Delta$-length) trajectory
-        segments to sample a finite reachability graph
-        $\scr{H}(\Delta)$.
-    \item Using regression, we enrich the graph relations and annotate
-        the edges with the estimated affine maps.
-    \item Interpreting the directed reachability graph as a transition
-        system, we use a bounded model checker to find fixed length
-        violations to safety properties.
-\end{enumerate}
-
-\section{An Example: Van der Pol Oscillator}
-
-We now describe the above using the Van der Pol oscillator benchmark
-presented in \secref{math-dyn-sys}. Simulations generated using
-uniformly random samples are shown in \figref{vdp-cont}. We want to
-check the property $P_3$, indicated by the red box, given the initial
-set indicated by the green box. The abstraction is defined by
-$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
-where each cell is of size $0.2 \times 0.2$ units.
-Scatter-and-simulate is then used to construct the abstract graph $G$,
-using $2$ samples per cell and the time step $\Delta = 0.1s$. The
-complete process follows.
-
-\begin{enumerate}
-    \item{\emph{Abstract}} Consider an implicit abstraction induced by the quantization
-    function $\quant_\epsilon(\x)$. The corresponding reachability
-    graph $\scr{H}(\Delta)$ (with time step $\Delta$) is given by
-    $\tupleof{\scr{C},\areach{\Delta}, \scr{C}_0, \scr{C}_u}$.
-\item{\emph{Discover}}: Using scatter-and-simulate, enumerate a finite number
-    of cells (vertices) and the relations (edges) of the graph
-    $\scr{H}(\Delta)$. Associate the set of generated trajectory
-    segments with their respective originating cells using a map $D:C
-    \mapsto \setof{\pi | start(\pi) \in C}$. The discovered
-    abstraction is show in \figref{vdp-abs}. As mentioned, red cells
-    are unsafe and green cells are initial cells.
-\item{\emph{Relationalize}} For each cell $C$, perform regression analysis on
-    the respective set of trajectory segments $D(C)$, and compute a
-    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
-    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(\Delta)$. Annotate
-    each edge in the graph $\scr{H}(\Delta)$ with the respective
-    relation.
-
-    \figref{vdp-map} shows the cells and the trajectory segments which
-    are part of the data sets constructed using the $1$-relational
-    modeling. Against each cell, its unique identifier (integer
-    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
-    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
-    relations. Note that self loops result when an observed trajectory
-    segment has its $start$ and $end$ states in the same cell.
-
-\item{\emph{Model Check}} The graph $\scr{H}(\Delta)$ can now be viewed as a
-    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
-    where $C \in vertices(\scr{H}(\Delta))$ and $\x \in
-    \HybridStates$. A transition $\tau \in \scrT$ is of the form
-    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
-    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
-    edge(\scr{H}(\Delta))}$.
-\item{\emph{Check Counter-example}} The infinite state (but finite location) transition system can
-    be model checked to find a concrete counter-example, which if
-    exists, can indicate the existence of a similar trace in the
-    original black-box system $\System$. The latter check is carried
-    out as before, using the numerical simulation function
-    $\simulate$. For the given example, the model checker is unable to
-    find a counter-example.
-\end{enumerate}
-
-\inclfig[width=0.75\textwidth]{vdp_cont_traces.png}{Van der
-Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}
-\label{fig:vdp-cont}
-
-\inclfig[width=0.75\textwidth]{vdp_abs.png}{The discovered
-abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
-initial cells.}
-\label{fig:vdp-abs}
-
-\inclfig[width=\textwidth]{vdp_disc_map.png}{Cells and
-trajectory segments used by $1$-relational modeling.}
-\label{fig:vdp-map}
-
-\inclfig[width=0.40\textwidth]{vdp_graph.png}{Enriched
-graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}
-\label{fig:vdp-graph}
-
-\begin{table*}[!htbp]
-\centering
-\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
-\label{tab:vdp-pwa}
-\begin{tabular}{@{}ccccc@{}}
-\toprule
-$C$ & $C'$ & $A$ & $b$ & $\delta$\\
-\midrule
-(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
-\midrule
-(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
-\midrule
-(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% (0, -1)   & (0, -1)   &  \begin{bmatrix}0.9944498 & 0.12911959 \\-0.12013468&1.63250531\end{bmatrix} &\begin{bmatrix}-0.00003683\\-0.00085411\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -1)   & (0, -2)   &  \begin{bmatrix}0.99523928& 0.1290447  \\-0.10236736&1.6314103 \end{bmatrix} &\begin{bmatrix}-0.00013019\\-0.00289192\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -2)   & (0, -2)   &  \begin{bmatrix}0.99565348& 0.12916002 \\-0.09352462&1.6341424 \end{bmatrix} &\begin{bmatrix}-0.00015574\\-0.00339664\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -2)   & (0, -4)   &  \begin{bmatrix}0.99668599& 0.12892455 \\-0.0725035 &1.62937613\end{bmatrix} &\begin{bmatrix}-0.00037391\\-0.00796189\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -9)   & (-1, -15) &  \begin{bmatrix}0.99985362& 0.12858788 \\-0.09447884&1.61476417\end{bmatrix} &\begin{bmatrix}-0.0018997 \\-0.04200363\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-1, -15) & (-3, -24) &  \begin{bmatrix}0.95473217& 0.12473489 \\-1.29198505&1.47155614\end{bmatrix} &\begin{bmatrix}-0.01248412\\-0.41100129\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-3, -24) & (-4, -29) &  \begin{bmatrix}1.11789749&-0.17261206 \\-1.71017779&0.66006597\end{bmatrix} &\begin{bmatrix}-1.0745344 \\-3.31246863\end{bmatrix}&\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}\\
-% \midrule
-% (-4, -29) & (-4, -29) &  \begin{bmatrix}0.99796588& 0.01012817 \\-0.41529747&1.02432211\end{bmatrix} &\begin{bmatrix}-0.00135199\\-0.28016983\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-1, -10) & (-2, -16) &  \begin{bmatrix}0.97381953& 0.12768367 \\-0.71242741&1.56764585\end{bmatrix} &\begin{bmatrix}-0.00320265\\-0.11964102\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% (-1, -10) & (-2, -15) &  \begin{bmatrix}0.97373058& 0.12859163 \\-0.70173855&1.58540872\end{bmatrix} &\begin{bmatrix}-0.0015415 \\-0.08616376\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-2, -16) & (-3, -23) &  \begin{bmatrix}0.92850123& 0.12242812 \\-1.84720528&1.39159826\end{bmatrix} &\begin{bmatrix}-0.02334236\\-0.73348865\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -23) & (-5, -29) &  \begin{bmatrix}3.14947585&-0.81150646 \\-0.96159406&0.23543254\end{bmatrix} &\begin{bmatrix}-3.13860326\\-4.99696623\end{bmatrix}&\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}\\
-% \midrule
-% (-2, -15) & (-3, -22) &  \begin{bmatrix}0.93156623& 0.12260264 \\-1.75117219&1.40104348\end{bmatrix} &\begin{bmatrix}-0.02214125\\-0.68423569\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -22) & (-5, -29) &  \begin{bmatrix}3.08806462&-0.72332776 \\-1.69271079&0.40308293\end{bmatrix} &\begin{bmatrix}-2.77908467\\-4.57411339\end{bmatrix}&\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}\\
-
-\subsection{Search Parameters}
-
-The search parameters for S3CAM-R include the parameters of S3CAM:
-$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
-maximum error budget $\delta_{max}$ for OLS and the maximum length of
-segmented trajectory for building $k$-relational models.
-
-We have already discussed the effects of $N$, $\epsilon$ and $\Delta$
-on S3CAM's performance. However, they also have an effect on
-relational modeling. A finer grid with small cells produces more
-accurate models than a coarser grid with large cells. Similarly,
-small time length trajectory segments result in more accurate
-modeling.
-%An increase in the number of samples can increases the
-%accuracy of regression.
-
-\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
-Given a $\delta_{max}$, we keep increasing $k$ during the
-$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
-satisfied. A $k_{max}$ is introduced to bound the longest
-segmented trajectory which can be considered.
-
-% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
-
-\subsection{Reasons for Failure}
-The approach can fail to find a counter-example, even when it exists, in
-one of three ways.
-
-\begin{itemize}
-
-    \item \emph{S3CAM Fails} No abstract counter-example is found by
-        S3CAM. We can remedy it by increasing search budgets and/or
-        restarting.
-
-    \item \emph{BMC Fails} An abstract counter-example is found, but
-        the BMC fails to find a concrete counter-example in the PWA
-        relational model. The failure can be attributed to either (a)
-        a spurious abstract counter-examples or (b) a poorly estimated
-        model. The former can be addressed by restarting but the
-        latter requires that the maximum model error $\delta_{max}$ be
-        decreased.
-
-    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
-        is found, and it is successfully concretized in the PWA
-        relational model by the BMC. However, it is not reproducible
-        in the black box system. This happens when the PWA relational
-        model is not precise enough.
-
-\end{itemize}
-
-\section{Implementation and Evaluation}
-
 The implementation was prototyped as S3CAM-R, an extension to our
@@
 solvers) can not handle under reasonable resources.
-
-\section{Conclusion and Future Work}
-
-We have presented another methodology to find falsifications in black
-box dynamical systems. Combining the ideas from abstraction based
-search (\chapref{abs}), with our previous relational
-abstractions~\cite{zutshi2012timed}, we proposed $k$-relational
-modeling to compute richer abstractions. Simple linear regression can
-be used to estimate the local dynamics of the trajectory segments and
-compute PWA relational models which can be interpreted as a transition
-system.  These can then be checked for safety violations using an
-off-the-shelf bounded model checker.  Finally, we implemented the
-ideas as a tool S3CAM-R and demonstrated its performance on a few examples.
-
-\subsection{Improvements}
-
-The approach seems promising as it avoids an expensive refinement
-step. But, before it can be applied to more complex systems, we need
-to overcome its shortcomings.  As a future extension, we are
-investigating ideas to improve the performance of the underlying three
-main sub-processes, namely, abstraction, modeling, and BMC. We detail
-these below.
-
-\begin{enumerate}
-
-\item \emph{Abstraction} The abstract domain we have used, has been
-    confined to the interval or rectangular domain. Such a domain is
-    very coarse when compared to more general polyhedral domains, using
-    which, we can gain better precision. Due to polyhedral domains
-    being computational expensive, their integration in S3CAM needs to
-    be explored thoroughly.
-
-\item \emph{Modeling} We have used $k$-relational modeling to increase
-    the precision of the enriched abstractions, however, there are
-    several other refinement strategies that remain to be explored.
-    This includes refining the abstract states by using guard
-    predicates (similar to predicate abstraction), using non-linear
-    templates in parametric regression to compute more precise
-    quadratic relations, and using an adaptive time step instead of a
-    fixed one to address the non-linearities due to a long time
-    horizon.
-
-\item \emph{Model Checking} As of now, we use SMT solvers to search
-    for concrete counter-examples. Although they are very efficient,
-    owing to extensive engineering effort, reachability of transition
-    systems resulting from dynamical systems remains a difficult
-    problem. As the time horizon of the safety property increases, the
-    possible combinations of discrete transitions increase
-    exponentially. Hence, to find a counterexample which is a sequence
-    of discrete transitions over a `long' time horizon is not
-    tractable for most, but the simplest of dynamical systems.
-    Instead, we can use linear programming solvers, by enumerating
-    each path in the graph $G^R$. Note that the constraints are all
-    linear (conjunctions) along an abstract path. However, in the
-    worst case, the number of paths in a graph can be of the order
-    $n!$ where $n$ is the number of vertices of a graph. Hence, this
-    will not be feasible, unless, we can prioritize paths by using a
-    triage process similar to one used by S3CAM and using a budget on
-    the maximum number of paths.
-
-    Another approach to address the issue would be to use an adaptive
-    time discretization technique, where relations over both shorter
-    and longer time steps are computed, and the SMT solver can
-    `select' the time step precise enough to find a counter-example.
-
-\item \emph{SMT Solvers} One major impediment to our approach is the
-    fact that SMT solvers use the theory of reals with \emph{exact
-    precision}. This is important for verification approaches, but can
-    be relaxed for the problem of falsification. An SMT solver which
-    uses approximate reasoning but returns robust counter-examples
-    will be as useful, and perhaps more efficient.
-
-\end{enumerate}
-
-\subsection{Data Driven Analysis}
-
-Finally, we would like to mention that our approach is `simulation'
-driven. It can be easily modified to be `data driven', by working with
-a fixed set of data. Given a data set, we then need to automatically
-find a good abstraction and a high fidelity PWA transition system
-using which, we can summarize the black box hybrid dynamical system.
-Apart from model checking the transition system, one can extend it to
-the analysis of SDCS. More specifically, by combining the transition
-system model of a plant with the control software, a model checker
-like CBMC \cite{kroening2014cbmc} can be used to do a closed loop
-symbolic analysis of the SDCS. This can be an alternative to S3CAM-X.
-
-%which will be
-%very useful to for analyzing data for potential property violations.

### 2017-02-08 intro.tex
.

@@
 % properties of SDCS.
-
-\section{Relational Modeling}
-
-We now describe how relational models can be computed for a given
-black box system by enriching the abstract reachability graph obtained
-by scatter-and-simulate. We first formalize the notion of an enriched
-graph, and then show how they can be interpreted as a PWA transition
-system using $k$-relational modeling.
-
-\subsection{Abstract Enriched Graph}
-
-The existential abstraction relation in \chapref{abs} was defined as
-follows:
-\[
-%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
-    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
-\]
-The abstract relation $\areach{t}$ can be enriched by incorporating
-the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
-system, such a relation can be rarely represented using an exact affine map.
-This is due to the presence of non-linear and hybrid behaviors. But,
-an affine map can always be estimated with an error.
-
-If the system dynamics are completely specified in the form of a white
-box model, we can use first order approximations to find the affine
-expressions for the relations. Using a tool like \flowstar, we can
-obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
-\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
-that, every element $b_i$ of the vector $\vb$ is contained in the
-respective scalar interval $b_i\in[b^l_i,b^h_i]$.
-
-For the case of black box systems, such a sound approximation is not
-possible. Instead, we rely on a statistical method like simple linear
-regression to estimate the affine map $\amap: A\x + \vb$. We then
-estimate the error $\delta$ and generalize $\amap$ to an interval
-affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
-abstraction with the below relation
-\[
-    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
-    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
-\]
-
-However, in the presence of complex non-linear behavior, using a
-single affine map can lead to a poor approximation. To increase the
-precision, one can use more than one affine map. Let us denote this set as
-$R_{(C,C')}$ (ref.~\secref{ovw}).
-\[
-    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
-\]
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
-    \draw[arw] (-3,0) -- (-1.2,0);
-    \draw[arw] (1.2,0) -- (3,0);
-
-\node at (-3.4,0) {$G$};
-\node at (3.4,0) {$G^R$};
-\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Using OLS, $G^R$ is computed by determining the appropriate
-$R_{(C,C')}$ for each edge of $G$.}
-\label{fig:enrichment}
-\vspace*{-.3cm}
-\end{figure}
-
-As shown in \figref{enrichment}, we use OLS to compute an enrichment
-of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
-the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
-of the enriched edge are clearly non-deterministic. Form a state $\x
-\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
-is satisfied. This interpretation results in an infinite state
-transition system. We now detail the construction of the set
-$R_{(C,C')}$ using $k$-relational modeling.
-
-% We now compute a PWA model for a given black box system by estimating
-% the dynamics using OLS. We can estimate the affine relations between
-% every cell, but as discussed in ~\chapref{abs}, it is a futile
-% approach.  Instead, we use the same heuristic as before:
-% scatter-and-simulate, to select the cell relations for estimation. We
-% build the reachability graph as before, but in addition, annotate each
-% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
-% respective relation. The reachability graph thus computed, is a
-% transition system. Using off-the-shelf bounded model checkers, we
-% reason about the system's safety properties.
-
-\subsection{$k$ - Relational Modeling}
-
-\begin{figure}[!htbp]
-\vspace*{-.2cm}
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-\begin{scope}[scale=0.8]
-\draw [line] (-1.0,0) circle (0.5);
-\draw [line] (-4.0,0) circle (0.5);
-\draw [line] (2.0,1) circle (0.5);
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-0.5,0) -- (1.5,1.0);
-\draw[arw] (-0.5,0) -- (1.5,-1.0);
-\draw[arw] (-3.5,0) -- (-1.5,0.0);
-\node at (-4.0,0) {$C$};
-\node at (-1.0,0) {$C'$};
-\node at (2.0,1.0) {$C''_1$};
-\node at (2.0,-1.0) {$C''_2$};
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
-\node at (-4.0,-2.5) {$k=0$};
-\node at (-1.0,-2.5) {$k=1$};
-\node at (+2.0,-2.5) {$k=2$};
-\draw [fill=black] (-4.0,-2.0) circle (0.05);
-\draw [fill=black] (-1.0,-2.0) circle (0.05);
-\draw [fill=black] (2.0,-2.0) circle (0.05);
-
-\end{scope}
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
-\label{fig:k-rel}
-\vspace*{-.3cm}
-\end{figure}
-
-Given a $G$, when computing the set of relations for an edge $R$, we
-consider only the set of abstract states reachable at a specific time
-step.  Intuitively, one might consider the states reachable in one
-time step $\Delta$.  However, such a process can be generalized by
-looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
-k\Delta}$ time steps.
-
-Using \figref{k-rel}, we illustrate this notion. To compute
-$R_{(C,C')}$, we
-split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
-$k\Delta$) as follows.  We observe the local behavior of the system by
-noting the evolution of the system $\System$ from a state $\x(t)
-\in C$ for a time length dependant on $k$. For
-\begin{itemize}
-    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
-    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
-    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
-        and split them in two sets (a) and (b) on the basis of the
-        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
-        \begin{enumerate}[(a)]
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
-        \end{enumerate}
-    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
-        and split them in to multiple sets on the basis of the cells
-        reached at time $(t+\Delta),\; (t+2\Delta),\;
-        \ldots,\;(t+n\Delta)$.
-\end{itemize}
-
-$k$-relational modeling can be understood as a heuristic, which uses
-the underlying abstraction to differentiate behaviors of the system
-which `diverge' (or are revealed to be `distinct' at a future time).
-Such a heuristic can be useful in increasing the precision of the
-learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
-$n$ and illustrate using examples.
-
-\mypara{$0$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
-\node at (0,-2.5) {$t=0$};
-\node at (4,-2.5) {$t=\Delta$};
-\draw [fill=black] (0.0,-2.0) circle (0.05);
-\draw [fill=black] (4.0,-2.0) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=3.0cm,scale=0.5]
-    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
-    \implies A_1\vx + \vb_1 + \delta_1$};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{All the trajectory segments will be used to construct the
-model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
-\label{fig:k0}
-\vspace*{-.3cm}
-\end{figure}
-
-When $k=0$, the $0$-relational model is a PWA transition system which only
-defines the evolution of states $\x$, and does not specify the
-reachable cell. The guard predicates of the relations are defined over
-cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.
-
-We use regression to estimate the dynamics for the outgoing trajectory
-segments form a cell $C$. Hence, the data set $\ds$ for the regression
-includes all trajectory segments beginning from the same cell
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C}.
-\]
-This includes trajectory segments ending in different cells, as shown
-in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
-\begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            \ldots \\
-            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
-        \end{array}
-    \right.
-\end{equation*}
-
-Note that this can be quite imprecise when the cells are big,
-containing regions of state-space with complex dynamics. This is true
-for both non-linear systems and hybrid dynamical system, where a cell
-can contain two or more modes with differing continuous dynamics.
-
-\mypara{$1$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
-    \node at (2,0) {\begin{minipage}{\linewidth}
-    \footnotesize \begin{equation*}
-        \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
-    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
-    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
-%Cardinality($R_{(C,C')}$) = $2$.}
-\label{fig:k1}
-\vspace*{-.3cm}
-\end{figure}
-
-% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-%     \node at (2,0) {\footnotesize $\begin{equation*}
-%     \scr{T} = \left\{
-%         \begin{array}{ll}
-%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-%         \end{array}
-%     \right.
-% \end{equation*}$};
-% \end{scope}
-
-To improve the preciseness of learnt dynamics, we include the
-reachability relation in the
-regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
-comprised only of trajectory segments $\pi_t$ which start and end in
-the same set of cells.
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
-\]
-For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
-$N$ transition relations.
-
-\figref{k1} illustrates the $k=1$ refinement of the case shown in
-\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
-= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
-relations are constructed. The resulting $1$-relational model
-is at least as precise as the corresponding $0$-relational model.
-
-\mypara{$k$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
-    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;
-
-%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-
-    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (8.0, 2.0) {$C''_1$};
-\node at (8.0, -2.0) {$C''_2$};
-%\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
-\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\node at (8,-4.0) {$t=2\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-\draw [fill=black] (8.0,-3.5) circle (0.05);
-\end{scope}
-
-\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
-    \node at (4,0) {\begin{minipage}{\linewidth}
-        \footnotesize \begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The $k=2$ refinement further splits the data set $\ds_1$
-    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
-    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
-    non-deterministically defining the system behavior.}
-\label{fig:k2}
-\vspace*{-.3cm}
-\end{figure}
-
-Finally, we generalize the relational models to $k$-relational PWA
-models. A $k$-relational model is constructed by using $k$ length
-\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
-\textit{connected} iff its cost $\cost(\SegTraj) = 0$.
-
-Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
-$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
-\textit{similar} if their trajectory segments have the same sequence
-of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
-\[
-    \forall i\in\setof{1 \ldots k}.
-    start(\pi_{t_i}) \in C_i
-    \land start(\pi'_{t_i}) \in C_i
-    \land end(\pi_{t_i}) \in C_{i+1}
-    \land end(\pi'_{t_i}) \in C_{i+1}
-\]
-
-For every cell $C$ in the reachability graph, we construct a data set
-$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
-length \textit{connected} and \textit{similar} segmented trajectories.
-%\[
-%    \ds = \setof{\pi_{\tau_i} |
-%        \pi_{\tau_i}
-%        start(\pi_{\tau_{i}}) \in C_i
-%        \land
-%        end(\pi_{\tau_{i}}) \in C_i}
-%\]
-
-\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.
-
-\section{Bounded Model Checking Black Box Systems}
-% include an algorithm
-We now describe the entire procedure in three steps.
-\begin{enumerate}
-    \item Given a black box system $\System$ specified by
-        $\simulate_\System$, we
-        use scatter-and-simulate with ($\Delta$-length) trajectory
-        segments to sample a finite reachability graph
-        $\scr{H}(\Delta)$.
-    \item Using regression, we enrich the graph relations and annotate
-        the edges with the estimated affine maps.
-    \item Interpreting the directed reachability graph as a transition
-        system, we use a bounded model checker to find fixed length
-        violations to safety properties.
-\end{enumerate}
-
-\section{An Example: Van der Pol Oscillator}
-
-We now describe the above using the Van der Pol oscillator benchmark
-presented in \secref{math-dyn-sys}. Simulations generated using
-uniformly random samples are shown in \figref{vdp-cont}. We want to
-check the property $P_3$, indicated by the red box, given the initial
-set indicated by the green box. The abstraction is defined by
-$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
-where each cell is of size $0.2 \times 0.2$ units.
-Scatter-and-simulate is then used to construct the abstract graph $G$,
-using $2$ samples per cell and the time step $\Delta = 0.1s$. The
-complete process follows.
-
-\begin{enumerate}
-    \item{\emph{Abstract}} Consider an implicit abstraction induced by the quantization
-    function $\quant_\epsilon(\x)$. The corresponding reachability
-    graph $\scr{H}(\Delta)$ (with time step $\Delta$) is given by
-    $\tupleof{\scr{C},\areach{\Delta}, \scr{C}_0, \scr{C}_u}$.
-\item{\emph{Discover}}: Using scatter-and-simulate, enumerate a finite number
-    of cells (vertices) and the relations (edges) of the graph
-    $\scr{H}(\Delta)$. Associate the set of generated trajectory
-    segments with their respective originating cells using a map $D:C
-    \mapsto \setof{\pi | start(\pi) \in C}$. The discovered
-    abstraction is show in \figref{vdp-abs}. As mentioned, red cells
-    are unsafe and green cells are initial cells.
-\item{\emph{Relationalize}} For each cell $C$, perform regression analysis on
-    the respective set of trajectory segments $D(C)$, and compute a
-    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
-    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(\Delta)$. Annotate
-    each edge in the graph $\scr{H}(\Delta)$ with the respective
-    relation.
-
-    \figref{vdp-map} shows the cells and the trajectory segments which
-    are part of the data sets constructed using the $1$-relational
-    modeling. Against each cell, its unique identifier (integer
-    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
-    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
-    relations. Note that self loops result when an observed trajectory
-    segment has its $start$ and $end$ states in the same cell.
-
-\item{\emph{Model Check}} The graph $\scr{H}(\Delta)$ can now be viewed as a
-    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
-    where $C \in vertices(\scr{H}(\Delta))$ and $\x \in
-    \HybridStates$. A transition $\tau \in \scrT$ is of the form
-    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
-    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
-    edge(\scr{H}(\Delta))}$.
-\item{\emph{Check Counter-example}} The infinite state (but finite location) transition system can
-    be model checked to find a concrete counter-example, which if
-    exists, can indicate the existence of a similar trace in the
-    original black-box system $\System$. The latter check is carried
-    out as before, using the numerical simulation function
-    $\simulate$. For the given example, the model checker is unable to
-    find a counter-example.
-\end{enumerate}
-
-\inclfig[width=0.75\linewidth]{vdp_cont_traces.png}{Van der
-Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}
-\label{fig:vdp-cont}
-
-\inclfig[width=0.75\linewidth]{vdp_abs.png}{The discovered
-abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
-initial cells.}
-\label{fig:vdp-abs}
-
-\inclfig[width=\linewidth]{vdp_disc_map.png}{Cells and
-trajectory segments used by $1$-relational modeling.}
-\label{fig:vdp-map}
-
-\inclfig[width=0.40\linewidth]{vdp_graph.png}{Enriched
-graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}
-\label{fig:vdp-graph}
-
-\begin{table*}[!htbp]
-\centering
-\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
-\label{tab:vdp-pwa}
-\begin{tabular}{@{}ccccc@{}}
-\toprule
-$C$ & $C'$ & $A$ & $b$ & $\delta$\\
-\midrule
-(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
-\midrule
-(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
-\midrule
-(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% (0, -1)   & (0, -1)   &  \begin{bmatrix}0.9944498 & 0.12911959 \\-0.12013468&1.63250531\end{bmatrix} &\begin{bmatrix}-0.00003683\\-0.00085411\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -1)   & (0, -2)   &  \begin{bmatrix}0.99523928& 0.1290447  \\-0.10236736&1.6314103 \end{bmatrix} &\begin{bmatrix}-0.00013019\\-0.00289192\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -2)   & (0, -2)   &  \begin{bmatrix}0.99565348& 0.12916002 \\-0.09352462&1.6341424 \end{bmatrix} &\begin{bmatrix}-0.00015574\\-0.00339664\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% (0, -2)   & (0, -4)   &  \begin{bmatrix}0.99668599& 0.12892455 \\-0.0725035 &1.62937613\end{bmatrix} &\begin{bmatrix}-0.00037391\\-0.00796189\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (0, -9)   & (-1, -15) &  \begin{bmatrix}0.99985362& 0.12858788 \\-0.09447884&1.61476417\end{bmatrix} &\begin{bmatrix}-0.0018997 \\-0.04200363\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-1, -15) & (-3, -24) &  \begin{bmatrix}0.95473217& 0.12473489 \\-1.29198505&1.47155614\end{bmatrix} &\begin{bmatrix}-0.01248412\\-0.41100129\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-3, -24) & (-4, -29) &  \begin{bmatrix}1.11789749&-0.17261206 \\-1.71017779&0.66006597\end{bmatrix} &\begin{bmatrix}-1.0745344 \\-3.31246863\end{bmatrix}&\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}\\
-% \midrule
-% (-4, -29) & (-4, -29) &  \begin{bmatrix}0.99796588& 0.01012817 \\-0.41529747&1.02432211\end{bmatrix} &\begin{bmatrix}-0.00135199\\-0.28016983\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-1, -10) & (-2, -16) &  \begin{bmatrix}0.97381953& 0.12768367 \\-0.71242741&1.56764585\end{bmatrix} &\begin{bmatrix}-0.00320265\\-0.11964102\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% (-1, -10) & (-2, -15) &  \begin{bmatrix}0.97373058& 0.12859163 \\-0.70173855&1.58540872\end{bmatrix} &\begin{bmatrix}-0.0015415 \\-0.08616376\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.01]\end{bmatrix}\\
-% \midrule
-% (-2, -16) & (-3, -23) &  \begin{bmatrix}0.92850123& 0.12242812 \\-1.84720528&1.39159826\end{bmatrix} &\begin{bmatrix}-0.02334236\\-0.73348865\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -23) & (-5, -29) &  \begin{bmatrix}3.14947585&-0.81150646 \\-0.96159406&0.23543254\end{bmatrix} &\begin{bmatrix}-3.13860326\\-4.99696623\end{bmatrix}&\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}\\
-% \midrule
-% (-2, -15) & (-3, -22) &  \begin{bmatrix}0.93156623& 0.12260264 \\-1.75117219&1.40104348\end{bmatrix} &\begin{bmatrix}-0.02214125\\-0.68423569\end{bmatrix}&\begin{bmatrix}[-0.0, 0.0]\\ [-0.0, 0.0]\end{bmatrix}\\
-% \midrule
-% (-3, -22) & (-5, -29) &  \begin{bmatrix}3.08806462&-0.72332776 \\-1.69271079&0.40308293\end{bmatrix} &\begin{bmatrix}-2.77908467\\-4.57411339\end{bmatrix}&\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}\\
-
-\subsection{Search Parameters}
-
-The search parameters for S3CAM-R include the parameters of S3CAM:
-$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
-maximum error budget $\delta_{max}$ for OLS and the maximum length of
-segmented trajectory for building $k$-relational models.
-
-We have already discussed the effects of $N$, $\epsilon$ and $\Delta$
-on S3CAM's performance. However, they also have an effect on
-relational modeling. A finer grid with small cells produces more
-accurate models than a coarser grid with large cells. Similarly,
-small time length trajectory segments result in more accurate
-modeling.
-%An increase in the number of samples can increases the
-%accuracy of regression.
-
-\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
-Given a $\delta_{max}$, we keep increasing $k$ during the
-$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
-satisfied. A $k_{max}$ is introduced to bound the longest
-segmented trajectory which can be considered.
-
-% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
-
-\subsection{Reasons for Failure}
-The approach can fail to find a counter-example, even when it exists, in
-one of three ways.
-
-\begin{itemize}
-
-    \item \emph{S3CAM Fails} No abstract counter-example is found by
-        S3CAM. We can remedy it by increasing search budgets and/or
-        restarting.
-
-    \item \emph{BMC Fails} An abstract counter-example is found, but
-        the BMC fails to find a concrete counter-example in the PWA
-        relational model. The failure can be attributed to either (a)
-        a spurious abstract counter-examples or (b) a poorly estimated
-        model. The former can be addressed by restarting but the
-        latter requires that the maximum model error $\delta_{max}$ be
-        decreased.
-
-    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
-        is found, and it is successfully concretized in the PWA
-        relational model by the BMC. However, it is not reproducible
-        in the black box system. This happens when the PWA relational
-        model is not precise enough.
-
-\end{itemize}
-
-\section{Implementation and Evaluation}
-
-The implementation was prototyped as S3CAM-R, an extension to our
-previously mentioned tool S3CAM (\chapref{case}). OLS regression
-routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
-Python module for machine learning. SAL~\cite{SAL-SRI}
-with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
-solver.
-
-We used S3CAM to discover the abstraction graph $G$, which was then
-trimmed of the nodes which did not contribute to the error search
-(nodes from which error nodes were not reachable). In our experiments we
-used $1$-relational modeling to create the transition system from $G$.
-Using SAL, we checked for the given safety property. If a concrete
-trace was obtained from the BMC, it was further checked for validity
-by simulating using $\simulate$ to see if it was indeed an error
-trace.  If not, $100$ samples from its associated abstract state was
-sampled to check the presence of a counter-example in it
-neighborhood. This can be further extended by obtaining different
-discrete sequences or discrete trace sequences from the model checker.
-
-We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-few of the benchmarks described in \chapref{case}, including the Van
-der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
-benchmark. As before, we ran S3CAM-R $10$ times with different seeds
-and averaged the results. We tabulate both the total time taken and
-the time taken by SAL to compute the counter-example and compare
-against a newer implementation of S3CAM. Note that different
-abstraction parameters are used for S3CAM and S3CAM-R, due to the
-differences in which they operate.
-
-\begin{table*}[!htbp]
-\centering
-\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
-time taken by the BMC engine. The total time in seconds (rounded off
-to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
-$TO$ signifies time $>5hr$, after which the search was killed.}
-\label{tab:res-rel}
-\begin{tabular}{@{}llll@{}}
-\toprule
-Benchmark & BMC & S3CAM-R & S3CAM\\
-\midrule
-Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
-Brusselator               &$<$1  & $4$    & $2$\\
-Lorenz                    &$<$1 & $122$  & $35$\\
-%B.Ball                    &  &   & \\
-Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
-Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
-Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
-Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
-Bouncing Ball             &-  &$TO$   & $450$\\
-
-\bottomrule
-\end{tabular}
-\end{table*}
-
-% 1m8s, 2m56.184s, 1m31.059s[f], 1m32.665s,  1m45.312s, 1m22.397s, 2m1.451s[f], 1m12.118s, 1m26.596s
-% 0.12, 0.06,         0.33       0.13,       0.22,     FAIL,       ,X           0.08        0.15
-
-The results show promise, but clearly S3CAM performs better. Also,
-S3CAM-R timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
-and the bouncing ball. However, we need to explore more benchmarks to
-be conclusive. S3CAM-R's performance can be explained by the
-difficulty in finding a good abstraction ($\quant_\epsilon$), which
-needs to be fine enough, to obtain good prediction models but coarse
-enough, to be manageable by the current SMT solvers. Specifically, to
-obtain a good quality counter-example from the model checker (and
-avoid false positives), the transition system should be created from
-models with high accuracy, for which a finer abstraction is required.
-However, the exploration of a finer abstraction is exponentially more
-complex and results in a much bigger transition system, which the
-current state of the art bounded model checkers (which use SMT
-solvers) can not handle under reasonable resources.
-
-\section{Conclusion and Future Work}
-
-We have presented another methodology to find falsifications in black
-box dynamical systems. Combining the ideas from abstraction based
-search (\chapref{abs}), with our previous relational
-abstractions~\cite{zutshi2012timed}, we proposed $k$-relational
-modeling to compute richer abstractions. Simple linear regression can
-be used to estimate the local dynamics of the trajectory segments and
-compute PWA relational models which can be interpreted as a transition
-system.  These can then be checked for safety violations using an
-off-the-shelf bounded model checker.  Finally, we implemented the
-ideas as a tool S3CAM-R and demonstrated its performance on a few examples.
-
-\subsection{Improvements}
-
-The approach seems promising as it avoids an expensive refinement
-step. But, before it can be applied to more complex systems, we need
-to overcome its shortcomings.  As a future extension, we are
-investigating ideas to improve the performance of the underlying three
-main sub-processes, namely, abstraction, modeling, and BMC. We detail
-these below.
-
-\begin{enumerate}
-
-\item \emph{Abstraction} The abstract domain we have used, has been
-    confined to the interval or rectangular domain. Such a domain is
-    very coarse when compared to more general polyhedral domains, using
-    which, we can gain better precision. Due to polyhedral domains
-    being computational expensive, their integration in S3CAM needs to
-    be explored thoroughly.
-
-\item \emph{Modeling} We have used $k$-relational modeling to increase
-    the precision of the enriched abstractions, however, there are
-    several other refinement strategies that remain to be explored.
-    This includes refining the abstract states by using guard
-    predicates (similar to predicate abstraction), using non-linear
-    templates in parametric regression to compute more precise
-    quadratic relations, and using an adaptive time step instead of a
-    fixed one to address the non-linearities due to a long time
-    horizon.
-
-\item \emph{Model Checking} As of now, we use SMT solvers to search
-    for concrete counter-examples. Although they are very efficient,
-    owing to extensive engineering effort, reachability of transition
-    systems resulting from dynamical systems remains a difficult
-    problem. As the time horizon of the safety property increases, the
-    possible combinations of discrete transitions increase
-    exponentially. Hence, to find a counterexample which is a sequence
-    of discrete transitions over a `long' time horizon is not
-    tractable for most, but the simplest of dynamical systems.
-    Instead, we can use linear programming solvers, by enumerating
-    each path in the graph $G^R$. Note that the constraints are all
-    linear (conjunctions) along an abstract path. However, in the
-    worst case, the number of paths in a graph can be of the order
-    $n!$ where $n$ is the number of vertices of a graph. Hence, this
-    will not be feasible, unless, we can prioritize paths by using a
-    triage process similar to one used by S3CAM and using a budget on
-    the maximum number of paths.
-
-    Another approach to address the issue would be to use an adaptive
-    time discretization technique, where relations over both shorter
-    and longer time steps are computed, and the SMT solver can
-    `select' the time step precise enough to find a counter-example.
-
-\item \emph{SMT Solvers} One major impediment to our approach is the
-    fact that SMT solvers use the theory of reals with \emph{exact
-    precision}. This is important for verification approaches, but can
-    be relaxed for the problem of falsification. An SMT solver which
-    uses approximate reasoning but returns robust counter-examples
-    will be as useful, and perhaps more efficient.
-
-\end{enumerate}
-
-\subsection{Data Driven Analysis}
-
-Finally, we would like to mention that our approach is `simulation'
-driven. It can be easily modified to be `data driven', by working with
-a fixed set of data. Given a data set, we then need to automatically
-find a good abstraction and a high fidelity PWA transition system
-using which, we can summarize the black box hybrid dynamical system.
-Apart from model checking the transition system, one can extend it to
-the analysis of SDCS. More specifically, by combining the transition
-system model of a plant with the control software, a model checker
-like CBMC \cite{kroening2014cbmc} can be used to do a closed loop
-symbolic analysis of the SDCS. This can be an alternative to S3CAM-X.
-
-%which will be
-%very useful to for analyzing data for potential property violations.

### 2017-02-08 pwarel.tex
.

@@
 discuss the background required to present our ideas.
+
+\section{Relational Modeling}
+
+We now describe how relational models can be computed for a given
+black box system by enriching the abstract reachability graph obtained
+by scatter-and-simulate. We first formalize the notion of an enriched
+graph, and then show how they can be interpreted as a PWA transition
+system using $k$-relational modeling.
+
+\subsection{Abstract Enriched Graph}
+
+The existential abstraction relation in \chapref{abs} was defined as
+follows:
+\[
+%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
+    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
+\]
+The abstract relation $\areach{t}$ can be enriched by incorporating
+the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
+system, such a relation can be rarely represented using an exact affine map.
+This is due to the presence of non-linear and hybrid behaviors. But,
+an affine map can always be estimated with an error.
+
+If the system dynamics are completely specified in the form of a white
+box model, we can use first order approximations to find the affine
+expressions for the relations. Using a tool like \flowstar, we can
+obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
+\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
+that, every element $b_i$ of the vector $\vb$ is contained in the
+respective scalar interval $b_i\in[b^l_i,b^h_i]$.
+
+For the case of black box systems, such a sound approximation is not
+possible. Instead, we rely on a statistical method like simple linear
+regression to estimate the affine map $\amap: A\x + \vb$. We then
+estimate the error $\delta$ and generalize $\amap$ to an interval
+affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
+abstraction with the below relation
+\[
+    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
+    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
+\]
+
+However, in the presence of complex non-linear behavior, using a
+single affine map can lead to a poor approximation. To increase the
+precision, one can use more than one affine map. Let us denote this set as
+$R_{(C,C')}$ (ref.~\secref{ovw}).
+\[
+    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
+\]
+
+\begin{figure}[!htbp]
+\begin{center}
+\tikzstyle{line} = [thick]
+\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
+\begin{tikzpicture}
+
+    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
+    \draw[arw] (-3,0) -- (-1.2,0);
+    \draw[arw] (1.2,0) -- (3,0);
+
+\node at (-3.4,0) {$G$};
+\node at (3.4,0) {$G^R$};
+\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
+
+\end{tikzpicture}
+\end{center}
+\vspace*{-.3cm}
+\caption{Using OLS, $G^R$ is computed by determining the appropriate
+$R_{(C,C')}$ for each edge of $G$.}
+\label{fig:enrichment}
+\vspace*{-.3cm}
+\end{figure}
+
+As shown in \figref{enrichment}, we use OLS to compute an enrichment
+of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
+the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
+of the enriched edge are clearly non-deterministic. Form a state $\x
+\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
+is satisfied. This interpretation results in an infinite state
+transition system. We now detail the construction of the set
+$R_{(C,C')}$ using $k$-relational modeling.
+
+% We now compute a PWA model for a given black box system by estimating
+% the dynamics using OLS. We can estimate the affine relations between
+% every cell, but as discussed in ~\chapref{abs}, it is a futile
+% approach.  Instead, we use the same heuristic as before:
+% scatter-and-simulate, to select the cell relations for estimation. We
+% build the reachability graph as before, but in addition, annotate each
+% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
+% respective relation. The reachability graph thus computed, is a
+% transition system. Using off-the-shelf bounded model checkers, we
+% reason about the system's safety properties.
+
+\subsection{$k$ - Relational Modeling}
+
+\begin{figure}[!htbp]
+\vspace*{-.2cm}
+\begin{center}
+\tikzstyle{line} = [thick]
+\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
+\begin{tikzpicture}
+
+\begin{scope}[scale=0.8]
+\draw [line] (-1.0,0) circle (0.5);
+\draw [line] (-4.0,0) circle (0.5);
+\draw [line] (2.0,1) circle (0.5);
+\draw [line] (2.0,-1) circle (0.5);
+\draw[arw] (-0.5,0) -- (1.5,1.0);
+\draw[arw] (-0.5,0) -- (1.5,-1.0);
+\draw[arw] (-3.5,0) -- (-1.5,0.0);
+\node at (-4.0,0) {$C$};
+\node at (-1.0,0) {$C'$};
+\node at (2.0,1.0) {$C''_1$};
+\node at (2.0,-1.0) {$C''_2$};
+\draw [line] (2.0,-1) circle (0.5);
+\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
+\node at (-4.0,-2.5) {$k=0$};
+\node at (-1.0,-2.5) {$k=1$};
+\node at (+2.0,-2.5) {$k=2$};
+\draw [fill=black] (-4.0,-2.0) circle (0.05);
+\draw [fill=black] (-1.0,-2.0) circle (0.05);
+\draw [fill=black] (2.0,-2.0) circle (0.05);
+
+\end{scope}
+\end{tikzpicture}
+\end{center}
+\vspace*{-.3cm}
+\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
+\label{fig:k-rel}
+\vspace*{-.3cm}
+\end{figure}
+
+Given a $G$, when computing the set of relations for an edge $R$, we
+consider only the set of abstract states reachable at a specific time
+step.  Intuitively, one might consider the states reachable in one
+time step $\Delta$.  However, such a process can be generalized by
+looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
+k\Delta}$ time steps.
+
+Using \figref{k-rel}, we illustrate this notion. To compute
+$R_{(C,C')}$, we
+split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
+$k\Delta$) as follows.  We observe the local behavior of the system by
+noting the evolution of the system $\System$ from a state $\x(t)
+\in C$ for a time length dependant on $k$. For
+\begin{itemize}
+    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
+    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
+    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
+        and split them in two sets (a) and (b) on the basis of the
+        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
+        \begin{enumerate}[(a)]
+            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
+            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
+        \end{enumerate}
+    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
+        and split them in to multiple sets on the basis of the cells
+        reached at time $(t+\Delta),\; (t+2\Delta),\;
+        \ldots,\;(t+n\Delta)$.
+\end{itemize}
+
+$k$-relational modeling can be understood as a heuristic, which uses
+the underlying abstraction to differentiate behaviors of the system
+which `diverge' (or are revealed to be `distinct' at a future time).
+Such a heuristic can be useful in increasing the precision of the
+learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
+$n$ and illustrate using examples.
+
+\mypara{$0$-Relational Model}
+
+\begin{figure}[!htbp]
+\begin{center}
+\tikzstyle{line} = [thick]
+\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
+\begin{tikzpicture}
+\begin{scope}[scale=0.5]
+
+    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
+    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
+    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
+    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
+    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
+    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
+
+\node at (0.0, 0.0) {$C$};
+\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
+\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
+\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
+\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
+\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
+\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
+\node at (2,1.5) {$\pi_1$};
+\node at (2,0.5) {$\pi_2$};
+\node at (1.7,-1) {$\pi_3$};
+\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
+\node at (0,-2.5) {$t=0$};
+\node at (4,-2.5) {$t=\Delta$};
+\draw [fill=black] (0.0,-2.0) circle (0.05);
+\draw [fill=black] (4.0,-2.0) circle (0.05);
+
+\end{scope}
+\begin{scope}[xshift=0,yshift=3.0cm,scale=0.5]
+    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
+    \implies A_1\vx + \vb_1 + \delta_1$};
+\end{scope}
+
+\end{tikzpicture}
+\end{center}
+\vspace*{-.3cm}
+\caption{All the trajectory segments will be used to construct the
+model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
+\label{fig:k0}
+\vspace*{-.3cm}
+\end{figure}
+
+When $k=0$, the $0$-relational model is a PWA transition system which only
+defines the evolution of states $\x$, and does not specify the
+reachable cell. The guard predicates of the relations are defined over
+cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.
+
+We use regression to estimate the dynamics for the outgoing trajectory
+segments form a cell $C$. Hence, the data set $\ds$ for the regression
+includes all trajectory segments beginning from the same cell
+\[
+    \ds = \setof{\pi_t | start(\pi_t) \in C}.
+\]
+This includes trajectory segments ending in different cells, as shown
+in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
+\begin{equation*}
+    \scr{T} = \left\{
+        \begin{array}{ll}
+            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
+            \ldots \\
+            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
+        \end{array}
+    \right.
+\end{equation*}
+
+Note that this can be quite imprecise when the cells are big,
+containing regions of state-space with complex dynamics. This is true
+for both non-linear systems and hybrid dynamical system, where a cell
+can contain two or more modes with differing continuous dynamics.
+
+\mypara{$1$-Relational Model}
+
+\begin{figure}[!htbp]
+\begin{center}
+\tikzstyle{line} = [thick]
+\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
+\begin{tikzpicture}
+\begin{scope}[scale=0.5]
+
+    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
+    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
+    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
+    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
+    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
+    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
+
+\node at (0.0, 0.0) {$C$};
+\node at (4.0, 2.0) {$C'_1$};
+\node at (4.0, -2.0) {$C'_2$};
+\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
+\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
+\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
+\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
+\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
+\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
+\node at (2,1.5) {$\pi_1$};
+\node at (2,0.5) {$\pi_2$};
+\node at (1.7,-1) {$\pi_3$};
+\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
+\node at (0,-4.0) {$t=0$};
+\node at (4,-4.0) {$t=\Delta$};
+\draw [fill=black] (0.0,-3.5) circle (0.05);
+\draw [fill=black] (4.0,-3.5) circle (0.05);
+
+\end{scope}
+\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
+    \node at (2,0) {\begin{minipage}{\linewidth}
+    \footnotesize \begin{equation*}
+        \scr{T} = \left\{
+        \begin{array}{ll}
+            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
+            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
+        \end{array}
+    \right.
+\end{equation*}
+\end{minipage}};
+\end{scope}
+
+\end{tikzpicture}
+\end{center}
+\vspace*{-.3cm}
+\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
+    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
+    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
+%Cardinality($R_{(C,C')}$) = $2$.}
+\label{fig:k1}
+\vspace*{-.3cm}
+\end{figure}
+
+% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
+%     \node at (2,0) {\footnotesize $\begin{equation*}
+%     \scr{T} = \left\{
+%         \begin{array}{ll}
+%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
+%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
+%         \end{array}
+%     \right.
+% \end{equation*}$};
+% \end{scope}
+
+To improve the preciseness of learnt dynamics, we include the
+reachability relation in the
+regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
+comprised only of trajectory segments $\pi_t$ which start and end in
+the same set of cells.
+\[
+    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
+\]
+For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
+$N$ transition relations.
+
+\figref{k1} illustrates the $k=1$ refinement of the case shown in
+\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
+= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
+relations are constructed. The resulting $1$-relational model
+is at least as precise as the corresponding $0$-relational model.
+
+\mypara{$k$-Relational Model}
+
+\begin{figure}[!htbp]
+\begin{center}
+\tikzstyle{line} = [thick]
+\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
+\begin{tikzpicture}
+\begin{scope}[scale=0.5]
+
+    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
+    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
+    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
+    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
+    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
+    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
+    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;
+
+%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
+%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
+
+    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
+
+\node at (0.0, 0.0) {$C$};
+\node at (4.0, 2.0) {$C'_1$};
+\node at (8.0, 2.0) {$C''_1$};
+\node at (8.0, -2.0) {$C''_2$};
+%\node at (4.0, -2.0) {$C'_2$};
+\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
+\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
+%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
+\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
+\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
+%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
+\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
+\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
+\node at (2,1.5) {$\pi_1$};
+\node at (2,0.5) {$\pi_2$};
+\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
+\node at (0,-4.0) {$t=0$};
+\node at (4,-4.0) {$t=\Delta$};
+\node at (8,-4.0) {$t=2\Delta$};
+\draw [fill=black] (0.0,-3.5) circle (0.05);
+\draw [fill=black] (4.0,-3.5) circle (0.05);
+\draw [fill=black] (8.0,-3.5) circle (0.05);
+\end{scope}
+
+\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
+    \node at (4,0) {\begin{minipage}{\linewidth}
+        \footnotesize \begin{equation*}
+    \scr{T} = \left\{
+        \begin{array}{ll}
+            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
+            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
+        \end{array}
+    \right.
+\end{equation*}
+\end{minipage}};
+\end{scope}
+
+\end{tikzpicture}
+\end{center}
+\vspace*{-.3cm}
+\caption{The $k=2$ refinement further splits the data set $\ds_1$
+    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
+    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
+    non-deterministically defining the system behavior.}
+\label{fig:k2}
+\vspace*{-.3cm}
+\end{figure}
+
+Finally, we generalize the relational models to $k$-relational PWA
+models. A $k$-relational model is constructed by using $k$ length
+\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
+\textit{connected} iff its cost $\cost(\SegTraj) = 0$.
+
+Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
+$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
+\textit{similar} if their trajectory segments have the same sequence
+of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
+\[
+    \forall i\in\setof{1 \ldots k}.
+    start(\pi_{t_i}) \in C_i
+    \land start(\pi'_{t_i}) \in C_i
+    \land end(\pi_{t_i}) \in C_{i+1}
+    \land end(\pi'_{t_i}) \in C_{i+1}
+\]
+
+For every cell $C$ in the reachability graph, we construct a data set
+$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
+length \textit{connected} and \textit{similar} segmented trajectories.
+%\[
+%    \ds = \setof{\pi_{\tau_i} |
+%        \pi_{\tau_i}
+%        start(\pi_{\tau_{i}}) \in C_i
+%        \land
+%        end(\pi_{\tau_{i}}) \in C_i}
+%\]
+
+\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.

### 2017-02-08 implementation.tex
.

@@
 The implementation was prototyped as S3CAM-R, an extension to our
-previously mentioned tool S3CAM (\chapref{case}). OLS regression
+previously mentioned tool S3CAM \cite{zutshi2014multiple}. OLS regression
 routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
@@
 We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-few of the benchmarks described in \chapref{case}, including the Van
+few of the benchmarks described in \cite{zutshi2014multiple}, including the Van
 der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
@@
 
-\begin{table*}[!htbp]
+\begin{table}[!htbp]
 \centering
@@
 \end{tabular}
-\end{table*}
+\end{table}
 

### 2017-02-08 prelims.tex
.

@@
 We define the PWA transition system as a transition system
-(ref.~\secref{DiscSys}) $\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
+%(ref.~\secref{DiscSys})
+$\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
 where each transition $\tau \in \scrT$, is associated with a

### 2017-02-08 pwarel.tex
.

@@
 scatter-and-simulate (parameterized by number of samples $n$ as
-described in \chapref{abs}) to explore the abstract graph
+described in \cite{zutshi2014multiple}) to explore the abstract graph
 $\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
@@
 
-Instead of using a CEGAR like loop (as in \chapref{abs}), we use the
+Instead of using a CEGAR like loop (as in \cite{zutshi2014multiple}), we use the
 generated trajectory segments to learn quantitative models describing
@@
 which non-deterministically describe relations between $\x \in C$ and $\x'
-\in C'$. This is illustrated in \figref{enriched-edge} (compare with
-\figref{segtraj}).
+\in C'$. This is illustrated in \figref{enriched-edge}.
+%(compare with \figref{segtraj}).
 
@@
 
-The existential abstraction relation in \chapref{abs} was defined as
+The existential abstraction relation in \cite{zutshi2014multiple} was defined as
 follows:
@@
 \end{scope}
-\begin{scope}[xshift=0,yshift=3.0cm,scale=0.5]
+\begin{scope}[xshift=0,yshift=2.0cm,scale=0.5]
     \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
@@
 \end{scope}
-\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
+\begin{scope}[xshift=0,yshift=2.5cm,scale=0.5]
     \node at (2,0) {\begin{minipage}{\linewidth}
@@
 
-\begin{scope}[xshift=0,yshift=4.0cm,scale=0.5]
+\begin{scope}[xshift=0,yshift=2.5cm,scale=0.5]
     \node at (4,0) {\begin{minipage}{\linewidth}
@@
 of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
-\[
+\begin{align*}
     \forall i\in\setof{1 \ldots k}.
-    start(\pi_{t_i}) \in C_i
-    \land start(\pi'_{t_i}) \in C_i
-    \land end(\pi_{t_i}) \in C_{i+1}
+    & start(\pi_{t_i}) \in C_i          
+    \land start(\pi'_{t_i}) \in C_i \land  \\
+    & end(\pi_{t_i}) \in C_{i+1}  
     \land end(\pi'_{t_i}) \in C_{i+1}
-\]
+\end{align*}
 

### 2017-04-07 abstract.tex
.

@@
-In this work we address the problem of finding unsafe behaviors with
-given a safety property for a hybrid dynamical system. We treat the
-system as a black box and approach the problem in two distinct steps.
-First, we model the behavior of the system using a piecewise affine
-discrete time model. Such a model is incomplete in nature and is
-computed with respect to the given property. Next, we encode the
-search for falsification as a bounded model checking query and use an
-SMT solver to find a counter example. If found, we check the existance
-of the violation in the original system to ensure reproducibility.
+Sampling based falsification techniques for black box hybrid dynamical
+systems use numerical simulations to search for violations of
+properties (specified in temporal logic). This constraints them to a
+`best effort' category unlike rigorous verification techniques, which
+however require a white box model. This work tries to bridge the gap
+by presenting an automatic falsification technique which models the
+black box system before searching for a violation. It proceeds by (a)
+modeling the behavior of the system using a piecewise affine (PWA)
+discrete time model and, (b) encoding the search for a violation as a
+standard bounded model checking problem. The resulting problem is a
+disjunction of linear programs and can be solved directly using
+off-the-shelf SMT solvers or by multiple runs of a linear programming
+solver. If a counter-example is found, the existence of a
+corresponding violation in the original system is checked for
+reproducibility.
+
+% In this work we address the problem of finding unsafe behaviors with
+% given a safety property for a hybrid dynamical system. We treat the
+% system as a black box and approach the problem in two distinct steps.
+% First, we model the behavior of the system using a piecewise affine
+% discrete time model. Such a model is incomplete in nature and is
+% computed with respect to the given property. Next, we encode the
+% search for falsification as a bounded model checking query and use an
+% SMT solver to find a counter example. If found, we check the existence
+% of the violation in the original system to ensure reproducibility.
 
 %Time bounded LTL property in
-%Models from verificaiton
+%Models from verification
 %model without data

### 2017-04-07 bmc.tex
.

@@
     \item Given a black box system $\System$ specified by
-        $\simulate_\System$, we
+        $\simulate^\System$, we
         use scatter-and-simulate with ($\Delta$-length) trajectory
@@
 We now describe the above using the Van der Pol oscillator benchmark
-presented in \secref{math-dyn-sys}. Simulations generated using
+presented in \cite{zutshi2014multiple}. Simulations generated using
 uniformly random samples are shown in \figref{vdp-cont}. We want to

### 2017-04-07 implementation.tex
.

@@
 \label{tab:res-rel}
-\begin{tabular}{@{}llll@{}}
+\begin{tabular}{@{}lllll@{}}
 \toprule
-Benchmark & BMC & S3CAM-R & S3CAM\\
+    Benchmark & BMC & LP & S3CAM-R & S3CAM\\
 \midrule
-Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
-Brusselator               &$<$1  & $4$    & $2$\\
-Lorenz                    &$<$1 & $122$  & $35$\\
+    Van der Pol ($\prop_3$)   &$<$1 & $0$ & $130$ & $24$\\
+    Brusselator               &$<$1 & $0$ & $4$    & $2$\\
+    Lorenz                    &$<$1 & $0$ & $122$  & $35$\\
 %B.Ball                    &  &   & \\
-Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
-Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
-Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
-Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
-Bouncing Ball             &-  &$TO$   & $450$\\
+    Nav ($\prop_P$)           &$4480$ & $0$ & $9423$   &$603$ \\
+    Nav ($\prop_Q$)           &$970$ & $0$ & $8105$   &$546$ \\
+    Nav ($\prop_R$)           &-  & $0$ &$TO$ & $2003$\\
+    Nav ($\prop_S$)           &-  & $0$ &$TO$ & $2100$\\
+    Bouncing Ball             &-  & $0$ &$TO$ & $450$\\
 

### 2017-04-07 prelims.tex
.

@@
-In this section, we present the basics of relational abstractions and
-fundamentals of regression analysis.
+In this section, we present the model of the hybrid dynamical system
+under test and the basics of pwa relational abstraction.
 
-\subsection{Relational Abstraction}
+We now use the family of reachability relations presented
+in~\cite{zutshi2014multiple} to charecterize the system $\System$
+under test. We denote the states of the hybrid system $\System$, by $\HybridStates \subseteq \Modes \times \Reals^n$ be the (infinite)
+set of hybrid states of the system $\System$, where $\Modes$ is a
+finite set of discrete modes, and $\Inputs$ be the set of input
+signals to $\System$ of the form $[0,T] \rightarrow \reals^k$ for some
+given time horizon $T$. The family of
+relations $\reach{t,u} \subseteq \HybridStates \times \HybridStates$,
+parameterized by time $t \geq 0$ and input signal $u \in \Inputs$
+defined over time $[0,T]$.  We use the notation $\vx\ \reach{t,u} \vy$
+to denote $(\vx,\vy) \in \reach{t,u}$. The relations $\reach{t,u}$
+satisfy the following basic properties:
 
-For general hybrid dynamical systems, algorithmic ways for computing
-analytical solutions do not exist. Discretization (along with suitable
-abstractions) is often employed to transform the systems into a
-discrete transition system. Relational abstractions is one such idea,
-which abstracts continuous dynamics by discrete relations using
-appropriate \textit{reachability invariants}. Both time independent
-and time dependent relations have been proposed; the former captures
-all reachable states over all time, whereas, the latter explicitly
-includes time by relating reachable states to time. Thus it
-can prove timing properties whereas, time independent relational
-abstractions can not. In both cases, the resulting abstractions can be
-interpreted as discrete transition systems and can be analyzed using
-model checkers. We briefly discuss these two types of abstractions.
+We assume a black-box model
+equipped with a forward simulation map $\simulate^{\System}$ that can
+be used to compute the discrete time traces of the system. Let
+$\reach{}$ Given the
+hybrid state space $\vx\in\HybridStates$ and input space $u
+\in\Inputs$ and time step $\Delta_t$, it computes the unique successor
+state $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u}
+\vy$.
 
-\mypara{Timeless Relational Abstraction}
+\begin{definition}
 
-Relational abstractions for hybrid systems were proposed in
-\cite{Sankaranarayanan+Tiwari/2011/Relational}. For a hybrid automaton
-model, they can summarize the continuous dynamics of each mode using a
-binary relation over continuous states. The resulting relations are
-timeless (independent) and hence valid for all time as long as
-the mode invariant is satisfied. The relations take the general form
-of $R(\x,\x') \bowtie 0$, where $\bowtie$ represents one of the
-relational operators $=, \ge, \le, <, >$. For example, an abstraction
-that captures the monotonicity with respect to time for the
-differential equation $\dot{x} = 2$ is $x' > x$. The abstraction
-capturing the relation between the set of ODEs: $\dot{x} = 2$,
-$\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
+    The system under test $\System$ evolves in the set of infinite
+    hybrid state-space $\HybridStates$ under the effect of inputs $\vu
+    \in \Inputs$. Its evolution is defined by a simulation function
+    $\simulate^\System : \HybridStates \times \Inputs \Reals^+ \mapsto
+    \HybridStates$, where $\simulate^\System(\x,\vu,\Delta_t)$ maps a
+    current state $\x$ at time $t$ to a new state $\x'$ at time
+    $t+\Delta_t$, under a constant input $\vu$.
 
-Such relation summaries, discretize the continuous evolution of a
-given system into an infinite state transition system, in which safety
-properties can be verified using $k$-induction, or falsified using
-bounded model checkers. The relationalization procedure involves
-finding suitable invariants in a chosen abstract domain, such as
-affine abstractions, eigen abstractions or box abstractions
-\cite{Sankaranarayanan+Tiwari/2011/Relational}.  For this, different
-techniques like template based invariant generation
-\cite{Gulwani+Tiwari/2008/Constraint,
-Colon+Sankaranarayanan+Sipma/03/Linear} can be used.
+\end{definition}
 
-\mypara{Timed and Time-Aware Relational Abstractions}
+The subscript $\System$ will be dropped whenever the system $\System$
+being referred to is clear from the context.
 
-As the above discussed relations are timeless, we cannot reason over
-the timing properties of the original system. Moreover, they are not
-suitable for time-triggered systems. Timed relational
-abstractions~\cite{zutshi2012timed} and time-aware relational
-abstractions~\cite{mover2013time} were proposed to overcome these
-shortcomings by explicitly including time in the relations. In
-addition, both can be more precise than their timeless counterpart,
-but unlike it, assume continuous affine dynamics.
-
-To analyze time-triggered systems like an SDCS modeled by an affine
-hybrid automata~\footnote{An affine hybrid automata is restricted to
-the ODEs, resets and guards being affine.}, timed relational
-abstractions can be computed directly from the solutions of the
-underlying affine ODEs.  Recall that the solution of an affine ODE
-$\dot{\x} = A\x$ can be represented using the matrix exponential as
-$\x(t) = e^{tA}\x(0)$. This gives us the timed relation $\x' =
-e^{tA}\x$.  Clearly, the relation is non-linear with respect to time.
-However, for the case of SDCS with a fixed sampling time period, we
-obtain linear relations in $\x$.  We demonstrated the usefulness of
-timed relational abstractions for the case of linear systems
-in \cite{zutshi2012timed}, and now incorporate the idea to discretize
-black box dynamical systems.
-
-Similar to timeless relational abstractions, time-aware relational
-abstractions \cite{mover2013time} construct binary relations between
-the current state $\x$ of the system and any future reachable state
-$\x'$. The difference lies with the latter also constructing relations
-between the current time $t$ and any future time $t'$. This is
-achieved by a case by case analysis of the eigen structure of the
-matrix $A$ (of an affine ODE). Separate abstractions are used for the
-case of linear systems with constant rate, real eigen values and complex eigenvalues.
+\myipara{Black-Box Assumptions:} For above properties of the
+$\reach{t,u}$ relation to hold, we require that $\System$ be
+deterministic. We also assume that $\System$ can be simulated from a
+given initial state using function $\simulate^{\System}$. This
+requires existence and uniqueness of trajectories over a finite time
+horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
+the vector field in each hybrid mode and ruling away issues such as
+finite escape times~\cite{Meiss/2007/Differential}. From a practical
+standpoint, these assumptions do not pose significant restrictions for
+the type of problems that we wish to address.  Finally, we assume full
+observability of the system state, ignore numerical errors due to
+simulation and {\em do not} assume that an analytic closed-form
+representation of the system dynamics is available.
 
@@
 
-\mypara{Regression Analysis}
-
-In statistics, regression is the problem of finding a
-\textit{predictor}, which can suitably predict the relationship
-between the given set of observed input $\x$ and output $\y$ vectors.
-In other words, assuming that $\y$ depends on $\x$, regression
-strategies find either a parameterized or a non-parameterized
-prediction function to explain the dependence. We now discuss simple
-linear regression, which is parametric in nature and searches for an
-affine predictor. It is also called \textit{ordinary least squares}.
+In statistics, linear regression is the problem of finding an affine
+function or \textit{predictor}, which can `best' summarize the
+relationship between the given set of observed input $\x$ and output
+$\y$ vectors. The notion of `best' is formally captured using a loss
+function which can be non-linear. We use the commonly used loss
+function \textit{ordinary least squares} (OLS) for this presentation.
 
@@
 
-Let the data set be comprised of $N$ input and output pairs $(\x,
-y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
-the case in the current context of \textit{finding the best fit}, the
-problem is over-determined; there are more equations than
-unknowns. Hence, a single affine function cannot be found which
-satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This is formally
+The
+Let the data set be comprised of $N$ input and output pairs $(\x, y)$,
+where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is the case in
+the current context of \textit{finding the best fit}, the problem is
+over-determined; there are more equations than unknowns. Hence, a
+single affine function cannot be found which satisfies the equation
+$\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we need
+to find the `best' choice for $\vec{a}$ and $b$. This is formally
 defined using a loss function. For the case of simple linear
@@
 
+% \mypara{Regression Analysis}
+
+% In statistics, regression is the problem of finding a
+% \textit{predictor}, which can suitably predict the relationship
+% between the given set of observed input $\x$ and output $\y$ vectors.
+% In other words, assuming that $\y$ depends on $\x$, regression
+% strategies find either a parameterized or a non-parameterized
+% prediction function to explain the dependence. We now discuss simple
+% linear regression, which is parametric in nature and searches for an
+% affine predictor. It is also called \textit{ordinary least squares}.
+
+% \mypara{Ordinary Least Squares (OLS)}
+
+% Let the data set be comprised of $N$ input and output pairs $(\x,
+% y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
+% the case in the current context of \textit{finding the best fit}, the
+% problem is over-determined; there are more equations than
+% unknowns. Hence, a single affine function cannot be found which
+% satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
+% need to find the `best' choice for $\vec{a}$ and $b$. This is formally
+% defined using a loss function. For the case of simple linear
+% regression or OLS, the loss function is the sum of squares of the
+% errors in prediction.  The task is then to determine the vector of
+% coefficients $\vec{a}$ and an offset constant $b$, such that the least
+% square error of the affine predictor is minimized for the given data
+% set.
+% \begin{equation}
+%     \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
+% \label{reg}
+% \end{equation}
+% To ease the presentation, we can rewrite the above as a homogeneous
+% expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
+% $\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
+% \[
+%     \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
+% \]
+% The solution of OLS can be analytically computed as
+% \[ Ab = (X^TX)^{-1}X^T\y\]
+% where $X$ is the matrix representing the horizontal stacking of all
+% $\hat{\x}$.
+% The details can be found in several texts on learning and statistics
+% \cite{friedman2001elements}.
+
+% Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
+% where $\x\in\reals^n$ we can use OLS to approximate its trajectories
+% at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
+% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
+% \begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
+%     global relational model for the system. The data set for the
+%     regression is a set of pairs $\setof{(\x,\x') | \x' =
+%     \simulate(\x, \Delta)}$ and can be generated on demand. Another
+%     set can be generated to compute the error $\delta$ as an interval
+%     of vectors, where each element is an interval $\delta_i \in
+%     [\delta_{min}, \delta_{max}]$.
+%
+%     Affine maps are poor
+%     approximations for arbitrary non-linear functions. Hence, we use a
+%     collection of affine maps to approximate the local behaviors (in
+%     state-space) of the system $\simulate$. This results in a piece
+%     wise approximation, as described in the next section.
+
+\subsection{Relational Abstractions}
+For a hybrid automaton model, they can summarize the continuous
+dynamics of each mode using a binary relation over continuous states.
+The resulting relations are timeless (independent) and hence valid for
+all time as long as the mode invariant is satisfied. The relations
+take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
+represents one of the relational operators $=, \ge, \le, <, >$. For
+example, an abstraction that captures the monotonicity with respect to
+time for the differential equation $\dot{x} = 2$ is $x' > x$. The
+abstraction capturing the relation between the set of ODEs: $\dot{x} =
+2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
+
+%%  Similar to timeless relational abstractions, time-aware relational
+%%  abstractions \cite{mover2013time} construct binary relations between
+%%  the current state $\x$ of the system and any future reachable state
+%%  $\x'$. The difference lies with the latter also constructing relations
+%%  between the current time $t$ and any future time $t'$. This is
+%%  achieved by a case by case analysis of the eigen structure of the
+%%  matrix $A$ (of an affine ODE). Separate abstractions are used for the
+%%  case of linear systems with constant rate, real eigen values and
+%%  complex eigenvalues.
+
 \subsection{Piecewise Affine (PWA) Transition System}
@@
 behavior of a hybrid dynamical system $\System$ defined by a simulator
-$\simulate_{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
+$\simulate^{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
 form, we can only approximate general hybrid systems.}. Each
@@
 % map $\gm$.
+
+\subsection{Learning Affine Function from Trajectory Segments}
+
+The $\simulate$ function can be used to generate a data set of
+relations parameterized by time $\Delta_t$, for different values of
+states $\x_i$ and inputs $\vu_i$ as $\ds_{\Delta_t} =
+\setof{(\x_i,\x'_i,\vu_i) | \x'_i = \simulate(\x_i,\vu_i)}$.
+Moreover, by repeatedly applying $\simulate$, a discrete time series
+of a finite length $N$ can be generated from a given state $\x_1$
+and a sequence of inputs $\tupleof{\vu_1, \vu_2,\ldots,\vu_n}$ as shown
+below.
+\[\tupleof{\x_1, \x_2, \ldots, \x_n} \;\text{where}\; \x_{i>1} =
+\simulate(\x_{i-1}, \vu_{i-1}, \Delta_t)\]
+
+We now introduce a refinement of the data set $\ds \subseteq \ds$ by
+incrementally observing the future states in the time series.
+
+\[\ds_{\Delta_t} = \setof{(\x_i,\x'_i,\vu_i) | \x'_i =
+\simulate(\x_i,\vu_i)}\]
+
+\[\ds^{C_i} \subseteq \ds_t\]
+\[\ds^{C_i}\]
+
+$k$-Refinement of the data set $\ds^k$.
+$\ds_k = \setof{(\x,\x',\vu) | \x' =  \simulate(\x,\vu,k\Delta_t)}$
+
+state-input pairs of the form
+$(\x,\x')$
+data set equivalence traj segment
+$D_c, $

### 2017-04-07 pwarel.tex
.

@@
-\subsection{Overview}
-\label{sec:ovw}
-
 We now briefly describe the approach. Given a numerical simulator
@@
 
-Instead of using a CEGAR like loop (as in \cite{zutshi2014multiple}), we use the
-generated trajectory segments to learn quantitative models describing
-the local behavior of the system. These models are defined by a set of
-relations $R \subseteq \HybridStates \times \HybridStates$ for each
-edge of the reachability graph.
+Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
+we use the generated trajectory segments to learn quantitative models
+describing the local behavior of the system. These models are defined
+by a set of relations $R \subseteq \HybridStates \times \HybridStates$
+for each edge of the reachability graph.
 

### 2017-04-07 abstract.tex
.

@@
-Sampling based falsification techniques for black box hybrid dynamical
-systems use numerical simulations to search for violations of
-properties (specified in temporal logic). This constraints them to a
-`best effort' category unlike rigorous verification techniques, which
-however require a white box model. This work tries to bridge the gap
-by presenting an automatic falsification technique which models the
-black box system before searching for a violation. It proceeds by (a)
-modeling the behavior of the system using a piecewise affine (PWA)
-discrete time model and, (b) encoding the search for a violation as a
-standard bounded model checking problem. The resulting problem is a
-disjunction of linear programs and can be solved directly using
-off-the-shelf SMT solvers or by multiple runs of a linear programming
-solver. If a counter-example is found, the existence of a
-corresponding violation in the original system is checked for
-reproducibility.
+% Sampling based falsification techniques for black box hybrid dynamical
+% systems use numerical simulations to search for violations of
+% properties (specified in temporal logic). This constraints them to a
+% `best effort' category unlike rigorous verification techniques, which
+% however require a white box model. This work tries to bridge the gap
+% by presenting an automatic falsification technique which models the
+% black box system before searching for a violation. It proceeds by (a)
+% modeling the behavior of the system using a piecewise affine (PWA)
+% discrete time model and, (b) encoding the search for a violation as a
+% standard bounded model checking problem. The resulting problem is a
+% disjunction of linear programs and can be solved directly using
+% off-the-shelf SMT solvers or by multiple runs of a linear programming
+% solver. If a counter-example is found, the existence of a
+% corresponding violation in the original system is checked for
+% reproducibility.
+
+Models of embedded control systems are often so complex that they can
+be treated as effectively black-box systems. Sampling-based methods
+for testing such systems use numerical simulations to search for
+violations of properties (specified in a suitable requirement
+formalism, such as temporal logic). This constrains them to a `best
+effort' category unlike rigorous verification techniques, which
+however require a model amenable to symbolic analysis. This work tries
+to bridge the gap by presenting an automatic falsification technique
+which learns a data-driven abstraction of the black box system before
+searching for a violation. It proceeds by (a) modeling the behavior of
+the system using a piecewise affine (PWA) discrete-time {\em
+relational} model and, (b) encoding the search for a violation in the
+abstract model as a bounded model checking problem. The bounded model
+checking query in this instance can be viewed as a disjunction of
+conjunctions of linear constraints (i.e. linear programs) and can be
+solved directly using off-the-shelf SMT solvers or by multiple runs of
+a linear programming solver. For any counter-example found in the
+abstraction, we check the existence of a corresponding violation in
+the original system. We demonstrate the efficacy of our technique on a
+few dynamical systems examples.
 

### 2017-04-07 implementation.tex
.

--- 2017-02-08 implementation.tex
@@
 \toprule
-Benchmark & BMC & S3CAM-R & S3CAM\\
+    Benchmark & PWA Rel + BMC & PWA Rel + LP & S3CAM\\
 \midrule
-Van der Pol ($\prop_3$)   &$<$1 & $130$ & $24$\\
-Brusselator               &$<$1  & $4$    & $2$\\
-Lorenz                    &$<$1 & $122$  & $35$\\
+    Van der Pol ($\prop_3$)   &$130$ & $0$ & $24$\\
+    Brusselator               &$4$ & $0$ & $2$\\
+    Lorenz                    &$122$ & $0$ & $35$\\
 %B.Ball                    &  &   & \\
-Nav ($\prop_P$)           &$4480$ &$9423$   &$603$ \\
-Nav ($\prop_Q$)           &$970$  &$8105$   &$546$ \\
-Nav ($\prop_R$)           &-  &$TO$   & $2003$\\
-Nav ($\prop_S$)           &-  &$TO$   & $2100$\\
-Bouncing Ball             &-  &$TO$   & $450$\\
+    Nav ($\prop_P$)           &$TO$ & $2100$  & $603$ \\%9423 + 4480
+    Nav ($\prop_Q$)           &$TO$ & $25m$   & $546$ \\%8105 + 970
+    Nav ($\prop_R$)           &$TO$ & $0$  & $2003$\\
+    Nav ($\prop_S$)           &$TO$ & $0$  & $2100$\\
+    Bouncing Ball             &$TO$ & $0$  & $450$\\
 

### 2017-04-07 prelims.tex
.

@@
 In this section, we present the model of the hybrid dynamical system
-under test and the basics of pwa relational abstraction.
+under test, and the background for pwa relational abstractions.
 
-We now use the family of reachability relations presented
-in~\cite{zutshi2014multiple} to charecterize the system $\System$
-under test. We denote the states of the hybrid system $\System$, by $\HybridStates \subseteq \Modes \times \Reals^n$ be the (infinite)
-set of hybrid states of the system $\System$, where $\Modes$ is a
-finite set of discrete modes, and $\Inputs$ be the set of input
-signals to $\System$ of the form $[0,T] \rightarrow \reals^k$ for some
-given time horizon $T$. The family of
-relations $\reach{t,u} \subseteq \HybridStates \times \HybridStates$,
-parameterized by time $t \geq 0$ and input signal $u \in \Inputs$
-defined over time $[0,T]$.  We use the notation $\vx\ \reach{t,u} \vy$
-to denote $(\vx,\vy) \in \reach{t,u}$. The relations $\reach{t,u}$
-satisfy the following basic properties:
+\mypara{The System under test $\System$} is assumed to be a hybrid
+dynamical system. Its state is denoted by $\HybridStates \subseteq
+\Modes \times \Reals^n$, where $\Modes$ is a finite set of discrete
+modes. Let $\Inputs: [0,T] \rightarrow \reals^k$ be the set of input
+signals to $\System$ over a finite time horizon $T$.
 
-We assume a black-box model
-equipped with a forward simulation map $\simulate^{\System}$ that can
-be used to compute the discrete time traces of the system. Let
-$\reach{}$ Given the
-hybrid state space $\vx\in\HybridStates$ and input space $u
-\in\Inputs$ and time step $\Delta_t$, it computes the unique successor
-state $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u}
+For the purpose of falsification, $\System$ is viewed as a black-box
+equipped with a forward simulation map $\simulate^{\System}:
+(\HybridStates, \vu, t) \mapsto \HybridStates$.  A simulation step
+computes the unique reachable state $\vx \in \HybridStates$ in time
+$t$ from a given state input pair $(\vx,\vu)$.  We formalize this
+using the notion of time parameterized reachability relations
+$\reach{t, \vu}_\System \subseteq \HybridStates \times \HybridStates$
+detailed in~\cite{zutshi2014multiple}.
+Hence, $\vy=\simulate^{\System}(\vx,u,t)$ \emph{iff} $\vx \reach{t,u}
 \vy$.
+% Given the hybrid state space $\vx\in\HybridStates$ and input space $u
+% \in\Inputs$ and time step $t$, it computes the unique successor state
+% $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u} \vy$.
+
+\myipara{System Assumptions:} We require that $\System$ be
+forward deterministic. This
+requires existence and uniqueness of trajectories over a finite time
+horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
+the vector field in each hybrid mode and ruling away issues such as
+finite escape times~\cite{Meiss/2007/Differential}. From a practical
+standpoint, these assumptions do not pose significant restrictions for
+the type of problems that we wish to address. Finally, we assume full
+observability of the system state, validate our results only against
+the $\simulate$ function and {\em do not} assume that an analytic closed-form
+representation of the system dynamics is available.
+
+As the relations abstract the solution of a continuous time,
+foward deterministic hybrid dynamical
+system, they are deterministic, causal and satisfy the semi-group property.
+
+From here on, we drop the subscript $\System$.
 
 \begin{definition}
-
     The system under test $\System$ evolves in the set of infinite
@@
     $t+\Delta_t$, under a constant input $\vu$.
-
 \end{definition}
 
-The subscript $\System$ will be dropped whenever the system $\System$
-being referred to is clear from the context.
+\subsection{Learning Dynamics using Simple Linear Regression}
 
-\myipara{Black-Box Assumptions:} For above properties of the
-$\reach{t,u}$ relation to hold, we require that $\System$ be
-deterministic. We also assume that $\System$ can be simulated from a
-given initial state using function $\simulate^{\System}$. This
-requires existence and uniqueness of trajectories over a finite time
-horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
-the vector field in each hybrid mode and ruling away issues such as
-finite escape times~\cite{Meiss/2007/Differential}. From a practical
-standpoint, these assumptions do not pose significant restrictions for
-the type of problems that we wish to address.  Finally, we assume full
-observability of the system state, ignore numerical errors due to
-simulation and {\em do not} assume that an analytic closed-form
-representation of the system dynamics is available.
-
-\subsection{Learning Dynamics using Simple Linear Regression}
+Best effor falsification works by enumerating the reachability
+relations $\reach{}$ within a given budget. Depending upon the
+specific instance of the problem, the budget can be prohibhitive,
+requiring exponential computational and storage resources. Instead, we
+proceed by summarizing these relations by their corresponding affine
+maps using linear regression. The learnt analytical forms are
+represented as $\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are
+matrices and $\delta$ the estimated error.
 
@@
 
-The
 Let the data set be comprised of $N$ input and output pairs $(\x, y)$,
@@
 the current context of \textit{finding the best fit}, the problem is
-over-determined; there are more equations than unknowns. Hence, a
+over-determined: there are more equations than unknowns. Hence, a
 single affine function cannot be found which satisfies the equation
 $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we need
-to find the `best' choice for $\vec{a}$ and $b$. This is formally
+to find the `best' choice for $\vec{a}$ and $b$. This choice is formally
 defined using a loss function. For the case of simple linear
@@
 $\hat{\x}$.
-The details can be found in several texts on learning and statistics
+The details can be found in standard texts on learning and statistics
 \cite{friedman2001elements}.
@@
 
-\subsection{Relational Abstractions}
-For a hybrid automaton model, they can summarize the continuous
-dynamics of each mode using a binary relation over continuous states.
-The resulting relations are timeless (independent) and hence valid for
-all time as long as the mode invariant is satisfied. The relations
-take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
-represents one of the relational operators $=, \ge, \le, <, >$. For
-example, an abstraction that captures the monotonicity with respect to
-time for the differential equation $\dot{x} = 2$ is $x' > x$. The
-abstraction capturing the relation between the set of ODEs: $\dot{x} =
-2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
+% \subsection{Relational Abstractions}
+% For a hybrid automaton model, they can summarize the continuous
+% dynamics of each mode using a binary relation over continuous states.
+% The resulting relations are timeless (independent) and hence valid for
+% all time as long as the mode invariant is satisfied. The relations
+% take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
+% represents one of the relational operators $=, \ge, \le, <, >$. For
+% example, an abstraction that captures the monotonicity with respect to
+% time for the differential equation $\dot{x} = 2$ is $x' > x$. The
+% abstraction capturing the relation between the set of ODEs: $\dot{x} =
+% 2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
 
@@
 \[
-    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
+    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,\pVar')) \;|\;
+    g_\tau(\pVar) \land g'_\tau(\pVar') \land f_{\tau}(\pVar, \pVar')}
 \]
-where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
-$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
-The assertion on the initial states $\Theta$ is also affine.
+where $f_{\tau}$ is an affine relation on the pre and post states
+$\pVar,\pVar'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
+post conditions) on the states and $\Theta$ is an affine assertion on
+the initial states.
 
@@
 where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
-$m$. The affine maps $f_\tau$ can be represented in matrix form as
-$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
-an error term $\delta$ which is defined by a vector of intervals
-$[\delta_{\min},\delta_{\max}]$. Hence,
+$m$. The affine relation $f_\tau$ can be represented using matrix $A$
+and an offset vector $\vb$ as $f_\tau \subseteq \setof{(\x,\x') | \x'
+= A\x+\vb}$. Finaly, we incorporte an approximate non-deterministic
+error term $\delta$, defined as a vector of intervals
+$[\delta_{\min},\delta_{\max}]$.
 \[
-f:\x \mapsto A\x + \vb + \delta
+    f_\tau \subseteq \setof{(\x,\x') | \x' = A\x+\vb}.
 \]
-defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
-\in f_\tau(\x)}$.
+%defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
+%\in f_\tau(\x)}$.
 
@@
         \begin{array}{ll}
-            g_1(\x) \land g'_1(\x') \implies \x' \in f_1(\x) \\
+            g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
             \ldots\\
-            g_n(\x) \land g'_n(\x') \implies \x' \in f_n(\x) \\
+            g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
         \end{array}
@@
 
-where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
-$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
-denotes the next state of the system.
+%where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
+%$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
+where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
+relations and $\x'$ denotes the next state of the system.
 
@@
 deadlock, with no further executions. This usually results from
-incorrect modeling of SDCS, and from here on, we do not consider such
+modeling errors, and from here on, we do not consider such
 cases.
@@
 % map $\gm$.
-
-\subsection{Learning Affine Function from Trajectory Segments}
-
-The $\simulate$ function can be used to generate a data set of
-relations parameterized by time $\Delta_t$, for different values of
-states $\x_i$ and inputs $\vu_i$ as $\ds_{\Delta_t} =
-\setof{(\x_i,\x'_i,\vu_i) | \x'_i = \simulate(\x_i,\vu_i)}$.
-Moreover, by repeatedly applying $\simulate$, a discrete time series
-of a finite length $N$ can be generated from a given state $\x_1$
-and a sequence of inputs $\tupleof{\vu_1, \vu_2,\ldots,\vu_n}$ as shown
-below.
-\[\tupleof{\x_1, \x_2, \ldots, \x_n} \;\text{where}\; \x_{i>1} =
-\simulate(\x_{i-1}, \vu_{i-1}, \Delta_t)\]
-
-We now introduce a refinement of the data set $\ds \subseteq \ds$ by
-incrementally observing the future states in the time series.
-
-\[\ds_{\Delta_t} = \setof{(\x_i,\x'_i,\vu_i) | \x'_i =
-\simulate(\x_i,\vu_i)}\]
-
-\[\ds^{C_i} \subseteq \ds_t\]
-\[\ds^{C_i}\]
-
-$k$-Refinement of the data set $\ds^k$.
-$\ds_k = \setof{(\x,\x',\vu) | \x' =  \simulate(\x,\vu,k\Delta_t)}$
-
-state-input pairs of the form
-$(\x,\x')$
-data set equivalence traj segment
-$D_c, $

### 2017-04-07 pwarel.tex
.

@@
-We now briefly describe the approach. Given a numerical simulator
+\subsection{PWA Relational Abstraction}
+
+Given an abstraction, we can define its enrichment
+
+We now briefly overview the approach, which consists of three
+steps
+
+Discover the abstraction 
+enriches the abstraction
+formulates a BMC query
+concretizes
+
+abstraction defined by a quantization function $\quant_\epsilon$
+time step $\Delta$
+
+Given a numerical simulator
 $\simulate$ and an initial abstraction defined by a quantization
@@
 for each edge of the reachability graph.
+
+\subsection{Graph Abstraction}
+DEFINITION 3.4 (ABSTRACT STATE GRAPH). Let ∆ > 0 be
+a fixed time step. The abstract state graph H(∆) for time step ∆
+is defined by the set of vertices C, and edges (Ci, Cj ) whenever
+Ci
+∆ Cj . Let C0 denote the collection of initial cells in C, i.e., cells
+Ci such that Ci ∩ X0 6= ∅. Further, let Cu denote the set of unsafe
+cells, or cells Cj such that there is a state xj ∈ Cj that reaches
+an unsafe state y ∈ Xf within time 0 ≤ tj < ∆. The abstraction
+H(∆) is given by D
+C,
+∆ , C0, Cu
+E
+.
+
+\subsection{Enriched Graph Abstraction}
 
@@
 We now describe how relational models can be computed for a given
-black box system by enriching the abstract reachability graph obtained
-by scatter-and-simulate. We first formalize the notion of an enriched
-graph, and then show how they can be interpreted as a PWA transition
-system using $k$-relational modeling.
+black box system by enriching the abstract reachability graph. We
+first formalize the notion of an enriched graph, and then show how
+they can be interpreted as a PWA transition system using
+$k$-relational modeling.
 
@@
 \figref{k2} illustrates the $k=2$ refinement on \figref{k1}.
+
+\subsection{Learning Affine Function from Trajectory Segments}
+
+The $\simulate$ function can be used to generate a data set of
+relations parameterized by time $\Delta_t$, for different values of
+states $\x_i$ and inputs $\vu_i$ as $\ds_{\Delta_t} =
+\setof{(\x_i,\x'_i,\vu_i) | \x'_i = \simulate(\x_i,\vu_i)}$.
+Moreover, by repeatedly applying $\simulate$, a discrete time series
+of a finite length $N$ can be generated from a given state $\x_1$
+and a sequence of inputs $\tupleof{\vu_1, \vu_2,\ldots,\vu_n}$ as shown
+below.
+\[\tupleof{\x_1, \x_2, \ldots, \x_n} \;\text{where}\; \x_{i>1} =
+\simulate(\x_{i-1}, \vu_{i-1}, \Delta_t)\]
+
+We now introduce a refinement of the data set $\ds \subseteq \ds$ by
+incrementally observing the future states in the time series.
+
+\[\ds_{\Delta_t} = \setof{(\x_i,\x'_i,\vu_i) | \x'_i =
+\simulate(\x_i,\vu_i)}\]
+
+\[\ds^{C_i} \subseteq \ds_t\]
+\[\ds^{C_i}\]
+
+$k$-Refinement of the data set $\ds^k$.
+$\ds_k = \setof{(\x,\x',\vu) | \x' =  \simulate(\x,\vu,k\Delta_t)}$
+
+state-input pairs of the form
+$(\x,\x')$
+data set equivalence traj segment
+$D_c, $

### 2017-04-07 prelims.tex
.

@@
 equipped with a forward simulation map $\simulate^{\System}:
-(\HybridStates, \vu, t) \mapsto \HybridStates$.  A simulation step
-computes the unique reachable state $\vx \in \HybridStates$ in time
+(\vx, \vu, t) \mapsto \vx$.  A simulation step
+involves computing the unique reachable state $\vx \in \HybridStates$ in time
 $t$ from a given state input pair $(\vx,\vu)$.  We formalize this
@@
 $\reach{t, \vu}_\System \subseteq \HybridStates \times \HybridStates$
-detailed in~\cite{zutshi2014multiple}.
-Hence, $\vy=\simulate^{\System}(\vx,u,t)$ \emph{iff} $\vx \reach{t,u}
-\vy$.
+detailed in~\cite{zutshi2014multiple}. Finally, $\vx'=\simulate^{\System}(\vx,u,t)$ \emph{iff} $\vx \reach{t,u} \vx'$.
 % Given the hybrid state space $\vx\in\HybridStates$ and input space $u
@@
 
-\myipara{System Assumptions:} We require that $\System$ be
-forward deterministic. This
-requires existence and uniqueness of trajectories over a finite time
-horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
-the vector field in each hybrid mode and ruling away issues such as
-finite escape times~\cite{Meiss/2007/Differential}. From a practical
-standpoint, these assumptions do not pose significant restrictions for
-the type of problems that we wish to address. Finally, we assume full
+\myipara{System Assumptions:} We require that $\System$ be forward
+deterministic. This requires existence and uniqueness of trajectories
+over a finite time horizon $[0,T]$, which can be guaranteed by
+Lipschitz continuity of the vector field in each hybrid mode and
+ruling away issues such as finite escape
+times~\cite{Meiss/2007/Differential}. From a practical standpoint,
+these assumptions do not pose significant restrictions for the type of
+problems that we wish to address. Finally, we assume full
 observability of the system state, validate our results only against
-the $\simulate$ function and {\em do not} assume that an analytic closed-form
-representation of the system dynamics is available.
+the $\simulate$ function and {\em do not} assume that an analytic
+closed-form representation of the system dynamics is available.
 

### 2017-04-07 prelims.tex

@@
 
-\myipara{System Assumptions:} We require that $\System$ be forward
-deterministic. This requires existence and uniqueness of trajectories
-over a finite time horizon $[0,T]$, which can be guaranteed by
-Lipschitz continuity of the vector field in each hybrid mode and
-ruling away issues such as finite escape
-times~\cite{Meiss/2007/Differential}. From a practical standpoint,
-these assumptions do not pose significant restrictions for the type of
-problems that we wish to address. Finally, we assume full
-observability of the system state, validate our results only against
-the $\simulate$ function and {\em do not} assume that an analytic
-closed-form representation of the system dynamics is available.
-
-As the relations abstract the solution of a continuous time,
-foward deterministic hybrid dynamical
-system, they are deterministic, causal and satisfy the semi-group property.
-
-From here on, we drop the subscript $\System$.
+\mypara{System Assumptions:} We require that $\System$ can always be
+simulated forward in time with deterministic results. This in turn
+requires existence and uniqueness of trajectories over a finite time
+horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
+the vector field in each underlying hybrid mode and ruling away issues
+such as finite escape times~\cite{Meiss/2007/Differential}.
+Furthermore, as the relations abstract the solution of a continuous
+time, foward deterministic hybrid dynamical system, they are
+deterministic, causal and satisfy the semi-group property.  These
+assumptions are justified as for the dynamical systems we address in
+this presentation. Finally, we assume full observability of the system
+state, validate our results only against the $\simulate$ function (as
+the analytic closed-form representation of the system dynamics is
+assumed to be unavailable).  From here on, we drop the subscript
+$\System$.
 

### 2017-04-07 prelims.tex
.

@@
 $\reach{t, \vu}_\System \subseteq \HybridStates \times \HybridStates$
-detailed in~\cite{zutshi2014multiple}. Finally, $\vx'=\simulate^{\System}(\vx,u,t)$ \emph{iff} $\vx \reach{t,u} \vx'$.
+detailed in~\cite{zutshi2014multiple}.
 % Given the hybrid state space $\vx\in\HybridStates$ and input space $u
@@
 % $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u} \vy$.
+
+\begin{definition}[System Evolution]
+    The system under test $\System$ evolves in the set of infinite
+    hybrid state-space $\HybridStates$ under the effect of inputs $\vu
+    \in \Inputs$. Its evolution is defined by the simulation function
+    $\simulate^\System : \HybridStates \times \Inputs \times \Reals^+ \mapsto
+    \HybridStates$, where $\simulate^\System(\x,\vu,\Delta_t)$ maps a
+    current state $\x$ at time $t$ to a new state $\x'$ at time
+    $t+\Delta_t$, under a constant input $\vu$.
+    Finally, $\vx'=\simulate^{\System}(\vx,u,t)$ \emph{iff} $\vx
+    \reach{t,u} \vx' \in \reach{t, \vu}_\System$.
+\end{definition}
 
@@
 $\System$.
-
-\begin{definition}
-    The system under test $\System$ evolves in the set of infinite
-    hybrid state-space $\HybridStates$ under the effect of inputs $\vu
-    \in \Inputs$. Its evolution is defined by a simulation function
-    $\simulate^\System : \HybridStates \times \Inputs \Reals^+ \mapsto
-    \HybridStates$, where $\simulate^\System(\x,\vu,\Delta_t)$ maps a
-    current state $\x$ at time $t$ to a new state $\x'$ at time
-    $t+\Delta_t$, under a constant input $\vu$.
-\end{definition}
 

### 2017-04-07 pwarel.tex
.

--- 2017-04-07 pwarel.tex
@@
 discuss the background required to present our ideas.
-
-\section{Relational Modeling}
-
-We now describe how relational models can be computed for a given
-black box system by enriching the abstract reachability graph obtained
-by scatter-and-simulate. We first formalize the notion of an enriched
-graph, and then show how they can be interpreted as a PWA transition
-system using $k$-relational modeling.
-
-\subsection{Abstract Enriched Graph}
-
-The existential abstraction relation in \cite{zutshi2014multiple} was defined as
-follows:
-\[
-%\exists \x\in\C. \exists \x'\in\C'. \x'=\simulate(\x, \tau) \iff \C\rel{\tau}\C'
-    C \areach{t} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x \areach{t} \x'
-\]
-The abstract relation $\areach{t}$ can be enriched by incorporating
-the affine relation between $\x$ and $\x'$. For an arbitrary dynamical
-system, such a relation can be rarely represented using an exact affine map.
-This is due to the presence of non-linear and hybrid behaviors. But,
-an affine map can always be estimated with an error.
-
-If the system dynamics are completely specified in the form of a white
-box model, we can use first order approximations to find the affine
-expressions for the relations. Using a tool like \flowstar, we can
-obtain sound over-approximate affine maps of the form $A\x + [\vb^l,
-\vb^h]$, where $[\vb^l,\vb^h]$ denotes the interval of vectors, such
-that, every element $b_i$ of the vector $\vb$ is contained in the
-respective scalar interval $b_i\in[b^l_i,b^h_i]$.
-
-For the case of black box systems, such a sound approximation is not
-possible. Instead, we rely on a statistical method like simple linear
-regression to estimate the affine map $\amap: A\x + \vb$. We then
-estimate the error $\delta$ and generalize $\amap$ to an interval
-affine map as before $\amap: A\x + \vb + \delta$.  Finally, we get an
-abstraction with the below relation
-\[
-    C \areach{f} C' \iff \exists \x \in C.\ \exists \x' \in C'.\ \x' \in A\x + \vb \pm \delta.
-    %C \areach{A\x+\vb+\delta} C' \iff \exists \x \in C.\ \exists \x' \in C'.\
-\]
-
-However, in the presence of complex non-linear behavior, using a
-single affine map can lead to a poor approximation. To increase the
-precision, one can use more than one affine map. Let us denote this set as
-$R_{(C,C')}$ (ref.~\secref{ovw}).
-\[
-    R_{(C,C')}(\x,\x'): \setof{f_i \;|\; \exists \x \in C.\ \exists \x' \in C'.\ \x' \in f_i(\x)}
-\]
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-    \draw [line] (-1.2,-1.2) rectangle (1.2,1.2);
-    \draw[arw] (-3,0) -- (-1.2,0);
-    \draw[arw] (1.2,0) -- (3,0);
-
-\node at (-3.4,0) {$G$};
-\node at (3.4,0) {$G^R$};
-\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Using OLS, $G^R$ is computed by determining the appropriate
-$R_{(C,C')}$ for each edge of $G$.}
-\label{fig:enrichment}
-\vspace*{-.3cm}
-\end{figure}
-
-As shown in \figref{enrichment}, we use OLS to compute an enrichment
-of the abstraction graph $G$. For every edge $(C,C') \in edges(G)$,
-the set of relations $R_{(C,C')}(\x,\x')$ is computed. The semantics
-of the enriched edge are clearly non-deterministic. Form a state $\x
-\in C$, any $f_i\in R_{(C,C')}$ can be taken as long as $\x' \in C$'
-is satisfied. This interpretation results in an infinite state
-transition system. We now detail the construction of the set
-$R_{(C,C')}$ using $k$-relational modeling.
-
-% We now compute a PWA model for a given black box system by estimating
-% the dynamics using OLS. We can estimate the affine relations between
-% every cell, but as discussed in ~\chapref{abs}, it is a futile
-% approach.  Instead, we use the same heuristic as before:
-% scatter-and-simulate, to select the cell relations for estimation. We
-% build the reachability graph as before, but in addition, annotate each
-% edge with the values of estimated $A$, $\vb$, and $\delta$ for the
-% respective relation. The reachability graph thus computed, is a
-% transition system. Using off-the-shelf bounded model checkers, we
-% reason about the system's safety properties.
-
-\subsection{$k$ - Relational Modeling}
-
-\begin{figure}[!htbp]
-\vspace*{-.2cm}
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-
-\begin{scope}[scale=0.8]
-\draw [line] (-1.0,0) circle (0.5);
-\draw [line] (-4.0,0) circle (0.5);
-\draw [line] (2.0,1) circle (0.5);
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-0.5,0) -- (1.5,1.0);
-\draw[arw] (-0.5,0) -- (1.5,-1.0);
-\draw[arw] (-3.5,0) -- (-1.5,0.0);
-\node at (-4.0,0) {$C$};
-\node at (-1.0,0) {$C'$};
-\node at (2.0,1.0) {$C''_1$};
-\node at (2.0,-1.0) {$C''_2$};
-\draw [line] (2.0,-1) circle (0.5);
-\draw[arw] (-5.0,-2.0) -- (3.0,-2.0);
-\node at (-4.0,-2.5) {$k=0$};
-\node at (-1.0,-2.5) {$k=1$};
-\node at (+2.0,-2.5) {$k=2$};
-\draw [fill=black] (-4.0,-2.0) circle (0.05);
-\draw [fill=black] (-1.0,-2.0) circle (0.05);
-\draw [fill=black] (2.0,-2.0) circle (0.05);
-
-\end{scope}
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Nodes/Cells of $G$ shown along with increasing values of $k$.}
-\label{fig:k-rel}
-\vspace*{-.3cm}
-\end{figure}
-
-Given a $G$, when computing the set of relations for an edge $R$, we
-consider only the set of abstract states reachable at a specific time
-step.  Intuitively, one might consider the states reachable in one
-time step $\Delta$.  However, such a process can be generalized by
-looking at the states reachable at ${0\Delta, 1\Delta, \ldots,
-k\Delta}$ time steps.
-
-Using \figref{k-rel}, we illustrate this notion. To compute
-$R_{(C,C')}$, we
-split the data set $D(C,C')$ consisting of trajectory segments (of time lengths
-$k\Delta$) as follows.  We observe the local behavior of the system by
-noting the evolution of the system $\System$ from a state $\x(t)
-\in C$ for a time length dependant on $k$. For
-\begin{itemize}
-    \item{$k=0$}: we observe all trajectory segments of length $\Delta$.
-    \item{$k=1$}: we observe trajectory segments of length $\Delta$, which satisfy $\x(t+\Delta) \in C'$.
-    \item{$k=2$}: we observe trajectory segments of length $2\Delta$,
-        and split them in two sets (a) and (b) on the basis of the
-        cells reached at time $(t+\Delta)$ and $(t+2\Delta)$.
-        \begin{enumerate}[(a)]
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_1$
-            \item $\x(t+\Delta) \in C' \land \x(t+2\Delta) \in C''_2$
-        \end{enumerate}
-    \item{$k=n$}: we observe trajectory segments of length $n\Delta$,
-        and split them in to multiple sets on the basis of the cells
-        reached at time $(t+\Delta),\; (t+2\Delta),\;
-        \ldots,\;(t+n\Delta)$.
-\end{itemize}
-
-$k$-relational modeling can be understood as a heuristic, which uses
-the underlying abstraction to differentiate behaviors of the system
-which `diverge' (or are revealed to be `distinct' at a future time).
-Such a heuristic can be useful in increasing the precision of the
-learnt affine maps.  We now formalize this notion for $k=0$, $1$, and
-$n$ and illustrate using examples.
-
-\mypara{$0$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    %\draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (3.7,2.2) {\scriptsize{$\x_0'$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (3.7,1.7) {\scriptsize{$\x_1'$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.7,-1.2) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-2.0) -- (4.5,-2.0);
-\node at (0,-2.5) {$t=0$};
-\node at (4,-2.5) {$t=\Delta$};
-\draw [fill=black] (0.0,-2.0) circle (0.05);
-\draw [fill=black] (4.0,-2.0) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=2.0cm,scale=0.5]
-    \node at (2,0) {\footnotesize $\scr{T} \equiv g_1(\x):\vx \in C
-    \implies A_1\vx + \vb_1 + \delta_1$};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{All the trajectory segments will be used to construct the
-model; $\ds = \setof{\pi_1, \pi_2, \pi_3}$.}
-\label{fig:k0}
-\vspace*{-.3cm}
-\end{figure}
-
-When $k=0$, the $0$-relational model is a PWA transition system which only
-defines the evolution of states $\x$, and does not specify the
-reachable cell. The guard predicates of the relations are defined over
-cells: $g_i(\x): \x \in C_i$ while $g'_i(\x'): True$.
-
-We use regression to estimate the dynamics for the outgoing trajectory
-segments form a cell $C$. Hence, the data set $\ds$ for the regression
-includes all trajectory segments beginning from the same cell
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C}.
-\]
-This includes trajectory segments ending in different cells, as shown
-in ~\figref{k0}. For $N$ cells, this results in a $\rho$ with $N$ transition relations.
-\begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            \ldots \\
-            g_n(\x):\vx \in C_n \implies A_n\vx + \vb_n + \delta_n\\
-        \end{array}
-    \right.
-\end{equation*}
-
-Note that this can be quite imprecise when the cells are big,
-containing regions of state-space with complex dynamics. This is true
-for both non-linear systems and hybrid dynamical system, where a cell
-can contain two or more modes with differing continuous dynamics.
-
-\mypara{$1$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    \draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-    \draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\node at (1.7,-1) {$\pi_3$};
-\draw[arw] (-1.5,-3.5) -- (5.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-
-\end{scope}
-\begin{scope}[xshift=0,yshift=2.5cm,scale=0.5]
-    \node at (2,0) {\begin{minipage}{\linewidth}
-    \footnotesize \begin{equation*}
-        \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The data gets split into two sets $\ds_1 = \setof{\pi_1,
-    \pi_2}$ and $\ds_2 = \setof{\pi_3}$ and two relations:
-    $R_{(C,C'_1)}$ and $R_{(C,C'_2)}$, each with one affine map, are constructed.}
-%Cardinality($R_{(C,C')}$) = $2$.}
-\label{fig:k1}
-\vspace*{-.3cm}
-\end{figure}
-
-% \begin{scope}[xshift=0,yshift=4.0cm,scale=1]
-%     \node at (2,0) {\footnotesize $\begin{equation*}
-%     \scr{T} = \left\{
-%         \begin{array}{ll}
-%             g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-%             g_2(\x):\vx \in C \land g'_2(\x):\vx \in C'_2 \implies A_2\vx + \vb_2 + \delta_2
-%         \end{array}
-%     \right.
-% \end{equation*}$};
-% \end{scope}
-
-To improve the preciseness of learnt dynamics, we include the
-reachability relation in the
-regression. For every relation $C\areach{t}C'$, the data set $\ds$ is
-comprised only of trajectory segments $\pi_t$ which start and end in
-the same set of cells.
-\[
-    \ds = \setof{\pi_t | start(\pi_t) \in C \land end(\pi_t) \in C'}.
-\]
-For $N$ edges in $G$, $1$-relationalization results in a $\rho$ with
-$N$ transition relations.
-
-\figref{k1} illustrates the $k=1$ refinement of the case shown in
-\figref{k0}. The data set $\ds$ is split into two data sets $\ds_1
-= \setof{\pi_1, \pi_2}$ and $\ds_2 = \setof{\pi_3}$, using which, two
-relations are constructed. The resulting $1$-relational model
-is at least as precise as the corresponding $0$-relational model.
-
-\mypara{$k$-Relational Model}
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\begin{tikzpicture}
-\begin{scope}[scale=0.5]
-
-    \draw [line] (-1.0,-1.0) rectangle (1.0,1.0);
-    \draw [line] (3.0,1.0) rectangle (5.0,3.0);
-    %\draw [line] (3.0,-3.0) rectangle (5.0,-1.0);
-    \draw [line] (7.0,1.0) rectangle (9.0,3.0);
-    \draw [line] (7.0,-3.0) rectangle (9.0,-1.0);
-    \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,2.2) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,1.8) ;
-    \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (3.5,1.7) .. controls +(2.0,2.0) and +(-2.0,-2.0) .. (8.5,-2.2) ;
-
-%     \draw [line] (-0.7,0.5) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,2.2);
-%     \draw [line] (-0.5,0.0) .. controls +(2.0,2.0) and +(-2.0,-2.0) ..  (3.5,1.7);
-
-    %\draw [line] (-0.0,-0.5) .. controls +(2.0,2.0) and +(-2.0,-1.0) ..  (3.5,-1.2);
-
-\node at (0.0, 0.0) {$C$};
-\node at (4.0, 2.0) {$C'_1$};
-\node at (8.0, 2.0) {$C''_1$};
-\node at (8.0, -2.0) {$C''_2$};
-%\node at (4.0, -2.0) {$C'_2$};
-\node at (-0.75,0.7) {\scriptsize{$\x_0$}};
-\node at (-0.7,0.0) {\scriptsize{$\x_1$}};
-%\node at (-0.2,-0.5) {\scriptsize{$\x_2$}};
-\node at (3.3,2.5) {\scriptsize{$\x_0'$}};
-\node at (3.7,1.5) {\scriptsize{$\x_1'$}};
-%\node at (3.7,-1.5) {\scriptsize{$\x_2'$}};
-\node at (8.7,1.6) {\scriptsize{$\x_0''$}};
-\node at (8.5,-2.7) {\scriptsize{$\x_1''$}};
-\node at (2,1.5) {$\pi_1$};
-\node at (2,0.5) {$\pi_2$};
-\draw[arw] (-1.5,-3.5) -- (9.5,-3.5);
-\node at (0,-4.0) {$t=0$};
-\node at (4,-4.0) {$t=\Delta$};
-\node at (8,-4.0) {$t=2\Delta$};
-\draw [fill=black] (0.0,-3.5) circle (0.05);
-\draw [fill=black] (4.0,-3.5) circle (0.05);
-\draw [fill=black] (8.0,-3.5) circle (0.05);
-\end{scope}
-
-\begin{scope}[xshift=0,yshift=2.5cm,scale=0.5]
-    \node at (4,0) {\begin{minipage}{\linewidth}
-        \footnotesize \begin{equation*}
-    \scr{T} = \left\{
-        \begin{array}{ll}
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_1\vx + \vb_1 + \delta_1\\
-            g_1(\x):\vx \in C \land g'_1(\x):\vx \in C'_1 \implies A_2\vx + \vb_2 + \delta_2
-        \end{array}
-    \right.
-\end{equation*}
-\end{minipage}};
-\end{scope}
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{The $k=2$ refinement further splits the data set $\ds_1$
-    into two sets $\ds_{11} = \setof{\pi_1}$ and $\ds_{12} =
-    \setof{\pi_2}$ and $R_{(C,C'_1)}$ now has two affine maps,
-    non-deterministically defining the system behavior.}
-\label{fig:k2}
-\vspace*{-.3cm}
-\end{figure}
-
-Finally, we generalize the relational models to $k$-relational PWA
-models. A $k$-relational model is constructed by using $k$ length
-\textit{connected} segmented trajectories. A segmented trajectory $\SegTraj$ is
-\textit{connected} iff its cost $\cost(\SegTraj) = 0$.
-
-Two $\SegTraj:\tupleof{\pi_{t_1}, \ldots, \pi_{t_k}}$ and
-$\SegTraj':\tupleof{\pi'_{t_1}, \ldots, \pi'_{t_k}}$ are
-\textit{similar} if their trajectory segments have the same sequence
-of cell traversals $\tupleof{C_1, C_2, \ldots, C_k}$, \ie
-\begin{align*}
-    \forall i\in\setof{1 \ldots k}.
-    & start(\pi_{t_i}) \in C_i          
-    \land start(\pi'_{t_i}) \in C_i \land  \\
-    & end(\pi_{t_i}) \in C_{i+1}  
-    \land end(\pi'_{t_i}) \in C_{i+1}
-\end{align*}
-
-For every cell $C$ in the reachability graph, we construct a data set
-$\ds$ by collecting $\pi_{t_i}$, such that they are the first segment of $k$
-length \textit{connected} and \textit{similar} segmented trajectories.
-%\[
-%    \ds = \setof{\pi_{\tau_i} |
-%        \pi_{\tau_i}
-%        start(\pi_{\tau_{i}}) \in C_i
-%        \land
-%        end(\pi_{\tau_{i}}) \in C_i}
-%\]
-
-\figref{k2} illustrates the $k=2$ refinement on \figref{k1}.

### 2017-04-08 prelims.tex
.

@@
 
+\subsection{Bounded Model Checking (BMC)}
+The PWA transition system is a finite location, infinite state
+transition system. As all the guards and transitions are affine, a
+bounded time reachability query is equivalent to checking all
+combinaions of discrete locations, where each combination can be
+summarized as a linear program. As the time is bounded, the number of
+combinations are finite, although exponential in number.
+
+It should be noted that an explicit execution of the transition system
+represents a directed acyclic graph, with each location being a node,
+and a directed edge (in the direction of time advance) from each node
+to its reachable node. A path on this graph is then a linear program,
+with each branching corresponsing to a logical disjunction.
+
+Temporal properties of pwa transitions systems can be checked using
+off-the-shelf model checkers like SAL~\cite{SAL-SRI} which can verify
+infinite state transitions systems.  Furthermore, lazy SMT solvers can
+be used for this specific problem~\cite{shoukry2017smc}. We now show
+how relational pwa abstractions can be viewed as pwa transition
+systems and subsequenly, the problem of falsification can be answered
+by a BMC query.
+
 % Given a dynamical system over a continuous state-space

### 2017-04-08 prelims.tex
.

--- 2017-04-07 prelims.tex
@@
 
+\subsection{Bounded Model Checking (BMC)}
+The PWA transition system is a finite location, infinite state
+transition system. As all the guards and transitions are affine, a
+bounded reachability query is equivalent to checking all
+combinaions of discrete locations, where each combination can be
+summarized as a linear program. As the time is bounded, the number of
+combinations are finite, but, exponential in number.
+
+It should be noted that an explicit execution of the transition system
+represents a directed acyclic graph, with each location being a node,
+and a directed edge from each node to all nodes reachable in forward
+time. A path on this graph is then a linear program, with each
+branching node corresponsing to a logical disjunction.
+
+Temporal properties of pwa transitions systems can be checked using
+off-the-shelf model checkers like SAL~\cite{SAL-SRI}, which can reason
+over infinite state transitions systems using SMT solvers.
+Furthermore, lazy SMT solvers can be employed for this specific
+problem instance to achieve better efficiency~\cite{shoukry2017smc}.
+We now show how relational pwa abstractions can be viewed as pwa
+transition systems, and how the problem of falsification can
+be answered by a BMC query.
+
 % Given a dynamical system over a continuous state-space

### 2017-04-08 pwarel.tex
.

@@
-\subsection{PWA Relational Abstraction}
-
-Given an abstraction, we can define its enrichment
-
-We now briefly overview the approach, which consists of three
-steps
-
-Discover the abstraction 
-enriches the abstraction
-formulates a BMC query
-concretizes
+A discrete abstraction over a continuous space uses relations to
+describe the behavior of the system over time. Such relations abstract
+the underlying relation between concrete states, and concrete paths
+can no longer be constructed directly. Instead, search procedueres
+like Counter-Example Guided Refinement must be used to find concrete
+paths. In general, search techniques are exponential with respect to
+computational resources and memory. As an alternative, we introduce
+data driven `enrichment' of abstractions to approximate the concrete
+relations underlying the abstract relations. Informally, we use linear
+regression to compute affine maps associated with the relations
+between two abstract states. We then show how associating the computed
+maps with the relations transforms a discrete abstraction into a PWA
+transitions system.
 
@@
 \subsection{Graph Abstraction}
-DEFINITION 3.4 (ABSTRACT STATE GRAPH). Let ∆ > 0 be
-a fixed time step. The abstract state graph H(∆) for time step ∆
-is defined by the set of vertices C, and edges (Ci, Cj ) whenever
-Ci
-∆ Cj . Let C0 denote the collection of initial cells in C, i.e., cells
-Ci such that Ci ∩ X0 6= ∅. Further, let Cu denote the set of unsafe
+The abstraction of the hybrid dynamical system is defined by
+paritioning the combiend state-input space $\HybridStates \times
+\Inputs$ into hyper-rectangles or cells $C_i \in \C$. The cells are
+pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
+their union $\bigcup_{C_i\in\C}C_i$ is the state-input
+space $\HybridStates \times \Inputs$. This discrete abstraction over
+the continuous space is implicit and defined using a quantization
+function $Q_d$ parameterized by the precision $d$.
+This further
+defines an equivalence relations such that
+\[
+    (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; Q()
+\]
+
+$(\vx_i, \vu_i) \in C_i \; \text{iff} \; \C_i = Q(\vx_i, \vu_i)$
+
+\begin{definition}[] Let C be a d-quantized tiling. The d-quantized
+    tiling-based plant abstraction (denoted Pd) is given by a graph
+    (V, E), where V = C, and the set of edges E is defined such that
+    (Cj, Ck) ∈ E iff: there is a state-input pair (x, u) ∈ Cj and a
+state-input pair (x 0 , u) ∈ Ck such that x 0 = SIM(x, u, τ ).
+\end{definition}
+
+\begin{definition}[Abstract State Graph]. Let ∆ > 0 be a fixed time
+    step. The abstract state graph H(∆) for time step ∆ is defined by
+    the set of vertices C, and edges (Ci, Cj ) whenever Ci ∆ Cj . Let
+    C0 denote the collection of initial cells in C, i.e., cells Ci
+    such that Ci ∩ X0 6= ∅. Further, let Cu denote the set of unsafe
     cells, or cells Cj such that there is a state xj ∈ Cj that reaches
     an unsafe state y ∈ Xf within time 0 ≤ tj < ∆. The abstraction
-H(∆) is given by D
-C,
-∆ , C0, Cu
-E
-.
+H(∆) is given by D C, ∆ , C0, Cu E .  \end{definition}
 
-\subsection{Enriched Graph Abstraction}
+\subsection{Enriched Abstraction}
 

### 2017-04-08 pwarel.tex
.

@@
 
-abstraction defined by a quantization function $\quant_\epsilon$
-time step $\Delta$
+\subsection{Graph Abstraction}
 
-Given a numerical simulator
-$\simulate$ and an initial abstraction defined by a quantization
-function $\quant_\epsilon$ and a time step $\Delta$, we use
-scatter-and-simulate (parameterized by number of samples $n$ as
-described in \cite{zutshi2014multiple}) to explore the abstract graph
-$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
-number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
-\areach{\Delta} C'$.
-
-Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
-we use the generated trajectory segments to learn quantitative models
-describing the local behavior of the system. These models are defined
-by a set of relations $R \subseteq \HybridStates \times \HybridStates$
-for each edge of the reachability graph.
-
-\subsection{Graph Abstraction}
 The abstraction of the hybrid dynamical system is defined by
@@
 the continuous space is implicit and defined using a quantization
-function $Q_d$ parameterized by the precision $d$.
-This further
-defines an equivalence relations such that
+function $Q_\epsilon$ parameterized by the precision $\epsilon$.
+This further defines an equivalence relations such that
 \[
@@
 \end{definition}
+
+Given a numerical simulator $\simulate$ and an initial abstraction
+defined by a quantization function $\quant_\epsilon$ and a time step
+$\Delta$, we pick samples and simulate them to explore the abstraction
+on the fly to obtain a graph $G$, which has a finite number of
+abstract states $C$ (or cells) and edges $(C,C')$ iff $C
+\areach{\Delta} C'$.
 
@@
 discuss the background required to present our ideas.
+
+Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
+we use the generated trajectory segments to learn quantitative models
+describing the local behavior of the system. These models are defined
+by a set of relations $R \subseteq \HybridStates \times \HybridStates$
+for each edge of the reachability graph.

### 2017-04-08 pwarel.tex
.

@@
 \[
-    (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; Q()
+    (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; Q_\epsilon(\vx_i,
+    \vu_i) = Q_\epsilon(\vx_j, \vu_j)
 \]
+In other words, $(\vx_i, \vu_i) \in C_i \; \text{iff} \; C_i =
+Q_\epsilon(\vx_i, \vu_i)$.
 
-$(\vx_i, \vu_i) \in C_i \; \text{iff} \; \C_i = Q(\vx_i, \vu_i)$
-
-\begin{definition}[] Let C be a d-quantized tiling. The d-quantized
-    tiling-based plant abstraction (denoted Pd) is given by a graph
-    (V, E), where V = C, and the set of edges E is defined such that
-    (Cj, Ck) ∈ E iff: there is a state-input pair (x, u) ∈ Cj and a
-state-input pair (x 0 , u) ∈ Ck such that x 0 = SIM(x, u, τ ).
+\begin{definition}[Graph Abstraction]
+    The quantization function $Q_\epsilon$ induces a DAG on the discrete
+    abstraction given by a graph $G(V, E)$, where each vertex
+    corresponds to a cell $V_i = C_i$, and an edge exists between two
+    cells $(C_j, C_k) \in E$ iff there exist state-input pairs
+    $(\vx,\vu) \in C_j$, $(\vx', \vu) \in C_k$ such that $\vx =
+    \simulate(\vx, \vu )$.
 \end{definition}
@@
 
-\begin{definition}[Abstract State Graph]. Let ∆ > 0 be a fixed time
-    step. The abstract state graph H(∆) for time step ∆ is defined by
-    the set of vertices C, and edges (Ci, Cj ) whenever Ci ∆ Cj . Let
-    C0 denote the collection of initial cells in C, i.e., cells Ci
-    such that Ci ∩ X0 6= ∅. Further, let Cu denote the set of unsafe
-    cells, or cells Cj such that there is a state xj ∈ Cj that reaches
-    an unsafe state y ∈ Xf within time 0 ≤ tj < ∆. The abstraction
-H(∆) is given by D C, ∆ , C0, Cu E .  \end{definition}
+% \begin{definition}[Abstract State Graph]. Let ∆ > 0 be a fixed time
+%     step. The abstract state graph H(∆) for time step ∆ is defined by
+%     the set of vertices C, and edges (Ci, Cj ) whenever Ci ∆ Cj . Let
+%     C0 denote the collection of initial cells in C, i.e., cells Ci
+%     such that Ci ∩ X0 6= ∅. Further, let Cu denote the set of unsafe
+%     cells, or cells Cj such that there is a state xj ∈ Cj that reaches
+%     an unsafe state y ∈ Xf within time 0 ≤ tj < ∆. The abstraction
+% H(∆) is given by D C, ∆ , C0, Cu E .  \end{definition}
 

### 2017-04-08 pwarel.tex
.

@@
     abstraction given by a graph $G(V, E)$, where each vertex
-    corresponds to a cell $V_i = C_i$, and an edge exists between two
-    cells $(C_j, C_k) \in E$ iff there exist state-input pairs
-    $(\vx,\vu) \in C_j$, $(\vx', \vu) \in C_k$ such that $\vx =
+    corresponds to a cell $V_i = C$, and an edge exists between two
+    cells $(C, C') \in E_i$ iff there exist state-input pairs
+    $(\vx,\vu) \in C$, $(\vx', \vu) \in C'$ such that $\vx' =
     \simulate(\vx, \vu )$.
@@
 Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
-trajectory segment between the respective cells. The graph abstraction
+trajectory between the respective cells. The graph abstraction
 $G$ only states that there exists a state $\x \in C$ from which the
@@
 only present the former. Using regression analysis on the $start$ and
-$end$ states of the witnessed trajectory segments between two cells,
+$end$ states of the witnessed trajectory between two cells,
 we compute an approximate discrete map along with an error estimate.
 Moreover, using the simulation function $\simulate$, additional
-trajectory segments (or data) can be generated if required. The data
+trajectories (or data) can be generated if required. The data
 can be separated into a training set and testing set to compute the
@@
 Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
-we use the generated trajectory segments to learn quantitative models
+we use the generated trajectories to learn quantitative models
 describing the local behavior of the system. These models are defined

### 2017-04-08 intro.tex
.
new intro

@@
-Model based design has been quite successful in the industry. 
-Tools like \SIMULINK are often used to create models describing hybrid
-dynamical systems. Testing such systems for safety properties remains
-a hard problem, and the current approaches consist of either guided
-testing using numerical simulations or translating to hybrid automata.
-We now suggest an approach which lies in the middle.
+% Testing of hybrid systems is a complex task, due to the interaction
+% between discrete and continuous dynamics. Industrially developed
+% hybrid systems, with embedded software interacting with physical
+% components, are typically safety-critical. Thus, it is important to
+% guarantee that the systems adhere to functional and safety
+% specifications. Since the reachability problem for hybrid systems is
+% known to in general be undecidable, exhaustive exploration of the
+% state-space is intractable, and thus testing is an approach to attempt
+% to achieve system correctness. However, to guarantee safety and other
+% requirements, the testing procedures need to rely on a rigorous
+% theoretical foundation. In this special session aims to collect and
+% connect researchers, within academia as well as industry, working on
+% model-based testing of hybrid systems. The session calls for papers
+% proposing new theoretical research directions in the field, as well as
+% papers describing practical applications of model-based testing of
+% hybrid systems.
 
-In this work we propose a two step process for searching violations of
-the properties. This involves (a) learning a piece-wise affine model
-and (b) exhaustively searching it for violations. It must be noted
-that the violation found in the model might not be valid in the
-original system, depending on the accuracy of the model.
-% concretizing the violation in the original system.
+Model-based development is rapidly becoming the preferred paradigm for
+developing embedded control software in an industrial context.  Tools
+like \SIMULINK are often used to create models of closed-loop
+dynamical systems, \ie, a plant model of the physical systems that we
+wish to control, and a model for the embedded controllers.  Automatic
+approaches for testing the safety and performance of such closed-loop
+systems is challenging, and current approaches focus mostly on guided
+testing using numerical simulations
+\cite{annpureddy2011s,donze2010breach,deshmukh2015stochastic,dreossi2015efficient,akazaki}.
+On the other hand, theoretical research on hybrid and timed systems
+focusses on formal guarantees for restricted classes of dynamical
+systems such as hybrid automata where the system state evolves
+according to a vector field that is either constant, affine
+\cite{frehse2011spaceex}, or polynomial \cite{chen2015reachability}.
+In this paper, we propose an approach that leverages useful features
+from both kinds of techniques.
 
-The piece-wise affine model is the aggregation of affine maps, each
-approximating observed local behavior of the system.  The model is
-discrete time and `relational'. It relates the states reachable in a
-fixed time step $\Delta$ to a given state.
+We propose a two step process for finding violations of safety
+properties. In the first step, we learn a piece-wise affine (PWA) {\em
+relational} model from simulations of the dynamical system. In the
+second step, we leverage symbolic techniques to exhaustively analyze
+the model learned in the first phase. We remark that the approach we
+propose is neither sound nor complete: a violation found by our
+technique may not exist in the original system, and our technique may
+miss violations that exist in the original system. A fair question is:
+why is such a technique useful? There are two perspectives from which
+we can answer this question.
 
-In our previous work~\cite{zutshi2014multiple}, we analyzed systems by
-restricting ourselves to their black box semantics. This enabled us to
-reason about the state-space reachability of the system without a
-direct analysis of its structure. Using a coarse abstraction which we
-searched on-the-fly, we could observe the local dynamics as required.
-This gave us an efficient procedure to find abstract counter-examples.
-However, to concretize the counter-examples, a `closer look' or
-refinement of the abstraction is required. The grid based state-space
-abstraction was refined by splitting the relevant cells (abstract
-states) into smaller ones. As noted previously, due to the curse of
-dimensionality, such a uniform splitting is an expensive operation,
-and can not scale to higher dimensions.
+In many settings, we may not have a physics-based model derived from
+first principles, but may just have time-series data of the system
+behavior. In such a setting, a data-driven model allows us to
+generalize individual time-series behaviors through an abstract, PWA
+relational model. There are several approaches in the literature to
+perform data-driven modeling, system identification techniques that
+learn dynamic models from data \cite{ljung1999system}, auto-regressive
+models for forecasting \cite{wei1994time}, machine learning techniques
+that learn static or dynamic models from data
+\cite{narendra1990identification,lu2009linear}. However, in the second
+step of our technique, \ie, symbolic analysis of the learned mode, we
+need to have models that can be digested by existing symbolic tools,
+which is harder to do with some of the aforementioned heavier
+data-driven modeling methods. Hence, we use a lighter flavor of
+data-driven modeling such as the one proposed for learning relational
+models from data
+\cite{zutshi2012timed,sankaranarayanan2011relational}. By learning
+what is essentially a discrete-time dynamical system with PWA updates,
+our technique permits us to use software analysis approaches such as
+bounded model checking assisted by SMT and linear programming solvers.
 
-We further our exploration of trajectory segment based methods; and
-explore an alternative approach to overcome the explosion in abstract
-states. Instead of selectively refining the abstraction, we compute a
-model of the black box system and use bounded model checking to find a
-concrete counter-example in the model. Due to modeling errors, this
-might not be reproducible in the original black-box system. We then
-use this counter-example to guide the search towards a counter-example
-in the original system.
+The other perspective is when we have a model that is highly complex.
+We note that the ultimate goal for our technique is to find violations
+of safety properties in the {\em actual embedded control system}, and
+not, per se, in the model. The aphorism attributed to mathematician
+George E.P. Box is worth repeating,  ``all models are wrong, some
+models are useful.'' Thus, we can sidestep analyzing a highly complex
+model by approximating its behavior with a simpler relational model,
+which in turn gives us promising tests to run on the actual system,
+then our technique has intrinsic value.
 
-More specifically, we use regression to quantitatively estimate the
-discovered relations, which were witnessed by trajectory segments, by
-affine maps. These maps approximate locally observed behaviors, which
-are incorporated into the sampled reachability graph as edge
-annotations.  The resulting graph or the Piece-Wise Affine (PWA)
-relational model can be interpreted as an infinite state discrete
-transition system and model checked for time bounded safety
-properties. A counter-example if found, can indicate the presence of a
-violation in the system.
+% In our previous work~\cite{zutshi2014multiple}, we analyzed systems by
+% restricting ourselves to their black box semantics. This enabled us to
+% reason about the state-space reachability of the system without a
+% direct analysis of its structure. Using a coarse abstraction which we
+% searched on-the-fly, we could observe the local dynamics as required.
+% This gave us an efficient procedure to find abstract counter-examples.
+% However, to concretize the counter-examples, a `closer look' or
+% refinement of the abstraction is required. The grid based state-space
+% abstraction was refined by splitting the relevant cells (abstract
+% states) into smaller ones. As noted previously, due to the curse of
+% dimensionality, such a uniform splitting is an expensive operation,
+% and can not scale to higher dimensions.
+
+In \cite{zutshi2014multiple}, the authors investigate an on-the-fly
+abstraction technique that is close to the approach we propose. A key
+difference is that the abstraction explored in this work is an {\em
+existential} abstraction: the approximate model learned is a directed
+acyclic graph, where nodes represent regions of the state-space, and
+an edge between two regions indicates that there is a transition
+between some states in the regions. Our approach can be viewed as a
+further improvement of the existential abstraction by annotating an
+edge in the graph-based abstraction with an affine relation between
+the source and destination states connected by the edge.
+
+% We further our exploration of trajectory segment based methods; and
+% explore an alternative approach to overcome the explosion in abstract
+% states. Instead of selectively refining the abstraction, we compute a
+% model of the black box system and use bounded model checking to find a
+% concrete counter-example in the model. Due to modeling errors, this
+% might not be reproducible in the original black-box system.
+
+More specifically, we use linear regression to quantitatively estimate
+the discovered edges in the graph-based abstraction of the
+state-space. These linear maps approximate locally observed behaviors,
+and the resulting graph or the Piece-Wise Affine (PWA) relational
+model can be interpreted as an infinite state discrete transition
+system and model checked for time bounded safety properties. A
+counter-example if found, can indicate the presence of a violation in
+the system.  We then use this counter-example to guide the search
+towards a counter-example in the original system.
+
 %Furthermore, using linear programming, we can over-approximate the
@@
 
-% Towards the end of the chapter, we discuss extensions to data driven
+% In the conclusion, we discuss extensions to data driven
 % approaches (instead of simulator driven), where instead of a
@@
 % properties of SDCS.
+% Bounded time behvaiors are modeled
+% only one violation is being searched for
+% mention explicitly the relation between older cegar work
+
+The layout of the paper is as follows: In Section~\ref{sec:prelims},
+we introduce the basic steps to obtain a PWA relational model. In
+Section~\ref{sec:pwa-rel}, we explore the connection between the PWA
+relational model and previous graph-based abstraction techniques such
+as those in \cite{zutshi2014multiple}. We present extensions of
+relational modeling in Section~\ref{sec:rel-mod}, and present
+experimental validation of our technique on a few textbook examples of
+dynamical systems with hybrid and polynomial dynamics in
+Section~\ref{sec:res}.

### 2017-04-08 pwarel.tex
.

@@
 
-The abstraction of the hybrid dynamical system is defined by
-paritioning the combiend state-input space $\HybridStates \times
+The abstraction of the hybrid dynamical system is a a
+paritioning of the combined state-input space $\HybridStates \times
 \Inputs$ into hyper-rectangles or cells $C_i \in \C$. The cells are
@@
 their union $\bigcup_{C_i\in\C}C_i$ is the state-input
-space $\HybridStates \times \Inputs$. This discrete abstraction over
-the continuous space is implicit and defined using a quantization
+space $\HybridStates \times \Inputs$. The paritions over
+the continuous input-state space are implicitly defined using a quantization
 function $Q_\epsilon$ parameterized by the precision $\epsilon$.

### 2017-04-08 pwarel.tex
.

@@
 Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
-trajectory between the respective cells. The graph abstraction
-$G$ only states that there exists a state $\x \in C$ from which the
-system can evolve to a future state $\x' \in C'$. To increase the
-precision we iteratively refined the abstraction by state
-splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
-set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
-which non-deterministically describe relations between $\x \in C$ and $\x'
+trajectory between the respective cells. The graph abstraction $G$
+only states that there exists a state $\x \in C$ from which the system
+can evolve to a future state $\x' \in C'$. To increase the precision,
+iterative refinement of the abstraction by state splitting was
+proposed in~\cite{zutshi2014multiple}. Due to the curse of
+dimensionality, state splitting is not scalable. Instead, we propose
+an `enrichment' $G^R$ of $G$ by computing a set of local relations
+$R_{(C,C')}(\x,\x')$ for every edge $(C,C')$, which
+non-deterministically describe relations between $\x \in C$ and $\x'
 \in C'$. This is illustrated in \figref{enriched-edge}.

### 2017-04-08 pwarel.tex
.

@@
 set computation methods. Because we assume black box semantics, we
-only present the former. Using regression analysis on the $start$ and
-$end$ states of the witnessed trajectory between two cells,
+only present the former. Using regression analysis on a witnessed trajectory between two cells,
 we compute an approximate discrete map along with an error estimate.
@@
 can be separated into a training set and testing set to compute the
-map and the error respectively.  The latter case can be explored if
-the symbolic dynamics of the system are known. A tool like
-\flowstar~\cite{chen2013flow} can be used to find the reachable set
-map.
+map and the error respectively.
+
+%The latter case can be explored if
+%the symbolic dynamics of the system are known. A tool like
+%\flowstar~\cite{chen2013flow} can be used to find the reachable set
+%map.
 

### 2017-04-08 intro.tex
intro cosmetic

@@
 behavior. In such a setting, a data-driven model allows us to
-generalize individual time-series behaviors through an abstract, PWA
-relational model. There are several approaches in the literature to
-perform data-driven modeling, system identification techniques that
-learn dynamic models from data \cite{ljung1999system}, auto-regressive
-models for forecasting \cite{wei1994time}, machine learning techniques
-that learn static or dynamic models from data
-\cite{narendra1990identification,lu2009linear}. However, in the second
-step of our technique, \ie, symbolic analysis of the learned mode, we
-need to have models that can be digested by existing symbolic tools,
-which is harder to do with some of the aforementioned heavier
-data-driven modeling methods. Hence, we use a lighter flavor of
-data-driven modeling such as the one proposed for learning relational
-models from data
-\cite{zutshi2012timed,sankaranarayanan2011relational}. By learning
-what is essentially a discrete-time dynamical system with PWA updates,
-our technique permits us to use software analysis approaches such as
-bounded model checking assisted by SMT and linear programming solvers.
+generalize individual time-series behaviors. But, there are several
+approaches in the literature to perform data-driven modeling; e.g.,
+system identification techniques that learn dynamic models from data
+\cite{ljung1999system}, auto-regressive models for forecasting
+\cite{wei1994time}, machine learning techniques that learn static or
+dynamic models from data
+\cite{narendra1990identification,lu2009linear}.  Why then do we need a
+new modeling technique? The answer to this question lies in our goal
+for the second step, \ie, symbolic analysis of the learned model. For
+this step, we need to have models that can be digested by existing
+symbolic tools, which is harder to do with some of the aforementioned
+heavier data-driven modeling methods. Hence, we use a lighter flavor
+of data-driven modeling such as the one proposed for learning
+relational models from data
+\cite{zutshi2012timed,sankaranarayanan2011relational}. By learning a
+{\em PWA relational model}, \ie, essentially a PWA discrete-time
+dynamical system, we can use software analysis approaches such as
+bounded model checking assisted by SMT solvers and linear programming
+solvers.
 
-The other perspective is when we have a model that is highly complex.
-We note that the ultimate goal for our technique is to find violations
-of safety properties in the {\em actual embedded control system}, and
-not, per se, in the model. The aphorism attributed to mathematician
-George E.P. Box is worth repeating,  ``all models are wrong, some
-models are useful.'' Thus, we can sidestep analyzing a highly complex
-model by approximating its behavior with a simpler relational model,
-which in turn gives us promising tests to run on the actual system,
-then our technique has intrinsic value.
+The other perspective for our technique is the context when we have an
+effectively black-box model, \ie, a model that has features such as
+nonlinearities, delays or look-up tables that make symbolic analysis
+rather challenging.  We note that the ultimate goal for our technique
+is to find violations of safety properties in the {\em actual embedded
+control system}, and not, per se, in the model. The aphorism
+attributed to mathematician George E.P. Box is quite apt for our
+context, and bears worth repeating,  ``all models are wrong, some
+models are useful.'' We can sidestep analyzing a highly complex model
+by approximating its behavior with a simpler relational model, which
+in turn can give us promising tests to run on the actual system. This
+gives our technique intrinsic value as a debugging tool.
 

### 2017-04-08 intro.tex
.

--- 2017-04-08 intro.tex
@@
 abstraction technique that is close to the approach we propose. A key
-difference is that the abstraction explored in this work is an {\em
-existential} abstraction: the approximate model learned is a directed
-acyclic graph, where nodes represent regions of the state-space, and
-an edge between two regions indicates that there is a transition
-between some states in the regions. Our approach can be viewed as a
-further improvement of the existential abstraction by annotating an
-edge in the graph-based abstraction with an affine relation between
-the source and destination states connected by the edge.
+difference is that the abstraction explored in this work is an
+existential {\em relational} abstraction. The approximate model
+learned is a directed acyclic graph, where nodes represent regions of
+the state-space, and an edge between two regions indicates a
+transition between states in the regions. Our approach further
+improves the existential abstraction by annotating the edges in the
+graph-based abstraction with affine relations between the source and
+destination states.
 

### 2017-04-08 implementation.tex
.

@@
 \midrule
-    Van der Pol ($\prop_3$)   &$130$ & $0$ & $24$\\
-    Brusselator               &$4$ & $0$ & $2$\\
-    Lorenz                    &$122$ & $0$ & $35$\\
+    Van der Pol ($\prop_3$)   &$14$ & $20$ & $24$\\
+    Brusselator               &$4$ & $4$ & $2$\\
+    Lorenz                    &$122$ & $120$ & $35$\\
 %B.Ball                    &  &   & \\
-    Nav ($\prop_P$)           &$TO$ & $2100$  & $603$ \\%9423 + 4480
-    Nav ($\prop_Q$)           &$TO$ & $25m$   & $546$ \\%8105 + 970
-    Nav ($\prop_R$)           &$TO$ & $0$  & $2003$\\
+    Nav ($\prop_P$)           &$TO$ & $1710$  & $603$ \\%9423 + 4480
+    Nav ($\prop_Q$)           &$TO$ & $1560$   & $546$ \\%8105 + 970
+    Nav ($\prop_R$)           &$TO$ & $2352$  & $2003$\\
     Nav ($\prop_S$)           &$TO$ & $0$  & $2100$\\

### 2017-04-08 implementation.tex
.

@@
     Nav ($\prop_S$)           &$TO$ & $0$  & $2100$\\
-    Bouncing Ball             &$TO$ & $0$  & $450$\\
+    Bouncing Ball             &$Safe$ & $Safe$  & $450$\\
 

### 2017-04-08 prelims.tex

@@
 In this section, we present the model of the hybrid dynamical system
-under test, and the background for pwa relational abstractions.
+under test, and the background for PWA relational abstractions.
 
-\mypara{The System under test $\System$} is assumed to be a hybrid
-dynamical system. Its state is denoted by $\HybridStates \subseteq
-\Modes \times \Reals^n$, where $\Modes$ is a finite set of discrete
-modes. Let $\Inputs: [0,T] \rightarrow \reals^k$ be the set of input
-signals to $\System$ over a finite time horizon $T$.
+\mypara{System under test ($\System$)}
 
-For the purpose of falsification, $\System$ is viewed as a black-box
-equipped with a forward simulation map $\simulate^{\System}:
-(\vx, \vu, t) \mapsto \vx$.  A simulation step
-involves computing the unique reachable state $\vx \in \HybridStates$ in time
-$t$ from a given state input pair $(\vx,\vu)$.  We formalize this
-using the notion of time parameterized reachability relations
-$\reach{t, \vu}_\System \subseteq \HybridStates \times \HybridStates$
-detailed in~\cite{zutshi2014multiple}.
+We assume that $\System$ is a hybrid dynamical system with exogenous
+inputs modeling an embedded control system. The system state-space
+(denoted $\HybridStates$) is a subset of $\Modes \times \Reals^n$,
+where $\Modes$ is a finite set of discrete modes, and $\Reals^n$
+represents the continuous-valued state-space. For ease of exposition,
+we assume that the hybrid system is effectively a switched-mode
+dynamical system, \ie, the mode of the system is completely determined
+by its state.  An input to $\System$ is assumed to be a time-bounded
+{\em signal}, \ie, a function from $[0,T]$ to $\Inputs$, with $T$ as
+the time horizon, and $\Inputs \subseteq \Reals^k$.  In this paper, we
+view $\System$ as effectively black-box equipped with a forward
+simulation map, in a simulation time-step of $\Delta_t$ from time $t$,
+allows computing the unique reachable state $\vx \in \HybridStates$
+given state-input pair $(\vx,\vu)$.
+
+% time and input parameterized reachability relation $\reach{t,
+% \vu}_\System \subseteq \HybridStates \times \HybridStates$ as follows:
+% $\simulate^{\System}:
+% (\vx, \vu, t) \mapsto \vx$, where $\vx \in \HybridStates$, $\vu \in
+% \Inputs$ and $t \in [0,T]$.
+
 % Given the hybrid state space $\vx\in\HybridStates$ and input space $u
@@
 \begin{definition}[System Evolution]
-    The system under test $\System$ evolves in the set of infinite
-    hybrid state-space $\HybridStates$ under the effect of inputs $\vu
-    \in \Inputs$. Its evolution is defined by the simulation function
-    $\simulate^\System : \HybridStates \times \Inputs \times \Reals^+ \mapsto
-    \HybridStates$, where $\simulate^\System(\x,\vu,\Delta_t)$ maps a
-    current state $\x$ at time $t$ to a new state $\x'$ at time
-    $t+\Delta_t$, under a constant input $\vu$.
-    Finally, $\vx'=\simulate^{\System}(\vx,u,t)$ \emph{iff} $\vx
-    \reach{t,u} \vx' \in \reach{t, \vu}_\System$.
+    The behavior of system $\System$ at time $t$ is described as an
+    evolution in the hybrid state-space $\HybridStates$ under the
+    effect of the input $\vu \in \Inputs$ at time $t$. The forward
+    simulation map $\simulate^\System : \HybridStates \times \Inputs
+    \times \Reals^+ \mapsto \HybridStates$, maps a state $\x$ at time
+    $t$ to a new state $\x'$ at time $t+\Delta_t$, under a constant
+    input $\vu$.  Finally, we define the time and input parameterized
+    reachability relation $\reach{t,u} \subseteq \HybridStates \times
+    \HybridStates$ such that $(\vx,\vx') \in \reach{t,u}$ iff
+    $\vx'=\simulate^{\System}(\vx,u,t)$.
 \end{definition}
@@
 horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
-the vector field in each underlying hybrid mode and ruling away issues
-such as finite escape times~\cite{Meiss/2007/Differential}.
-Furthermore, as the relations abstract the solution of a continuous
-time, foward deterministic hybrid dynamical system, they are
-deterministic, causal and satisfy the semi-group property.  These
-assumptions are justified as for the dynamical systems we address in
-this presentation. Finally, we assume full observability of the system
-state, validate our results only against the $\simulate$ function (as
-the analytic closed-form representation of the system dynamics is
-assumed to be unavailable).  From here on, we drop the subscript
-$\System$.
+the vector field in each underlying hybrid mode, ruling away issues
+such as finite escape times~\cite{Meiss/2007/Differential}, and having
+well-defined semantics for simulations at switching boundaries. While
+these assumptions hold for the dynamical system examples which we
+consider in this paper, such assumptions can be restrictive for
+general control system models. When analyzing a system that does not
+obey theoretical conditions that allow deterministic simulation, we
+assume that the underlying tool used for numerical approximations of
+system behaviors uses deterministic simulation semantics.  Finally, we
+assume full observability of the system state, validate our results
+only against the $\simulate$ function (as the analytic closed-form
+representation of the system dynamics is assumed to be unavailable).
+From here on, we drop the subscript $\System$ whenever it is clear
+from the context.
 
@@
 
-Best effor falsification works by enumerating the reachability
-relations $\reach{}$ within a given budget. Depending upon the
-specific instance of the problem, the budget can be prohibhitive,
-requiring exponential computational and storage resources. Instead, we
-proceed by summarizing these relations by their corresponding affine
-maps using linear regression. The learnt analytical forms are
-represented as $\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are
-matrices and $\delta$ the estimated error.
+The approach in \cite{zutshi2014multiple} uses the forward simulation
+map to explore a fixed number of tuples in the reachability relation
+$\reach{t,\vu}$ specified by a user-provided budget.  In general, the
+budget required to adequately explore the transition space requires
+exploring an exponential (in the dimension of $\HybridStates$) number
+of such tuples. This can cause undue burden in terms of computation
+time as well as storage.  The central idea in this paper is to {\em
+summarize} several tuples as a single affine map obtained using linear
+regression.  In other words, we can replace several ``proximal''
+tuples $(\vx,\vx')$ in $\reach{t,u}$ by an affine map of the form
+$\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are matrices and
+$\delta$ is a worst-case error estimate.
 
@@
 relationship between the given set of observed input $\x$ and output
-$\y$ vectors. The notion of `best' is formally captured using a loss
+$\x'$ vectors. The notion of `best' is formally captured using a loss
 function which can be non-linear. We use the commonly used loss
@@
 \mypara{Ordinary Least Squares (OLS)}
+For rest of the section, we elide the discrete mode from the hybrid
+state of the system, and only focus on the continuous state. Note that
+we can do this because of our assumption that the system is a
+switched-mode dynamical system.  Suppose we have $N$ tuples of the
+form $(\x,\x')$ in the reachability relation that we have explored,
+where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn an
+affine function described as 
+\begin{equation}
+\label{eq:affinemap}
+\x' = A\x + B
+\end{equation}
+Here, $A$ is a $n\times n$ matrix, and $B$ is a $n\times 1$ matrix.
 
-Let the data set be comprised of $N$ input and output pairs $(\x, y)$,
-where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is the case in
-the current context of \textit{finding the best fit}, the problem is
-over-determined: there are more equations than unknowns. Hence, a
-single affine function cannot be found which satisfies the equation
-$\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we need
-to find the `best' choice for $\vec{a}$ and $b$. This choice is formally
-defined using a loss function. For the case of simple linear
+If $N>n$, which is the case in the current context of \textit{finding
+the best fit}, the problem is over-determined: there are more
+equations than unknowns. Hence, a single affine function cannot be
+found which satisfies the equation \eqref{eq:affinemap}. Instead, we
+need to find the `best' choice for $\vec{a}$ and $b$. This choice is
+formally defined using a loss function. For the case of simple linear
 regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine the vector of
-coefficients $\vec{a}$ and an offset constant $b$, such that the least
-square error of the affine predictor is minimized for the given data
-set.
+errors in prediction.  The task is then to determine $A$ and $B$, such
+that the least square error of the affine predictor is minimized for
+the given data set.
+
 \begin{equation}
-    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
+    \operatornamewithlimits{argmin}_{A,B}\displaystyle\sum_{i=1}^{N}{\left(\x' - (A\x + B)\right)^2}
 \label{reg}
 \end{equation}
+
 To ease the presentation, we can rewrite the above as a homogeneous
-expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
-$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
+expression by using $\hat{\x} = [\x^T\ 1]^T$ and replacing $A$ and $B$ by the vector
+$\vec{A} = [A\ B]$. The equation~\ref{reg} now becomes
 \[
-    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
+\operatornamewithlimits{argmin}_{\vec{A}}\displaystyle\sum_{i=1}^{N}{(\x' - \vec{A}^T \hat{\x})^2}
 \]
-The solution of OLS can be analytically computed as
-\[ Ab = (X^TX)^{-1}X^T\y\]
-where $X$ is the matrix representing the horizontal stacking of all
-$\hat{\x}$.
-The details can be found in standard texts on learning and statistics
-\cite{friedman2001elements}.
 
-Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
-where $\x\in\reals^n$ we can use OLS to approximate its trajectories
-at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
-$A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
-\begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
-    global relational model for the system. The data set for the
-    regression is a set of pairs $\setof{(\x,\x') | \x' =
-    \simulate(\x, \Delta)}$ and can be generated on demand. Another
-    set can be generated to compute the error $\delta$ as an interval
-    of vectors, where each element is an interval $\delta_i \in
-    [\delta_{min}, \delta_{max}]$.
+The solution of OLS can be analytically computed as the Moore-Penrose
+pseudo-inverse: $\vec{A} = (X^TX)^{-1}X^T X'$, where $X$ is the matrix
+representing the horizontal stacking of all $\hat{\x}$ and $X'$
+represents the stacking of all $\x'$ values.  The details can be found
+in standard texts on learning and statistics
+\cite{friedman2001elements}. Alternative methods based on gradient
+descent can also be used to find a suitable $A$ matrix.
 
-    Affine maps are poor
-    approximations for arbitrary non-linear functions. Hence, we use a
-    collection of affine maps to approximate the local behaviors (in
-    state-space) of the system $\simulate$. This results in a piece
-    wise approximation, as described in the next section.
+In this fashion, we can we can use OLS to approximate its trajectories
+of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
+= \vec{A}\hat{\x}$.
+
+% where
+% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T
+% \end{bmatrix}$ and $\vb = \begin{bmatrix}b_1 \\ \vdots \\ b_n
+% \end{bmatrix}$.
+
+We note that linear regression tools are often able to provide an
+$n\times 1$ vector $\vec{\delta}$ of error intervals, where the
+$i^{th}$ entry of the form $[{\delta_i}_{\min},{\delta_i}_{\max}]$
+indicates the best and worst possible error in computation of the
+$i^{th}$ state in $\x'$.
+
+Affine maps are poor approximations for arbitrary non-linear
+functions. Hence, we use a collection of affine maps to approximate
+the local behaviors (in state-space) of the system $\simulate$. This
+results in a piece wise approximation, as described in the next
+section.
 

### 2017-04-08 prelims.tex

@@
 functions. Hence, we use a collection of affine maps to approximate
-the local behaviors (in state-space) of the system $\simulate$. This
+the local behaviors (in state-space) of the system $\System$. This
 results in a piece wise approximation, as described in the next
@@
 We define the PWA transition system as a transition system
-%(ref.~\secref{DiscSys})
-$\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
-where each transition $\tau \in \scrT$, is associated with a
-transition relation of the form
+$\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
+a nonempty set of locations, $\pVars$ is a set of variables, $\scrT$
+is a finite subset of $\pLocs \times \pLocs$, and each transition
+$\tau \in \scrT$ is associated with a transition relation of the form
 \[
@@
 $\pVar,\pVar'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
-post conditions) on the states and $\Theta$ is an affine assertion on
-the initial states.
+post conditions) on the states. Finally, $\pLocI$ is an initial
+location $\Theta$ is an affine assertion characterizing the initial
+states included in $\pLocI$.
 
@@
 
-We now formalize the PWA affine transition system for a dynamical system.
-\begin{definition}[PWA Transition System]
+We now formalize the PWA relational model for a dynamical system.
+\begin{definition}[PWA Relational Model]
 
-Given a hybrid dynamical system over a state-space
-$\HybridStates^n$, a PWA transition system $\rho$ is given by the
-tuple $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in
-\scrT$ are affine transition relations and $\Theta$ is an affine
-predicate over $\x$ and the states $\x\in\HybridStates^n$. The
-transition relation is then defined by $\scrT$ with $n$ transition
-relations as follows
+Given a hybrid dynamical system over a state-space $\reals^n$, a PWA
+relational model $\rho$ is given by the tuple
+$\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in \scrT$ is a
+collection of affine transitions  and $\Theta$ is an affine predicate
+over $\x\in\real^n$. The transition relation is then defined by
+$\scrT$ with $n$ transition relations as follows
 
@@
 \end{equation}
-
 \end{definition}
 
-%where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
-%$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
 where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
@@
 A PWA model is \textit{deterministic}, iff for every state
-$\x\in\HybridStates^n$, a unique guarded affine map is satisfied and
-all errors $\delta$ are singletons. However, in practice such a case
-rarely exists. A PWA model will be usually non-deterministic, both due
-to multiple choice of transition relations and the reachable states at
-the $k^{th}$ time step (or after $k\Delta$ units of time) being a set.
-Abusing the notation, we denote the set of states reachable in a
-single step in $\rho$ by $\x' = \rho(\x)$.
+$\x\in\reals^n$, a unique guarded affine map is satisfied and if all
+entries in the error interval vector $\vec{\delta}$ are singletons.
+However, in practice such a case rarely exists. A PWA model will be
+usually non-deterministic, both due to multiple choices of transitions
+from any location and the reachable states at the $k^{th}$ time step
+(or after $k\Delta$ units of time) being a set.  Abusing the notation,
+we denote the set of states reachable in a single step in $\rho$ by
+$\x' = \rho(\x)$.
 

### 2017-04-08 bmc.tex
.

@@
-% include an algorithm
-We now describe the entire procedure in three steps.
-\begin{enumerate}
-    \item Given a black box system $\System$ specified by
-        $\simulate^\System$, we
-        use scatter-and-simulate with ($\Delta$-length) trajectory
-        segments to sample a finite reachability graph
-        $\scr{H}(\Delta)$.
-    \item Using regression, we enrich the graph relations and annotate
-        the edges with the estimated affine maps.
-    \item Interpreting the directed reachability graph as a transition
-        system, we use a bounded model checker to find fixed length
-        violations to safety properties.
-\end{enumerate}
+\section{Model Checking}
+
+As a final step, the enriched graph is model checked for the given
+safety property. If a counter-example is found, it is simulated using
+the $\simulate$ function and if reproducible, a violation is
+concluded. In all other cases the falsification results in an
+`unknown' result.
+
+% % include an algorithm
+% We now describe the entire procedure in three steps.
+% \begin{enumerate}
+%     \item Given a black box system $\System$ specified by
+%         $\simulate^\System$, we
+%         use scatter-and-simulate with ($\Delta$-length) trajectory
+%         segments to sample a finite reachability graph
+%         $\scr{H}(\Delta)$.
+%     \item Using regression, we enrich the graph relations and annotate
+%         the edges with the estimated affine maps.
+%     \item Interpreting the directed reachability graph as a transition
+%         system, we use a bounded model checker to find fixed length
+%         violations to safety properties.
+% \end{enumerate}
 
@@
 
-We now describe the above using the Van der Pol oscillator benchmark
-presented in \cite{zutshi2014multiple}. Simulations generated using
-uniformly random samples are shown in \figref{vdp-cont}. We want to
-check the property $P_3$, indicated by the red box, given the initial
-set indicated by the green box. The abstraction is defined by
-$\quant_{0.2}(\x)$, which results in an evenly gridded state-space,
-where each cell is of size $0.2 \times 0.2$ units.
-Scatter-and-simulate is then used to construct the abstract graph $G$,
-using $2$ samples per cell and the time step $\Delta = 0.1s$. The
-complete process follows.
+We now describe the pesented technique using the Van der Pol
+oscillator benchmark from~\cite{zutshi2014multiple}. Simulations
+generated using uniformly random samples are shown in
+\figref{vdp-cont}. We want to check the property $P_3$, indicated by
+the red box, given the initial set indicated by the green box. The
+abstraction is defined by $\quant_{0.2}(\x)$, which results in an
+evenly gridded state-space, where each cell is of size $0.2 \times
+0.2$ units.  Scatter-and-simulate is then used to construct the
+abstract graph $G$, using $2$ samples per cell and the time step
+$\Delta = 0.1s$. The complete process follows.
 
 \begin{enumerate}
-    \item{\emph{Abstract}} Consider an implicit abstraction induced by the quantization
-    function $\quant_\epsilon(\x)$. The corresponding reachability
-    graph $\scr{H}(\Delta)$ (with time step $\Delta$) is given by
-    $\tupleof{\scr{C},\areach{\Delta}, \scr{C}_0, \scr{C}_u}$.
+    \item{\emph{Abstract}}: Assume an implicit abstraction using the quantization
+        function $\quant_{\epsilon=0.2}(\x)$. The corresponding reachability
+    graph is $\scr{H}(\Delta=0.1)$.
 \item{\emph{Discover}}: Using scatter-and-simulate, enumerate a finite number
     of cells (vertices) and the relations (edges) of the graph
-    $\scr{H}(\Delta)$. Associate the set of generated trajectory
-    segments with their respective originating cells using a map $D:C
-    \mapsto \setof{\pi | start(\pi) \in C}$. The discovered
-    abstraction is show in \figref{vdp-abs}. As mentioned, red cells
+    $\scr{H}(0.1)$. Associate the set of generated trajectory
+    segments with their respective originating cells using a map $\ds:C
+        \mapsto \setof{(\vx,\vx') |\vx \in C}$. The discovered
+    abstraction is show in \figref{vdp-abs}. The red cells
     are unsafe and green cells are initial cells.
-\item{\emph{Relationalize}} For each cell $C$, perform regression analysis on
+\item{\emph{Relationalize}}: For each cell $C$, perform regression analysis on
     the respective set of trajectory segments $D(C)$, and compute a
     set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
-    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(\Delta)$. Annotate
-    each edge in the graph $\scr{H}(\Delta)$ with the respective
+    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(0.1)$. Annotate
+    each edge in the graph $\scr{H}(0.1)$ with the respective
     relation.
@@
 
-\item{\emph{Model Check}} The graph $\scr{H}(\Delta)$ can now be viewed as a
+\item{\emph{Model Check}}: The graph $\scr{H}(0.1)$ can now be viewed as a
     transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
-    where $C \in vertices(\scr{H}(\Delta))$ and $\x \in
+    where $C \in vertices(\scr{H}(0.1))$ and $\x \in
     \HybridStates$. A transition $\tau \in \scrT$ is of the form
@@
     \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
-    edge(\scr{H}(\Delta))}$.
-\item{\emph{Check Counter-example}} The infinite state (but finite location) transition system can
+    edge(\scr{H}(0.1))}$.
+\item{\emph{Check Counter-example}}: The infinite state (but finite location) transition system can
     be model checked to find a concrete counter-example, which if
@@
 
-\inclfig[width=0.75\linewidth]{vdp_cont_traces.png}{Van der
+\inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
 Pol: continuous trajectories. Red and green boxes indicate unsafe and
@@
 
-\inclfig[width=0.75\linewidth]{vdp_abs.png}{The discovered
+\inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
 abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
@@
 
-\inclfig[width=\linewidth]{vdp_disc_map.png}{Cells and
+\inclfig[width=0.95\linewidth]{vdp_disc_map.pdf}{Cells and
 trajectory segments used by $1$-relational modeling.}
@@
 
-\inclfig[width=0.40\linewidth]{vdp_graph.png}{Enriched
+\inclfig[width=0.50\linewidth]{vdp_graph.pdf}{Enriched
 graph $G^R$. The affine maps $f$ for the transition relations are show

### 2017-04-08 prelims.tex
.

@@
 
+% A configuration of the PWATS is
+% a pair (l,\nu(v)), where \nu is a valuation function assigning values to
+% variables.
+
+% A run of the PWA TS is as follows: the PWA TS starts in the initial
+% location from a valuation \nu(v) \in theta.
+
+% A move of the PWA TS is described by the function mu: from (L,R^n) to
+% (L,R^n), and mu follows the rules that
+% mu(l,\nu(v)) = (l',\nu'(v)), iff (l,g,g',f,l') in T, and
+% \nu(v) satisfies g, \nu(v') satisfies g', and \nu(v') = f(\nu(v)).
+
 \subsection{Piecewise Affine (PWA) Transition System}
 
-We define the PWA transition system as a transition system
-$\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
-a nonempty set of locations, $\pVars$ is a set of variables, $\scrT$
-is a finite subset of $\pLocs \times \pLocs$, and each transition
-$\tau \in \scrT$ is associated with a transition relation of the form
+We define the PWA transition system as a discrete transition system
+$\rho:\tupleof{\pLocs,\vx,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
+a nonempty set of locations, $\vx$ is a set of variables, $\scrT$
+is a finite set of transitions of the form $\tupleof{l, \tau,
+l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
+post-location and $\tau(\vx,\vx') \in \scrT$ is the transition. $\tau$
+is labled by the transition relation $\rho_{\tau}$
+over the current states and the next states. The transition relation
+is defined in terms of affine predicates on the current and next
+states and an affine update between them.
 \[
-    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,\pVar')) \;|\;
-    g_\tau(\pVar) \land g'_\tau(\pVar') \land f_{\tau}(\pVar, \pVar')}
+    \rho_{\tau}(\vx,\vx') \subseteq \setof{(\vx,\vx')) \;|\;
+    g_\tau(\vx) \land g'_\tau(\vx') \land f_{\tau}(\vx, \vx')}
 \]
 where $f_{\tau}$ is an affine relation on the pre and post states
-$\pVar,\pVar'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
+$\vx,\vx'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
 post conditions) on the states. Finally, $\pLocI$ is an initial

### 2017-04-08 prelims.tex
.

@@
 
-% A configuration of the PWATS is
-% a pair (l,\nu(v)), where \nu is a valuation function assigning values to
-% variables.
-
-% A run of the PWA TS is as follows: the PWA TS starts in the initial
-% location from a valuation \nu(v) \in theta.
-
-% A move of the PWA TS is described by the function mu: from (L,R^n) to
-% (L,R^n), and mu follows the rules that
-% mu(l,\nu(v)) = (l',\nu'(v)), iff (l,g,g',f,l') in T, and
-% \nu(v) satisfies g, \nu(v') satisfies g', and \nu(v') = f(\nu(v)).
-
 \subsection{Piecewise Affine (PWA) Transition System}
@@
 l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
-post-location and $\tau(\vx,\vx') \in \scrT$ is the transition. $\tau$
+post-location and $\tau \in \scrT$ is the transition. $\tau$
 is labled by the transition relation $\rho_{\tau}$
@@
 states included in $\pLocI$.
+
+% A configuration of the PWATS is
+% a pair (l,\nu(v)), where \nu is a valuation function assigning values to
+% variables.
+
+% A run of the PWA TS is as follows: the PWA TS starts in the initial
+% location from a valuation \nu(v) \in theta.
+
+% A move of the PWA TS is described by the function mu: from (L,R^n) to
+% (L,R^n), and mu follows the rules that
+% mu(l,\nu(v)) = (l',\nu'(v)), iff (l,g,g',f,l') in T, and
+% \nu(v) satisfies g, \nu(v') satisfies g', and \nu(v') = f(\nu(v)).
 

### 2017-04-08 bmc.tex
.

@@
 
-\inclfig[width=0.95\linewidth]{vdp_disc_map.pdf}{Cells and
+\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
 trajectory segments used by $1$-relational modeling.}
@@
 
-\begin{table}[!htbp]
+\begin{table}[h]
 \centering
@@
 
-The search parameters for S3CAM-R include the parameters of S3CAM:
+The search parameters for PWA-Rel include the parameters of S3CAM:
 $N$, $\epsilon$ and $\Delta$. Additionally, they also include the
@@
 
-We have already discussed the effects of $N$, $\epsilon$ and $\Delta$
-on S3CAM's performance. However, they also have an effect on
+The effects of $N$, $\epsilon$ and $\Delta$
+on S3CAM's performance were detailed in~\cite{zutshi2014multiple}. However, they also have an effect on
 relational modeling. A finer grid with small cells produces more

### 2017-04-08 implementation.tex
.

@@
-The implementation was prototyped as S3CAM-R, an extension to our
-previously mentioned tool S3CAM \cite{zutshi2014multiple}. OLS regression
+The implementation was prototyped as PWA-Rel,
+%an extension to our
+%previously mentioned tool S3CAM \cite{zutshi2014multiple}.
+OLS regression
 routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
@@
 We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-few of the benchmarks described in \cite{zutshi2014multiple}, including the Van
-der Pol oscillator, Brusselator, Lorenz attractor and the Navigation
-benchmark. As before, we ran S3CAM-R $10$ times with different seeds
-and averaged the results. We tabulate both the total time taken and
-the time taken by SAL to compute the counter-example and compare
-against a newer implementation of S3CAM. Note that different
-abstraction parameters are used for S3CAM and S3CAM-R, due to the
-differences in which they operate.
+few of the benchmarks described in \cite{zutshi2014multiple},
+including the Van der Pol oscillator, Brusselator, Lorenz attractor
+and the Navigation benchmark. As before, we ran PWA-Rel $10$ times
+with different seeds and averaged the results. We tabulate both the
+total time taken and the time taken by SAL to compute the
+counter-example and compare against an implementation of S3CAM. Note
+that different abstraction parameters are used for S3CAM and
+PWA-Rel, due to the differences in which they operate.
 
@@
 time taken by the BMC engine. The total time in seconds (rounded off
-to an integer) is noted under \textbf{S3CAM-R} and \textbf{S3CAM}.
+to an integer) is noted under \textbf{PWA-Rel} and \textbf{S3CAM}.
 $TO$ signifies time $>5hr$, after which the search was killed.}
@@
 
-The results show promise, but clearly S3CAM performs better. Also,
-S3CAM-R timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
+The results show promise, but the performance is comparable. Also,
+PWA-Rel timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
 and the bouncing ball. However, we need to explore more benchmarks to
-be conclusive. S3CAM-R's performance can be explained by the
+be conclusive. PWA-Rel's performance can be explained by the
 difficulty in finding a good abstraction ($\quant_\epsilon$), which

### 2017-04-08 bmc.tex
.

@@
 
-    \figref{vdp-map} shows the cells and the trajectory segments which
+    \figref{vdp-map} shows the cells and the trajectory segments, which
     are part of the data sets constructed using the $1$-relational
@@
 Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}
-\label{fig:vdp-cont}
+initial sets.}{vdp-cont}
 
@@
 abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
-initial cells.}
-\label{fig:vdp-abs}
+initial cells.}{vdp-abs}
 
 \inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
-trajectory segments used by $1$-relational modeling.}
-\label{fig:vdp-map}
+trajectory segments used by $1$-relational modeling.}{vdp-map}
 
@@
 graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}
-\label{fig:vdp-graph}
+in \tabref{vdp-pwa}.}{vdp-graph}
 

### 2017-04-08 implementation.tex
.

@@
-The implementation was prototyped as PWA-Rel,
-%an extension to our
-%previously mentioned tool S3CAM \cite{zutshi2014multiple}.
-OLS regression
+The presented technique was prototyped as a tool: PWA-Rel.
+We compared it against S3CAM~\cite{zutshi2014multiple} on several academic
+benchmarks.
+
+PWA-Rel was implemented in Python. Its different parts comprised of
+the (a) abstraction graph generation, (b) regeression analysis and (c)
+BMC. S3CAM was used to compute (a). To compute (b), OLS regression
 routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
-Python module for machine learning. SAL~\cite{SAL-SRI}
-with Yices2~\cite{dutertre2014yices} was the model checker and the SMT
-solver.
+Python module for machine learning. Finally, the The bounded model
+checking (c) was done using the SAL model checker~\cite{SAL-SRI},
+which used Yices2~\cite{dutertre2014yices} as the SMT solver. A naive
+alternative linear programming (LP) based model checker specific to the PWA
+relational abstractions was implemented utilizing PyGLPK.
 
 We used S3CAM to discover the abstraction graph $G$, which was then
-trimmed of the nodes which did not contribute to the error search
-(nodes from which error nodes were not reachable). In our experiments we
-used $1$-relational modeling to create the transition system from $G$.
-Using SAL, we checked for the given safety property. If a concrete
-trace was obtained from the BMC, it was further checked for validity
-by simulating using $\simulate$ to see if it was indeed an error
-trace.  If not, $100$ samples from its associated abstract state was
-sampled to check the presence of a counter-example in it
-neighborhood. This can be further extended by obtaining different
-discrete sequences or discrete trace sequences from the model checker.
+filtered to remove the nodes from which error nodes were not
+reachable. In our experiments we used $1$-relational modeling to
+create the transition system from $G$. Using the BMC, we checked for
+the given safety property. If a concrete trace was obtained, it was
+further checked for reproducibility by sampling and simulating its `neighborhood' with the
+$\simulate$ function. We define `neighborhood' as the abstract states
+corresponding to the returned concrete trace. Specifically, $100$
+samples were uniformly randomly chosen and simulated and tested for
+error traces. This procedure can also be extended to obtaining
+different discrete sequences or discrete trace sequences from the
+model checker by requesting distinct counter-examples.
+
+% abstract state was sampled to check the presence of a counter-example
+% in it neighborhood. This can be further extended by obtaining
+% different discrete sequences or discrete trace sequences from the
+% model checker.
 

### 2017-04-08 implementation.tex
.

@@
 We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-few of the benchmarks described in \cite{zutshi2014multiple},
+the more difficult benchmarks described in \cite{zutshi2014multiple},
 including the Van der Pol oscillator, Brusselator, Lorenz attractor
-and the Navigation benchmark. As before, we ran PWA-Rel $10$ times
-with different seeds and averaged the results. We tabulate both the
-total time taken and the time taken by SAL to compute the
-counter-example and compare against an implementation of S3CAM. Note
-that different abstraction parameters are used for S3CAM and
-PWA-Rel, due to the differences in which they operate.
+and the Navigation benchmark. PWA-Rel and S3CAM were ran $10$ times
+with different seeds and the running times were averaged. We tabulate
+both the total time taken by PWA-Rel+SAL and PWA-Rel+LP to compute the
+counter-example and compare against an implementation of S3CAM.
+Different abstraction parameters are used for S3CAM and PWA-Rel, due
+to the differences in which they operate. PWA-Rel requires a finer
+abstraction to begni with in order to minimize the modeling errors.
+On the other hand, a coarser abstraction as the intital abstraction is
+works better for S3CAM due to the expensive refinement procedure
+involved.
 
@@
 \toprule
-    Benchmark & PWA Rel + BMC & PWA Rel + LP & S3CAM\\
+    Benchmark & PWA Rel + SAL & PWA Rel + LP & S3CAM\\
 \midrule
-    Van der Pol ($\prop_3$)   &$14$ & $20$ & $24$\\
-    Brusselator               &$4$ & $4$ & $2$\\
-    Lorenz                    &$122$ & $120$ & $35$\\
+    Van der Pol ($\prop_3$)   &$0.2$ & $0.34$ & $0.4$\\
+    Brusselator               &$0.06$ & $0.06$ & $0.03$\\
+    Lorenz                    &$0.4$ & $0.33$ & $0.58$\\
 %B.Ball                    &  &   & \\
-    Nav ($\prop_P$)           &$TO$ & $1710$  & $603$ \\%9423 + 4480
-    Nav ($\prop_Q$)           &$TO$ & $1560$   & $546$ \\%8105 + 970
-    Nav ($\prop_R$)           &$TO$ & $2352$  & $2003$\\
-    Nav ($\prop_S$)           &$TO$ & $0$  & $2100$\\
-    Bouncing Ball             &$Safe$ & $Safe$  & $450$\\
+    Nav ($\prop_P$)           &$TO$ & $28.5$  & $10$ \\%9423 + 4480
+    Nav ($\prop_Q$)           &$TO$ & $26$   & $9.1$ \\%8105 + 970
+    Nav ($\prop_R$)           &$TO$ & $39.2$  & $33.3$\\
+    Nav ($\prop_S$)           &$TO$ & $71$  & $35$\\
+    Bouncing Ball             &$Safe$ & $Safe$  & $7.5$\\
 
@@
 
-The results show promise, but the performance is comparable. Also,
-PWA-Rel timed out on some benchmarks like Nav $\prop_R$, Nav $\prop_S$
-and the bouncing ball. However, we need to explore more benchmarks to
+As apparent form the results, the performance is comparable in all
+benchmarks, but the last one: bouncing ball. PWA-Rel determines it to
+be `safe' as it can not find a counter-example. Also, PWA-Rel performs
+better with the LP solver as PWA-Rel+SAL times out on most of the
+benchmarks. However, we need to explore more benchmarks to
 be conclusive. PWA-Rel's performance can be explained by the

### 2017-04-08 pwarel.tex

@@
 paritioning of the combined state-input space $\HybridStates \times
-\Inputs$ into hyper-rectangles or cells $C_i \in \C$. The cells are
+\Inputs$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
 pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
-their union $\bigcup_{C_i\in\C}C_i$ is the state-input
+their union $\bigcup_{C_i\in\Cells}C_i$ is the state-input
 space $\HybridStates \times \Inputs$. The paritions over

### 2017-04-08 implementation.tex
.

@@
 benchmarks, but the last one: bouncing ball. PWA-Rel determines it to
-be `safe' as it can not find a counter-example. Also, PWA-Rel performs
-better with the LP solver as PWA-Rel+SAL times out on most of the
-benchmarks. However, we need to explore more benchmarks to
-be conclusive. PWA-Rel's performance can be explained by the
-difficulty in finding a good abstraction ($\quant_\epsilon$), which
-needs to be fine enough, to obtain good prediction models but coarse
-enough, to be manageable by the current SMT solvers. Specifically, to
-obtain a good quality counter-example from the model checker (and
-avoid false positives), the transition system should be created from
-models with high accuracy, for which a finer abstraction is required.
-However, the exploration of a finer abstraction is exponentially more
-complex and results in a much bigger transition system, which the
-current state of the art bounded model checkers (which use SMT
-solvers) can not handle under reasonable resources.
+be `safe' as it can not find a counter-example in the PWA relational
+model. Also, apparent froom the results, PWA-Rel is more scalale with
+the LP solver and PWA-Rel+SAL times out on most of the benchmarks.
+The efficiency of the naive LP solver can be attributed to the careful
+selection of the pareters for S3CAM (finer abstraction), resulting in
+a small subgraph which can be easily explored for all paths.
+
+However, we need to explore more benchmarks to be conclusive on the
+tool's performance on real world examples. Although, PWA-Rel does not
+need to refine like S3CAM, and should be more scalable, it is still
+constrained by the accuracy of the pwa models. Its performance can be
+explained by the difficulty in finding an abstraction
+($\quant_\epsilon$), which is fine enough, to result in accurate
+prediction models but coarse enough to be manageable by the current
+bmc tools.  To obtain a reproducible counter-example from the model
+checker (and avoid false positives), the transition system should be
+created from models with high accuracy, for which a finer abstraction
+is required.  However, the exploration of a finer abstraction is
+exponentially more complex and results in a much bigger transition
+system, which the current state of the art bounded model checkers
+(which use generic SMT solvers) can not handle under reasonable
+resources.

### 2017-04-08 bmc.tex
.

@@
 
-    \item \emph{S3CAM Fails} No abstract counter-example is found by
+    \item \emph{S3CAM Fails}: No abstract counter-example is found by
         S3CAM. We can remedy it by increasing search budgets and/or
@@
 
-    \item \emph{BMC Fails} An abstract counter-example is found, but
+    \item \emph{BMC Fails}: An abstract counter-example is found, but
         the BMC fails to find a concrete counter-example in the PWA
@@
 
-    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
+    \item \emph{Inaccurate PWA Modeling}: An abstract counter-example
         is found, and it is successfully concretized in the PWA
@@
         model is not precise enough.
-
 \end{itemize}

### 2017-04-08 bmc.tex
.

--- 2017-04-08 bmc.tex
@@
 % (-3, -22) & (-5, -29) &  \begin{bmatrix}3.08806462&-0.72332776 \\-1.69271079&0.40308293\end{bmatrix} &\begin{bmatrix}-2.77908467\\-4.57411339\end{bmatrix}&\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}\\
-
-\subsection{Search Parameters}
-
-The search parameters for PWA-Rel include the parameters of S3CAM:
-$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
-maximum error budget $\delta_{max}$ for OLS and the maximum length of
-segmented trajectory for building $k$-relational models.
-
-The effects of $N$, $\epsilon$ and $\Delta$
-on S3CAM's performance were detailed in~\cite{zutshi2014multiple}. However, they also have an effect on
-relational modeling. A finer grid with small cells produces more
-accurate models than a coarser grid with large cells. Similarly,
-small time length trajectory segments result in more accurate
-modeling.
-%An increase in the number of samples can increases the
-%accuracy of regression.
-
-\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
-Given a $\delta_{max}$, we keep increasing $k$ during the
-$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
-satisfied. A $k_{max}$ is introduced to bound the longest
-segmented trajectory which can be considered.
-
-% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
-
-\subsection{Reasons for Failure}
-The approach can fail to find a counter-example, even when it exists, in
-one of three ways.
-
-\begin{itemize}
-
-    \item \emph{S3CAM Fails} No abstract counter-example is found by
-        S3CAM. We can remedy it by increasing search budgets and/or
-        restarting.
-
-    \item \emph{BMC Fails} An abstract counter-example is found, but
-        the BMC fails to find a concrete counter-example in the PWA
-        relational model. The failure can be attributed to either (a)
-        a spurious abstract counter-examples or (b) a poorly estimated
-        model. The former can be addressed by restarting but the
-        latter requires that the maximum model error $\delta_{max}$ be
-        decreased.
-
-    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
-        is found, and it is successfully concretized in the PWA
-        relational model by the BMC. However, it is not reproducible
-        in the black box system. This happens when the PWA relational
-        model is not precise enough.
-
-\end{itemize}

### 2017-04-08 implementation.tex
.

@@
 resources.
+
+% more scalable 
+\subsection{Search Parameters}
+
+The search parameters for PWA-Rel include the parameters of S3CAM:
+$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
+maximum error budget $\delta_{max}$ for OLS and the maximum length of
+segmented trajectory for building $k$-relational models.
+
+The effects of $N$, $\epsilon$ and $\Delta$
+on S3CAM's performance were detailed in~\cite{zutshi2014multiple}. However, they also have an effect on
+relational modeling. A finer grid with small cells produces more
+accurate models than a coarser grid with large cells. Similarly,
+small time length trajectory segments result in more accurate
+modeling.
+%An increase in the number of samples can increases the
+%accuracy of regression.
+
+\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
+Given a $\delta_{max}$, we keep increasing $k$ during the
+$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
+satisfied. A $k_{max}$ is introduced to bound the longest
+segmented trajectory which can be considered.
+
+% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
+
+\subsection{Reasons for Failure}
+The approach can fail to find a counter-example, even when it exists, in
+one of three ways.
+
+\begin{itemize}
+
+    \item \emph{S3CAM Fails}: No abstract counter-example is found by
+        S3CAM. We can remedy it by increasing search budgets and/or
+        restarting.
+
+    \item \emph{BMC Fails}: An abstract counter-example is found, but
+        the BMC fails to find a concrete counter-example in the PWA
+        relational model. The failure can be attributed to either (a)
+        a spurious abstract counter-examples or (b) a poorly estimated
+        model. The former can be addressed by restarting but the
+        latter requires that the maximum model error $\delta_{max}$ be
+        decreased.
+
+    \item \emph{Inaccurate PWA Modeling}: An abstract counter-example
+        is found, and it is successfully concretized in the PWA
+        relational model by the BMC. However, it is not reproducible
+        in the black box system. This happens when the PWA relational
+        model is not precise enough.
+\end{itemize}

### 2017-04-08 bmc.tex
.

@@
     out as before, using the numerical simulation function
-    $\simulate$. For the given example, the model checker is unable to
-    find a counter-example.
+        $\simulate$~\footnote{For the given example, the model checker is unable to
+        find a counter-example.}.
 \end{enumerate}

### 2017-04-08 pwarel.tex
.

@@
 like Counter-Example Guided Refinement must be used to find concrete
-paths. In general, search techniques are exponential with respect to
+paths. In general, such search techniques are exponential with respect to
 computational resources and memory. As an alternative, we introduce
@@
 maps with the relations transforms a discrete abstraction into a PWA
-transitions system.
+transition system.
 

### 2017-04-08 pwarel.tex
.

@@
 
+% The abstraction of the hybrid dynamical system is a a
+% paritioning of the combined state-input space $\HybridStates \times
+% \Inputs$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
+% pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
+% their union $\bigcup_{C_i\in\Cells}C_i$ is the state-input
+% space $\HybridStates \times \Inputs$. The paritions over
+% the continuous input-state space are implicitly defined using a quantization
+% function $Q_\epsilon$ parameterized by the precision $\epsilon$.
+% This further defines an equivalence relations such that
+% \[
+%     (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; Q_\epsilon(\vx_i,
+%     \vu_i) = Q_\epsilon(\vx_j, \vu_j)
+% \]
+% In other words, $(\vx_i, \vu_i) \in C_i \; \text{iff} \; C_i =
+% Q_\epsilon(\vx_i, \vu_i)$.
+
+% \begin{definition}[Graph Abstraction]
+%     The quantization function $Q_\epsilon$ induces a DAG on the discrete
+%     abstraction given by a graph $G(V, E)$, where each vertex
+%     corresponds to a cell $V_i = C$, and an edge exists between two
+%     cells $(C, C') \in E_i$ iff there exist state-input pairs
+%     $(\vx,\vu) \in C$, $(\vx', \vu) \in C'$ such that $\vx' =
+%     \simulate(\vx, \vu )$.
+% \end{definition}
+
 The abstraction of the hybrid dynamical system is a a
-paritioning of the combined state-input space $\HybridStates \times
-\Inputs$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
+paritioning of the hybrid state space $\HybridStates$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
 pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
-their union $\bigcup_{C_i\in\Cells}C_i$ is the state-input
-space $\HybridStates \times \Inputs$. The paritions over
-the continuous input-state space are implicitly defined using a quantization
+their union $\bigcup_{C_i\in\Cells}C_i$ is the state space $\HybridStates$. The paritions over
+the continuous state space are implicitly defined using a quantization
 function $Q_\epsilon$ parameterized by the precision $\epsilon$.
@@
 \[
-    (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; Q_\epsilon(\vx_i,
-    \vu_i) = Q_\epsilon(\vx_j, \vu_j)
+    (\vx_i) \equiv (\vx_j) \;\text{iff}\; Q_\epsilon(\vx_i) = Q_\epsilon(\vx_j)
 \]
-In other words, $(\vx_i, \vu_i) \in C_i \; \text{iff} \; C_i =
-Q_\epsilon(\vx_i, \vu_i)$.
+In other words, $(\vx_i) \in C_i \; \text{iff} \; C_i = Q_\epsilon(\vx_i)$.
 
@@
     cells $(C, C') \in E_i$ iff there exist state-input pairs
-    $(\vx,\vu) \in C$, $(\vx', \vu) \in C'$ such that $\vx' =
-    \simulate(\vx, \vu )$.
+    $(\vx) \in C$, $(\vx') \in C'$ such that $\vx' =
+    \simulate(\vx)$.
 \end{definition}

### 2017-04-08 pwarel.tex
.

@@
 function $Q_\epsilon$ parameterized by the precision $\epsilon$.
-This further defines an equivalence relations such that
-\[
-    (\vx_i) \equiv (\vx_j) \;\text{iff}\; Q_\epsilon(\vx_i) = Q_\epsilon(\vx_j)
-\]
+This further defines an equivalence relations such that $(\vx_i)
+\equiv (\vx_j) \;\text{iff}\; Q_\epsilon(\vx_i) = Q_\epsilon(\vx_j)$.
 In other words, $(\vx_i) \in C_i \; \text{iff} \; C_i = Q_\epsilon(\vx_i)$.
@@
 \begin{definition}[Graph Abstraction]
-    The quantization function $Q_\epsilon$ induces a DAG on the discrete
+    The quantization function $Q_\epsilon$ induces a directed acyclic
+    graph (DAG) on the discrete
     abstraction given by a graph $G(V, E)$, where each vertex

### 2017-04-08 bmc.tex
prelims changes

--- 2017-04-08 bmc.tex
@@
 
-    \item \emph{S3CAM Fails}: No abstract counter-example is found by
+    \item \emph{S3CAM Fails} No abstract counter-example is found by
         S3CAM. We can remedy it by increasing search budgets and/or
@@
 
-    \item \emph{BMC Fails}: An abstract counter-example is found, but
+    \item \emph{BMC Fails} An abstract counter-example is found, but
         the BMC fails to find a concrete counter-example in the PWA
@@
 
-    \item \emph{Inaccurate PWA Modeling}: An abstract counter-example
+    \item \emph{Inaccurate PWA Modeling} An abstract counter-example
         is found, and it is successfully concretized in the PWA
@@
         model is not precise enough.
+
 \end{itemize}

### 2017-04-08 intro.tex
prelims changes

@@
 \cite{ljung1999system}, auto-regressive models for forecasting
-\cite{wei1994time}, machine learning techniques that learn static or
-dynamic models from data
-\cite{narendra1990identification,lu2009linear}.  Why then do we need a
-new modeling technique? The answer to this question lies in our goal
-for the second step, \ie, symbolic analysis of the learned model. For
-this step, we need to have models that can be digested by existing
-symbolic tools, which is harder to do with some of the aforementioned
-heavier data-driven modeling methods. Hence, we use a lighter flavor
-of data-driven modeling such as the one proposed for learning
-relational models from data
+\cite{wei1994time}, bounded-error hybrid systems identification
+methods \cite{bemporad2003greedy,bemporad2005bounded}, clustering
+based methods \cite{ferrari2003clustering}, methods to identify
+piecewise affine dynamics \cite{paoletti2007identification}, and
+machine learning techniques that learn static or dynamic models from
+data
+\cite{narendra1990identification,lu2009linear,juloski2005bayesian}.
+Why then do we need a new modeling technique? The answer to this
+question lies in our goal for the second step, \ie, symbolic analysis
+of the learned model. For this step, we need to have models that can
+be digested by existing symbolic tools, which is harder to do with
+some of the aforementioned heavier data-driven modeling methods.
+Hence, we use a lighter flavor of data-driven modeling such as the one
+proposed for learning relational models from data
 \cite{zutshi2012timed,sankaranarayanan2011relational}. By learning a
@@
 In \cite{zutshi2014multiple}, the authors investigate an on-the-fly
-abstraction technique that is close to the approach we propose. A key
-difference is that the abstraction explored in this work is an
-existential {\em relational} abstraction. The approximate model
-learned is a directed acyclic graph, where nodes represent regions of
-the state-space, and an edge between two regions indicates a
-transition between states in the regions. Our approach further
-improves the existential abstraction by annotating the edges in the
-graph-based abstraction with affine relations between the source and
-destination states.
+abstraction technique that is close to the approach we propose.  Both
+approaches use an existential abstraction of the state-space of the
+hybrid dynamical system. The abstract model learned is a directed
+acyclic graph, where nodes represent regions of the state-space, and
+an edge between two regions indicates a transition between states in
+the regions.  A key difference in this work is that the model being
+learned in addition to being an existential graph-based abstraction is
+also a {\em relational} abstraction.  In other words, the approach
+presented here improves the existential abstraction by annotating the
+edges in the graph-based abstraction with affine relations between the
+source and destination states.  More specifically, we use linear
+regression to quantitatively estimate the discovered edges in the
+graph-based abstraction of the state-space. These linear maps
+approximate locally observed behaviors, and the resulting graph or the
+Piece-Wise Affine (PWA) relational model can be interpreted as an
+infinite state discrete transition system and model checked for time
+bounded safety properties. A counter-example if found, can indicate
+the presence of a violation in the system.  We then use this
+counter-example to guide the search towards a counter-example in the
+original system.
 
-% We further our exploration of trajectory segment based methods; and
-% explore an alternative approach to overcome the explosion in abstract
-% states. Instead of selectively refining the abstraction, we compute a
-% model of the black box system and use bounded model checking to find a
-% concrete counter-example in the model. Due to modeling errors, this
-% might not be reproducible in the original black-box system.
-
-More specifically, we use linear regression to quantitatively estimate
-the discovered edges in the graph-based abstraction of the
-state-space. These linear maps approximate locally observed behaviors,
-and the resulting graph or the Piece-Wise Affine (PWA) relational
-model can be interpreted as an infinite state discrete transition
-system and model checked for time bounded safety properties. A
-counter-example if found, can indicate the presence of a violation in
-the system.  We then use this counter-example to guide the search
-towards a counter-example in the original system.
+We remark that our work focusses on building discrete PWA models, and
+is unrelated to work on analyzing dynamical systems that have PWA
+vector fields. Techniques to analyze PWA dynamical systems using
+bisimulation relations translate the system into a non-deterministic
+infinite state transition system, and use model checking approaches to
+verify them \cite{asarin2000approximate}. Related approaches using
+bisimulation relations \cite{pappas2003bisimilar,tabuada2006linear,
+yordanov2007model} have interesting parallels with the work presented
+here.
 

### 2017-04-08 prelims.tex
emsoft formatvi main.tex
.
prelims changes

@@
 
-\subsection{Piecewise Affine (PWA) Transition System}
+\subsection{PWA Transition System}
 
-We define the PWA transition system as a discrete transition system
-$\rho:\tupleof{\pLocs,\vx,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
-a nonempty set of locations, $\vx$ is a set of variables, $\scrT$
-is a finite set of transitions of the form $\tupleof{l, \tau,
-l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
-post-location and $\tau \in \scrT$ is the transition. $\tau$
-is labled by the transition relation $\rho_{\tau}$
-over the current states and the next states. The transition relation
-is defined in terms of affine predicates on the current and next
-states and an affine update between them.
-\[
-    \rho_{\tau}(\vx,\vx') \subseteq \setof{(\vx,\vx')) \;|\;
-    g_\tau(\vx) \land g'_\tau(\vx') \land f_{\tau}(\vx, \vx')}
-\]
-where $f_{\tau}$ is an affine relation on the pre and post states
-$\vx,\vx'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
-post conditions) on the states. Finally, $\pLocI$ is an initial
-location $\Theta$ is an affine assertion characterizing the initial
-states included in $\pLocI$.
+We define a PWA transition system $A$ as a tuple
+$\tupleof{\pLocs,\vx,\scrT, G, U, \pLocI,\Theta}$, where:
+\begin{itemize}[noitemsep, leftmargin= 1.5 em]
+\item
+$\pLocs$ is a set of locations,
+\item
+$\vx$ is a variable that takes values in $\reals^m$,
+\item
+$\scrT$ is a finite subset of $\pLocs \times G \times U \times G \times \pLocs$, 
+\item
+$G$ is a set of conjunctions of affine predicates, 
+\item
+$U$ is a set of affine relations,
+\item
+$\pLocI$ is an initial location, and,
+\item
+$\Theta$ is an initial affine predicate.
+\end{itemize}
+
+The operation of $A$ is described in terms of its configuration and
+its moves. A configuration of $A$ is a pair $(\ell, \nu)$, where $\ell
+\in \pLocs$, and $\nu$ is a valuation function assigning values in
+$\reals^m$ to the variable $\vx$.  The move function $\mu$ maps a
+configuration $(\ell,\nu)$ to the next configuration $(\ell',\nu')$,
+iff there is a transition $\tau$ of the form $(\ell,g,f,g',\ell')$,
+and $\nu(\vx)$ satisfies predicate $g$, $\nu'(\vx)$ satisfies
+predicate $g'$ and $\nu'(\vx) \in f(\nu(\vx))$.
+
+For a given $\tau$, we call the affine relation $f_\tau$ its update
+relation, and the predicates $g_\tau$ and $g'_\tau$ as pre- and
+post-conditions on the transition. The machine starts its operation in
+location $\pLocI$ with an initial valuation $\nu_0$ to the variable
+$\vx$ such that $\nu_0(\vx)$ satisfies $\Theta$.
+
+% We define the PWA transition system as a discrete transition system
+% $\rho:\tupleof{\pLocs,\vx,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
+% a nonempty set of locations, $\vx$ is a set of variables, $\scrT$
+% is a finite set of transitions of the form $\tupleof{l, \tau,
+% l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
+% post-location and $\tau \in \scrT$ is the transition. $\tau$
+% is labled by the transition relation $\rho_{\tau}$
+% over the current states and the next states. The transition relation
+% is defined in terms of affine predicates on the current and next
+% states and an affine update between them.
+% \[
+%     \rho_{\tau}(\vx,\vx') \subseteq \setof{(\vx,\vx')) \;|\;
+%     g_\tau(\vx) \land g'_\tau(\vx') \land f_{\tau}(\vx, \vx')}
+% \]
+% where $f_{\tau}$ is an affine relation on the pre and post states
+% $\vx,\vx'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
+% post conditions) on the states. Finally, $\pLocI$ is an initial
+% location $\Theta$ is an affine assertion characterizing the initial
+% states included in $\pLocI$.
 
@@
 
-We can now use a PWA transition system $\rho$ to approximate the
-behavior of a hybrid dynamical system $\System$ defined by a simulator
-$\simulate^{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
-form, we can only approximate general hybrid systems.}. Each
-transition of $\rho$ represents a discrete time step $\Delta$. Let the
-state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
-states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
-predicates $g_\tau$ are given by a conjunction of hyper-planes and
-hence can be represented in matrix form as a polyhedron (defined by
-$m$ constraints) in the state-space
-\[
-g:C\x - \vd \le 0
-\]
-where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
-$m$. The affine relation $f_\tau$ can be represented using matrix $A$
-and an offset vector $\vb$ as $f_\tau \subseteq \setof{(\x,\x') | \x'
-= A\x+\vb}$. Finaly, we incorporte an approximate non-deterministic
-error term $\delta$, defined as a vector of intervals
-$[\delta_{\min},\delta_{\max}]$.
-\[
-    f_\tau \subseteq \setof{(\x,\x') | \x' = A\x+\vb}.
-\]
+We how we can use the formal machinery of a PWA transition system $A$
+to approximate the behavior of a hybrid dynamical system $\System$
+defined by a simulator $\simulate^{\System}$~\footnote{Due to $f_\tau$
+being restricted to an affine form, we can only approximate general
+hybrid systems.}.
+
 %defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
@@
 
-Given a hybrid dynamical system over a state-space $\reals^n$, a PWA
-relational model $\rho$ is given by the tuple
-$\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in \scrT$ is a
-collection of affine transitions  and $\Theta$ is an affine predicate
-over $\x\in\real^n$. The transition relation is then defined by
-$\scrT$ with $n$ transition relations as follows
+A PWA relational model $A(\System)$ of a hybrid dynamical system
+$\System$ is simply a PWA transition system, where:
 
-\begin{equation}
-    \scrT = \left\{
-        \begin{array}{ll}
-            g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
-            \ldots\\
-            g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
-        \end{array}
-    \right.
-\end{equation}
+\begin{itemize}[noitemsep, leftmargin= 1.5 em]
+\item
+each transition of $A$ represents a discrete time step $\Delta$ of the
+system $\System$,
+\item
+if the state-space of $\System$ is $\reals^n$, the the valuation
+function for variable $\x$ in $A$ assigns $\x$ some value in
+$\reals^n$,
+\item
+for the $N$ tuples $(v,v')$ discovered by simulation, if $v' = Av
++ B + \vec{\delta}$ represents the affine relation discovered using
+linear regression, and if $h$ is the convex hull of all the $v$
+states across all tuples and $h'$ is the convex hull of all the $v'$
+states across all tuples, then $f_\tau \subseteq
+\setof{(\nu(\x),\nu'(\x)) \mid \nu'(\x) \in A\x + B + \vec{\delta}}$,
+and $g$ states that $\nu(\x)$ is in $h$, and $g'$ states that
+$\nu'(\x) \in h'$.
+\end{itemize}
+
+Note that the guard predicates $g_\tau$ represent a conjunction of $m$
+affine inequalities, then we can represent $g_\tau$ as follows:
+\[
+g_\tau  \equiv C\x - \vd \le 0
+\]
+where $C$ is an $m \times n$ matrix and $\vd$ is $m \times 1$ matrix.
+
+% Given a hybrid dynamical system over a state-space $\reals^n$, a PWA
+% relational model $\rho$ is given by the tuple
+% $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in \scrT$ is a
+% collection of affine transitions  and $\Theta$ is an affine predicate
+% over $\x\in\real^n$. The transition relation is then defined by
+% $\scrT$ with $n$ transition relations as follows
+% 
+% \begin{equation}
+%     \scrT = \left\{
+%         \begin{array}{ll}
+%             g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
+%             \ldots\\
+%             g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
+%         \end{array}
+%     \right.
+% \end{equation}
 \end{definition}
 
-where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
-relations and $\x'$ denotes the next state of the system.
+% where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
+% relations and $\x'$ denotes the next state of the system.
 
-A PWA model is \textit{deterministic}, iff for every state
+A PWA relational model is \textit{deterministic}, iff for every state
 $\x\in\reals^n$, a unique guarded affine map is satisfied and if all
@@
 
-A PWA transition system is \textit{complete}, iff for every state
-$\x\in\HybridStates^n$, there exists at least one satisfied
+A PWA relational model system is \textit{complete}, iff for every
+state $\x\in\HybridStates^n$, there exists at least one satisfied
 transition relation $\scrT$. If this is not the case, the system can
 deadlock, with no further executions. This usually results from
-modeling errors, and from here on, we do not consider such
-cases.
+modeling errors, and from here on, we do not consider such cases.
 

### 2017-04-08 prelims.tex
.

@@
 From here on, we drop the subscript $\System$ whenever it is clear
-from the context.
+from the context. Also, to ease the presentation, we conflate the
+inputs with the state and denote them by $\vx$.
 
@@
 % post-location and $\tau \in \scrT$ is the transition. $\tau$
-% is labled by the transition relation $\rho_{\tau}$
+% is labeled by the transition relation $\rho_{\tau}$
 % over the current states and the next states. The transition relation
@@
 bounded reachability query is equivalent to checking all
-combinaions of discrete locations, where each combination can be
+combinations of discrete locations, where each combination can be
 summarized as a linear program. As the time is bounded, the number of
@@
 time. A path on this graph is then a linear program, with each
-branching node corresponsing to a logical disjunction.
+branching node corresponding to a logical disjunction.
 

### 2017-04-08 pwarel.tex
.

@@
 the underlying relation between concrete states, and concrete paths
-can no longer be constructed directly. Instead, search procedueres
+can no longer be constructed directly. Instead, search procedures
 like Counter-Example Guided Refinement must be used to find concrete
@@
 % The abstraction of the hybrid dynamical system is a a
-% paritioning of the combined state-input space $\HybridStates \times
+% partitioning of the combined state-input space $\HybridStates \times
 % \Inputs$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
@@
 % their union $\bigcup_{C_i\in\Cells}C_i$ is the state-input
-% space $\HybridStates \times \Inputs$. The paritions over
+% space $\HybridStates \times \Inputs$. The partitions over
 % the continuous input-state space are implicitly defined using a quantization
@@
 
+Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
+we use the generated trajectories to learn quantitative models
+describing the local behavior of the system. These models are defined
+by a set of relations $R \subseteq \HybridStates \times \HybridStates$
+for each edge of the reachability graph. In the next section, we
+summarize this idea.
+
 \subsection{Enriched Abstraction}
@@
 set computation methods. Because we assume black box semantics, we
-only present the former. Using regression analysis on a witnessed trajectory between two cells,
-we compute an approximate discrete map along with an error estimate.
-Moreover, using the simulation function $\simulate$, additional
-trajectories (or data) can be generated if required. The data
-can be separated into a training set and testing set to compute the
-map and the error respectively.
+only present the former. Using regression analysis on a witnessed
+trajectory between two cells, we compute an approximate discrete map
+along with an error estimate.  Moreover, using the simulation function
+$\simulate$, additional trajectories (or data) can be generated if
+required. The data can be separated into a training set and testing
+set to compute the map and the error respectively.
 
@@
 off-the-shelf bounded model checkers to find concrete violations of a
-given safety property, and even other temporal properties.  We now
-discuss the background required to present our ideas.
-
-Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
-we use the generated trajectories to learn quantitative models
-describing the local behavior of the system. These models are defined
-by a set of relations $R \subseteq \HybridStates \times \HybridStates$
-for each edge of the reachability graph.
+given safety property, and even other temporal properties.

### 2017-04-08 prelims.tex
simplifying notation and consistency

--- 2017-04-08 prelims.tex
@@
 location $\pLocI$ with an initial valuation $\nu_0$ to the variable
-$\vx$ such that $\nu_0(\vx)$ satisfies $\Theta$.
+$\vx$ such that $\nu_0(\vx)$ satisfies $\Theta$. For convenience, we
+abuse notation, and use the name a variable as a proxy for a certain
+valuation, and use the primed version of the variable as a proxy for
+its valuation after a transition, \ie, $\vx$ is used instead of
+$\nu(\vx)$ and $\vx'$ is used instead of $\nu'(\vx)$.
 
@@
 states across all tuples, then $f_\tau \subseteq
-\setof{(\nu(\x),\nu'(\x)) \mid \nu'(\x) \in A\x + B + \vec{\delta}}$,
-and $g$ states that $\nu(\x)$ is in $h$, and $g'$ states that
-$\nu'(\x) \in h'$.
+\setof{(\x,\x') \mid \x' \in A\x + B + \vec{\delta}}$,
+and $g$ states that $\x$ is in $h$, and $g'$ states that
+$\x' \in h'$.
 \end{itemize}

### 2017-04-08 bmc.tex
.
simplifying notation and consistency

--- 2017-04-08 bmc.tex
@@
-\section{Model Checking}
+% \section{Model Checking}
 
+\mypara{Model Checking}
 As a final step, the enriched graph is model checked for the given
@@
 
-We now describe the pesented technique using the Van der Pol
+We now describe the presented technique using the Van der Pol
 oscillator benchmark from~\cite{zutshi2014multiple}. Simulations

### 2017-04-08 implementation.tex

@@
 PWA-Rel was implemented in Python. Its different parts comprised of
-the (a) abstraction graph generation, (b) regeression analysis and (c)
+the (a) abstraction graph generation, (b) regression analysis and (c)
 BMC. S3CAM was used to compute (a). To compute (b), OLS regression
@@
 to the differences in which they operate. PWA-Rel requires a finer
-abstraction to begni with in order to minimize the modeling errors.
-On the other hand, a coarser abstraction as the intital abstraction is
+abstraction to begin with in order to minimize the modeling errors.
+On the other hand, a coarser abstraction as the initial abstraction is
 works better for S3CAM due to the expensive refinement procedure
@@
 be `safe' as it can not find a counter-example in the PWA relational
-model. Also, apparent froom the results, PWA-Rel is more scalale with
+model. Also, apparent from the results, PWA-Rel is more scalable with
 the LP solver and PWA-Rel+SAL times out on most of the benchmarks.
 The efficiency of the naive LP solver can be attributed to the careful
-selection of the pareters for S3CAM (finer abstraction), resulting in
+selection of the parameters for S3CAM (finer abstraction), resulting in
 a small subgraph which can be easily explored for all paths.

### 2017-04-08 prelims.tex
lazy fix

--- 2017-04-08 prelims.tex
@@
 modeling errors, and from here on, we do not consider such cases.
+
+In what follows, we use the terms PWA relational model and PWA
+transition system interchangeably to mean PWA relational model.
 

### 2017-04-08 implementation.tex
.

@@
-The presented technique was prototyped as a tool: PWA-Rel.
-We compared it against S3CAM~\cite{zutshi2014multiple} on several academic
-benchmarks.
+The presented technique was prototyped as a tool: PWA-Rel.  We
+compared it against S3CAM~\cite{zutshi2014multiple} on several
+academic benchmarks. Below, we present the details of the
+implementation and the benchmarks used.
 
@@
 
-We used S3CAM to discover the abstraction graph $G$, which was then
-filtered to remove the nodes from which error nodes were not
-reachable. In our experiments we used $1$-relational modeling to
-create the transition system from $G$. Using the BMC, we checked for
-the given safety property. If a concrete trace was obtained, it was
-further checked for reproducibility by sampling and simulating its `neighborhood' with the
-$\simulate$ function. We define `neighborhood' as the abstract states
+PWA-Rel uses S3CAM to discover the abstraction graph $G$, which is
+then filtered to remove the nodes from which error nodes are not
+reachable. In all our experiments we used $1$-relational modeling to
+create the transition system from $G$. This was determined to be a
+good trade-off between longer simulations and increase in precision,
+with respect to the mentioned benchmarks.
+
+Using the BMC, we checked for the given safety property. If a concrete
+trace was obtained, it was further checked for reproducibility by
+sampling and simulating its `neighborhood' with the $\simulate$
+function. We define `neighborhood' as the abstract states
 corresponding to the returned concrete trace.  Specifically, $100$

### 2017-04-08 implementation.tex
.

@@
 
+\subsection{Implementation}
 PWA-Rel was implemented in Python. Its different parts comprised of
@@
 which used Yices2~\cite{dutertre2014yices} as the SMT solver. A naive
-alternative linear programming (LP) based model checker specific to the PWA
-relational abstractions was implemented utilizing PyGLPK.
+alternative linear programming (LP) based model checker, specific to
+the PWA relational abstractions was implemented utilizing PyGLPK.
+Specifically, each path of the graph $G^R$ was sent to the lp solver
+PyGLPK and checked for feasibility. If feasible, the checker returned
+a concrete counter-example.
 
@@
 
+\subsection{Evaluation}
 We tabulate our preliminary evaluation in \tabref{res-rel}. We used
@@
 including the Van der Pol oscillator, Brusselator, Lorenz attractor
-and the Navigation benchmark. PWA-Rel and S3CAM were ran $10$ times
-with different seeds and the running times were averaged. We tabulate
+and the Navigation benchmark. PWA-Rel and S3CAM were tested $10$ times
+with different seeds and their running times were averaged. We tabulate
 both the total time taken by PWA-Rel+SAL and PWA-Rel+LP to compute the

### 2017-04-08 pwarel.tex
.

@@
 The abstraction of the hybrid dynamical system is a a
-paritioning of the hybrid state space $\HybridStates$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
+partitioning of the hybrid state space $\HybridStates$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
 pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
-their union $\bigcup_{C_i\in\Cells}C_i$ is the state space $\HybridStates$. The paritions over
+their union $\bigcup_{C_i\in\Cells}C_i$ is the state space $\HybridStates$. The partitions over
 the continuous state space are implicitly defined using a quantization

### 2017-04-08 implementation.tex
.

@@
 \caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
-time taken by the BMC engine. The total time in seconds (rounded off
-to an integer) is noted under \textbf{PWA-Rel} and \textbf{S3CAM}.
+time taken by the BMC engine. The total time in minutes is noted under
+    \textbf{PWA-Rel+BMC solver} and \textbf{S3CAM}.
 $TO$ signifies time $>5hr$, after which the search was killed.}
@@
 
-As apparent form the results, the performance is comparable in all
-benchmarks, but the last one: bouncing ball. PWA-Rel determines it to
-be `safe' as it can not find a counter-example in the PWA relational
-model. Also, apparent from the results, PWA-Rel is more scalable with
-the LP solver and PWA-Rel+SAL times out on most of the benchmarks.
-The efficiency of the naive LP solver can be attributed to the careful
-selection of the parameters for S3CAM (finer abstraction), resulting in
-a small subgraph which can be easily explored for all paths.
+In all the runs of both the tools, a falsification was found. As
+apparent form the results, the performance is comparable in all
+benchmarks, but the last benchmark: bouncing ball. PWA-Rel determines
+it to be `safe' as it can not find a counter-example in the PWA
+relational model. Also, apparent from the results, PWA-Rel is more
+scalable with the LP solver and PWA-Rel+SAL times out on most of the
+benchmarks.  The efficiency of the naive LP solver can be attributed
+to the careful selection of the parameters for S3CAM (finer
+abstraction), resulting in a small subgraph which can be easily
+explored for all paths.
 

### 2017-05-26 abstract.tex
changed abstract to clarify the confusion apparent in reviews

@@
+Models of embedded control systems are often so complex that
+falsification and testing approaches can scale only by numerically
+simulating them, and hence treating the systems as black-boxes. Common
+approaches like sampling-based falsification, use numerical
+simulations to search for violations of safety properties. However,
+this relegates them to the `best effort' category. On the other hand,
+\emph{rigorous} verification (and falsification) techniques require
+symbolic models and are comparatively less scalable. We try to bridge
+this gap by incorporating data-driven models in a sampling based
+search. Using an efficient, on-the-fly discrete abstraction
+search procedure, we compute models only for selective regions of the state
+space, thus `enriching' the abstraction. We then show how the
+`enriched' abstraction can be used to guide the search for violations
+without refinement.
+%We use grid-based discrete abstractions to partition
+%the state-space of the black-box hybrid dynamical system. We then use
+%an efficient search procedure to find abstract violations of safety
+%properties
+%Assuming a grid-based implicit abstraction over the
+%state-space of the black-box hybrid dynamical system, we enumerate
+%abstract violations of the safety property. We then model the system
+%behavior along these violations and search for concrete violations.
+The automatic falsification proceeds by (a) randomly exploring the
+state-space abstraction using numerically computed trajectories and
+enumerating abstract violations, (b) modeling the behavior of the
+system along these abstract violations by using discrete-time
+piecewise affine (PWA) \emph{relational} models, and finally, (c)
+encoding the search for a concrete violation as a bounded model
+checking (BMC) problem.  The BMC query is then a disjunction of
+conjunctions of linear constraints (i.e. combinations of linear
+programs). It can be solved using off-the-shelf SMT solvers or by
+multiple runs of a linear programming solver. For any counter-example
+found in this `enriched' abstraction, we check the existence of a
+corresponding violation in the original system. We demonstrate the
+efficacy of our technique on a few benchmarks.
+
+% Models of embedded control systems are often so complex that they can
+% be treated as effectively black-box systems. Sampling-based methods
+% for testing such systems use numerical simulations to search for
+% violations of properties (specified in a suitable requirement
+% formalism, such as temporal logic). This constrains them to a `best
+% effort' category unlike rigorous verification techniques, which
+% however require a model amenable to symbolic analysis. This work tries
+% to bridge the gap by presenting an automatic falsification technique
+% which learns a data-driven abstraction of the black box system before
+% searching for a violation. It proceeds by (a) modeling the behavior of
+% the system using a piecewise affine (PWA) discrete-time {\em
+% relational} model and, (b) encoding the search for a violation in the
+% abstract model as a bounded model checking problem. The bounded model
+% checking query in this instance can be viewed as a disjunction of
+% conjunctions of linear constraints (i.e. linear programs) and can be
+% solved directly using off-the-shelf SMT solvers or by multiple runs of
+% a linear programming solver. For any counter-example found in the
+% abstraction, we check the existence of a corresponding violation in
+% the original system. We demonstrate the efficacy of our technique on a
+% few dynamical systems examples.
+
+% =========================================================================
 % Sampling based falsification techniques for black box hybrid dynamical
@@
 
-Models of embedded control systems are often so complex that they can
-be treated as effectively black-box systems. Sampling-based methods
-for testing such systems use numerical simulations to search for
-violations of properties (specified in a suitable requirement
-formalism, such as temporal logic). This constrains them to a `best
-effort' category unlike rigorous verification techniques, which
-however require a model amenable to symbolic analysis. This work tries
-to bridge the gap by presenting an automatic falsification technique
-which learns a data-driven abstraction of the black box system before
-searching for a violation. It proceeds by (a) modeling the behavior of
-the system using a piecewise affine (PWA) discrete-time {\em
-relational} model and, (b) encoding the search for a violation in the
-abstract model as a bounded model checking problem. The bounded model
-checking query in this instance can be viewed as a disjunction of
-conjunctions of linear constraints (i.e. linear programs) and can be
-solved directly using off-the-shelf SMT solvers or by multiple runs of
-a linear programming solver. For any counter-example found in the
-abstraction, we check the existence of a corresponding violation in
-the original system. We demonstrate the efficacy of our technique on a
-few dynamical systems examples.
-
 % In this work we address the problem of finding unsafe behaviors with

### 2017-05-30 graph_abs.tex
.

--- 2017-04-08 pwarel.tex
@@
+In this section, we present the model of the hybrid dynamical system
+under test, and the background for PWA relational abstractions.
+
+\mypara{System under test ($\System$)}
+
+We assume that $\System$ is a hybrid dynamical system with exogenous
+inputs modeling an embedded control system. The system state-space
+(denoted $\HybridStates$) is a subset of $\Modes \times \Reals^n$,
+where $\Modes$ is a finite set of discrete modes, and $\Reals^n$
+represents the continuous-valued state-space. For ease of exposition,
+we assume that the hybrid system is effectively a switched-mode
+dynamical system, \ie, the mode of the system is completely determined
+by its state.  An input to $\System$ is assumed to be a time-bounded
+{\em signal}, \ie, a function from $[0,T]$ to $\Inputs$, with $T$ as
+the time horizon, and $\Inputs \subseteq \Reals^k$.  In this paper, we
+view $\System$ as effectively black-box equipped with a forward
+simulation map, in a simulation time-step of $\Delta_t$ from time $t$,
+allows computing the unique reachable state $\vx \in \HybridStates$
+given state-input pair $(\vx,\vu)$.
+
+% time and input parameterized reachability relation $\reach{t,
+% \vu}_\System \subseteq \HybridStates \times \HybridStates$ as follows:
+% $\simulate^{\System}:
+% (\vx, \vu, t) \mapsto \vx$, where $\vx \in \HybridStates$, $\vu \in
+% \Inputs$ and $t \in [0,T]$.
+
+% Given the hybrid state space $\vx\in\HybridStates$ and input space $u
+% \in\Inputs$ and time step $t$, it computes the unique successor state
+% $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u} \vy$.
+
+\begin{definition}[System Evolution]
+    The behavior of system $\System$ at time $t$ is described as an
+    evolution in the hybrid state-space $\HybridStates$ under the
+    effect of the input $\vu \in \Inputs$ at time $t$. The forward
+    simulation map $\simulate^\System : \HybridStates \times \Inputs
+    \times \Reals^+ \mapsto \HybridStates$, maps a state $\x$ at time
+    $t$ to a new state $\x'$ at time $t+\Delta_t$, under a constant
+    input $\vu$.  Finally, we define the time and input parameterized
+    reachability relation $\reach{t,u} \subseteq \HybridStates \times
+    \HybridStates$ such that $(\vx,\vx') \in \reach{t,u}$ iff
+    $\vx'=\simulate^{\System}(\vx,u,t)$.
+\end{definition}
+
+\mypara{System Assumptions:} We require that $\System$ can always be
+simulated forward in time with deterministic results. This in turn
+requires existence and uniqueness of trajectories over a finite time
+horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
+the vector field in each underlying hybrid mode, ruling away issues
+such as finite escape times~\cite{Meiss/2007/Differential}, and having
+well-defined semantics for simulations at switching boundaries. While
+these assumptions hold for the dynamical system examples which we
+consider in this paper, such assumptions can be restrictive for
+general control system models. When analyzing a system that does not
+obey theoretical conditions that allow deterministic simulation, we
+assume that the underlying tool used for numerical approximations of
+system behaviors uses deterministic simulation semantics.  Finally, we
+assume full observability of the system state, validate our results
+only against the $\simulate$ function (as the analytic closed-form
+representation of the system dynamics is assumed to be unavailable).
+From here on, we drop the subscript $\System$ whenever it is clear
+from the context. Also, to ease the presentation, we conflate the
+inputs with the state and denote them by $\vx$.
+
+% Abstractions for black box systems
+% graph based
+% enrichment procedure
+
 A discrete abstraction over a continuous space uses relations to
@@
 given safety property, and even other temporal properties.
+
+\todo{Add figure? graph-abs -enrich-> pwa-transition-system -bmc->
+concrete-CEX}

### 2017-05-30 prelims.tex
.

@@
-In this section, we present the model of the hybrid dynamical system
-under test, and the background for PWA relational abstractions.
-
-\mypara{System under test ($\System$)}
-
-We assume that $\System$ is a hybrid dynamical system with exogenous
-inputs modeling an embedded control system. The system state-space
-(denoted $\HybridStates$) is a subset of $\Modes \times \Reals^n$,
-where $\Modes$ is a finite set of discrete modes, and $\Reals^n$
-represents the continuous-valued state-space. For ease of exposition,
-we assume that the hybrid system is effectively a switched-mode
-dynamical system, \ie, the mode of the system is completely determined
-by its state.  An input to $\System$ is assumed to be a time-bounded
-{\em signal}, \ie, a function from $[0,T]$ to $\Inputs$, with $T$ as
-the time horizon, and $\Inputs \subseteq \Reals^k$.  In this paper, we
-view $\System$ as effectively black-box equipped with a forward
-simulation map, in a simulation time-step of $\Delta_t$ from time $t$,
-allows computing the unique reachable state $\vx \in \HybridStates$
-given state-input pair $(\vx,\vu)$.
-
-% time and input parameterized reachability relation $\reach{t,
-% \vu}_\System \subseteq \HybridStates \times \HybridStates$ as follows:
-% $\simulate^{\System}:
-% (\vx, \vu, t) \mapsto \vx$, where $\vx \in \HybridStates$, $\vu \in
-% \Inputs$ and $t \in [0,T]$.
-
-% Given the hybrid state space $\vx\in\HybridStates$ and input space $u
-% \in\Inputs$ and time step $t$, it computes the unique successor state
-% $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u} \vy$.
-
-\begin{definition}[System Evolution]
-    The behavior of system $\System$ at time $t$ is described as an
-    evolution in the hybrid state-space $\HybridStates$ under the
-    effect of the input $\vu \in \Inputs$ at time $t$. The forward
-    simulation map $\simulate^\System : \HybridStates \times \Inputs
-    \times \Reals^+ \mapsto \HybridStates$, maps a state $\x$ at time
-    $t$ to a new state $\x'$ at time $t+\Delta_t$, under a constant
-    input $\vu$.  Finally, we define the time and input parameterized
-    reachability relation $\reach{t,u} \subseteq \HybridStates \times
-    \HybridStates$ such that $(\vx,\vx') \in \reach{t,u}$ iff
-    $\vx'=\simulate^{\System}(\vx,u,t)$.
-\end{definition}
-
-\mypara{System Assumptions:} We require that $\System$ can always be
-simulated forward in time with deterministic results. This in turn
-requires existence and uniqueness of trajectories over a finite time
-horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
-the vector field in each underlying hybrid mode, ruling away issues
-such as finite escape times~\cite{Meiss/2007/Differential}, and having
-well-defined semantics for simulations at switching boundaries. While
-these assumptions hold for the dynamical system examples which we
-consider in this paper, such assumptions can be restrictive for
-general control system models. When analyzing a system that does not
-obey theoretical conditions that allow deterministic simulation, we
-assume that the underlying tool used for numerical approximations of
-system behaviors uses deterministic simulation semantics.  Finally, we
-assume full observability of the system state, validate our results
-only against the $\simulate$ function (as the analytic closed-form
-representation of the system dynamics is assumed to be unavailable).
-From here on, we drop the subscript $\System$ whenever it is clear
-from the context. Also, to ease the presentation, we conflate the
-inputs with the state and denote them by $\vx$.
-
-\subsection{Learning Dynamics using Simple Linear Regression}
-
-The approach in \cite{zutshi2014multiple} uses the forward simulation
-map to explore a fixed number of tuples in the reachability relation
-$\reach{t,\vu}$ specified by a user-provided budget.  In general, the
-budget required to adequately explore the transition space requires
-exploring an exponential (in the dimension of $\HybridStates$) number
-of such tuples. This can cause undue burden in terms of computation
-time as well as storage.  The central idea in this paper is to {\em
-summarize} several tuples as a single affine map obtained using linear
-regression.  In other words, we can replace several ``proximal''
-tuples $(\vx,\vx')$ in $\reach{t,u}$ by an affine map of the form
-$\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are matrices and
-$\delta$ is a worst-case error estimate.
-
-In statistics, linear regression is the problem of finding an affine
-function or \textit{predictor}, which can `best' summarize the
-relationship between the given set of observed input $\x$ and output
-$\x'$ vectors. The notion of `best' is formally captured using a loss
-function which can be non-linear. We use the commonly used loss
-function \textit{ordinary least squares} (OLS) for this presentation.
-
-\mypara{Ordinary Least Squares (OLS)}
-For rest of the section, we elide the discrete mode from the hybrid
-state of the system, and only focus on the continuous state. Note that
-we can do this because of our assumption that the system is a
-switched-mode dynamical system.  Suppose we have $N$ tuples of the
-form $(\x,\x')$ in the reachability relation that we have explored,
-where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn an
-affine function described as 
-\begin{equation}
-\label{eq:affinemap}
-\x' = A\x + B
-\end{equation}
-Here, $A$ is a $n\times n$ matrix, and $B$ is a $n\times 1$ matrix.
-
-If $N>n$, which is the case in the current context of \textit{finding
-the best fit}, the problem is over-determined: there are more
-equations than unknowns. Hence, a single affine function cannot be
-found which satisfies the equation \eqref{eq:affinemap}. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This choice is
-formally defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine $A$ and $B$, such
-that the least square error of the affine predictor is minimized for
-the given data set.
-
-\begin{equation}
-    \operatornamewithlimits{argmin}_{A,B}\displaystyle\sum_{i=1}^{N}{\left(\x' - (A\x + B)\right)^2}
-\label{reg}
-\end{equation}
-
-To ease the presentation, we can rewrite the above as a homogeneous
-expression by using $\hat{\x} = [\x^T\ 1]^T$ and replacing $A$ and $B$ by the vector
-$\vec{A} = [A\ B]$. The equation~\ref{reg} now becomes
-\[
-\operatornamewithlimits{argmin}_{\vec{A}}\displaystyle\sum_{i=1}^{N}{(\x' - \vec{A}^T \hat{\x})^2}
-\]
-
-The solution of OLS can be analytically computed as the Moore-Penrose
-pseudo-inverse: $\vec{A} = (X^TX)^{-1}X^T X'$, where $X$ is the matrix
-representing the horizontal stacking of all $\hat{\x}$ and $X'$
-represents the stacking of all $\x'$ values.  The details can be found
-in standard texts on learning and statistics
-\cite{friedman2001elements}. Alternative methods based on gradient
-descent can also be used to find a suitable $A$ matrix.
-
-In this fashion, we can we can use OLS to approximate its trajectories
-of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
-= \vec{A}\hat{\x}$.
-
-% where
-% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T
-% \end{bmatrix}$ and $\vb = \begin{bmatrix}b_1 \\ \vdots \\ b_n
-% \end{bmatrix}$.
-
-We note that linear regression tools are often able to provide an
-$n\times 1$ vector $\vec{\delta}$ of error intervals, where the
-$i^{th}$ entry of the form $[{\delta_i}_{\min},{\delta_i}_{\max}]$
-indicates the best and worst possible error in computation of the
-$i^{th}$ state in $\x'$.
-
-Affine maps are poor approximations for arbitrary non-linear
-functions. Hence, we use a collection of affine maps to approximate
-the local behaviors (in state-space) of the system $\System$. This
-results in a piece wise approximation, as described in the next
-section.
-
-% \mypara{Regression Analysis}
-
-% In statistics, regression is the problem of finding a
-% \textit{predictor}, which can suitably predict the relationship
-% between the given set of observed input $\x$ and output $\y$ vectors.
-% In other words, assuming that $\y$ depends on $\x$, regression
-% strategies find either a parameterized or a non-parameterized
-% prediction function to explain the dependence. We now discuss simple
-% linear regression, which is parametric in nature and searches for an
-% affine predictor. It is also called \textit{ordinary least squares}.
-
-% \mypara{Ordinary Least Squares (OLS)}
-
-% Let the data set be comprised of $N$ input and output pairs $(\x,
-% y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
-% the case in the current context of \textit{finding the best fit}, the
-% problem is over-determined; there are more equations than
-% unknowns. Hence, a single affine function cannot be found which
-% satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
-% need to find the `best' choice for $\vec{a}$ and $b$. This is formally
-% defined using a loss function. For the case of simple linear
-% regression or OLS, the loss function is the sum of squares of the
-% errors in prediction.  The task is then to determine the vector of
-% coefficients $\vec{a}$ and an offset constant $b$, such that the least
-% square error of the affine predictor is minimized for the given data
-% set.
-% \begin{equation}
-%     \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
-% \label{reg}
-% \end{equation}
-% To ease the presentation, we can rewrite the above as a homogeneous
-% expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
-% $\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
-% \[
-%     \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
-% \]
-% The solution of OLS can be analytically computed as
-% \[ Ab = (X^TX)^{-1}X^T\y\]
-% where $X$ is the matrix representing the horizontal stacking of all
-% $\hat{\x}$.
-% The details can be found in several texts on learning and statistics
-% \cite{friedman2001elements}.
-
-% Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
-% where $\x\in\reals^n$ we can use OLS to approximate its trajectories
-% at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
-% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
-% \begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
-%     global relational model for the system. The data set for the
-%     regression is a set of pairs $\setof{(\x,\x') | \x' =
-%     \simulate(\x, \Delta)}$ and can be generated on demand. Another
-%     set can be generated to compute the error $\delta$ as an interval
-%     of vectors, where each element is an interval $\delta_i \in
-%     [\delta_{min}, \delta_{max}]$.
-%
-%     Affine maps are poor
-%     approximations for arbitrary non-linear functions. Hence, we use a
-%     collection of affine maps to approximate the local behaviors (in
-%     state-space) of the system $\simulate$. This results in a piece
-%     wise approximation, as described in the next section.
-
-% \subsection{Relational Abstractions}
-% For a hybrid automaton model, they can summarize the continuous
-% dynamics of each mode using a binary relation over continuous states.
-% The resulting relations are timeless (independent) and hence valid for
-% all time as long as the mode invariant is satisfied. The relations
-% take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
-% represents one of the relational operators $=, \ge, \le, <, >$. For
-% example, an abstraction that captures the monotonicity with respect to
-% time for the differential equation $\dot{x} = 2$ is $x' > x$. The
-% abstraction capturing the relation between the set of ODEs: $\dot{x} =
-% 2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
-
-%%  Similar to timeless relational abstractions, time-aware relational
-%%  abstractions \cite{mover2013time} construct binary relations between
-%%  the current state $\x$ of the system and any future reachable state
-%%  $\x'$. The difference lies with the latter also constructing relations
-%%  between the current time $t$ and any future time $t'$. This is
-%%  achieved by a case by case analysis of the eigen structure of the
-%%  matrix $A$ (of an affine ODE). Separate abstractions are used for the
-%%  case of linear systems with constant rate, real eigen values and
-%%  complex eigenvalues.
-
 \subsection{PWA Transition System}
@@
 % map $\gm$.
+
+\section{Learning Dynamics using Simple Linear Regression}
+
+The approach in \cite{zutshi2014multiple} uses the forward simulation
+map to explore a fixed number of tuples in the reachability relation
+$\reach{t,\vu}$ specified by a user-provided budget.  In general, the
+budget required to adequately explore the transition space requires
+exploring an exponential (in the dimension of $\HybridStates$) number
+of such tuples. This can cause undue burden in terms of computation
+time as well as storage.  The central idea in this paper is to {\em
+summarize} several tuples as a single affine map obtained using linear
+regression.  In other words, we can replace several ``proximal''
+tuples $(\vx,\vx')$ in $\reach{t,u}$ by an affine map of the form
+$\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are matrices and
+$\delta$ is a worst-case error estimate.
+
+In statistics, linear regression is the problem of finding an affine
+function or \textit{predictor}, which can `best' summarize the
+relationship between the given set of observed input $\x$ and output
+$\x'$ vectors. The notion of `best' is formally captured using a loss
+function which can be non-linear. We use the commonly used loss
+function \textit{ordinary least squares} (OLS) for this presentation.
+
+\mypara{Ordinary Least Squares (OLS)}
+For rest of the section, we elide the discrete mode from the hybrid
+state of the system, and only focus on the continuous state. Note that
+we can do this because of our assumption that the system is a
+switched-mode dynamical system.  Suppose we have $N$ tuples of the
+form $(\x,\x')$ in the reachability relation that we have explored,
+where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn an
+affine function described as 
+\begin{equation}
+\label{eq:affinemap}
+\x' = A\x + B
+\end{equation}
+Here, $A$ is a $n\times n$ matrix, and $B$ is a $n\times 1$ matrix.
+
+If $N>n$, which is the case in the current context of \textit{finding
+the best fit}, the problem is over-determined: there are more
+equations than unknowns. Hence, a single affine function cannot be
+found which satisfies the equation \eqref{eq:affinemap}. Instead, we
+need to find the `best' choice for $\vec{a}$ and $b$. This choice is
+formally defined using a loss function. For the case of simple linear
+regression or OLS, the loss function is the sum of squares of the
+errors in prediction.  The task is then to determine $A$ and $B$, such
+that the least square error of the affine predictor is minimized for
+the given data set.
+
+\begin{equation}
+    \operatornamewithlimits{argmin}_{A,B}\displaystyle\sum_{i=1}^{N}{\left(\x' - (A\x + B)\right)^2}
+\label{reg}
+\end{equation}
+
+To ease the presentation, we can rewrite the above as a homogeneous
+expression by using $\hat{\x} = [\x^T\ 1]^T$ and replacing $A$ and $B$ by the vector
+$\vec{A} = [A\ B]$. The equation~\ref{reg} now becomes
+\[
+\operatornamewithlimits{argmin}_{\vec{A}}\displaystyle\sum_{i=1}^{N}{(\x' - \vec{A}^T \hat{\x})^2}
+\]
+
+The solution of OLS can be analytically computed as the Moore-Penrose
+pseudo-inverse: $\vec{A} = (X^TX)^{-1}X^T X'$, where $X$ is the matrix
+representing the horizontal stacking of all $\hat{\x}$ and $X'$
+represents the stacking of all $\x'$ values.  The details can be found
+in standard texts on learning and statistics
+\cite{friedman2001elements}. Alternative methods based on gradient
+descent can also be used to find a suitable $A$ matrix.
+
+In this fashion, we can we can use OLS to approximate its trajectories
+of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
+= \vec{A}\hat{\x}$.
+
+% where
+% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T
+% \end{bmatrix}$ and $\vb = \begin{bmatrix}b_1 \\ \vdots \\ b_n
+% \end{bmatrix}$.
+
+We note that linear regression tools are often able to provide an
+$n\times 1$ vector $\vec{\delta}$ of error intervals, where the
+$i^{th}$ entry of the form $[{\delta_i}_{\min},{\delta_i}_{\max}]$
+indicates the best and worst possible error in computation of the
+$i^{th}$ state in $\x'$.
+
+Affine maps are poor approximations for arbitrary non-linear
+functions. Hence, we use a collection of affine maps to approximate
+the local behaviors (in state-space) of the system $\System$. This
+results in a piece wise approximation, as described in the next
+section.
+
+% \mypara{Regression Analysis}
+
+% In statistics, regression is the problem of finding a
+% \textit{predictor}, which can suitably predict the relationship
+% between the given set of observed input $\x$ and output $\y$ vectors.
+% In other words, assuming that $\y$ depends on $\x$, regression
+% strategies find either a parameterized or a non-parameterized
+% prediction function to explain the dependence. We now discuss simple
+% linear regression, which is parametric in nature and searches for an
+% affine predictor. It is also called \textit{ordinary least squares}.
+
+% \mypara{Ordinary Least Squares (OLS)}
+
+% Let the data set be comprised of $N$ input and output pairs $(\x,
+% y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
+% the case in the current context of \textit{finding the best fit}, the
+% problem is over-determined; there are more equations than
+% unknowns. Hence, a single affine function cannot be found which
+% satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
+% need to find the `best' choice for $\vec{a}$ and $b$. This is formally
+% defined using a loss function. For the case of simple linear
+% regression or OLS, the loss function is the sum of squares of the
+% errors in prediction.  The task is then to determine the vector of
+% coefficients $\vec{a}$ and an offset constant $b$, such that the least
+% square error of the affine predictor is minimized for the given data
+% set.
+% \begin{equation}
+%     \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
+% \label{reg}
+% \end{equation}
+% To ease the presentation, we can rewrite the above as a homogeneous
+% expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
+% $\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
+% \[
+%     \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
+% \]
+% The solution of OLS can be analytically computed as
+% \[ Ab = (X^TX)^{-1}X^T\y\]
+% where $X$ is the matrix representing the horizontal stacking of all
+% $\hat{\x}$.
+% The details can be found in several texts on learning and statistics
+% \cite{friedman2001elements}.
+
+% Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
+% where $\x\in\reals^n$ we can use OLS to approximate its trajectories
+% at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
+% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
+% \begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
+%     global relational model for the system. The data set for the
+%     regression is a set of pairs $\setof{(\x,\x') | \x' =
+%     \simulate(\x, \Delta)}$ and can be generated on demand. Another
+%     set can be generated to compute the error $\delta$ as an interval
+%     of vectors, where each element is an interval $\delta_i \in
+%     [\delta_{min}, \delta_{max}]$.
+%
+%     Affine maps are poor
+%     approximations for arbitrary non-linear functions. Hence, we use a
+%     collection of affine maps to approximate the local behaviors (in
+%     state-space) of the system $\simulate$. This results in a piece
+%     wise approximation, as described in the next section.
+
+% \subsection{Relational Abstractions}
+% For a hybrid automaton model, they can summarize the continuous
+% dynamics of each mode using a binary relation over continuous states.
+% The resulting relations are timeless (independent) and hence valid for
+% all time as long as the mode invariant is satisfied. The relations
+% take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
+% represents one of the relational operators $=, \ge, \le, <, >$. For
+% example, an abstraction that captures the monotonicity with respect to
+% time for the differential equation $\dot{x} = 2$ is $x' > x$. The
+% abstraction capturing the relation between the set of ODEs: $\dot{x} =
+% 2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
+
+%%  Similar to timeless relational abstractions, time-aware relational
+%%  abstractions \cite{mover2013time} construct binary relations between
+%%  the current state $\x$ of the system and any future reachable state
+%%  $\x'$. The difference lies with the latter also constructing relations
+%%  between the current time $t$ and any future time $t'$. This is
+%%  achieved by a case by case analysis of the eigen structure of the
+%%  matrix $A$ (of an affine ODE). Separate abstractions are used for the
+%%  case of linear systems with constant rate, real eigen values and
+%%  complex eigenvalues.

### 2017-05-31 graph_abs.tex
rearranged sections

@@
 
-\mypara{System under test ($\System$)}
+\subsection{System under test ($\System$)}
 
@@
 
-\todo{Add figure? graph-abs -enrich-> pwa-transition-system -bmc->
-concrete-CEX}
+\begin{figure}[!htbp]
+\begin{center}
+\tikzstyle{line} = [thick]
+\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
+\tikzstyle{block} = [rectangle, minimum width=0.5cm, minimum height=1.0cm, text centered, draw=black, align=center]%, fill=blue!25]
+\begin{tikzpicture}[scale=1.0, on grid, auto]
+
+    \node (G) at (-4,0) {$G$};
+    \node (GR) at (0,0) {$G^R$};
+    \node (CEX) at (4,0) {$CEx$};
+
+    \node (enrich) [block] at (-2, 0) {\scriptsize Enrich using \\ \scriptsize affine maps};
+    \node (BMC) [block] at (2, 0) {\scriptsize Find CEx \\ \scriptsize using BMC};
+
+    %\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
+
+    %\draw [line] (-2.0,-0.5) rectangle (-1.0,0.5);
+    \draw[arw] (G.east) -- (enrich.west);
+    \draw[arw] (enrich.east) -- (GR.west);
+    \draw[arw] (GR.east) -- (BMC.west);
+    \draw[arw] (BMC.east) -- (CEX.west);
+
+\end{tikzpicture}
+\end{center}
+\vspace*{-.3cm}
+\caption{Finding counter-examples (CEx) in a graph abstraction using PWA models.}
+\label{fig:overview}
+\vspace*{-.3cm}
+\end{figure}

### 2017-05-31 graph_abs.tex
.

@@
     effect of the input $\vu \in \Inputs$ at time $t$. The forward
-    simulation map $\simulate^\System : \HybridStates \times \Inputs
-    \times \Reals^+ \mapsto \HybridStates$, maps a state $\x$ at time
-    $t$ to a new state $\x'$ at time $t+\Delta_t$, under a constant
+    simulation map $\mathsc{sim}^\System : \HybridStates \times \Inputs
+    \times \Reals^+ \mapsto \HybridStates$, maps a state $\x\in\HybridStates$ at time
+    $t$ to a new state $\x'\in\HybridStates$ at time $t+\Delta\in\Reals^+$, under a constant
     input $\vu$.  Finally, we define the time and input parameterized
     reachability relation $\reach{t,u} \subseteq \HybridStates \times
-    \HybridStates$ such that $(\vx,\vx') \in \reach{t,u}$ iff
-    $\vx'=\simulate^{\System}(\vx,u,t)$.
+    \HybridStates$ such that
+    \begin{equation}
+        (\vx,\vx') \in \reach{t,u} \;\text{iff}\; \vx'=\mathsc{sim}^{\System}(\vx,u,t).
+    \end{equation}
 \end{definition}
@@
 assume full observability of the system state, validate our results
-only against the $\simulate$ function (as the analytic closed-form
+only against the $\mathsc{sim}^\System$ function (as the analytic closed-form
 representation of the system dynamics is assumed to be unavailable).
@@
 
-% Abstractions for black box systems
-% graph based
-% enrichment procedure
+\subsection{Graph Abstraction}
 
-A discrete abstraction over a continuous space uses relations to
-describe the behavior of the system over time. Such relations abstract
-the underlying relation between concrete states, and concrete paths
-can no longer be constructed directly. Instead, search procedures
-like Counter-Example Guided Refinement must be used to find concrete
-paths. In general, such search techniques are exponential with respect to
+We base our methodology on the implicit discrete abstraction for
+black-box dynamical systems~\cite{zutshi2014multiple}. The authors
+showed how this abstraction is a graph abstraction, and reduced the
+search for abstract violations to finding paths in a graph.
+
+% Such an abstraction is defined by partitioning the
+% state-space of the system using a grid, and hence each abstract state
+% is a hyper-rectangle. The relations between abstract states are
+% obtained by suitably lifting $(\vx,\vx') \in\reach{t,u}$. Considering
+% the abstract states as vertices of a graph and the relations as edges
+% between nodes, the problem of finding an abstract counter-example
+% (CEx) can be cast as finding a path in the graph. Such an abstraction
+% can be explored on-the-fly by enumerating the abstract relations
+% between a pair of states.
+
+\mypara{A discrete abstraction} over a continuous space uses relations
+to describe the behavior of the system over time. Such relations
+abstract the underlying relation between concrete states, and concrete
+paths can no longer be constructed directly. Instead, search
+procedures like Counter-Example Guided Refinement (CEGAR) must be used
+to find concrete paths. In general, such search techniques are based
+on refinement of the state-space and exponential with respect to
 computational resources and memory. As an alternative, we introduce
 data driven `enrichment' of abstractions to approximate the concrete
-relations underlying the abstract relations. Informally, we use linear
-regression to compute affine maps associated with the relations
+relations underlying the abstract relations.  Specifically, we use
+linear regression to compute affine maps associated with the relations
 between two abstract states. We then show how associating the computed
@@
 transition system.
-
-\subsection{Graph Abstraction}
 

### 2017-05-31 prelims.tex
.

@@
 
-Temporal properties of pwa transitions systems can be checked using
+Temporal properties of PWA transitions systems can be checked using
 off-the-shelf model checkers like SAL~\cite{SAL-SRI}, which can reason
@@
 problem instance to achieve better efficiency~\cite{shoukry2017smc}.
-We now show how relational pwa abstractions can be viewed as pwa
+We now show how relational PWA abstractions can be viewed as PWA
 transition systems, and how the problem of falsification can

### 2017-05-31 graph_abs.tex
.

@@
 
-The abstraction of the hybrid dynamical system is a a
-partitioning of the hybrid state space $\HybridStates$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
-pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
-their union $\bigcup_{C_i\in\Cells}C_i$ is the state space $\HybridStates$. The partitions over
-the continuous state space are implicitly defined using a quantization
-function $Q_\epsilon$ parameterized by the precision $\epsilon$.
-This further defines an equivalence relations such that $(\vx_i)
-\equiv (\vx_j) \;\text{iff}\; Q_\epsilon(\vx_i) = Q_\epsilon(\vx_j)$.
-In other words, $(\vx_i) \in C_i \; \text{iff} \; C_i = Q_\epsilon(\vx_i)$.
+The abstraction of the hybrid dynamical system is a a partitioning of
+the hybrid state space $\HybridStates$ into hyper-rectangles or cells
+$C_i \in \Cells$. The cells are pairwise disjoint $C_i \cap C_j \neq
+\emptyset$ iff $i \neq j$ and their union $\bigcup_{C_i\in\Cells}C_i$
+is the state space $\HybridStates$. The partitions over the continuous
+state space are implicitly defined using a quantization function
+$Q_\epsilon$ parameterized by the precision $\epsilon$.  This further
+defines an equivalence relations such that $(\vx_i) \equiv (\vx_j)
+\;\text{iff}\; Q_\epsilon(\vx_i) = Q_\epsilon(\vx_j)$.  In other
+words, $(\vx_i) \in C_i \; \text{iff} \; C_i = Q_\epsilon(\vx_i)$. The
+abstraction can be interpreted as a graph and explored using fixed
+time $(\Delta\in\Reals^+)$ relations obtained using
+$\simulate(\vx)$~\footnote{$\simulate(\vx)$ is a fixed
+parameterization of $\mathsc{sim}(\vx,\Delta)$}.
 

### 2017-06-05 graph_abs.tex
incomplete commit of new abstraction section

@@
 
-\subsection{System under test ($\System$)}
+\subsection{System-under-test ($\System$)}
 
-We assume that $\System$ is a hybrid dynamical system with exogenous
-inputs modeling an embedded control system. The system state-space
-(denoted $\HybridStates$) is a subset of $\Modes \times \Reals^n$,
-where $\Modes$ is a finite set of discrete modes, and $\Reals^n$
-represents the continuous-valued state-space. For ease of exposition,
-we assume that the hybrid system is effectively a switched-mode
-dynamical system, \ie, the mode of the system is completely determined
-by its state.  An input to $\System$ is assumed to be a time-bounded
-{\em signal}, \ie, a function from $[0,T]$ to $\Inputs$, with $T$ as
-the time horizon, and $\Inputs \subseteq \Reals^k$.  In this paper, we
-view $\System$ as effectively black-box equipped with a forward
-simulation map, in a simulation time-step of $\Delta_t$ from time $t$,
-allows computing the unique reachable state $\vx \in \HybridStates$
-given state-input pair $(\vx,\vu)$.
+We assume that the system-under-test is a hybrid dynamical system, \ie, a
+system which has discrete modes and a continuous state-space, and which
+evolvution is either (1) continuous in time according to the dynamics
+associated with its current discrete mode (typically described using
+differential equations over its continuous state), or (2) through
+discrete jumps that possibly change the discrete mode of the system and
+possibly reset its continuous state to some value.  However, we assume
+that the system is {\em effectively black-box}. This means that we do not
+assume any knowledge of the symbolic dynamical equations describing the
+system behavior. We do assume some knowledge of the system: an {\em
+interface} view, in which we are allowed to stimulate the system with
+inputs and observe its outputs.
+
+We formally define a system-under-test $\System$ as a tuple $(\Modes,
+\ContStates, \initmode, \initstate, \Delta, \Inputs, \Outputs)$.  Here,
+system state-space (denoted $\HybridStates$) is a subset of $\Modes \times
+\ContStates$, where $\Modes$ is a finite set of discrete modes, and
+$\ContStates$ is a compact subset of $\Reals^n$ representing the set of
+continuous states.  We assume that the system is {\em initialized} to an
+initial mode $\initmode$ and an initial state $\initstate$ at time $t=0$.
+We assume that the system has $m$ exogenuous inputs which take values from
+the set $\Inputs$ which is a compact subset of $\Reals^m$. We assume that
+the system has $\ell \le n$ outputs, where $\Outputs \subseteq
+\HybridStates$. In this paper, we assume that the entire system state is
+observable.  In other words, $\Outputs = \HybridStates$.
+
+Finally, we restrict our attention to switched-mode systems, \ie, those in
+which there is a partition imposed on $\ContStates$, and there is a
+bijective map between each element $\ContStates_i$ in the partition and
+each mode in $\Modes$.
+
+The operational semantics of $\System$ is defined in terms of a
+forward-simulation map $\simmap_\Delta$, which is a function that maps a
+time $t$, the current state of the system, and an input value in $\Inputs$
+to the state of the system at time $t+\Delta$.
+
+% We assume that $\System$ is a hybrid dynamical system with exogenous
+% inputs modeling an embedded control system. 
+% 
+% we assume that the hybrid system is effectively a switched-mode
+% dynamical system, \ie, the mode of the system is completely determined
+% by its state.  An input to $\System$ is assumed to be a time-bounded
+% {\em signal}, \ie, a function from $[0,T]$ to $\Inputs$, with $T$ as
+% the time horizon, and $\Inputs \subseteq \Reals^k$.  In this paper, we
+% view $\System$ as effectively black-box equipped with a forward
+% simulation map, in a simulation time-step of $\Delta_t$ from time $t$,
+% allows computing the unique reachable state $\vx \in \HybridStates$
+% given state-input pair $(\vx,\vu)$.
 

### 2017-06-08 graph_abs.tex
transition system abstraction

@@
 We formally define a system-under-test $\System$ as a tuple $(\Modes,
-\ContStates, \initmode, \initstate, \Delta, \Inputs, \Outputs)$.  Here,
-system state-space (denoted $\HybridStates$) is a subset of $\Modes \times
-\ContStates$, where $\Modes$ is a finite set of discrete modes, and
-$\ContStates$ is a compact subset of $\Reals^n$ representing the set of
-continuous states.  We assume that the system is {\em initialized} to an
-initial mode $\initmode$ and an initial state $\initstate$ at time $t=0$.
-We assume that the system has $m$ exogenuous inputs which take values from
-the set $\Inputs$ which is a compact subset of $\Reals^m$. We assume that
-the system has $\ell \le n$ outputs, where $\Outputs \subseteq
-\HybridStates$. In this paper, we assume that the entire system state is
-observable.  In other words, $\Outputs = \HybridStates$.
+\ContStates, \InitModes, \InitStates, \Delta, \Inputs, \Outputs)$.
+Here, system state-space (denoted $\HybridStates$) is a subset of
+$\Modes \times \ContStates$, where $\Modes$ is a finite set of
+discrete modes, and $\ContStates$ is a compact subset of $\Reals^n$
+representing the set of continuous states.  The sets $\InitModes
+\subseteq \Modes$ and $\InitStates \subseteq \ContStates$ represent
+initial modes and states respectively.  We assume that the system is
+{\em initialized} to some initial mode $\initmode \in \InitModes$ and
+an initial state $\initstate \in \InitStates$ at time $t=0$.  We
+assume that the system has $m$ exogenuous inputs which take values
+from the set $\Inputs$ which is a compact subset of $\Reals^m$. We
+assume that the system has $\ell \le n$ outputs, where $\Outputs
+\subseteq \HybridStates$. In this paper, we assume that the entire
+system state is observable.  In other words, $\Outputs =
+\HybridStates$.
 
@@
 bijective map between each element $\ContStates_i$ in the partition and
-each mode in $\Modes$.
+each mode in $\Modes$. Thus, we can omit the discrete modes from the
+system description, and set $\HybridStates$ = $\ContStates$.
 
-The operational semantics of $\System$ is defined in terms of a
-forward-simulation map $\simmap_\Delta$, which is a function that maps a
-time $t$, the current state of the system, and an input value in $\Inputs$
-to the state of the system at time $t+\Delta$.
+The operational semantics of $\System$ is defined in terms of a {\em
+forward simulation map}\footnote{We require that $\System$ can always
+be simulated forward in time with deterministic results. A few
+technical conditions are required on the system dynamics are required
+to ensure that this assumption is valid, see for
+example~\cite{Meiss/2007/Differential} and \cite{goebel2012hybrid} for
+further details. In practice, such conditions are usually impossible
+to check for a black-box model. We then assume that the black-box
+model is described as a model in a deterministic simulation framework
+such as Simulink\textregistered.} $\simmap_\Delta$, which is a
+function that maps a time $t$, the current state of the system, and an
+input value in $\Inputs$ to the state of the system at time
+$t+\Delta$, \ie, $\simmap_\Delta$ is a function from $\ContStates
+\times \Inputs \times \Reals^{\ge 0}$ to $\ContStates$.
 
-% We assume that $\System$ is a hybrid dynamical system with exogenous
+Let $\inputsignal$ = $(t_0, u_0)$, $\ldots$, $(t_N,u_N)$ be a
+timed-sequence of inputs, where $t_0 = 0$, and for all $i > 0$, $t_i -
+t_{i-1} = \Delta$, and $u_i \in \Inputs$.  We can extend the forward
+simulation map to a sequence of states to describe the application of
+an input signal to excite the system as follows:
+$\simmap_\Delta(\initstate, \mathbf{u} = \contState_N$, if for all $i  \in
+[0,N-1]$, $\simmap_\Delta(\contState_i, u_i,t_i) = \contState_{i+1}$.
+
+% An alternative view of the forward simulation-map is as a
+% parameterized reachability relation $\reach_\Delta{t,u} \subseteq
+% \ContStates \times \ContStates$ defined as below:
+% \[
+% (\contState,\contState') \in \reach_\Delta{t,u}\text{ iff } \contState' = \simmap_\Delta(\contState,t,u).
+% \]
+
 % inputs modeling an embedded control system. 
@@
 
-\begin{definition}[System Evolution]
-    The behavior of system $\System$ at time $t$ is described as an
-    evolution in the hybrid state-space $\HybridStates$ under the
-    effect of the input $\vu \in \Inputs$ at time $t$. The forward
-    simulation map $\mathsc{sim}^\System : \HybridStates \times \Inputs
-    \times \Reals^+ \mapsto \HybridStates$, maps a state $\x\in\HybridStates$ at time
-    $t$ to a new state $\x'\in\HybridStates$ at time $t+\Delta\in\Reals^+$, under a constant
-    input $\vu$.  Finally, we define the time and input parameterized
-    reachability relation $\reach{t,u} \subseteq \HybridStates \times
-    \HybridStates$ such that
-    \begin{equation}
-        (\vx,\vx') \in \reach{t,u} \;\text{iff}\; \vx'=\mathsc{sim}^{\System}(\vx,u,t).
-    \end{equation}
-\end{definition}
+% \begin{definition}[System Evolution]
+%     The behavior of system $\System$ at time $t$ is described as an
+%     evolution in the hybrid state-space $\HybridStates$ under the
+%     effect of the input $\vu \in \Inputs$ at time $t$. The forward
+%     simulation map $\mathsc{sim}^\System : \HybridStates \times \Inputs
+%     \times \Reals^+ \mapsto \HybridStates$, maps a state $\x\in\HybridStates$ at time
+%     $t$ to a new state $\x'\in\HybridStates$ at time $t+\Delta\in\Reals^+$, under a constant
+%     input $\vu$.  Finally, we define the time and input parameterized
+%     reachability relation $\reach{t,u} \subseteq \HybridStates \times
+%     \HybridStates$ such that
+%     \begin{equation}
+%         (\vx,\vx') \in \reach{t,u} \;\text{iff}\; \vx'=\mathsc{sim}^{\System}(\vx,u,t).
+%     \end{equation}
+% \end{definition}
 
-\mypara{System Assumptions:} We require that $\System$ can always be
-simulated forward in time with deterministic results. This in turn
-requires existence and uniqueness of trajectories over a finite time
-horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
-the vector field in each underlying hybrid mode, ruling away issues
-such as finite escape times~\cite{Meiss/2007/Differential}, and having
-well-defined semantics for simulations at switching boundaries. While
-these assumptions hold for the dynamical system examples which we
-consider in this paper, such assumptions can be restrictive for
-general control system models. When analyzing a system that does not
-obey theoretical conditions that allow deterministic simulation, we
-assume that the underlying tool used for numerical approximations of
-system behaviors uses deterministic simulation semantics.  Finally, we
-assume full observability of the system state, validate our results
-only against the $\mathsc{sim}^\System$ function (as the analytic closed-form
-representation of the system dynamics is assumed to be unavailable).
-From here on, we drop the subscript $\System$ whenever it is clear
-from the context. Also, to ease the presentation, we conflate the
-inputs with the state and denote them by $\vx$.
-
+% \mypara{System Assumptions:} 
+% We require that $\System$ can always be
+% simulated forward in time with deterministic results. This in turn
+% requires existence and uniqueness of trajectories over a finite time
+% horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
+% the vector field in each underlying hybrid mode, ruling away issues
+% such as finite escape times~\cite{Meiss/2007/Differential}, and having
+% well-defined semantics for simulations at switching boundaries. While
+% these assumptions hold for the dynamical system examples which we
+% consider in this paper, such assumptions can be restrictive for
+% general control system models. When analyzing a system that does not
+% obey theoretical conditions that allow deterministic simulation, we
+% assume that the underlying tool used for numerical approximations of
+% system behaviors uses deterministic simulation semantics.  Finally, we
+% assume full observability of the system state, validate our results
+% only against the $\mathsc{sim}^\System$ function (as the analytic closed-form
+% representation of the system dynamics is assumed to be unavailable).
+% From here on, we drop the subscript $\System$ whenever it is clear
+% from the context. Also, to ease the presentation, we conflate the
+% inputs with the state and denote them by $\vx$.
+% 
 \subsection{Graph Abstraction}
 
-We base our methodology on the implicit discrete abstraction for
-black-box dynamical systems~\cite{zutshi2014multiple}. The authors
-showed how this abstraction is a graph abstraction, and reduced the
-search for abstract violations to finding paths in a graph.
+We observe that the parameterized reachability relation described in
+the previous section induces an existential discrete abstraction of
+the system similar to the one considered in \cite{zutshi2014multiple}.
+We briefly review this discrete abstraction and show how falsification
+of systems reduces to a path-search in the resulting transition
+system. We make the observation that this discrete abstraction is
+actually a Kripke structure, \ie, a transition system defined as the
+tuple $(\Cells, \CellEdges, \CellLabelingFunction)$, where $\Cells$ is
+a finite set of cells, $\CellEdges \subseteq \Cells \times \Cells$ is
+an unlabeled transition relation, and $\CellLabelingFunction$ is a
+function that assigns one or more labels to the cells from a finite
+set of labels.
 
-% Such an abstraction is defined by partitioning the
-% state-space of the system using a grid, and hence each abstract state
-% is a hyper-rectangle. The relations between abstract states are
-% obtained by suitably lifting $(\vx,\vx') \in\reach{t,u}$. Considering
-% the abstract states as vertices of a graph and the relations as edges
-% between nodes, the problem of finding an abstract counter-example
-% (CEx) can be cast as finding a path in the graph. Such an abstraction
-% can be explored on-the-fly by enumerating the abstract relations
-% between a pair of states.
+In what follows, we review an $\epsilon$-precision transition system
+abstraction of a system $\System$ (denoted as
+$\TSAbstraction{\epsilon}{\System}$) as follows. We define a
+quantization function $\quant_\epsilon$ that maps a state $\contState$ in
+$\ContStates$ to $10^{-\epsilon}*\lfloor \contState*10^\epsilon \rfloor$.
+This function induces an equivalence relation $\contState_i
+\equiv_\epsilon \contState_j$ iff $\quant_\epsilon(\contState_i) =
+\quant_\epsilon(\contState_j)$. Now consider the quotient structure
+$\ContStates/\equiv_\epsilon$. We treat each equivalence class in the
+quotient structure as a cell. We discover the transitions in the
+abstraction using simulations. Given a numerical simulator for the
+model with a forward simulation map $\simmap_\Delta$, we add an edge
+from cell $C$ to cell $C'$ iff:
+\begin{itemize}
+\item
+$\simmap_\Delta(\contState, u, t)$ = $\contState'$, 
+\item
+$\quant_\epsilon(\contState)$ = $C$, and,
+\item
+$\quant_\epsilon(\contState') = C'$.
+\end{itemize}
 
-\mypara{A discrete abstraction} over a continuous space uses relations
-to describe the behavior of the system over time. Such relations
-abstract the underlying relation between concrete states, and concrete
-paths can no longer be constructed directly. Instead, search
-procedures like Counter-Example Guided Refinement (CEGAR) must be used
-to find concrete paths. In general, such search techniques are based
-on refinement of the state-space and exponential with respect to
-computational resources and memory. As an alternative, we introduce
-data driven `enrichment' of abstractions to approximate the concrete
-relations underlying the abstract relations.  Specifically, we use
-linear regression to compute affine maps associated with the relations
-between two abstract states. We then show how associating the computed
-maps with the relations transforms a discrete abstraction into a PWA
-transition system.
+While our framework is general enough to reason about arbitrary
+temporal logic properties of trajectories, we restrict our attention
+to safety properties, where we assume that there is a specified set of
+states $\UnsafeStates \subseteq \ContStates$ that are unsafe.  Thus,
+our label set has only two elements:
+$\setof{\init,\unsafe}$. We then define $\CellLabelingFunction$ as
+follows: 
+\begin{equation*}
+\CellLabelingFunction(C) = \left\{
+\begin{array}{l@{\hspace{1em}}l}
+\init   & \exists \contState \in \InitStates\text{ s.t.  }\quant_\epsilon(\contState) = C  \\
+\unsafe & \exists \contState \in \UnsafeStates\text{ s.t.  }\quant_\epsilon(\contState) = C  \\
+\{\}    & \text{otherwise.}    
+\end{array}\right.
+\end{equation*}
+
+\subsection{CEGAR for falsifying $\TSAbstraction{\epsilon}{\System}$}
+
+We remark that our abstraction is {\em neither an overapproximation
+nor an underapproximation} of the continuous-time continuous-space
+reachability relation between states. The purpose of this abstraction
+is to discover falsifying trajectories; thus, this technique explores
+the abstraction to get abstract falsifying trajectories starting at an
+initial abstract state that lead to error, and to confirm that such an
+abstract trajectory is not spurious, it performs a concretization step
+by performing an actual simulation of the system. If the abstract
+counterexample trajectories are found to be spurious, then
+$\TSAbstraction{\epsilon}{\System}$ can be refined, basically by
+increasing $\epsilon$. We summarize this procedure in
+Algorithm~\ref{algo:cegar}.
+
+The biggest drawback of Algorithm~\ref{algo:cegar} is that with
+increasing $\epsilon$, the size of $\Cells$ in
+$\TSAbstraction{\epsilon}{\System}$ increases exponentially.  Thus,
+every subsequent run of the falsification algorithm is slower. In the
+next section, we propose an alternative abstraction that enriches the
+basic abstraction proposed here by adding labels to the transitions.
+We can then perform refinement by enriching the transition relation,
+rather than increasing $\epsilon$.
+
+% The abstraction of the hybrid dynamical system is a partitioning of
+% the hybrid state space $\HybridStates$ into hyper-rectangles or cells
+% $C_i \in \Cells$. The cells are pairwise disjoint $C_i \cap C_j \neq
+% \emptyset$ iff $i \neq j$ and their union $\bigcup_{C_i\in\Cells}C_i$
+% is the state space $\HybridStates$. The partitions over the continuous
+% state space are implicitly defined using a quantization function
+% $\quant_\epsilon$ parameterized by the precision $\epsilon$.  This further
+% defines an equivalence relations such that $(\vx_i) \equiv (\vx_j)
+% \;\text{iff}\; \quant_\epsilon(\vx_i) = \quant_\epsilon(\vx_j)$.  In other
+% words, $(\vx_i) \in C_i \; \text{iff} \; C_i = \quant_\epsilon(\vx_i)$. The
+% abstraction can be interpreted as a graph and explored using fixed
+% time $(\Delta\in\Reals^+)$ relations obtained using
+% $\simulate(\vx)$~\footnote{$\simulate(\vx)$ is a fixed
+% parameterization of $\mathsc{sim}(\vx,\Delta)$}.
+% 
+% % Such an abstraction is defined by partitioning the
+% % state-space of the system using a grid, and hence each abstract state
+% % is a hyper-rectangle. The relations between abstract states are
+% % obtained by suitably lifting $(\vx,\vx') \in\reach{t,u}$. Considering
+% % the abstract states as vertices of a graph and the relations as edges
+% % between nodes, the problem of finding an abstract counter-example
+% % (CEx) can be cast as finding a path in the graph. Such an abstraction
+% % can be explored on-the-fly by enumerating the abstract relations
+% % between a pair of states.
+% 
+% \mypara{A discrete abstraction} over a continuous space uses relations
+% to describe the behavior of the system over time. Such relations
+% abstract the underlying dense-time trajectories of the given dynamical
+% system to a transition system.
+
+% relation between concrete states, and concrete
+% paths can no longer be constructed directly. Instead, search
+% procedures like Counter-Example Guided Refinement (CEGAR) must be used
+% to find concrete paths. In general, such search techniques are based
+% on refinement of the state-space and exponential with respect to
+% computational resources and memory. As an alternative, we introduce
+% data driven `enrichment' of abstractions to approximate the concrete
+% relations underlying the abstract relations.  Specifically, we use
+% linear regression to compute affine maps associated with the relations
+% between two abstract states. We then show how associating the computed
+% maps with the relations transforms a discrete abstraction into a PWA
+% transition system.
 
@@
 % the continuous input-state space are implicitly defined using a quantization
-% function $Q_\epsilon$ parameterized by the precision $\epsilon$.
+% function $\quant_\epsilon$ parameterized by the precision $\epsilon$.
 % This further defines an equivalence relations such that
 % \[
-%     (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; Q_\epsilon(\vx_i,
-%     \vu_i) = Q_\epsilon(\vx_j, \vu_j)
+%     (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; \quant_\epsilon(\vx_i,
+%     \vu_i) = \quant_\epsilon(\vx_j, \vu_j)
 % \]
 % In other words, $(\vx_i, \vu_i) \in C_i \; \text{iff} \; C_i =
-% Q_\epsilon(\vx_i, \vu_i)$.
+% \quant_\epsilon(\vx_i, \vu_i)$.
 
 % \begin{definition}[Graph Abstraction]
-%     The quantization function $Q_\epsilon$ induces a DAG on the discrete
+%     The quantization function $\quant_\epsilon$ induces a DAG on the discrete
 %     abstraction given by a graph $G(V, E)$, where each vertex
@@
 
-The abstraction of the hybrid dynamical system is a a partitioning of
-the hybrid state space $\HybridStates$ into hyper-rectangles or cells
-$C_i \in \Cells$. The cells are pairwise disjoint $C_i \cap C_j \neq
-\emptyset$ iff $i \neq j$ and their union $\bigcup_{C_i\in\Cells}C_i$
-is the state space $\HybridStates$. The partitions over the continuous
-state space are implicitly defined using a quantization function
-$Q_\epsilon$ parameterized by the precision $\epsilon$.  This further
-defines an equivalence relations such that $(\vx_i) \equiv (\vx_j)
-\;\text{iff}\; Q_\epsilon(\vx_i) = Q_\epsilon(\vx_j)$.  In other
-words, $(\vx_i) \in C_i \; \text{iff} \; C_i = Q_\epsilon(\vx_i)$. The
-abstraction can be interpreted as a graph and explored using fixed
-time $(\Delta\in\Reals^+)$ relations obtained using
-$\simulate(\vx)$~\footnote{$\simulate(\vx)$ is a fixed
-parameterization of $\mathsc{sim}(\vx,\Delta)$}.
-
-\begin{definition}[Graph Abstraction]
-    The quantization function $Q_\epsilon$ induces a directed acyclic
-    graph (DAG) on the discrete
-    abstraction given by a graph $G(V, E)$, where each vertex
-    corresponds to a cell $V_i = C$, and an edge exists between two
-    cells $(C, C') \in E_i$ iff there exist state-input pairs
-    $(\vx) \in C$, $(\vx') \in C'$ such that $\vx' =
-    \simulate(\vx)$.
-\end{definition}
-
-Given a numerical simulator $\simulate$ and an initial abstraction
-defined by a quantization function $\quant_\epsilon$ and a time step
-$\Delta$, we pick samples and simulate them to explore the abstraction
-on the fly to obtain a graph $G$, which has a finite number of
-abstract states $C$ (or cells) and edges $(C,C')$ iff $C
-\areach{\Delta} C'$.
+% \begin{definition}[Graph Abstraction]
+%     The quantization function $\quant_\epsilon$ induces a directed acyclic
+%     graph (DAG) on the discrete
+%     abstraction given by a graph $G(V, E)$, where each vertex
+%     corresponds to a cell $V_i = C$, and an edge exists between two
+%     cells $(C, C') \in E_i$ iff there exist state-input pairs
+%     $(\vx) \in C$, $(\vx') \in C'$ such that $\vx' =
+%     \simulate(\vx)$.
+% \end{definition}
+% 
+% 
+% 
+% Given a numerical simulator $\simulate$ and an initial abstraction
+% defined by a quantization function $\quant_\epsilon$ and a time step
+% $\Delta$, we pick samples and simulate them to explore the abstraction
+% on the fly to obtain a graph $G$, which has a finite number of
+% abstract states $C$ (or cells) and edges $(C,C')$ iff $C
+% \areach{\Delta} C'$.
 
@@
 
-Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
-we use the generated trajectories to learn quantitative models
-describing the local behavior of the system. These models are defined
-by a set of relations $R \subseteq \HybridStates \times \HybridStates$
-for each edge of the reachability graph. In the next section, we
-summarize this idea.
+% As 
+% Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
+% we use the generated trajectories to learn quantitative models
+% describing the local behavior of the system. These models are defined
+% by a set of relations $R \subseteq \HybridStates \times \HybridStates$
+% for each edge of the reachability graph. In the next section, we
+% summarize this idea.
 
-\subsection{Enriched Abstraction}
+% \subsection{Enriched Abstraction}
+% 
+% \input{figs/segtraj_rel_enriched.tex}
+% 
+% Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
+% trajectory between the respective cells. The graph abstraction $G$
+% only states that there exists a state $\x \in C$ from which the system
+% can evolve to a future state $\x' \in C'$. To increase the precision,
+% iterative refinement of the abstraction by state splitting was
+% proposed in~\cite{zutshi2014multiple}. Due to the curse of
+% dimensionality, state splitting is not scalable. Instead, we propose
+% an `enrichment' $G^R$ of $G$ by computing a set of local relations
+% $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$, which
+% non-deterministically describe relations between $\x \in C$ and $\x'
+% \in C'$. This is illustrated in \figref{enriched-edge}.
+% %(compare with \figref{segtraj}).
+% 
+% The enriched graph $G^R$ captures the underlying local forward
+% dynamics describing the evolution of the system in each abstract
+% state. We represent the dynamics using an affine model with an
+% interval error. Such a model can either be approximated using learning
+% methods or computed as a sound (over) approximation using reachability
+% set computation methods. Because we assume black box semantics, we
+% only present the former. Using regression analysis on a witnessed
+% trajectory between two cells, we compute an approximate discrete map
+% along with an error estimate.  Moreover, using the simulation function
+% $\simulate$, additional trajectories (or data) can be generated if
+% required. The data can be separated into a training set and testing
+% set to compute the map and the error respectively.
+% 
+% %The latter case can be explored if
+% %the symbolic dynamics of the system are known. A tool like
+% %\flowstar~\cite{chen2013flow} can be used to find the reachable set
+% %map.
+% 
+% Observe that $G^R$, a directed reachability graph, is rich enough to
+% search for concrete behaviors in the system. We call it a time
+% parameterized PWA relational abstraction. It can be interpreted as an
+% infinite state discrete transition system, and we can use
+% off-the-shelf bounded model checkers to find concrete violations of a
+% given safety property, and even other temporal properties.
 
-\input{figs/segtraj_rel_enriched.tex}
-
-Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
-trajectory between the respective cells. The graph abstraction $G$
-only states that there exists a state $\x \in C$ from which the system
-can evolve to a future state $\x' \in C'$. To increase the precision,
-iterative refinement of the abstraction by state splitting was
-proposed in~\cite{zutshi2014multiple}. Due to the curse of
-dimensionality, state splitting is not scalable. Instead, we propose
-an `enrichment' $G^R$ of $G$ by computing a set of local relations
-$R_{(C,C')}(\x,\x')$ for every edge $(C,C')$, which
-non-deterministically describe relations between $\x \in C$ and $\x'
-\in C'$. This is illustrated in \figref{enriched-edge}.
-%(compare with \figref{segtraj}).
-
-The enriched graph $G^R$ captures the underlying local forward
-dynamics describing the evolution of the system in each abstract
-state. We represent the dynamics using an affine model with an
-interval error. Such a model can either be approximated using learning
-methods or computed as a sound (over) approximation using reachability
-set computation methods. Because we assume black box semantics, we
-only present the former. Using regression analysis on a witnessed
-trajectory between two cells, we compute an approximate discrete map
-along with an error estimate.  Moreover, using the simulation function
-$\simulate$, additional trajectories (or data) can be generated if
-required. The data can be separated into a training set and testing
-set to compute the map and the error respectively.
-
-%The latter case can be explored if
-%the symbolic dynamics of the system are known. A tool like
-%\flowstar~\cite{chen2013flow} can be used to find the reachable set
-%map.
-
-Observe that $G^R$, a directed reachability graph, is rich enough to
-search for concrete behaviors in the system. We call it a time
-parameterized PWA relational abstraction. It can be interpreted as an
-infinite state discrete transition system, and we can use
-off-the-shelf bounded model checkers to find concrete violations of a
-given safety property, and even other temporal properties.
-
-\begin{figure}[!htbp]
-\begin{center}
-\tikzstyle{line} = [thick]
-\tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-\tikzstyle{block} = [rectangle, minimum width=0.5cm, minimum height=1.0cm, text centered, draw=black, align=center]%, fill=blue!25]
-\begin{tikzpicture}[scale=1.0, on grid, auto]
-
-    \node (G) at (-4,0) {$G$};
-    \node (GR) at (0,0) {$G^R$};
-    \node (CEX) at (4,0) {$CEx$};
-
-    \node (enrich) [block] at (-2, 0) {\scriptsize Enrich using \\ \scriptsize affine  maps};
-    \node (BMC) [block] at (2, 0) {\scriptsize Find CEx \\ \scriptsize using BMC};
-
-    %\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-
-    %\draw [line] (-2.0,-0.5) rectangle (-1.0,0.5);
-    \draw[arw] (G.east) -- (enrich.west);
-    \draw[arw] (enrich.east) -- (GR.west);
-    \draw[arw] (GR.east) -- (BMC.west);
-    \draw[arw] (BMC.east) -- (CEX.west);
-
-\end{tikzpicture}
-\end{center}
-\vspace*{-.3cm}
-\caption{Finding counter-examples (CEx) in a graph abstraction using PWA models.}
-\label{fig:overview}
-\vspace*{-.3cm}
-\end{figure}
+% \begin{figure}[!htbp]
+% \begin{center}
+% \tikzstyle{line} = [thick]
+% \tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
+% \tikzstyle{block} = [rectangle, minimum width=0.5cm, minimum height=1.0cm, text centered, draw=black, align=center]%, fill=blue!25]
+% \begin{tikzpicture}[scale=1.0, on grid, auto]
+% 
+%     \node (G) at (-4,0) {$G$};
+%     \node (GR) at (0,0) {$G^R$};
+%     \node (CEX) at (4,0) {$CEx$};
+% 
+%     \node (enrich) [block] at (-2, 0) {\scriptsize Enrich using \\ \scriptsize affine  maps};
+%     \node (BMC) [block] at (2, 0) {\scriptsize Find CEx \\ \scriptsize using BMC};
+% 
+%     %\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
+% 
+%     %\draw [line] (-2.0,-0.5) rectangle (-1.0,0.5);
+%     \draw[arw] (G.east) -- (enrich.west);
+%     \draw[arw] (enrich.east) -- (GR.west);
+%     \draw[arw] (GR.east) -- (BMC.west);
+%     \draw[arw] (BMC.east) -- (CEX.west);
+% 
+% \end{tikzpicture}
+% \end{center}
+% \vspace*{-.3cm}
+% \caption{Finding counter-examples (CEx) in a graph abstraction using PWA models.}
+% \label{fig:overview}
+% \vspace*{-.3cm}
+% \end{figure}

### 2017-06-08 graph_abs.tex
stuff
section 3 done

@@
 
-Let $\inputsignal$ = $(t_0, u_0)$, $\ldots$, $(t_N,u_N)$ be a
-timed-sequence of inputs, where $t_0 = 0$, and for all $i > 0$, $t_i -
-t_{i-1} = \Delta$, and $u_i \in \Inputs$.  We can extend the forward
-simulation map to a sequence of states to describe the application of
-an input signal to excite the system as follows:
-$\simmap_\Delta(\initstate, \mathbf{u} = \contState_N$, if for all $i  \in
-[0,N-1]$, $\simmap_\Delta(\contState_i, u_i,t_i) = \contState_{i+1}$.
-
+% Let $\inputsignal$ = $(t_0, u_0)$, $\ldots$, $(t_N,u_N)$ be a
+% timed-sequence of inputs, where $t_0 = 0$, and for all $i > 0$, $t_i -
+% t_{i-1} = \Delta$, and $u_i \in \Inputs$.  We can extend the forward
+% simulation map to a sequence of states to describe the application of
+% an input signal to excite the system as follows:
+% $\simmap_\Delta(\initstate, \mathbf{u}) = \contState_N$, if for all $i
+% \in [0,N-1]$, $\simmap_\Delta(\contState_i, u_i,t_i) =
+% \contState_{i+1}$.
+% 
 % An alternative view of the forward simulation-map is as a
@@
 % 
-\subsection{Graph Abstraction}
+\subsection{$\epsilon$-precision Transition System Abstraction}
 
-We observe that the parameterized reachability relation described in
-the previous section induces an existential discrete abstraction of
-the system similar to the one considered in \cite{zutshi2014multiple}.
-We briefly review this discrete abstraction and show how falsification
-of systems reduces to a path-search in the resulting transition
-system. We make the observation that this discrete abstraction is
-actually a Kripke structure, \ie, a transition system defined as the
-tuple $(\Cells, \CellEdges, \CellLabelingFunction)$, where $\Cells$ is
-a finite set of cells, $\CellEdges \subseteq \Cells \times \Cells$ is
-an unlabeled transition relation, and $\CellLabelingFunction$ is a
-function that assigns one or more labels to the cells from a finite
-set of labels.
+We now show how we can abstract a hybrid dynamical system $\System$
+which has trajectories in dense-time by a discrete transition system.
+A discrete transition system $\mathcal{T}$ can be defined as the tuple
+$(\Cells, \CellEdges, \Cells_\init, \CellLabelingFunction)$, where
+$\Cells$ is a finite set of cells, $\CellEdges \subseteq \Cells \times
+\Cells$ is an unlabeled transition relation, and
+$\CellLabelingFunction$ is a function that assigns one or more labels
+to the cells from a finite set $L$ of labels. The operational
+semantics of the transition system are as follows: the system starts
+in some cell in $\Cells_\init$. At each step, if the transition system
+is in cell $C$, it transitions to cell $C'$ if $(C,C') \in
+\CellEdges$. A finite-time trace property $\varphi$ of the system is
+some subset of the free monoid $(2^L)^*$. We say that $\mathcal{T}
+\not\models \varphi$ if there exists some trace of $\mathcal{T}$ such
+that the sequence of labels in that trace does not belong to
+$\varphi$.
 
-In what follows, we review an $\epsilon$-precision transition system
-abstraction of a system $\System$ (denoted as
-$\TSAbstraction{\epsilon}{\System}$) as follows. We define a
-quantization function $\quant_\epsilon$ that maps a state $\contState$ in
-$\ContStates$ to $10^{-\epsilon}*\lfloor \contState*10^\epsilon \rfloor$.
+We now give an example of a specific kind of transition system
+abstraction that we can build given $\System$ and the simulation map
+$\simmap_\Delta$.  We call the transition system an
+$\epsilon$-precision transition system and denote it by
+$\TSAbstraction{\epsilon}{\System}$.  Let $\epsilon  = 10^{-k}$, for
+some positive integer $k$.  We define a quantization function
+$\quant_\epsilon$ that maps a state $\contState$ in $\ContStates$ to
+$10^{-k}*\lfloor \contState*10^k \rfloor$. This transformation
+essentially truncates the decimal representation of the numeric value
+of the state $\state$ to at most $k$ digits after the decimal point.
 This function induces an equivalence relation $\contState_i
 \equiv_\epsilon \contState_j$ iff $\quant_\epsilon(\contState_i) =
-\quant_\epsilon(\contState_j)$. Now consider the quotient structure
-$\ContStates/\equiv_\epsilon$. We treat each equivalence class in the
-quotient structure as a cell. We discover the transitions in the
-abstraction using simulations. Given a numerical simulator for the
-model with a forward simulation map $\simmap_\Delta$, we add an edge
-from cell $C$ to cell $C'$ iff:
-\begin{itemize}
+\quant_\epsilon(\contState_j)$.
+
+Using the equivalence relation $\equiv_\epsilon$, we can now define
+the tuple $(\Cells, \CellEdges, \CellLabelingFunction)$ for
+$\TSAbstraction{\epsilon}{\System}$ by considering the quotient
+structure $\ContStates/\equiv_\epsilon$.  Essentially, each
+equivalence class in the quotient structure is a cell in $\Cells$.
+For cells $C$ and $C'$ in $\TSAbstraction{\epsilon}{\System}$, the
+edge $(C,C')$ can be an element of $\CellEdges$ iff:
+$\simmap_\Delta(\contState, u, t)$ = $\contState'$,
+$\quant_\epsilon(\contState)$ = $C$, and,
+$\quant_\epsilon(\contState') = C'$.
+
+We assume that there is a specified set of states $\UnsafeStates
+\subseteq \ContStates$ that are unsafe\footnote{Our framework is
+general enough to reason about richer class of correctness properties
+for trajectories such as bounded-time LTL properties. In this case, we
+assume the LTL property is defined over Boolean-valued predicates over
+$\ContStates$. The labeling function $\CellLabelingFunction$ would
+then label cells with the predicates that evaluate to $\mathit{true}$
+for some state in the cell.}, and thus restrict our attention to
+simple safety properties. Thus, our label set has only two elements:
+$\setof{\init,\unsafe}$. We then define $\CellLabelingFunction$ as
+$\CellLabelingFunction(C) = \init$ if $C$ contains some initial state,
+$\CellLabelingFunction(C) = \unsafe$ if $C$ contains some unsafe
+state, and empty otherwise.
+
+In what follows, we review the procedure
+$\mathsf{ScatterAndSimulate}$($\System$,$\simmap_\Delta$,$\epsilon$)
+from \cite{zutshi2014multiple} that, given the $\System$, the forward
+simulation map $\simmap_\Delta$, and a precision $\epsilon$ creates an
+$\epsilon$-precision transition system abstraction of $\System$.  The
+essence of the procedure is a breadth-first exploration of
+$\ContStates$ to build $\TSAbstraction{\epsilon}{\System}$ =
+$(\Cells,\CellEdges,\CellLabelingFunction)$. It consists of the
+following steps:
+\begin{enumerate}
 \item
-$\simmap_\Delta(\contState, u, t)$ = $\contState'$, 
+Add a set of random samples from the initial set of states to a queue
+(\ie some states $\contState$ at time $0$).
 \item
-$\quant_\epsilon(\contState)$ = $C$, and,
+Pop the head of the queue (say state $\contState$ at time $t$), identify
+the corresponding cell by applying $\quant_\epsilon$, and add it to
+$\Cells$. 
 \item
-$\quant_\epsilon(\contState') = C'$.
-\end{itemize}
-
-While our framework is general enough to reason about arbitrary
-temporal logic properties of trajectories, we restrict our attention
-to safety properties, where we assume that there is a specified set of
-states $\UnsafeStates \subseteq \ContStates$ that are unsafe.  Thus,
-our label set has only two elements:
-$\setof{\init,\unsafe}$. We then define $\CellLabelingFunction$ as
-follows: 
-\begin{equation*}
-\CellLabelingFunction(C) = \left\{
-\begin{array}{l@{\hspace{1em}}l}
-\init   & \exists \contState \in \InitStates\text{ s.t.  }\quant_\epsilon(\contState) = C  \\
-\unsafe & \exists \contState \in \UnsafeStates\text{ s.t.  }\quant_\epsilon(\contState) = C  \\
-\{\}    & \text{otherwise.}    
-\end{array}\right.
-\end{equation*}
+Obtain $\contState' = \simmap_\Delta(\contState, t, u)$. Obtain a a
+fixed number of states in the neighborhood of $\contState'$, and add
+$\contState'$ and these states (all time $t+\Delta$) to the queue. 
+\item
+Add an edge between the cells corresponding to $\contState$ and 
+$\contState'$.
+\item
+Go to step 2 till either the queue is empty or a fixed number of cells
+or edges is added to $\TSAbstraction{\epsilon}{\System}$.
+\end{enumerate}
 
@@
 nor an underapproximation} of the continuous-time continuous-space
-reachability relation between states. The purpose of this abstraction
-is to discover falsifying trajectories; thus, this technique explores
-the abstraction to get abstract falsifying trajectories starting at an
-initial abstract state that lead to error, and to confirm that such an
-abstract trajectory is not spurious, it performs a concretization step
-by performing an actual simulation of the system. If the abstract
-counterexample trajectories are found to be spurious, then
+reachability relation between states. The purpose of the
+$\epsilon$-precision transition system is to discover falsifying
+trajectories. This can be accomplished by
+Algorithm~\ref{algo:s3camBMC} that uses CounterExample-Guided
+Abstraction and Refinement (CEGAR).
+
+Here, we construct $\TSAbstraction{\epsilon}{\System}$
+(Line~\ref{algoline:construct}), we check if there is a path in
+$\TSAbstraction{\epsilon}{\System}$ that leads to a cell labeled
+unsafe. This can be done by a bounded model checking procedure
+(Line~\ref{algoline:bmc}) that explores at most $\ell$ transitions of
+$\TSAbstraction{\epsilon}{\System}$ (\ie, traces that are at most
+$\ell\Delta$ seconds in length). The model checker may return a set of
+counterexample trajectories $\Pi$. If this set is empty
+(Line~\ref{algoline:nocex}), then either the exploration performed by
+$\mathsf{ScatterAndSimulate}$ is not enough or the system could be
+safe. In the former case, the user can increase the scope of
+exploration for $\mathsf{ScatterAndSimulate}$.  The next step is to
+check if the abstract trajectories are spurious, by performing a
+concretization step by performing an actual simulation of the system
+(Line~\ref{algoline:checkspurious}). If all abstract counterexample
+trajectories are found to be spurious, then
 $\TSAbstraction{\epsilon}{\System}$ can be refined, basically by
-increasing $\epsilon$. We summarize this procedure in
-Algorithm~\ref{algo:cegar}.
+increasing $\epsilon$.
+\begin{algorithm}[t] 
+\DontPrintSemicolon
+\caption{CEGAR for $\TSAbstraction{\epsilon}{\System}$\label{algo:s3camBMC}}
+\KwIn{$\System$, $\simmap_\Delta$, $\epsilon_0$ , Smallest Precision $\epsilonmin$} 
+$\epsilon \assign \epsilon_0$ \;
+\While{$\epsilon \ge \epsilonmin$}{
+    $\TSAbstraction{\epsilon}{\System}$ $\assign$ 
+    $\mathsf{ScatterAndSimulate}$($\System$,$\simmap_\Delta$,$\epsilon$)
+    \nllabel{algoline:construct} \;
+    $\Pi$ $\assign$ $\mathsf{BoundedModelCheck}$($\TSAbstraction{\epsilon}{\System}$) 
+    \nllabel{algoline:bmc} \;
+    \lIf {$\Pi = \emptyset$}{ return FAIL \nllabel{algoline:nocex} } 
+    \lIf {$\mathsf{checkIfConcretizable}(\Pi)$}{ return VIOLATION \nllabel{algoline:checkspurious} }
+    $\epsilon$ $\assign$ $\frac{\epsilon}{10}$ \nllabel{algoline:refine} \;
+}
+\end{algorithm}
 
-The biggest drawback of Algorithm~\ref{algo:cegar} is that with
+The biggest drawback of Algorithm~\ref{algo:s3camBMC} is that with
 increasing $\epsilon$, the size of $\Cells$ in

### 2017-06-09 scam.tex
started adding relational models

--- 2017-06-08 graph_abs.tex
@@
 A discrete transition system $\mathcal{T}$ can be defined as the tuple
-$(\Cells, \CellEdges, \Cells_\init, \CellLabelingFunction)$, where
+$(\Cells, \CellEdges, \InitCells, \CellLabelingFunction)$, where
 $\Cells$ is a finite set of cells, $\CellEdges \subseteq \Cells \times

### 2017-06-09 implementation.tex
.

@@
 
+    insulin-st-1             &$Safe$ & $Safe$  & $0.0$\\
+    insulin-st-2             &$Safe$ & $Safe$  & $0.0$\\
+    lane change              &$Safe$ & $Safe$  & $0.0$\\
+    HVAC                     &$Safe$ & $Safe$  & $0.0$\\
+
 \bottomrule
@@
 need to refine like S3CAM, and should be more scalable, it is still
-constrained by the accuracy of the pwa models. Its performance can be
+constrained by the accuracy of the PWA models. Its performance can be
 explained by the difficulty in finding an abstraction

### 2017-06-10 learning.tex
relational models

--- 2017-05-31 prelims.tex
@@
-\subsection{PWA Transition System}
-
-We define a PWA transition system $A$ as a tuple
-$\tupleof{\pLocs,\vx,\scrT, G, U, \pLocI,\Theta}$, where:
-\begin{itemize}[noitemsep, leftmargin= 1.5 em]
-\item
-$\pLocs$ is a set of locations,
-\item
-$\vx$ is a variable that takes values in $\reals^m$,
-\item
-$\scrT$ is a finite subset of $\pLocs \times G \times U \times G \times \pLocs$, 
-\item
-$G$ is a set of conjunctions of affine predicates, 
-\item
-$U$ is a set of affine relations,
-\item
-$\pLocI$ is an initial location, and,
-\item
-$\Theta$ is an initial affine predicate.
-\end{itemize}
-
-The operation of $A$ is described in terms of its configuration and
-its moves. A configuration of $A$ is a pair $(\ell, \nu)$, where $\ell
-\in \pLocs$, and $\nu$ is a valuation function assigning values in
-$\reals^m$ to the variable $\vx$.  The move function $\mu$ maps a
-configuration $(\ell,\nu)$ to the next configuration $(\ell',\nu')$,
-iff there is a transition $\tau$ of the form $(\ell,g,f,g',\ell')$,
-and $\nu(\vx)$ satisfies predicate $g$, $\nu'(\vx)$ satisfies
-predicate $g'$ and $\nu'(\vx) \in f(\nu(\vx))$.
-
-For a given $\tau$, we call the affine relation $f_\tau$ its update
-relation, and the predicates $g_\tau$ and $g'_\tau$ as pre- and
-post-conditions on the transition. The machine starts its operation in
-location $\pLocI$ with an initial valuation $\nu_0$ to the variable
-$\vx$ such that $\nu_0(\vx)$ satisfies $\Theta$. For convenience, we
-abuse notation, and use the name a variable as a proxy for a certain
-valuation, and use the primed version of the variable as a proxy for
-its valuation after a transition, \ie, $\vx$ is used instead of
-$\nu(\vx)$ and $\vx'$ is used instead of $\nu'(\vx)$.
-
-% We define the PWA transition system as a discrete transition system
-% $\rho:\tupleof{\pLocs,\vx,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
-% a nonempty set of locations, $\vx$ is a set of variables, $\scrT$
-% is a finite set of transitions of the form $\tupleof{l, \tau,
-% l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
-% post-location and $\tau \in \scrT$ is the transition. $\tau$
-% is labeled by the transition relation $\rho_{\tau}$
-% over the current states and the next states. The transition relation
-% is defined in terms of affine predicates on the current and next
-% states and an affine update between them.
-% \[
-%     \rho_{\tau}(\vx,\vx') \subseteq \setof{(\vx,\vx')) \;|\;
-%     g_\tau(\vx) \land g'_\tau(\vx') \land f_{\tau}(\vx, \vx')}
-% \]
-% where $f_{\tau}$ is an affine relation on the pre and post states
-% $\vx,\vx'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
-% post conditions) on the states. Finally, $\pLocI$ is an initial
-% location $\Theta$ is an affine assertion characterizing the initial
-% states included in $\pLocI$.
-
-% A configuration of the PWATS is
-% a pair (l,\nu(v)), where \nu is a valuation function assigning values to
-% variables.
-
-% A run of the PWA TS is as follows: the PWA TS starts in the initial
-% location from a valuation \nu(v) \in theta.
-
-% A move of the PWA TS is described by the function mu: from (L,R^n) to
-% (L,R^n), and mu follows the rules that
-% mu(l,\nu(v)) = (l',\nu'(v)), iff (l,g,g',f,l') in T, and
-% \nu(v) satisfies g, \nu(v') satisfies g', and \nu(v') = f(\nu(v)).
-
-We how we can use the formal machinery of a PWA transition system $A$
-to approximate the behavior of a hybrid dynamical system $\System$
-defined by a simulator $\simulate^{\System}$~\footnote{Due to $f_\tau$
-being restricted to an affine form, we can only approximate general
-hybrid systems.}.
-
-%defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
-%\in f_\tau(\x)}$.
-
-% Given a state vector $\x\in\reals^n$, a guarded affine map
-% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
-% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
-% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
-% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
-% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
-% guard is satisfied.
-
-We now formalize the PWA relational model for a dynamical system.
-\begin{definition}[PWA Relational Model]
-
-A PWA relational model $A(\System)$ of a hybrid dynamical system
-$\System$ is simply a PWA transition system, where:
-
-\begin{itemize}[noitemsep, leftmargin= 1.5 em]
-\item
-each transition of $A$ represents a discrete time step $\Delta$ of the
-system $\System$,
-\item
-if the state-space of $\System$ is $\reals^n$, the the valuation
-function for variable $\x$ in $A$ assigns $\x$ some value in
-$\reals^n$,
-\item
-for the $N$ tuples $(v,v')$ discovered by simulation, if $v' = Av
-+ B + \vec{\delta}$ represents the affine relation discovered using
-linear regression, and if $h$ is the convex hull of all the $v$
-states across all tuples and $h'$ is the convex hull of all the $v'$
-states across all tuples, then $f_\tau \subseteq
-\setof{(\x,\x') \mid \x' \in A\x + B + \vec{\delta}}$,
-and $g$ states that $\x$ is in $h$, and $g'$ states that
-$\x' \in h'$.
-\end{itemize}
-
-Note that the guard predicates $g_\tau$ represent a conjunction of $m$
-affine inequalities, then we can represent $g_\tau$ as follows:
-\[
-g_\tau  \equiv C\x - \vd \le 0
-\]
-where $C$ is an $m \times n$ matrix and $\vd$ is $m \times 1$ matrix.
-
-% Given a hybrid dynamical system over a state-space $\reals^n$, a PWA
-% relational model $\rho$ is given by the tuple
-% $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in \scrT$ is a
-% collection of affine transitions  and $\Theta$ is an affine predicate
-% over $\x\in\real^n$. The transition relation is then defined by
-% $\scrT$ with $n$ transition relations as follows
-% 
-% \begin{equation}
-%     \scrT = \left\{
-%         \begin{array}{ll}
-%             g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
-%             \ldots\\
-%             g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
-%         \end{array}
-%     \right.
-% \end{equation}
-\end{definition}
-
-% where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
-% relations and $\x'$ denotes the next state of the system.
-
-A PWA relational model is \textit{deterministic}, iff for every state
-$\x\in\reals^n$, a unique guarded affine map is satisfied and if all
-entries in the error interval vector $\vec{\delta}$ are singletons.
-However, in practice such a case rarely exists. A PWA model will be
-usually non-deterministic, both due to multiple choices of transitions
-from any location and the reachable states at the $k^{th}$ time step
-(or after $k\Delta$ units of time) being a set.  Abusing the notation,
-we denote the set of states reachable in a single step in $\rho$ by
-$\x' = \rho(\x)$.
-
-A PWA relational model system is \textit{complete}, iff for every
-state $\x\in\HybridStates^n$, there exists at least one satisfied
-transition relation $\scrT$. If this is not the case, the system can
-deadlock, with no further executions. This usually results from
-modeling errors, and from here on, we do not consider such cases.
-
-In what follows, we use the terms PWA relational model and PWA
-transition system interchangeably to mean PWA relational model.
-
-\subsection{Bounded Model Checking (BMC)}
-The PWA transition system is a finite location, infinite state
-transition system. As all the guards and transitions are affine, a
-bounded reachability query is equivalent to checking all
-combinations of discrete locations, where each combination can be
-summarized as a linear program. As the time is bounded, the number of
-combinations are finite, but, exponential in number.
-
-It should be noted that an explicit execution of the transition system
-represents a directed acyclic graph, with each location being a node,
-and a directed edge from each node to all nodes reachable in forward
-time. A path on this graph is then a linear program, with each
-branching node corresponding to a logical disjunction.
-
-Temporal properties of PWA transitions systems can be checked using
-off-the-shelf model checkers like SAL~\cite{SAL-SRI}, which can reason
-over infinite state transitions systems using SMT solvers.
-Furthermore, lazy SMT solvers can be employed for this specific
-problem instance to achieve better efficiency~\cite{shoukry2017smc}.
-We now show how relational PWA abstractions can be viewed as PWA
-transition systems, and how the problem of falsification can
-be answered by a BMC query.
-
-% Given a dynamical system over a continuous state-space
-% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
-% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
-% dynamical system to a finite set of guarded affine maps
-% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
-% rule by the affine map of the satisfied $\gm$.
-
-% \begin{equation}
-%     \pwa = \left\{
-%         \begin{array}{ll}
-%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
-%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
-%             \ldots & \ldots\\
-%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
-%         \end{array}
-%     \right.
-% \end{equation}
-
-% \end{definition}
-
-% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
-% Abusing the notation, we denote the next state computed by the PWA
-% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
-% for every state $\x\in\ContStates$, a unique guarded affine map is
-% satisfied. A PWA model is \textit{complete}, iff for every state
-% $\x\in\ContStates$, there exists at least one satisfied guarded affine
-% map $\gm$.
-
-\section{Learning Dynamics using Simple Linear Regression}
-
 The approach in \cite{zutshi2014multiple} uses the forward simulation

### 2017-06-10 bmc_section.tex
Move bmc in a section
I create bmc_section.tex, the file bmc.tex contains the experiment on
van der pol.

--- 2017-05-31 prelims.tex
@@
-\subsection{PWA Transition System}
-
-We define a PWA transition system $A$ as a tuple
-$\tupleof{\pLocs,\vx,\scrT, G, U, \pLocI,\Theta}$, where:
-\begin{itemize}[noitemsep, leftmargin= 1.5 em]
-\item
-$\pLocs$ is a set of locations,
-\item
-$\vx$ is a variable that takes values in $\reals^m$,
-\item
-$\scrT$ is a finite subset of $\pLocs \times G \times U \times G \times \pLocs$, 
-\item
-$G$ is a set of conjunctions of affine predicates, 
-\item
-$U$ is a set of affine relations,
-\item
-$\pLocI$ is an initial location, and,
-\item
-$\Theta$ is an initial affine predicate.
-\end{itemize}
-
-The operation of $A$ is described in terms of its configuration and
-its moves. A configuration of $A$ is a pair $(\ell, \nu)$, where $\ell
-\in \pLocs$, and $\nu$ is a valuation function assigning values in
-$\reals^m$ to the variable $\vx$.  The move function $\mu$ maps a
-configuration $(\ell,\nu)$ to the next configuration $(\ell',\nu')$,
-iff there is a transition $\tau$ of the form $(\ell,g,f,g',\ell')$,
-and $\nu(\vx)$ satisfies predicate $g$, $\nu'(\vx)$ satisfies
-predicate $g'$ and $\nu'(\vx) \in f(\nu(\vx))$.
-
-For a given $\tau$, we call the affine relation $f_\tau$ its update
-relation, and the predicates $g_\tau$ and $g'_\tau$ as pre- and
-post-conditions on the transition. The machine starts its operation in
-location $\pLocI$ with an initial valuation $\nu_0$ to the variable
-$\vx$ such that $\nu_0(\vx)$ satisfies $\Theta$. For convenience, we
-abuse notation, and use the name a variable as a proxy for a certain
-valuation, and use the primed version of the variable as a proxy for
-its valuation after a transition, \ie, $\vx$ is used instead of
-$\nu(\vx)$ and $\vx'$ is used instead of $\nu'(\vx)$.
-
-% We define the PWA transition system as a discrete transition system
-% $\rho:\tupleof{\pLocs,\vx,\scrT, \pLocI,\Theta}$, where $\pLocs$ is
-% a nonempty set of locations, $\vx$ is a set of variables, $\scrT$
-% is a finite set of transitions of the form $\tupleof{l, \tau,
-% l'}$, where $l\in\pLocs$ is the pre-location, $l'\in\pLocs$ the
-% post-location and $\tau \in \scrT$ is the transition. $\tau$
-% is labeled by the transition relation $\rho_{\tau}$
-% over the current states and the next states. The transition relation
-% is defined in terms of affine predicates on the current and next
-% states and an affine update between them.
-% \[
-%     \rho_{\tau}(\vx,\vx') \subseteq \setof{(\vx,\vx')) \;|\;
-%     g_\tau(\vx) \land g'_\tau(\vx') \land f_{\tau}(\vx, \vx')}
-% \]
-% where $f_{\tau}$ is an affine relation on the pre and post states
-% $\vx,\vx'$, $g_{\tau}$ and $g'_{\tau}$ are affine guards (pre and
-% post conditions) on the states. Finally, $\pLocI$ is an initial
-% location $\Theta$ is an affine assertion characterizing the initial
-% states included in $\pLocI$.
-
-% A configuration of the PWATS is
-% a pair (l,\nu(v)), where \nu is a valuation function assigning values to
-% variables.
-
-% A run of the PWA TS is as follows: the PWA TS starts in the initial
-% location from a valuation \nu(v) \in theta.
-
-% A move of the PWA TS is described by the function mu: from (L,R^n) to
-% (L,R^n), and mu follows the rules that
-% mu(l,\nu(v)) = (l',\nu'(v)), iff (l,g,g',f,l') in T, and
-% \nu(v) satisfies g, \nu(v') satisfies g', and \nu(v') = f(\nu(v)).
-
-We how we can use the formal machinery of a PWA transition system $A$
-to approximate the behavior of a hybrid dynamical system $\System$
-defined by a simulator $\simulate^{\System}$~\footnote{Due to $f_\tau$
-being restricted to an affine form, we can only approximate general
-hybrid systems.}.
-
-%defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
-%\in f_\tau(\x)}$.
-
-% Given a state vector $\x\in\reals^n$, a guarded affine map
-% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
-% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
-% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
-% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
-% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
-% guard is satisfied.
-
-We now formalize the PWA relational model for a dynamical system.
-\begin{definition}[PWA Relational Model]
-
-A PWA relational model $A(\System)$ of a hybrid dynamical system
-$\System$ is simply a PWA transition system, where:
-
-\begin{itemize}[noitemsep, leftmargin= 1.5 em]
-\item
-each transition of $A$ represents a discrete time step $\Delta$ of the
-system $\System$,
-\item
-if the state-space of $\System$ is $\reals^n$, the the valuation
-function for variable $\x$ in $A$ assigns $\x$ some value in
-$\reals^n$,
-\item
-for the $N$ tuples $(v,v')$ discovered by simulation, if $v' = Av
-+ B + \vec{\delta}$ represents the affine relation discovered using
-linear regression, and if $h$ is the convex hull of all the $v$
-states across all tuples and $h'$ is the convex hull of all the $v'$
-states across all tuples, then $f_\tau \subseteq
-\setof{(\x,\x') \mid \x' \in A\x + B + \vec{\delta}}$,
-and $g$ states that $\x$ is in $h$, and $g'$ states that
-$\x' \in h'$.
-\end{itemize}
-
-Note that the guard predicates $g_\tau$ represent a conjunction of $m$
-affine inequalities, then we can represent $g_\tau$ as follows:
-\[
-g_\tau  \equiv C\x - \vd \le 0
-\]
-where $C$ is an $m \times n$ matrix and $\vd$ is $m \times 1$ matrix.
-
-% Given a hybrid dynamical system over a state-space $\reals^n$, a PWA
-% relational model $\rho$ is given by the tuple
-% $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in \scrT$ is a
-% collection of affine transitions  and $\Theta$ is an affine predicate
-% over $\x\in\real^n$. The transition relation is then defined by
-% $\scrT$ with $n$ transition relations as follows
-% 
-% \begin{equation}
-%     \scrT = \left\{
-%         \begin{array}{ll}
-%             g_1(\x) \land g'_1(\x') \implies f_1(\x,\x') \\
-%             \ldots\\
-%             g_n(\x) \land g'_n(\x') \implies f_n(\x,\x') \\
-%         \end{array}
-%     \right.
-% \end{equation}
-\end{definition}
-
-% where $g_i$ and $g'_i$ are affine predicates and $f_i$ are affine
-% relations and $\x'$ denotes the next state of the system.
-
-A PWA relational model is \textit{deterministic}, iff for every state
-$\x\in\reals^n$, a unique guarded affine map is satisfied and if all
-entries in the error interval vector $\vec{\delta}$ are singletons.
-However, in practice such a case rarely exists. A PWA model will be
-usually non-deterministic, both due to multiple choices of transitions
-from any location and the reachable states at the $k^{th}$ time step
-(or after $k\Delta$ units of time) being a set.  Abusing the notation,
-we denote the set of states reachable in a single step in $\rho$ by
-$\x' = \rho(\x)$.
-
-A PWA relational model system is \textit{complete}, iff for every
-state $\x\in\HybridStates^n$, there exists at least one satisfied
-transition relation $\scrT$. If this is not the case, the system can
-deadlock, with no further executions. This usually results from
-modeling errors, and from here on, we do not consider such cases.
-
-In what follows, we use the terms PWA relational model and PWA
-transition system interchangeably to mean PWA relational model.
-
-\subsection{Bounded Model Checking (BMC)}
 The PWA transition system is a finite location, infinite state
@@
 be answered by a BMC query.
-
-% Given a dynamical system over a continuous state-space
-% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
-% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
-% dynamical system to a finite set of guarded affine maps
-% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
-% rule by the affine map of the satisfied $\gm$.
-
-% \begin{equation}
-%     \pwa = \left\{
-%         \begin{array}{ll}
-%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
-%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
-%             \ldots & \ldots\\
-%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
-%         \end{array}
-%     \right.
-% \end{equation}
-
-% \end{definition}
-
-% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
-% Abusing the notation, we denote the next state computed by the PWA
-% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
-% for every state $\x\in\ContStates$, a unique guarded affine map is
-% satisfied. A PWA model is \textit{complete}, iff for every state
-% $\x\in\ContStates$, there exists at least one satisfied guarded affine
-% map $\gm$.
-
-\section{Learning Dynamics using Simple Linear Regression}
-
-The approach in \cite{zutshi2014multiple} uses the forward simulation
-map to explore a fixed number of tuples in the reachability relation
-$\reach{t,\vu}$ specified by a user-provided budget.  In general, the
-budget required to adequately explore the transition space requires
-exploring an exponential (in the dimension of $\HybridStates$) number
-of such tuples. This can cause undue burden in terms of computation
-time as well as storage.  The central idea in this paper is to {\em
-summarize} several tuples as a single affine map obtained using linear
-regression.  In other words, we can replace several ``proximal''
-tuples $(\vx,\vx')$ in $\reach{t,u}$ by an affine map of the form
-$\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are matrices and
-$\delta$ is a worst-case error estimate.
-
-In statistics, linear regression is the problem of finding an affine
-function or \textit{predictor}, which can `best' summarize the
-relationship between the given set of observed input $\x$ and output
-$\x'$ vectors. The notion of `best' is formally captured using a loss
-function which can be non-linear. We use the commonly used loss
-function \textit{ordinary least squares} (OLS) for this presentation.
-
-\mypara{Ordinary Least Squares (OLS)}
-For rest of the section, we elide the discrete mode from the hybrid
-state of the system, and only focus on the continuous state. Note that
-we can do this because of our assumption that the system is a
-switched-mode dynamical system.  Suppose we have $N$ tuples of the
-form $(\x,\x')$ in the reachability relation that we have explored,
-where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn an
-affine function described as 
-\begin{equation}
-\label{eq:affinemap}
-\x' = A\x + B
-\end{equation}
-Here, $A$ is a $n\times n$ matrix, and $B$ is a $n\times 1$ matrix.
-
-If $N>n$, which is the case in the current context of \textit{finding
-the best fit}, the problem is over-determined: there are more
-equations than unknowns. Hence, a single affine function cannot be
-found which satisfies the equation \eqref{eq:affinemap}. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This choice is
-formally defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine $A$ and $B$, such
-that the least square error of the affine predictor is minimized for
-the given data set.
-
-\begin{equation}
-    \operatornamewithlimits{argmin}_{A,B}\displaystyle\sum_{i=1}^{N}{\left(\x' - (A\x + B)\right)^2}
-\label{reg}
-\end{equation}
-
-To ease the presentation, we can rewrite the above as a homogeneous
-expression by using $\hat{\x} = [\x^T\ 1]^T$ and replacing $A$ and $B$ by the vector
-$\vec{A} = [A\ B]$. The equation~\ref{reg} now becomes
-\[
-\operatornamewithlimits{argmin}_{\vec{A}}\displaystyle\sum_{i=1}^{N}{(\x' - \vec{A}^T \hat{\x})^2}
-\]
-
-The solution of OLS can be analytically computed as the Moore-Penrose
-pseudo-inverse: $\vec{A} = (X^TX)^{-1}X^T X'$, where $X$ is the matrix
-representing the horizontal stacking of all $\hat{\x}$ and $X'$
-represents the stacking of all $\x'$ values.  The details can be found
-in standard texts on learning and statistics
-\cite{friedman2001elements}. Alternative methods based on gradient
-descent can also be used to find a suitable $A$ matrix.
-
-In this fashion, we can we can use OLS to approximate its trajectories
-of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
-= \vec{A}\hat{\x}$.
-
-% where
-% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T
-% \end{bmatrix}$ and $\vb = \begin{bmatrix}b_1 \\ \vdots \\ b_n
-% \end{bmatrix}$.
-
-We note that linear regression tools are often able to provide an
-$n\times 1$ vector $\vec{\delta}$ of error intervals, where the
-$i^{th}$ entry of the form $[{\delta_i}_{\min},{\delta_i}_{\max}]$
-indicates the best and worst possible error in computation of the
-$i^{th}$ state in $\x'$.
-
-Affine maps are poor approximations for arbitrary non-linear
-functions. Hence, we use a collection of affine maps to approximate
-the local behaviors (in state-space) of the system $\System$. This
-results in a piece wise approximation, as described in the next
-section.
-
-% \mypara{Regression Analysis}
-
-% In statistics, regression is the problem of finding a
-% \textit{predictor}, which can suitably predict the relationship
-% between the given set of observed input $\x$ and output $\y$ vectors.
-% In other words, assuming that $\y$ depends on $\x$, regression
-% strategies find either a parameterized or a non-parameterized
-% prediction function to explain the dependence. We now discuss simple
-% linear regression, which is parametric in nature and searches for an
-% affine predictor. It is also called \textit{ordinary least squares}.
-
-% \mypara{Ordinary Least Squares (OLS)}
-
-% Let the data set be comprised of $N$ input and output pairs $(\x,
-% y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
-% the case in the current context of \textit{finding the best fit}, the
-% problem is over-determined; there are more equations than
-% unknowns. Hence, a single affine function cannot be found which
-% satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
-% need to find the `best' choice for $\vec{a}$ and $b$. This is formally
-% defined using a loss function. For the case of simple linear
-% regression or OLS, the loss function is the sum of squares of the
-% errors in prediction.  The task is then to determine the vector of
-% coefficients $\vec{a}$ and an offset constant $b$, such that the least
-% square error of the affine predictor is minimized for the given data
-% set.
-% \begin{equation}
-%     \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
-% \label{reg}
-% \end{equation}
-% To ease the presentation, we can rewrite the above as a homogeneous
-% expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
-% $\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
-% \[
-%     \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
-% \]
-% The solution of OLS can be analytically computed as
-% \[ Ab = (X^TX)^{-1}X^T\y\]
-% where $X$ is the matrix representing the horizontal stacking of all
-% $\hat{\x}$.
-% The details can be found in several texts on learning and statistics
-% \cite{friedman2001elements}.
-
-% Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
-% where $\x\in\reals^n$ we can use OLS to approximate its trajectories
-% at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
-% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
-% \begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
-%     global relational model for the system. The data set for the
-%     regression is a set of pairs $\setof{(\x,\x') | \x' =
-%     \simulate(\x, \Delta)}$ and can be generated on demand. Another
-%     set can be generated to compute the error $\delta$ as an interval
-%     of vectors, where each element is an interval $\delta_i \in
-%     [\delta_{min}, \delta_{max}]$.
-%
-%     Affine maps are poor
-%     approximations for arbitrary non-linear functions. Hence, we use a
-%     collection of affine maps to approximate the local behaviors (in
-%     state-space) of the system $\simulate$. This results in a piece
-%     wise approximation, as described in the next section.
-
-% \subsection{Relational Abstractions}
-% For a hybrid automaton model, they can summarize the continuous
-% dynamics of each mode using a binary relation over continuous states.
-% The resulting relations are timeless (independent) and hence valid for
-% all time as long as the mode invariant is satisfied. The relations
-% take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
-% represents one of the relational operators $=, \ge, \le, <, >$. For
-% example, an abstraction that captures the monotonicity with respect to
-% time for the differential equation $\dot{x} = 2$ is $x' > x$. The
-% abstraction capturing the relation between the set of ODEs: $\dot{x} =
-% 2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
-
-%%  Similar to timeless relational abstractions, time-aware relational
-%%  abstractions \cite{mover2013time} construct binary relations between
-%%  the current state $\x$ of the system and any future reachable state
-%%  $\x'$. The difference lies with the latter also constructing relations
-%%  between the current time $t$ and any future time $t'$. This is
-%%  achieved by a case by case analysis of the eigen structure of the
-%%  matrix $A$ (of an affine ODE). Separate abstractions are used for the
-%%  case of linear systems with constant rate, real eigen values and
-%%  complex eigenvalues.

### 2017-06-11 bmc_section.tex
Add bmc encoding

@@
+We describe the the procedure $\mathsf{BoundedModelCheck}$ used at
+Line~\ref{algoline:bmc} in Algorithm~\ref{algo:s3camBMC}
+that checks if the $\epsilon$-precision Relational Transition System
+$\RTSAbstraction{\epsilon}{\System}$ violates the safety property
+$\psi(\vx)$ in at most $k$ steps.
+
+We describe two possible implementations:
+(1) a symbolic Bounded Model Checking~\cite{Biere+Others/99/Symbolic} encoding of
+the paths of $\RTSAbstraction{\epsilon}{\System}$ of length $k$ that 
+violates $\psi$, and (2) an algorithm that explicitly enumerates the
+``discrete'' paths formed cells $\Cells$ and the transition
+$\CellEdges$, and then checks if the relations of the variables $\vx$
+on the visited edges hold for for each path (this is in similar in
+spirit to~\cite{DBLP:conf/fmcad/BuLWL08}).
+
+%% BMC macro
+\newcommand{\bvars}[1]{\ensuremath{{B}^{#1}}}
+\newcommand{\rvars}[1]{\ensuremath{{\vx}^{#1}}}
+\newcommand{\vars}[1]{\ensuremath{{V}^{#1}}}
+\newcommand{\cellvar}[1]{\ensuremath{{cell}^{#1}}}
+\newcommand{\initbmc}{\ensuremath{INIT}}
+\newcommand{\encbmc}{\ensuremath{BMC}}
+\newcommand{\errorbmc}{\ensuremath{ERROR}}
+\newcommand{\transbmc}{\ensuremath{TRANS}}
+\newcommand{\edgebmc}{\ensuremath{EDGE}}
+\newcommand{\subs}[3]{\ensuremath{{#1}[{#2}/{#3}]}}
+
+Given a set of variables $X$, we denote with $X^i=\{x^i | x \in X\}$
+the copy of the variables $X$ renamed with the $^i$ superscript.
+Given a Boolean formula $\exprs{V}$
+%% $\subs{\exprs{V}}{z}{x}$ is the formula $\exprs{V}$ where all the
+%% occurrences of the variable $x$ are replaced with $z$.
+We use the shorthand $\subs{\exprs{V}}{X^i}{X}$ to denote the formula
+where every variable $x$ is substituted to $x^+$
+
+We encode the finite set of cells $\Cells$ with a set of Boolean
+variables $\bvars{}$. For clarity, given the cell $C \in \Cells$ we
+write $\cellvar{} = C$, instead of showing its encoding with $\bvars{}$.
+%
+In the following, we interpret each variable $x \in \vx$ as a Real
+valued variable and we refer to all the Boolean and Real variables
+with the set $\vars{} = \rvars{} \vee \bvars{}$
+%
+We encode the set of paths of length $k$ of
+$\RTSAbstraction{\epsilon}{\System}$ that violates the property $\psi$
+as the quantifier first-order logic formula:
+\begin{eqnarray*}
+\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k) & := &
+\initbmc(\vars{0}) \wedge
+\bigwedge_{i=1}^{k}{\transbmc(\vars{i-1}, \vars{i})} \wedge
+\neg \subs{\psi}{\rvars{i}}{\rvars{}}\\
+% init
+\initbmc(\vars{i}) & := &
+\bigwedge_{C \in \InitCells}{(\cellvar{i} = C) \wedge
+\subs{\CellLabelingFunction(C)}{\rvars{i}}{\rvars{}}
+}
+\\
+% trans
+\transbmc(\vars{i}, \vars{i+1}) & := &
+  \bigvee_{(C,C^+) \in \CellLabelingFunction}{
+    \edgebmc_{(C,C^+)}(\vars{i}, \vars{i+1})
+  } \\
+\edgebmc_{(C,C^+)}(\vars{i}, \vars{i+1}) & := &
+(\cellvar{i} = C) \wedge (\cellvar{i+1} = C^+) \wedge  \\
+&& \subs{\CellLabelingFunction(C)}{\rvars{i}}{\rvars{}} \wedge
+   \subs{\CellLabelingFunction(C^+)}{\rvars{i+1}}{\rvars{}} \wedge \\
+&& \subs{\subs{\EdgeLabelingFunction((C,C^+))}{\rvars{i}}{\rvars{}}}{\rvars{i+1}}{{\rvars{+}}} \\
+\end{eqnarray*}
+
+The formula
+$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k)$ is satisfiable
+if and only if there exists a path of length $k$.
+The procedure
+$\mathsf{BoundedModelCheck}(\RTSAbstraction{\epsilon}{\System}, \psi, k)$
+checks the satisfiability of the sequence of formulas 
+$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, 0)$, 
+$\ldots$,
+$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k)$
+with a SMT solver, returning as soon as one of the formula is
+satisfiable (i.e. as soon as it found a path in
+$\RTSAbstraction{\epsilon}{\System}$).
+
 The PWA transition system is a finite location, infinite state
@@
 be answered by a BMC query.
+
+% To clarify:
+%  - is there a maximum k that we look at? Where do we guarnatee that?
+%

### 2017-06-11 bmc_section.tex
Add the explicit enumeration of paths.

@@
-We describe the the procedure $\mathsf{BoundedModelCheck}$ used at
-Line~\ref{algoline:bmc} in Algorithm~\ref{algo:s3camBMC}
-that checks if the $\epsilon$-precision Relational Transition System
-$\RTSAbstraction{\epsilon}{\System}$ violates the safety property
-$\psi(\vx)$ in at most $k$ steps.
-
-We describe two possible implementations:
-(1) a symbolic Bounded Model Checking~\cite{Biere+Others/99/Symbolic} encoding of
-the paths of $\RTSAbstraction{\epsilon}{\System}$ of length $k$ that 
-violates $\psi$, and (2) an algorithm that explicitly enumerates the
-``discrete'' paths formed cells $\Cells$ and the transition
-$\CellEdges$, and then checks if the relations of the variables $\vx$
-on the visited edges hold for for each path (this is in similar in
-spirit to~\cite{DBLP:conf/fmcad/BuLWL08}).
-
 %% BMC macro
+\newcommand{\unsafebmc}{\ensuremath{{\psi}_{\UnsafeStates}}}
 \newcommand{\bvars}[1]{\ensuremath{{B}^{#1}}}
@@
 \newcommand{\encbmc}{\ensuremath{BMC}}
-\newcommand{\errorbmc}{\ensuremath{ERROR}}
 \newcommand{\transbmc}{\ensuremath{TRANS}}
@@
 
+\newcommand{\abspathbmc}[0]{\ensuremath{\Pi_{abs}}}
+\newcommand{\concbmc}[1]{\ensuremath{CONC_{{#1}}}}
+%
+We describe the the procedure $\mathsf{BoundedModelCheck}$ used at
+Line~\ref{algoline:bmc} in Algorithm~\ref{algo:s3camBMC}
+that checks if the $\epsilon$-precision Relational Transition System
+$\RTSAbstraction{\epsilon}{\System}$ reaches an error condition
+$\UnsafeStates$ in at most $k$ steps. We assume that the set of unsafe
+states $\UnsafeStates$ can be represented with a formula
+$\unsafebmc(\vx)$
+
+We describe two possible implementations:
+(1) a symbolic Bounded Model Checking~\cite{Biere+Others/99/Symbolic} encoding of
+the paths of $\RTSAbstraction{\epsilon}{\System}$ of length $k$ that 
+reaches $\unsafebmc$, and (2) an algorithm that explicitly enumerates the
+``discrete'' paths formed cells $\Cells$ and the transition
+$\CellEdges$, and then checks if the relations of the variables $\vx$
+on the visited edges hold for for each path (this is in similar in
+spirit to~\cite{DBLP:conf/fmcad/BuLWL08}).
+
+\paragraph{Symbolic path exploration.}
+%
+%
 Given a set of variables $X$, we denote with $X^i=\{x^i | x \in X\}$
@@
 We encode the set of paths of length $k$ of
-$\RTSAbstraction{\epsilon}{\System}$ that violates the property $\psi$
+$\RTSAbstraction{\epsilon}{\System}$ that reaches $\unsafebmc$
 as the quantifier first-order logic formula:
 \begin{eqnarray*}
-\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k) & := &
+\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k) & := &
 \initbmc(\vars{0}) \wedge
 \bigwedge_{i=1}^{k}{\transbmc(\vars{i-1}, \vars{i})} \wedge
-\neg \subs{\psi}{\rvars{i}}{\rvars{}}\\
+\subs{\unsafebmc}{\rvars{k}}{\rvars{}}\\
 % init
@@
    \subs{\CellLabelingFunction(C^+)}{\rvars{i+1}}{\rvars{}} \wedge \\
-&& \subs{\subs{\EdgeLabelingFunction((C,C^+))}{\rvars{i}}{\rvars{}}}{\rvars{i+1}}{{\rvars{+}}} \\
+&& \subs{\subs{\EdgeLabelingFunction((C,C^+))}{\rvars{i}}{\rvars{}}}{\rvars{i+1}}{{\rvars{+}}}
 \end{eqnarray*}
@@
 The formula
-$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k)$ is satisfiable
-if and only if there exists a path of length $k$.
+$\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$ is satisfiable
+if and only if there exists a path of length $k$ in 
+$\RTSAbstraction{\epsilon}{\System}$ that reaches a state in $\unsafebmc$.
+
 The procedure
-$\mathsf{BoundedModelCheck}(\RTSAbstraction{\epsilon}{\System}, \psi, k)$
+$\mathsf{BoundedModelCheck}(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$
 checks the satisfiability of the sequence of formulas 
-$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, 0)$, 
+$\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, 0)$, 
 $\ldots$,
-$\encbmc(\RTSAbstraction{\epsilon}{\System}, \psi, k)$
+$\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$
 with a SMT solver, returning as soon as one of the formula is
-satisfiable (i.e. as soon as it found a path in
-$\RTSAbstraction{\epsilon}{\System}$).
+satisfiable, returning a path $\Pi$ of
+$\RTSAbstraction{\epsilon}{\System}$.
 
-The PWA transition system is a finite location, infinite state
-transition system. As all the guards and transitions are affine, a
-bounded reachability query is equivalent to checking all
-combinations of discrete locations, where each combination can be
-summarized as a linear program. As the time is bounded, the number of
-combinations are finite, but, exponential in number.
+\paragraph{Explicit path exploration.}
+The second algorithm does not encode the reachability problem
+$\mathsf{BoundedModelCheck}$ of a fixed length path as a single
+satisfiability query, but it explicitly enumerates
+the (abstract) paths in the graph that has $\Cells$
+as nodes, and $\CellEdges$ as edges, and may end in an unsafe state
+$\UnsafeStates$.
+%
+Such paths may not be real paths of
+$\RTSAbstraction{\epsilon}{\System}$, since they may not satisfy the
+conditions imposed by the labeling function $\CellLabelingFunction$
+and the relation $\EdgeLabelingFunction$, and hence the algorithm must
+further check, for each abstract path $\abspathbmc$, if tere exists a 
+concrete path $\Pi$ of
+$\RTSAbstraction{\epsilon}{\System}$.
 
-It should be noted that an explicit execution of the transition system
-represents a directed acyclic graph, with each location being a node,
-and a directed edge from each node to all nodes reachable in forward
-time. A path on this graph is then a linear program, with each
-branching node corresponding to a logical disjunction.
+$\abspathbmc := c_0, \ldots, c_l$ is an abstract path of
+ $\RTSAbstraction{\epsilon}{\System}$ where:
+\begin{itemize}[label=--,leftmargin=1em,labelsep=*]
+\item for $i \in [0,l]$, $c_i \in \Cells$,
+\item $c_0 \in \InitCells$,
+\item for $i \in [0,l-1]$, $(c_i, c_{i+1}) \in \EdgeLabelingFunction$,
+\item $c_l$ is such that
+$\CellLabelingFunction(c_l) \cap \UnsafeStates \neq \emptyset$.
+\end{itemize}
+%
+There exists a concrete path $\Pi$ of
+$\RTSAbstraction{\epsilon}{\System}$ that traverses the sequences of
+cells of an abstract path $\abspathbmc$ if and only if the formula
+\begin{eqnarray}
+\label{eq:simulateabs}
+\concbmc{\abspathbmc} & := &
+\bigg( \bigwedge_{i=0}^{l}{
+  \subs{\CellLabelingFunction(c_i)}{\rvars{i}}{\rvars{}}
+} \bigg)
+\wedge
+\subs{\unsafebmc}{\rvars{l}}{\rvars{}} \wedge \\
+&&
+\bigwedge_{i=1}^{l}{
+  \subs{
+  \subs{\CellEdges((c_i,c_{i+1}))}{\rvars{i}}{\rvars{}}
+  }{\rvars{i+1}}{\rvars{+}}
+}
+\end{eqnarray}
+is satisfiable\footnote{Formula~\ref{eq:simulateabs} is a conjunction
+of linear predicates and hence its satisfiability can be casted as the
+feasibility of a linear program}.
+%
+This can proved showing that each model to the formula
+$\concbmc{\abspathbmc}$  ``corresponds'' to a path of
+$\RTSAbstraction{\epsilon}{\System}$, and each path of
+$\RTSAbstraction{\epsilon}{\System}$ that traverse the sequence
+$\abspathbmc$ ``corresponds'' to a model of
+Formula~\ref{eq:simulateabs}.
 
-Temporal properties of PWA transitions systems can be checked using
-off-the-shelf model checkers like SAL~\cite{SAL-SRI}, which can reason
-over infinite state transitions systems using SMT solvers.
-Furthermore, lazy SMT solvers can be employed for this specific
-problem instance to achieve better efficiency~\cite{shoukry2017smc}.
-We now show how relational PWA abstractions can be viewed as PWA
-transition systems, and how the problem of falsification can
-be answered by a BMC query.
+The explicit path exploration enumerates all the abstract
+paths $\abspathbmc$ of length at most $k$. The enumeration can either
+be a depth-first search (DFS) or a breadth-first search (BFS). As soon
+as $\abspathbmc$ is found, the algorithm checks for the satisfiability
+of $\concbmc{\abspathbmc}$. If  $\concbmc{\abspathbmc}$ is
+satisfiable, the algorithm returns a concrete path $\Pi$, otherwise
+the search continue trying to enumerate another abstract path. 
+The search is guaranteed to terminate, since the length of the
+abstract paths is at most $k$.
+
+\todo{Tell that we use a DFS in the implementation section}
+
+%% Temporal properties of PWA transitions systems can be checked using
+%% off-the-shelf model checkers like SAL~\cite{SAL-SRI}, which can reason
+%% over infinite state transitions systems using SMT solvers.
+%% Furthermore, lazy SMT solvers can be employed for this specific
+%% problem instance to achieve better efficiency~\cite{shoukry2017smc}.
+%% We now show how relational PWA abstractions can be viewed as PWA
+%% transition systems, and how the problem of falsification can
+%% be answered by a BMC query.
 

### 2017-06-11 bmc_section.tex
Fix typo

@@
 and the relation $\EdgeLabelingFunction$, and hence the algorithm must
-further check, for each abstract path $\abspathbmc$, if tere exists a 
+further check, for each abstract path $\abspathbmc$, if there exists a 
 concrete path $\Pi$ of

### 2017-06-11 intro.tex
.
massaged intro

@@
-% Testing of hybrid systems is a complex task, due to the interaction
-% between discrete and continuous dynamics. Industrially developed
-% hybrid systems, with embedded software interacting with physical
-% components, are typically safety-critical. Thus, it is important to
-% guarantee that the systems adhere to functional and safety
-% specifications. Since the reachability problem for hybrid systems is
-% known to in general be undecidable, exhaustive exploration of the
-% state-space is intractable, and thus testing is an approach to attempt
-% to achieve system correctness. However, to guarantee safety and other
-% requirements, the testing procedures need to rely on a rigorous
-% theoretical foundation. In this special session aims to collect and
-% connect researchers, within academia as well as industry, working on
-% model-based testing of hybrid systems. The session calls for papers
-% proposing new theoretical research directions in the field, as well as
-% papers describing practical applications of model-based testing of
-% hybrid systems.
-
-Model-based development is rapidly becoming the preferred paradigm for
-developing embedded control software in an industrial context.  Tools
-like \SIMULINK are often used to create models of closed-loop
-dynamical systems, \ie, a plant model of the physical systems that we
-wish to control, and a model for the embedded controllers.  Automatic
-approaches for testing the safety and performance of such closed-loop
-systems is challenging, and current approaches focus mostly on guided
+Simulation-driven model-based development is rapidly becoming the
+preferred paradigm for developing embedded control software in an
+industrial context.  Tools like \SIMULINK are often used to create
+models of closed-loop dynamical systems, \ie, a plant model of the
+physical systems that we wish to control, and a model for the embedded
+controllers.  Automatic approaches for testing the safety and
+performance of such closed-loop systems largely focus on guided
 testing using numerical simulations
 \cite{annpureddy2011s,donze2010breach,deshmukh2015stochastic,dreossi2015efficient,akazaki}.
-On the other hand, theoretical research on hybrid and timed systems
-focusses on formal guarantees for restricted classes of dynamical
-systems such as hybrid automata where the system state evolves
-according to a vector field that is either constant, affine
-\cite{frehse2011spaceex}, or polynomial \cite{chen2015reachability}.
-In this paper, we propose an approach that leverages useful features
-from both kinds of techniques.
+On the other hand, a common approach taken in formal verification
+approaches for hybrid and timed systems is to model such closed-loop
+systems as restricted classes of dynamical systems such as those with
+constant, affine \cite{frehse2011spaceex}, or polynomial
+\cite{chen2015reachability} dynamics. Constructing such conservative
+abstractions of real-world closed-loop systems involves a difficult
+tradeoff between keeping the computed abstraction small in size but
+high in fidelity.
 
-We propose a two step process for finding violations of safety
-properties. In the first step, we learn a piece-wise affine (PWA) {\em
-relational} model from simulations of the dynamical system. In the
-second step, we leverage symbolic techniques to exhaustively analyze
-the model learned in the first phase. We remark that the approach we
-propose is neither sound nor complete: a violation found by our
-technique may not exist in the original system, and our technique may
-miss violations that exist in the original system. A fair question is:
-why is such a technique useful? There are two perspectives from which
-we can answer this question.
+In this paper, we wish to find inputs to the system and system initial
+conditions that falsify safety properties of effectively black-box
+systems, \ie, systems where the dynamics are highly complex, or not
+expressed in a closed-form. We thus investigate a falsification
+approach that operates on an on-the-fly abstraction of the system
+constructed by assuming that the system dynamics have a certain form.
+Our approach has two main steps: In the first step, we learn a
+piece-wise {\em relational model} from simulations of the given
+dynamical system. In the second step, we leverage symbolic techniques
+to exhaustively analyze the model learned in the first phase. We
+remark that the approach we propose is neither sound nor complete: a
+violation found by our technique may not exist in the original system,
+and our technique may miss violations that exist in the original
+system. A fair question is: why is such a technique useful? There are
+two perspectives from which we can answer this question.
 
@@
 \cite{zutshi2012timed,sankaranarayanan2011relational}. By learning a
-{\em PWA relational model}, \ie, essentially a PWA discrete-time
-dynamical system, we can use software analysis approaches such as
-bounded model checking assisted by SMT solvers and linear programming
-solvers.
+{\em kernelized relational model}, \ie, essentially a discrete-time
+dynamical system with a chosen form for the dynamics, we can use
+software analysis approaches such as bounded model checking assisted
+by SMT solvers and linear programming solvers.
 
-The other perspective for our technique is the context when we have an
-effectively black-box model, \ie, a model that has features such as
-nonlinearities, delays or look-up tables that make symbolic analysis
-rather challenging.  We note that the ultimate goal for our technique
-is to find violations of safety properties in the {\em actual embedded
-control system}, and not, per se, in the model. The aphorism
-attributed to mathematician George E.P. Box is quite apt for our
-context, and bears worth repeating,  ``all models are wrong, some
-models are useful.'' We can sidestep analyzing a highly complex model
-by approximating its behavior with a simpler relational model, which
-in turn can give us promising tests to run on the actual system. This
-gives our technique intrinsic value as a debugging tool.
-
-% In our previous work~\cite{zutshi2014multiple}, we analyzed systems by
-% restricting ourselves to their black box semantics. This enabled us to
-% reason about the state-space reachability of the system without a
-% direct analysis of its structure. Using a coarse abstraction which we
-% searched on-the-fly, we could observe the local dynamics as required.
-% This gave us an efficient procedure to find abstract counter-examples.
-% However, to concretize the counter-examples, a `closer look' or
-% refinement of the abstraction is required. The grid based state-space
-% abstraction was refined by splitting the relevant cells (abstract
-% states) into smaller ones. As noted previously, due to the curse of
-% dimensionality, such a uniform splitting is an expensive operation,
-% and can not scale to higher dimensions.
+The other perspective for our technique is that we are often faced
+with black-box models, where symbolic analysis is rather challenging.
+We note that the ultimate goal is to find violations of safety
+properties in the {\em actual embedded control system}, and not, per
+se, in the model. The aphorism attributed to mathematician George E.P.
+Box is quite apt for our context, and bears worth repeating,  ``all
+models are wrong, some models are useful.'' We can sidestep analyzing
+a highly complex model by approximating its behavior with a simpler
+relational model, which in turn can give us promising tests to run on
+the actual system. This gives our technique intrinsic value as a
+debugging tool.
 
 In \cite{zutshi2014multiple}, the authors investigate an on-the-fly
-abstraction technique that is close to the approach we propose.  Both
-approaches use an existential abstraction of the state-space of the
-hybrid dynamical system. The abstract model learned is a directed
-acyclic graph, where nodes represent regions of the state-space, and
-an edge between two regions indicates a transition between states in
-the regions.  A key difference in this work is that the model being
-learned in addition to being an existential graph-based abstraction is
-also a {\em relational} abstraction.  In other words, the approach
-presented here improves the existential abstraction by annotating the
-edges in the graph-based abstraction with affine relations between the
-source and destination states.  More specifically, we use linear
-regression to quantitatively estimate the discovered edges in the
-graph-based abstraction of the state-space. These linear maps
-approximate locally observed behaviors, and the resulting graph or the
-Piece-Wise Affine (PWA) relational model can be interpreted as an
-infinite state discrete transition system and model checked for time
-bounded safety properties. A counter-example if found, can indicate
-the presence of a violation in the system.  We then use this
-counter-example to guide the search towards a counter-example in the
-original system.
+abstraction technique that is close to the approach we propose.  This
+approach use an existential abstraction of the state-space of a given
+hybrid dynamical system. The abstract model learned can be viewed as a
+discrete-time transition system, where states of the transition system
+are cells in a tiling-based abstraction, and an edge between two cells
+indicates a concrete transition between the states of the actual
+system contained in the cell.  A key difference in this paper is that
+the model being learned in addition to being such a transition system
+is that it is now enriched with labels on the states and transitions.
+These labels capture {\em relations} between source and destination
+states. Thus an edge between two cells can now represent the {\em
+guard} region within the cell where the source states lie, and the
+(symbolic) {\em update} relation between the destination and the
+source states.  The relations that we learn depend on user-specified
+{\em kernels}, \ie, if the user specifies that the relations should be
+linear, we can use techniques such as linear regression or convex hull
+computation to obtain the update and guard relations respectively.
+Other kernels such as polynomials can also be included.  The relations
+discovered approximate locally observed behaviors, and the resulting
+kernelized transition system can be interpreted as a symbolic,
+discrete labeled transition system. We can apply time-bounded model
+checking to find counterexamples to bounded safety properties on such
+a system. A counterexample if found, can indicate the presence of a
+violation in the system.  We can then use this counterexample to guide
+the search towards a violation in the original system.
 
-We remark that our work focusses on building discrete PWA models, and
-is unrelated to work on analyzing dynamical systems that have PWA
-vector fields. Techniques to analyze PWA dynamical systems using
-bisimulation relations translate the system into a non-deterministic
-infinite state transition system, and use model checking approaches to
-verify them \cite{asarin2000approximate}. Related approaches using
-bisimulation relations \cite{pappas2003bisimilar,tabuada2006linear,
-yordanov2007model} have interesting parallels with the work presented
-here.
+We remark that our work focusses on building piecewise relational
+models, and is unrelated to work on analyzing dynamical systems that
+have piecewise-affine (PWA) vector fields. Techniques to analyze PWA
+dynamical systems using bisimulation relations translate the system
+into a non-deterministic infinite state transition system that is
+required to be an overapproximation of the original system and use
+model checking approaches to verify them \cite{asarin2000approximate}.
 
-%Furthermore, using linear programming, we can over-approximate the
-%neighbhourhood of the counter-example in the model.
+% Related approaches using bisimulation relations
+%\cite{pappas2003bisimilar,tabuada2006linear, yordanov2007model} have
+% interesting parallels with the work presented here.
 
-% In the conclusion, we discuss extensions to data driven
-% approaches (instead of simulator driven), where instead of a
-% simulator, a fixed set of data is provided as the behavioral
-% description of the system. Using a combination of relational modeling,
-% and program analysis we outline the future work for falsifying
-% properties of SDCS.
-% Bounded time behvaiors are modeled
-% only one violation is being searched for
-% mention explicitly the relation between older cegar work
-
-The layout of the paper is as follows: In Section~\ref{sec:prelims},
-we introduce the basic steps to obtain a PWA relational model. In
-Section~\ref{sec:pwa-rel}, we explore the connection between the PWA
-relational model and previous graph-based abstraction techniques such
-as those in \cite{zutshi2014multiple}. We present extensions of
-relational modeling in Section~\ref{sec:rel-mod}, and present
-experimental validation of our technique on a few textbook examples of
-dynamical systems with hybrid and polynomial dynamics in
-Section~\ref{sec:res}.
+The layout of the paper is as follows: In Section~\ref{sec:scam}, we
+review the discrete-time on-the-fly transition system abstraction
+introduced by \cite{zutshi2014multiple}.  We show how to enrich this
+abstraction to kernelized transition systems in
+Section~\ref{sec:scamr}.  We show how such systems can be analyzed
+using symbolic bounded model checking in Sec.~\ref{sec:bmc}.  We
+present an approach to refine the relational transition systems in
+Sec.~\ref{sec:rel-mod}, and present experimental validation of our
+technique on a few examples systems in Section~\ref{sec:res}.

### 2017-06-11 bmc_section.tex

@@
 \newcommand{\unsafebmc}{\ensuremath{{\psi}_{\UnsafeStates}}}
-\newcommand{\bvars}[1]{\ensuremath{{B}^{#1}}}
+\newcommand{\bvars}[1]{\ensuremath{{\vec{B}}^{#1}}}
 \newcommand{\rvars}[1]{\ensuremath{{\vx}^{#1}}}
-\newcommand{\vars}[1]{\ensuremath{{V}^{#1}}}
+\newcommand{\vars}[1]{\ensuremath{{\vec{V}}^{#1}}}
 \newcommand{\cellvar}[1]{\ensuremath{{cell}^{#1}}}
@@
 %
-We describe the the procedure $\mathsf{BoundedModelCheck}$ used at
-Line~\ref{algoline:bmc} in Algorithm~\ref{algo:s3camBMC}
+We describe the implementation of the procedure
+$\mathsf{BoundedModelCheck}$ used at Line~\ref{algoline:bmc} in
+Algorithm~\ref{algo:s3camBMC},
 that checks if the $\epsilon$-precision Relational Transition System
 $\RTSAbstraction{\epsilon}{\System}$ reaches an error condition
-$\UnsafeStates$ in at most $k$ steps. We assume that the set of unsafe
-states $\UnsafeStates$ can be represented with a formula
-$\unsafebmc(\vx)$
+$\UnsafeStates$ in at most $k$ steps returning, in that case, a path
+$\Pi$. In the following we assume that the set of unsafe states
+$\UnsafeStates$ can be represented with a Boolean formula
+$\unsafebmc(\vx)$.
 
-We describe two possible implementations:
-(1) a symbolic Bounded Model Checking~\cite{Biere+Others/99/Symbolic} encoding of
-the paths of $\RTSAbstraction{\epsilon}{\System}$ of length $k$ that 
-reaches $\unsafebmc$, and (2) an algorithm that explicitly enumerates the
-``discrete'' paths formed cells $\Cells$ and the transition
-$\CellEdges$, and then checks if the relations of the variables $\vx$
-on the visited edges hold for for each path (this is in similar in
-spirit to~\cite{DBLP:conf/fmcad/BuLWL08}).
+We describe two implementations:
+(1) a symbolic Bounded Model Checking~\cite{Biere+Others/99/Symbolic}
+(BMC) algorithm and (2) an algorithm (similarly
+to e.g.~\cite{DBLP:conf/fmcad/BuLWL08}) that explicitly enumerates the
+abstract paths formed by cells $\Cells$ and edges $\CellEdges$,
+checking for their feasibility in $\RTSAbstraction{\epsilon}{\System}$.
 
@@
 Given a set of variables $X$, we denote with $X^i=\{x^i | x \in X\}$
-the copy of the variables $X$ renamed with the $^i$ superscript.
-Given a Boolean formula $\exprs{V}$
+the copy of the variables $X$ renamed with the ``$^i$'' superscript.
+Given a Boolean formula $\exprs{\vars{}}$,
 %% $\subs{\exprs{V}}{z}{x}$ is the formula $\exprs{V}$ where all the
 %% occurrences of the variable $x$ are replaced with $z$.
-We use the shorthand $\subs{\exprs{V}}{X^i}{X}$ to denote the formula
-where every variable $x$ is substituted to $x^+$
+$\subs{\exprs{\vars{}}}{\vars{i}}{\vars{}}$ is the formula $\exprs{\vars{}}$
+where every variable $v \in \vars{}$ is substituted with $v^+ \in \vars{+}$.
 
@@
 $\RTSAbstraction{\epsilon}{\System}$ that reaches $\unsafebmc$
-as the quantifier first-order logic formula:
+as the quantifier-free first-order logic formula:
 \begin{eqnarray*}
@@
 The procedure
-$\mathsf{BoundedModelCheck}(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$
+$\mathsf{BoundedModelCheck}(\RTSAbstraction{\epsilon}{\System})$
 checks the satisfiability of the sequence of formulas 
@@
 $\encbmc(\RTSAbstraction{\epsilon}{\System}, \unsafebmc, k)$
-with a SMT solver, returning as soon as one of the formula is
-satisfiable, returning a path $\Pi$ of
-$\RTSAbstraction{\epsilon}{\System}$.
+with an SMT solver, returning a path $\Pi$ of
+$\RTSAbstraction{\epsilon}{\System}$ as soon as one of the formula is
+satisfiable, or no paths otherwise.
 
 \paragraph{Explicit path exploration.}
-The second algorithm does not encode the reachability problem
-$\mathsf{BoundedModelCheck}$ of a fixed length path as a single
-satisfiability query, but it explicitly enumerates
+The second algorithm explicitly enumerates
 the (abstract) paths in the graph that has $\Cells$
@@
 
-$\abspathbmc := c_0, \ldots, c_l$ is an abstract path of
+$\abspathbmc := C_0, \ldots, C_l$ is an abstract path of
  $\RTSAbstraction{\epsilon}{\System}$ where:
 \begin{itemize}[label=--,leftmargin=1em,labelsep=*]
-\item for $i \in [0,l]$, $c_i \in \Cells$,
-\item $c_0 \in \InitCells$,
-\item for $i \in [0,l-1]$, $(c_i, c_{i+1}) \in \EdgeLabelingFunction$,
-\item $c_l$ is such that
-$\CellLabelingFunction(c_l) \cap \UnsafeStates \neq \emptyset$.
+\item for $i \in [0,l]$, $C_i \in \Cells$,
+\item $C_0 \in \InitCells$,
+\item for $i \in [0,l-1]$, $(C_i, C_{i+1}) \in \EdgeLabelingFunction$,
+\item $C_l$ is such that
+$\CellLabelingFunction(C_l) \land \unsafebmc(\vx)$ is satisfiable
+(i.e. the cell $C_l$ contains an unsafe state).
 \end{itemize}
@@
 \bigg( \bigwedge_{i=0}^{l}{
-  \subs{\CellLabelingFunction(c_i)}{\rvars{i}}{\rvars{}}
+  \subs{\CellLabelingFunction(C_i)}{\rvars{i}}{\rvars{}}
 } \bigg)
@@
   \subs{
-  \subs{\CellEdges((c_i,c_{i+1}))}{\rvars{i}}{\rvars{}}
+  \subs{\CellEdges((C_i,C_{i+1}))}{\rvars{i}}{\rvars{}}
   }{\rvars{i+1}}{\rvars{+}}
@@
 of linear predicates and hence its satisfiability can be casted as the
-feasibility of a linear program}.
+feasibility of a linear program.}.
 %
-This can proved showing that each model to the formula
+This can be proved by showing that each model of the formula
 $\concbmc{\abspathbmc}$  ``corresponds'' to a path of
 $\RTSAbstraction{\epsilon}{\System}$, and each path of
-$\RTSAbstraction{\epsilon}{\System}$ that traverse the sequence
-$\abspathbmc$ ``corresponds'' to a model of
-Formula~\ref{eq:simulateabs}.
+$\RTSAbstraction{\epsilon}{\System}$ that traverses the sequence of
+cells of $\abspathbmc$ ``corresponds'' to a model of 
+$\concbmc{\abspathbmc}$.
 
@@
 be a depth-first search (DFS) or a breadth-first search (BFS). As soon
-as $\abspathbmc$ is found, the algorithm checks for the satisfiability
+as a $\abspathbmc$ is found, the algorithm checks the satisfiability
 of $\concbmc{\abspathbmc}$. If  $\concbmc{\abspathbmc}$ is
 satisfiable, the algorithm returns a concrete path $\Pi$, otherwise
-the search continue trying to enumerate another abstract path. 
+the search continues trying to enumerate another abstract path. 
 The search is guaranteed to terminate, since the length of the
 abstract paths is at most $k$.
+
+On the one hand, a possible advantage of the explicit enumeration (2)
+over the symbolic BMC (1) is that it can explore the paths of
+$\RTSAbstraction{\epsilon}{\System}$ using different heuristics
+(e.g. a DFS instead of a BFS), and  
+that each concretization check is ``simpler'' than a single symbolic
+BMC query.
+%
+On the other hand, the symbolic BMC implicitly exploits SMT solvers,
+that automatically prune the search space (and intuitively, also the number 
+of abstract paths that must be visited) by learning conflict clauses.
 

### 2017-06-11 implementation.tex
.

@@
 compared it against S3CAM~\cite{zutshi2014multiple} on several
-academic benchmarks. Below, we present the details of the
-implementation and the benchmarks used.
+benchmarks. Below, we present the details of the implementation and
+the benchmarks used.
 
 \subsection{Implementation}
-PWA-Rel was implemented in Python. Its different parts comprised of
-the (a) abstraction graph generation, (b) regression analysis and (c)
-BMC. S3CAM was used to compute (a). To compute (b), OLS regression
-routines were used from Scikit-learn~\cite{pedregosa2011scikit}, a
-Python module for machine learning. Finally, the The bounded model
-checking (c) was done using the SAL model checker~\cite{SAL-SRI},
-which used Yices2~\cite{dutertre2014yices} as the SMT solver. A naive
-alternative linear programming (LP) based model checker, specific to
-the PWA relational abstractions was implemented utilizing PyGLPK.
-Specifically, each path of the graph $G^R$ was sent to the lp solver
-PyGLPK and checked for feasibility. If feasible, the checker returned
-a concrete counter-example.
 
-PWA-Rel uses S3CAM to discover the abstraction graph $G$, which is
-then filtered to remove the nodes from which error nodes are not
-reachable. In all our experiments we used $1$-relational modeling to
-create the transition system from $G$. This was determined to be a
-good trade-off between longer simulations and increase in precision,
-with respect to the mentioned benchmarks.
+PWA-Rel is implemented in Python. It uses S3CAM to generate the graph
+abstraction and regression routines from
+Scikit-learn~\cite{pedregosa2011scikit} to compute
+$\RTSAbstraction{\epsilon}{\System}$ of the given system $\System$.
+It then constructs relational abstraction and uses a bounded model
+checker to find a counter-example to the given property in
+$\RTSAbstraction{\epsilon}{\System}$. Finally, using simulations, it
+confirms if the counter-example is valid for $\System$, in which case
+a successfull falsification run is concluded.
 
-Using the BMC, we checked for the given safety property. If a concrete
+To check the transition system, PWA-Rel can either invoke symbolic
+bounded model checkers or explicit model checkers. For symbolic BMC,
+the off-the-shelf SAL modelchecker~\cite{SAL-SRI} is used. It is
+configured to encode the problem as a single query for a given depth
+$k$, where $k$ is the bound on the length of the counter-example. It
+was computed by the length of the abstract counter-examples found. SAL
+uses the SMT solver Yices2~\cite{dutertre2014yices} to solve the
+encoded SMT constraints.
+
+% Two symbolic BMC were tried to compare scalability; the off-the-shelf
+% SAL model checker~\cite{SAL-SRI} and SatExBMC. Both model checkers
+% were configured to encode the problem as a single query for the given
+% depth. SAL uses the SMT solver Yices2~\cite{dutertre2014yices}, while
+% SatExBMC uses SatEx~\cite{shoukry2017smc}. SatEx differs from Yices2
+% (and other sound SMT solvers) in two aspects, (a) it is a lazy SMT
+% solver and (b) uses approximate convex optimization routines (via
+% CPLEX). Although (b) makes it unsound, it is not a concern for our
+% falsification procedure as it is based off approximate models.
+
+Explicit model checking was implemented out using a naive DFS
+(depth-first-exploration) stratergy, wherein, each abstract path is
+enumerated and checked for feasibility using a an optimization
+routine. Intefaces to various solvers were implemented, including GLPK
+and Ipopt. If feasible, the returned concrete counter-example is
+simulated and checked for existance in $\System$. Also, PWA-Rel
+extracts the bounds on the initial parameters by formulating multiple
+optimization problems. The bounds can be regarded as an over-appr
+
+% If the existance check fails, multiple
+% parameter values can be sampled within the determined bounds to search
+% for counter-examples in the neighbhourhood of the returned concrete
+% counter-example.
+
+% PWA-Rel uses S3CAM to discover the abstraction graph $G$, which is
+% then filtered to remove the nodes from which error nodes are not
+% reachable. In all our experiments we used $1$-relational modeling to
+% create the transition system from $G$. This was determined to be a
+% good trade-off between longer simulations and increase in precision,
+% with respect to the mentioned benchmarks.
+% \todo{$1$-relational modeling was used}
+
+Using BMC, we checked for the given safety property. If a concrete
 trace was obtained, it was further checked for reproducibility by
@@
 \subsection{Evaluation}
-We tabulate our preliminary evaluation in \tabref{res-rel}. We used
-the more difficult benchmarks described in \cite{zutshi2014multiple},
-including the Van der Pol oscillator, Brusselator, Lorenz attractor
-and the Navigation benchmark. PWA-Rel and S3CAM were tested $10$ times
-with different seeds and their running times were averaged. We tabulate
-both the total time taken by PWA-Rel+SAL and PWA-Rel+LP to compute the
-counter-example and compare against an implementation of S3CAM.
-Different abstraction parameters are used for S3CAM and PWA-Rel, due
-to the differences in which they operate. PWA-Rel requires a finer
-abstraction to begin with in order to minimize the modeling errors.
-On the other hand, a coarser abstraction as the initial abstraction is
-works better for S3CAM due to the expensive refinement procedure
-involved.
+
+We benchmarked our implementation and compared it with S3CAM.  The
+preliminary results are tabulated in~\tabref{res-rel}. We used few of
+the benchmarks described in \cite{zutshi2014multiple}, including the
+Van der Pol oscillator, Brusselator, Lorenz attractor and the
+Navigation benchmark. We also tried PWA-Rel on the insulin
+benchmark~\cite{chenformal} and lane change maneuver
+benchmark~\cite{althoff2012reachability}. PWA-Rel was compared with
+S3CAM by running both $10$ times with different seeds (as both use
+randomized algorithms), and their running times were averaged. PWA-Rel
+was evluated with both the symbolic BMC SAL (PWA-Rel+SAL) and the
+explicit BMC (PWA-Rel+OPT).  \todo{SatEx was also tried, but ...}. We
+tabulate both the total time taken by PWA-Rel+SAL and PWA-Rel+OPT to
+compute the counter-example and compare against an implementation of
+S3CAM.  Different abstraction parameters are used for S3CAM and
+PWA-Rel, due to the differences in which they operate. PWA-Rel
+requires a finer abstraction to begin with in order to minimize the
+modeling errors. On the other hand, a coarser abstraction as the
+initial abstraction works better for S3CAM due to the expensive
+refinement procedure involved.~\footnote{In all our experiments we
+used $1$-relational modeling to create the transition system from $G$.
+This was determined to be a good trade-off between longer simulations
+and increase in precision, with respect to the mentioned benchmarks.}
 

### 2017-06-11 prelims2.tex

--- 2017-04-08 prelims.tex
@@
 
-\subsection{Learning Dynamics using Simple Linear Regression}
+\subsection{Learning Dynamics using Linear Regression}
 
@@
 $\x'$ vectors. The notion of `best' is formally captured using a loss
-function which can be non-linear. We use the commonly used loss
-function \textit{ordinary least squares} (OLS) for this presentation.
+function (which can be non-linear), resulting in different kinds of
+regression analysis including ridge, lasso, huber, generelized and
+many other regressions. For simplicity, we present the commonly used
+loss function \textit{ordinary least squares} (OLS) in the next
+section.
+~\footnote{The choice of regression analysis should ideally be based on the
+quality of data. As we assume the data (from the simulator) to be
+noise free, the affecte on result of the falsification search is
+minimal.
+Though it is important,  not central to the
+discussion as long as the learnt models  can be used to learn different
+models}.
 
@@
 need to find the `best' choice for $\vec{a}$ and $b$. This choice is
-formally defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine $A$ and $B$, such
-that the least square error of the affine predictor is minimized for
-the given data set.
+formally defined using a loss function. For the case of OLS, the loss
+function is the sum of squares of the errors in prediction.  The task
+is then to determine $A$ and $B$, such that the least square error of
+the affine predictor is minimized for the given data set.
 

### 2017-06-11 implementation.tex
.

@@
 % model checker.
+
+\mypara{The search parameters} for PWA-Rel include the parameters of
+S3CAM: $N$, $\epsilon$ and $\Delta$. Additionally, they also include
+the maximum error budget $\delta_{max}$ for OLS and the maximum length
+of segmented trajectory for building $k$-relational models.  The
+effects of $N$, $\epsilon$ and $\Delta$ on S3CAM's performance were
+detailed in~\cite{zutshi2014multiple}. However, they also have an
+effect on relational modeling. A finer grid with small cells produces
+more accurate models than a coarser grid with large cells. Similarly,
+small time length trajectory segments result in more accurate
+modeling.
+
+%An increase in the number of samples can increases the
+%accuracy of regression.
+
+% \paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
+% Given a $\delta_{max}$, we keep increasing $k$ during the
+% $k$-relational modeling process till $\delta_i \le \delta_{max}$ is
+% satisfied. A $k_{max}$ is introduced to bound the longest
+% segmented trajectory which can be considered.
+
+% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
 
@@
 
-% more scalable 
-\subsection{Search Parameters}
+\subsection{Reasons for Failure}
 
-The search parameters for PWA-Rel include the parameters of S3CAM:
-$N$, $\epsilon$ and $\Delta$. Additionally, they also include the
-maximum error budget $\delta_{max}$ for OLS and the maximum length of
-segmented trajectory for building $k$-relational models.
-
-The effects of $N$, $\epsilon$ and $\Delta$
-on S3CAM's performance were detailed in~\cite{zutshi2014multiple}. However, they also have an effect on
-relational modeling. A finer grid with small cells produces more
-accurate models than a coarser grid with large cells. Similarly,
-small time length trajectory segments result in more accurate
-modeling.
-%An increase in the number of samples can increases the
-%accuracy of regression.
-
-\paragraph{Maximum Model Error ($\boldsymbol{\delta}_{\bold{max}}$) .}
-Given a $\delta_{max}$, we keep increasing $k$ during the
-$k$-relational modeling process till $\delta_i \le \delta_{max}$ is
-satisfied. A $k_{max}$ is introduced to bound the longest
-segmented trajectory which can be considered.
-
-% \paragraph{Bounded k relations ($\boldsymbol{k}_{\bold{max}}$.) }
-
-\subsection{Reasons for Failure}
-The approach can fail to find a counter-example, even when it exists, in
-one of three ways.
-
-\begin{itemize}
-
-    \item \emph{S3CAM Fails}: No abstract counter-example is found by
-        S3CAM. We can remedy it by increasing search budgets and/or
-        restarting.
-
-    \item \emph{BMC Fails}: An abstract counter-example is found, but
-        the BMC fails to find a concrete counter-example in the PWA
-        relational model. The failure can be attributed to either (a)
-        a spurious abstract counter-examples or (b) a poorly estimated
-        model. The former can be addressed by restarting but the
-        latter requires that the maximum model error $\delta_{max}$ be
-        decreased.
-
-    \item \emph{Inaccurate PWA Modeling}: An abstract counter-example
-        is found, and it is successfully concretized in the PWA
-        relational model by the BMC. However, it is not reproducible
-        in the black box system. This happens when the PWA relational
-        model is not precise enough.
-\end{itemize}
+The approach can fail to find a counter-example, even when it exists,
+in one of three ways. As it is dependant on S3CAM to find abstract
+paths, it can fail when (a) S3CAM fails to find abstract
+counter-examples. We can remedy it by increasing search budgets and/or
+restarting. Even when abstract counter-examples are found, they may
+not be \emph{concretizable}. In such a case, the (b) BMC search fails.
+The failure can be attributed to either spurious abstract
+counter-examples or a poorly estimated model. The former can be
+addressed by restarting but the latter requires that the maximum model
+error $\delta_{max}$ be decreased. Finally, the (c) counter-example
+from the BMC might not be reproducible in the black box system. This
+happens if the relational model is not precise enough.

### 2017-06-12 bmc.tex
rearranged sections
.

@@
 
-% % include an algorithm
+\todo{include an algorithm}
 % We now describe the entire procedure in three steps.
@@
 % \end{enumerate}
+
+\begin{algorithm}[t]
+\DontPrintSemicolon
+\caption{PWA-Rel\label{algo:pwa-rel}}
+\KwIn{$\System$, $\simmap_\Delta$, $\epsilon_0$ , Smallest Precision $\epsilonmin$}
+$\epsilon \assign \epsilon_0$ \;
+\While{$\epsilon \ge \epsilonmin$}{
+    $\TSAbstraction{\epsilon}{\System}$ $\assign$
+    $\mathsf{ScatterAndSimulate}$($\System$,$\simmap_\Delta$,$\epsilon$)
+    \nllabel{algoline:construct} \;
+    $\Pi$ $\assign$ $\mathsf{BoundedModelCheck}$($\TSAbstraction{\epsilon}{\System}$)
+    \nllabel{algoline:bmc} \;
+    \lIf {$\Pi = \emptyset$}{ return FAIL \nllabel{algoline:nocex} }
+    \lIf {$\mathsf{checkIfConcretizable}(\Pi)$}{ return VIOLATION \nllabel{algoline:checkspurious} }
+    $\epsilon$ $\assign$ $\frac{\epsilon}{10}$ \nllabel{algoline:refine} \;
+}
+\end{algorithm}
 
@@
 
-\begin{table}[h]
-\centering
-\caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
-\label{tab:vdp-pwa}
-\begin{tabular}{@{}ccccc@{}}
-\toprule
-$C$ & $C'$ & $A$ & $b$ & $\delta$\\
-\midrule
-(0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-(0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
-\midrule
-(-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-(-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
-\midrule
-(-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
-\midrule
-(-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
-\midrule
-(-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
-\bottomrule
-\end{tabular}
-\end{table}
+% Does table add anything to the discussion?
+%
+% \begin{table}[h]
+% \centering
+% \caption{PWA model computed using OLS. The affine model for each edge $(C,C')$ in the graph \figref{vdp-graph} is given by $x'\in Ax + b + \delta$, where $\delta$ is a vector of intervals.}
+% \label{tab:vdp-pwa}
+% \begin{tabular}{@{}ccccc@{}}
+% \toprule
+% $C$ & $C'$ & $A$ & $b$ & $\delta$\\
+% \midrule
+% (0, -1)   & (0, -1)   & $\begin{bmatrix}0.99& 0.12 \\-0.12&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
+% (0, -1)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.10&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
+% \midrule
+% (0, -2)   & (0, -2)   & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.63\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
+% (0, -2)   & (0, -4)   & $\begin{bmatrix}0.99& 0.12 \\-0.07&1.62\end{bmatrix}$&$\begin{bmatrix}0\\0\end{bmatrix}  $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
+% \midrule
+% (0, -9)   & (-1, -15) & $\begin{bmatrix}0.99& 0.12 \\-0.09&1.61\end{bmatrix}$&$\begin{bmatrix}0\\-0.04\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
+% \midrule
+% (-1, -15) & (-3, -24) & $\begin{bmatrix}0.95& 0.12 \\-1.29&1.47\end{bmatrix}$&$\begin{bmatrix}-0.01\\-0.41\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
+% \midrule
+% (-3, -24) & (-4, -29) & $\begin{bmatrix}1.11&-0.17 \\-1.71&0.66\end{bmatrix}$&$\begin{bmatrix}-1.07\\-3.31\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.04]\\ [-0.07, 0.09]\end{bmatrix}$\\
+% \midrule
+% (-4, -29) & (-4, -29) & $\begin{bmatrix}0.99& 0.01 \\-0.41&1.02\end{bmatrix}$&$\begin{bmatrix}0\\-0.28\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
+% \midrule
+% (-1, -10) & (-2, -16) & $\begin{bmatrix}0.97& 0.12 \\-0.71&1.56\end{bmatrix}$&$\begin{bmatrix}0\\-0.11\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
+% (-1, -10) & (-2, -15) & $\begin{bmatrix}0.97& 0.12 \\-0.70&1.58\end{bmatrix}$&$\begin{bmatrix}0\\-0.08\end{bmatrix} $&$\begin{bmatrix}[0, 0]\\ [0, 0.01]\end{bmatrix}$\\
+% \midrule
+% (-2, -16) & (-3, -23) & $\begin{bmatrix}0.92& 0.12 \\-1.84&1.39\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.73\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
+% \midrule
+% (-3, -23) & (-5, -29) & $\begin{bmatrix}3.14&-0.81 \\-0.96&0.23\end{bmatrix}$&$\begin{bmatrix}-3.13\\-4.99\end{bmatrix}$&$\begin{bmatrix}[-0.04, 0.03]\\ [-0.03, 0.04]\end{bmatrix}$\\
+% \midrule
+% (-2, -15) & (-3, -22) & $\begin{bmatrix}0.93& 0.12 \\-1.75&1.40\end{bmatrix}$&$\begin{bmatrix}-0.02\\-0.68\end{bmatrix}$&$\begin{bmatrix}[0, 0]\\ [0, 0]\end{bmatrix}$\\
+% \midrule
+% (-3, -22) & (-5, -29) & $\begin{bmatrix}3.08&-0.72 \\-1.69&0.40\end{bmatrix}$&$\begin{bmatrix}-2.77\\-4.57\end{bmatrix}$&$\begin{bmatrix}[-0.02, 0.02]\\ [-0.01, 0.02]\end{bmatrix}$\\
+% \bottomrule
+% \end{tabular}
+% \end{table}
 

### 2017-06-12 learning.tex
.

@@
 time as well as storage.  The central idea in this paper is to {\em
-summarize} several tuples as a single affine map obtained using linear
-regression.  In other words, we can replace several ``proximal''
-tuples $(\vx,\vx')$ in $\reach{t,u}$ by an affine map of the form
-$\vx' = A\vx + B\vu + \delta$, where $A$ and $B$ are matrices and
-$\delta$ is a worst-case error estimate.
+summarize} several tuples as a single map obtained using regression.
+In other words, we can replace several ``proximal'' tuples
+$(\vx,\vx')$ in $\reach{t,u}$ by a map $F$ of the form $\vx' = F(\vx)
++ \delta$, where $\delta$ is a worst-case error estimate. The form of
+$F$ depends on the kernel used for regression. For the special case of
+linear regression, the kernel is affine and the resulting map can be
+represented by a matrix $A$ as $\vx' = A\vx + \delta$.
 
-In statistics, linear regression is the problem of finding an affine
+In statistics, regression is the problem of finding an affine
 function or \textit{predictor}, which can `best' summarize the
@@
 function \textit{ordinary least squares} (OLS) for this presentation.
+
+If $N>n$, which is the case in the current context of \textit{finding
+the best fit}, the problem is over-determined: there are more
+equations than unknowns. Hence, a single affine function cannot be
+found which satisfies the equation \eqref{eq:affinemap}. Instead, we
+need to find the `best' choice for $\vec{a}$ and $b$. This choice is
+formally defined using a loss function. For the case of simple linear
+regression or OLS, the loss function is the sum of squares of the
+errors in prediction.  The task is then to determine $A$ and $B$, such
+that the least square error of the affine predictor is minimized for
+the given data set.
 
@@
 Here, $A$ is a $n\times n$ matrix, and $B$ is a $n\times 1$ matrix.
-
-If $N>n$, which is the case in the current context of \textit{finding
-the best fit}, the problem is over-determined: there are more
-equations than unknowns. Hence, a single affine function cannot be
-found which satisfies the equation \eqref{eq:affinemap}. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This choice is
-formally defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine $A$ and $B$, such
-that the least square error of the affine predictor is minimized for
-the given data set.
-
-\begin{equation}
-    \operatornamewithlimits{argmin}_{A,B}\displaystyle\sum_{i=1}^{N}{\left(\x' - (A\x + B)\right)^2}
-\label{reg}
-\end{equation}
-
-To ease the presentation, we can rewrite the above as a homogeneous
-expression by using $\hat{\x} = [\x^T\ 1]^T$ and replacing $A$ and $B$ by the vector
-$\vec{A} = [A\ B]$. The equation~\ref{reg} now becomes
-\[
-\operatornamewithlimits{argmin}_{\vec{A}}\displaystyle\sum_{i=1}^{N}{(\x' - \vec{A}^T \hat{\x})^2}
-\]
-
-The solution of OLS can be analytically computed as the Moore-Penrose
-pseudo-inverse: $\vec{A} = (X^TX)^{-1}X^T X'$, where $X$ is the matrix
-representing the horizontal stacking of all $\hat{\x}$ and $X'$
-represents the stacking of all $\x'$ values.  The details can be found
-in standard texts on learning and statistics
-\cite{friedman2001elements}. Alternative methods based on gradient
-descent can also be used to find a suitable $A$ matrix.
+The details can be found in standard texts on learning and
+statistics~\cite{friedman2001elements}.
 
@@
 = \vec{A}\hat{\x}$.
-
-% where
-% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T
-% \end{bmatrix}$ and $\vb = \begin{bmatrix}b_1 \\ \vdots \\ b_n
-% \end{bmatrix}$.
 
@@
 section.
+
+% \begin{equation}
+%     \operatornamewithlimits{argmin}_{A,B}\displaystyle\sum_{i=1}^{N}{\left(\x' - (A\x + B)\right)^2}
+% \label{reg}
+% \end{equation}
+
+% To ease the presentation, we can rewrite the above as a homogeneous
+% expression by using $\hat{\x} = [\x^T\ 1]^T$ and replacing $A$ and $B$ by the vector
+% $\vec{A} = [A\ B]$. The equation~\ref{reg} now becomes
+% \[
+% \operatornamewithlimits{argmin}_{\vec{A}}\displaystyle\sum_{i=1}^{N}{(\x' - \vec{A}^T \hat{\x})^2}
+% \]
+
+% The solution of OLS can be analytically computed as the Moore-Penrose
+% pseudo-inverse: $\vec{A} = (X^TX)^{-1}X^T X'$, where $X$ is the matrix
+% representing the horizontal stacking of all $\hat{\x}$ and $X'$
+% represents the stacking of all $\x'$ values.  The details can be found
+% in standard texts on learning and statistics
+% \cite{friedman2001elements}. Alternative methods based on gradient
+% descent can also be used to find a suitable $A$ matrix.
+
+% In this fashion, we can we can use OLS to approximate its trajectories
+% of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
+% = \vec{A}\hat{\x}$.
+
+% where
+% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T
+% \end{bmatrix}$ and $\vb = \begin{bmatrix}b_1 \\ \vdots \\ b_n
+% \end{bmatrix}$.
+
+% We note that linear regression tools are often able to provide an
+% $n\times 1$ vector $\vec{\delta}$ of error intervals, where the
+% $i^{th}$ entry of the form $[{\delta_i}_{\min},{\delta_i}_{\max}]$
+% indicates the best and worst possible error in computation of the
+% $i^{th}$ state in $\x'$.
 

### 2017-06-12 bmc.tex
rough restructuring

@@
 \end{algorithm}
-
-\section{An Example: Van der Pol Oscillator}
-
-We now describe the presented technique using the Van der Pol
-oscillator benchmark from~\cite{zutshi2014multiple}. Simulations
-generated using uniformly random samples are shown in
-\figref{vdp-cont}. We want to check the property $P_3$, indicated by
-the red box, given the initial set indicated by the green box. The
-abstraction is defined by $\quant_{0.2}(\x)$, which results in an
-evenly gridded state-space, where each cell is of size $0.2 \times
-0.2$ units.  Scatter-and-simulate is then used to construct the
-abstract graph $G$, using $2$ samples per cell and the time step
-$\Delta = 0.1s$. The complete process follows.
-
-\begin{enumerate}
-\item{\emph{Abstract}}: Assume an implicit abstraction using the quantization
-        function $\quant_{\epsilon=0.2}(\x)$. The corresponding reachability
-    graph is $\scr{H}(\Delta=0.1)$.
-\item{\emph{Discover}}: Using scatter-and-simulate, enumerate a finite number
-    of cells (vertices) and the relations (edges) of the graph
-    $\scr{H}(0.1)$. Associate the set of generated trajectory
-    segments with their respective originating cells using a map $\ds:C
-        \mapsto \setof{(\vx,\vx') |\vx \in C}$. The discovered
-    abstraction is show in \figref{vdp-abs}. The red cells
-    are unsafe and green cells are initial cells.
-\item{\emph{Relationalize}}: For each cell $C$, perform regression analysis on
-    the respective set of trajectory segments $D(C)$, and compute a
-    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
-    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(0.1)$. Annotate
-    each edge in the graph $\scr{H}(0.1)$ with the respective
-    relation.
-
-    \figref{vdp-map} shows the cells and the trajectory segments, which
-    are part of the data sets constructed using the $1$-relational
-    modeling. Against each cell, its unique identifier (integer
-    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
-    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
-    relations. Note that self loops result when an observed trajectory
-    segment has its $start$ and $end$ states in the same cell.
-
-\item{\emph{Model Check}}: The graph $\scr{H}(0.1)$ can now be viewed as a
-    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
-    where $C \in vertices(\scr{H}(0.1))$ and $\x \in
-    \HybridStates$. A transition $\tau \in \scrT$ is of the form
-    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
-    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
-    edge(\scr{H}(0.1))}$.
-\item{\emph{Check Counter-example}}: The infinite state (but finite location) transition system can
-    be model checked to find a concrete counter-example, which if
-    exists, can indicate the existence of a similar trace in the
-    original black-box system $\System$. The latter check is carried
-    out as before, using the numerical simulation function
-        $\simulate$~\footnote{For the given example, the model checker is unable to
-        find a counter-example.}.
-\end{enumerate}
-
-\inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
-Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}{vdp-cont}
-
-\inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
-abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
-initial cells.}{vdp-abs}
-
-\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
-trajectory segments used by $1$-relational modeling.}{vdp-map}
-
-\inclfig[width=0.50\linewidth]{vdp_graph.pdf}{Enriched
-graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}{vdp-graph}
 

### 2017-06-12 bmc_section.tex
rough restructuring

@@
 
+\section{An Example: Van der Pol Oscillator}
+
+\mypara{\emph{Model Check}}: The graph $\scr{H}(0.1)$ can now be viewed as a
+    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
+    where $C \in vertices(\scr{H}(0.1))$ and $\x \in
+    \HybridStates$. A transition $\tau \in \scrT$ is of the form
+    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
+    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
+    edge(\scr{H}(0.1))}$.
+\mypara{\emph{Check Counter-example}}: The infinite state (but finite location) transition system can
+    be model checked to find a concrete counter-example, which if
+    exists, can indicate the existence of a similar trace in the
+    original black-box system $\System$. The latter check is carried
+    out as before, using the numerical simulation function
+        $\simulate$~\footnote{For the given example, the model checker is unable to
+        find a counter-example.}.
+
+\inclfig[width=0.50\linewidth]{vdp_graph.pdf}{Enriched
+graph $G^R$. The affine maps $f$ for the transition relations are show
+in \tabref{vdp-pwa}.}{vdp-graph}
+
 %% Temporal properties of PWA transitions systems can be checked using

### 2017-06-12 scam.tex
rough restructuring

@@
 
+\section{An Example: Van der Pol Oscillator}
+\inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
+Pol: continuous trajectories. Red and green boxes indicate unsafe and
+initial sets.}{vdp-cont}
+\inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
+abstraction $\scr{H}(0.1)$. Red cells are unsafe cells and green cells are
+initial cells.}{vdp-abs}
+
+We now describe the presented technique using the Van der Pol
+oscillator benchmark from~\cite{zutshi2014multiple}. Simulations
+generated using uniformly random samples are shown in
+\figref{vdp-cont}. We want to check the property $P_3$, indicated by
+the red box, given the initial set indicated by the green box. The
+abstraction is defined by $\quant_{0.2}(\x)$, which results in an
+evenly gridded state-space, where each cell is of size $0.2 \times
+0.2$ units.  Scatter-and-simulate is then used to construct the
+abstract graph $G$, using $2$ samples per cell and the time step
+$\Delta = 0.1s$. The complete process follows.
+
+\mypara{Abstract}: Assume an implicit abstraction using the quantization
+        function $\quant_{\epsilon=0.2}(\x)$. The corresponding reachability
+    graph is $\scr{H}(\Delta=0.1)$.
+
+\mypara{Discover}: Using scatter-and-simulate, enumerate a finite number
+    of cells (vertices) and the relations (edges) of the graph
+    $\scr{H}(0.1)$. Associate the set of generated trajectory
+    segments with their respective originating cells using a map $\ds:C
+        \mapsto \setof{(\vx,\vx') |\vx \in C}$. The discovered
+    abstraction is show in \figref{vdp-abs}. The red cells
+    are unsafe and green cells are initial cells.
+
 \subsection{CEGAR for falsifying $\TSAbstraction{\epsilon}{\System}$}
@@
 rather than increasing $\epsilon$.
+
+\section{An Example: Van der Pol Oscillator}
+\todo{takes four iterations}
+\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
+trajectory segments used by $1$-relational modeling.}{vdp-map}
 

### 2017-06-12 abstract.tex
.

@@
 falsification and testing approaches can scale only by numerically
-simulating them, and hence treating the systems as black-boxes. Common
-approaches like sampling-based falsification, use numerical
-simulations to search for violations of safety properties. However,
-this relegates them to the `best effort' category. On the other hand,
-\emph{rigorous} verification (and falsification) techniques require
-symbolic models and are comparatively less scalable. We try to bridge
-this gap by incorporating data-driven models in a sampling based
-search. Using an efficient, on-the-fly discrete abstraction
-search procedure, we compute models only for selective regions of the state
-space, thus `enriching' the abstraction. We then show how the
-`enriched' abstraction can be used to guide the search for violations
-without refinement.
+simulating them, and hence treat the systems as black-boxes. Common
+approaches like sampling-based falsification use numerical simulations
+to search for violations of safety properties. However, this relegates
+them to the `best effort' category. On the other hand, \emph{rigorous}
+verification (and falsification) techniques require symbolic models
+and are comparatively less scalable. We try to bridge this gap by
+incorporating data-driven models in a sampling based search. Using an
+efficient, on-the-fly discrete abstraction search procedure and
+regression, we selectively model the behavior of the system in
+sub-regions of the state space. We then incorporate the model in the
+discrete abstractio, thus, `enriching' the abstraction. We show how
+this `enriched' abstraction can be used to guide the search for
+violations without refinement.  The automatic falsification proceeds
+by (a) randomly exploring the state-space abstraction using
+numerically computed trajectories and enumerating abstract violations,
+(b) modeling the behavior of the system along these abstract
+violations by using discrete-time piecewise \emph{relational} models,
+and finally, (c) encoding the search for a concrete violation as a
+bounded model checking (BMC) problem.  The BMC query is a disjunction
+of conjunctions of constraints (i.e. a combination of constraint
+problems over real arithmetic). It can be solved using off-the-shelf
+SMT solvers or by multiple runs of an optimization solver (which
+checks feasibility). For any counter-example found in this `enriched'
+abstraction, we check the existence of a corresponding violation in
+the original system. We demonstrate the technique on a few benchmarks.
+
 %We use grid-based discrete abstractions to partition
@@
 %behavior along these violations and search for concrete violations.
-The automatic falsification proceeds by (a) randomly exploring the
-state-space abstraction using numerically computed trajectories and
-enumerating abstract violations, (b) modeling the behavior of the
-system along these abstract violations by using discrete-time
-piecewise affine (PWA) \emph{relational} models, and finally, (c)
-encoding the search for a concrete violation as a bounded model
-checking (BMC) problem.  The BMC query is then a disjunction of
-conjunctions of linear constraints (i.e. combinations of linear
-programs). It can be solved using off-the-shelf SMT solvers or by
-multiple runs of a linear programming solver. For any counter-example
-found in this `enriched' abstraction, we check the existence of a
-corresponding violation in the original system. We demonstrate the
-efficacy of our technique on a few benchmarks.
 

### 2017-06-12 appendix.tex
added reviewers response as appendix

Plan of action:
---------------------------------------------

- What is the paper saying/should be saying?
- Simplify notations....pwa relational and k-relational

- Modify ordering of sections: Abstraction -> PWA enrichment
- Simplify notations
- Fold multiple PWA relational/transitions system into one and use the
  transitions system as a semantic interpretation for model checking.
  Can be eliminated if automatons/kripke structures are used instead.
- Fix the `convex-hull' in the definition 2.2 PWA Relational Model.
  Showcase clustering using a benchmark. Add it to future work
  and cite Nimit's  paper. Or, retract it entirely?
- Clearly state that the approach is not a generic data-drive system
  identification technique, but focused on the falsification aspect.
  We only model selective state-space regions: abstract violations.
- De-emphasize OLS (and the process and the cost functions used for
  modeling). Add that the idea hinges on selectively identifying the
  state-space for modeling and not how the models might be created.
  Emphasize again that the requirement is only on the linearity of
  the individual models and hence, OLS is one way. Trade-off between
  efficiency and complexity can be achieved by selecting the `right'
  modeling template.
- Add benchmarks: hybrid with multiple modes.

Summary of the reviews
---------------------------------------------

R1: Like the idea but the presentation and results need improvement

- Clarify that the plant is inherently a hybrid dynamical system
  and in the future (or if time permits) a controller can be added and
  analyzed together as a infinite state transition system.
- Clarify the clustering done.

R2: Seems the most positive, but focuses on PWA identification.

- Clarify the purpose of the approach is to use sub-optimal
  identification techniques but still provide a `reasonable' and
  scalable falsification technique.

R3: Poor benchmarks: use automotive control examples, try
polynomial kernels (with Z3), compare with simulation based
falsification (S-Taliro?).

- Try polynomial kernels on a benchmark?
- Clarify S3CAM was already compared with S-Taliro.

R4: Probably the most negative review.
Sloppy presentation; high level concepts are not well presented
and contributions are not clear. Benchmarks are too simple. Why OLS?
Contributions are not significant/novel.

- Spell out the contributions clearly.
- Use Simulink models.
- Cite papers which learn hybrid automata from i/o traces, and clarify
  the difference.
  \cite{medhat2015framework} and many others
- Clarify that OLS is just one choice out of many.

R5: Probably a borderline.

- Add a motivation section showcasing the contributions.

--
============================================================================
EMSOFT 2017 Reviews for Submission 159
============================================================================

Title: Piece-Wise Relational Models for Falsifying Safety Properties of Hybrid Systems

Authors: Aditya Zutshi, Sergio Mover, Sriram Sankaranarayanan and Jyotirmoy Deshmukh
============================================================================
                            REVIEWER 1
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 2
                       Originality (1-4): 4
               Technical Soundness (1-4): 2
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

This paper is about checking safety properties on continuous models of an
embedded controller's environment. The continuous model is provided as a
black-box function capable of simulating Δ seconds from a given state with
constant input values. From what I can understand: (i) the state-space of
the system is explored by running multiple simulations; (ii) the resulting
states are clustered based on a quantization function;

(iii) an affine
relation (the convex hull) is calculated to characterize each cluster;

[TODO: Fix this]

(iv)
linear regression analysis is used to characterize the transitions between
clusters as affine relations, (iv) A finer abstraction is produced by
distinguish trajectories that initially pass through the same (k - 1)
clusters but then end up in different clusters, and (v) safety properties of
the resulting discrete "relational" abstraction are analyzed using SMT-based
bounded model checking. The work does not aim at soundness or completeness,
but rather to find problems in real embedded systems. The technique is a
variation of one proposed at EMSOFT in 2014 ([28]). It is demonstrated on
the academic examples from [28] with mixed results.

   ============
   [TODO: Add an explicit controller+plant benchmark: fuzzy control?]
   ============

I am not an expert in this area, but it seems to me that this paper presents
some novel and interesting ideas. I found, however, that after a great
introduction, the presentation becomes quite frustrating. There are a
succession of different formalisms and notations, which often lack precision
and explanations, and the links between them are not clearly described (I
had to battle to write the above summary and I am not certain that it is
correct). And, even after a careful reading of the text, I am still unsure
about one or two fundamental details:

1. The title states that the paper addresses "hybrid systems" and the system
   model includes discrete states, but thereafter the paper only treats
   continuous systems. Apparently this is reasonable because of an
   assumption that the model is a "switched-mode dynamical system".

   It would be nice to define exactly what you mean by "switched-mode
   dynamical system". Does it mean that only the controller model makes
   discrete transitions and that the environment is a continuous model with
   no discrete events or discontinuities?

   ============
   The systems under test are assumed to be switched-mode to ease the
   presentation, but it is not a limitation of the presented
   technique. Black box systems which have hybrid dynamics due to
   multiple modes, are dealt in the same fashion.
   [TODO: $K$-refinement to differentiate the dynamics of different modes.
   We have added a benchmark to showcase this.
   [TODO: which one?]
   [TODO: clarify the absence of any controllers]
   [TODO: remove or clarify switched-mode systems]
   [TODO: add a hybrid benchmark]
   ============

   I imagine that the idea is to combine the generated discrete abstraction
   with a discrete controller model to verify both together. It would be
   good to state this clearly and to treat the resulting issues throughout
   the paper. I note that the experimental results do not treat this issue.
   Otherwise, why not change the title and focus the text on what you
   actually did?

   ============
   We will explore this in the future, but the current presentation
   only works with the plant, which is itself a hybrid dynamical
   system.
   [TODO: Add this in the conclusion/future work]
   ============

   (Similar remarks hold at a smaller scale. The discussion often describes
    what "can be" done, e.g., in sections 3.2 and 4.1. I would rather read
    about what "has been" done and what the result was!)

   ============
   [TODO]
   ============

2. Why do you need a bounded model checker to check safety invariants when
   you already enumerate and characterize the reachable states? That is, why
   not just test the invariant on the convex hulls as they are generated?

   My guess is that you later want to restrict the state-space by composing
   your abstraction with a controller model. Is that right? This probably
   seems obvious to you, but please put yourself in the place of someone who
   is new to the work and not an expert in the domain! In any case, this
   idea is not clearly stated, not addressed by the text, and not considered
   in the experimental evaluation.

   ============
   [TODO: I don't understand this comment, but it is related to
   point (iii) of the summary]
   Although we enumerate the abstract space, we require the model
   checker to find a concrete counter-example.
   ============

   Is this why the PWA transition system characterizes states by a valuation
   function v and not simply as tuples? If yes, how will this work? I do not
   understand why you have a valuation function when the system only has one
   variable (x).
   ============
   [TODO]
   ============

Other comments and suggestions

* The paper puts a certain emphasis on "learning" a model. Would it not be
  more accurate to talk about "approximating" or "abstracting"?
   ============
   [TODO: Seems reasonable to me. Any suggestions??]
   ============

* It would be great to include an "overview figure" that lists all of the
  different formalisms and explains how they fit together. The enumeration
  in section 5 is very helpful, but it introduces even more notation and
  does not completely describe how all of the other graphs and transition
  systems are related.
   ============
   [TODO]
   ============

INTRODUCTION

* "Automatic approaches for testing the safety and performance of such
closed-loop systems is challenging," ->
  "Developing automatic approaches for testing the safety and performance of
such closed-loop systems is challenging," ?

* "Related approaches using bisimulation relations [20, 25, 27] have
interesting parallels with the work presented here.": What does this sentence
mean exactly?

LEARNING PWA RELATIONAL MODELS

* "The system statespace (denoted X) is a subset of $Q × R^n"$
-> "The system statespace, denoted X, is a subset of $Q × R^n"$

* "An input to S is assumed to be a time-bounded signal, i.e., a function
  from [0,T] to U, with T as the time horizon, and $U \subseteq R^k$.": Why not just
  use functions from [0,T] to $R^k$? What do you gain by introduction U (it is
  not exploited in the rest of the paper)?

* "we view S as effectively black-box"
-> "we view S as a black-box"

* "equipped with a forward simulation map, in a simulation time-step"
-> "equipped with a forward simulation map, which in a simulation time-step"

* "allows computing the unique reachable state x ∈ X"
-> "allows computing the unique reachable state x' ∈ X" (?)

* "the forward simulation map $SIMS^S_Δ : X × U × R^+ ↦ X": why R^+$?

* Why is Δ sometimes subscripted with t? Do you use a constant step-size or
  not? The text could be clearer.

* "Also, to ease the presentation, we conflate the inputs with the state and
  denote them by x.": This sentence could be clearer. I think you mean to
  say that you consider that the input values are included in the continuous
  state x. That said, you keep using u at the start of section 2.1.

* "several “proximal” tuples": "proximate"?

* "In statistics, linear regression is the problem of finding"
-> "Linear regression is the problem of finding"

* "For rest of the section, we elide" -> "For the rest of the paper, we
  elide" ?

* "The equation 2 now becomes"
-> "Equation 2 now becomes"

* "In this fashion, we can we can use OLS to approximate its trajectories of
  system S at fixed time step"
-> "In this fashion, we can use OLS to approximate the trajectories of
system S at a fixed time step"

* The definition of a "PWA transition system" is not a numbered definition
  (while the definition of a PWA relation model is). The "move function"
  seems to define the semantics of the transition systems. This could be
  more clear. You use G here for a set of conjunctions of affine predicates,
  and later it is used for graphs.

* "and use the name a variable as a"
-> "and use the name of the variable as a"

* "We how we can use the formal machinery of a PWA transition"?

* "S defined by a simulator SIMS∆~\footnote{...}."
-> "S defined by a simulator SIMS∆.\footnote{...}"

* Definition 2.2 is not precise enough. Why not systematically state the
  values of $L, x, T, G, U, l_0, Θ$? This would aid the poor reader by
  specifying the initial state and the set of locations. What is the set of
  locations?

* "the the valuation function for variable x"
-> "the valuation function for variable x"

* "A PWA model will be usually non-deterministic,"
-> "A PWA model will usually be non-deterministic,"

* "Abusing the notation,"
-> "Abusing notation,"

* "a single step in ρ by x' = ρ(x)": this abuse of notation is regrettable.
  It would be clearer and more precise, I believe, to avoid it.

* "iff for every state $x ∈ X^n": Why "^n"$?

* "In what follows, we use the terms PWA relational model and PWA
   transition system interchangeably to mean PWA relational model.": Please
   choose one term and use it consistently.

* "As all the guards and transitions are affine, a bounded reachability
  query is equivalent to checking all combinations of discrete locations,
  where each combination can be summarized as a linear program.": I did not
  understand what you are trying to say here.

* "Temporal properties of pwa transitions systems"
  -> "Temporal properties of PWA transitions systems" (and elsewhere).

CONNECTION WITH GRAPH-BASED ABSTRACTION

* I found the transition from the previous section very abrupt and it was
  unclear where all this was going.

* "As an alternative, we introduce data driven ‘enrichment’ of abstractions
  to approximate the concrete relations underlying the abstract relations.
  Informally, we use linear regression to compute affine maps associated
  with the relations between two abstract states.": I would say that the
  first statement is informal and the second is formal. I prefer the second
  as it says precisely what you do. "data driven enrichment of abstractions"
  is a great sounding phrase, but I think it is essentially empty.

* "The abstraction of the hybrid dynamical system is a a partitioning"
-> "The abstraction of a hybrid dynamical system is a partitioning"

* "This further defines an equivalence relations"
-> "This further defines an equivalence relation"

* "We call it a time parameterized PWA relational abstraction": Is there
  a real need to introduce more terminology? This name is never used again.

RELATIONAL MODELING EXTENSIONS

* "such a relation can be rarely represented"
-> "such a relation can rarely be represented"

* "Using a tool like FLOW*": missing citation or URL.

* "we rely on statistical methods like simple linear": which other
  statistical methods did you actually use?

* "with the below relation"
-> "with the following relation"

* "C ~ C' ⟺  ∃x ∈ C. ∃x' ∈ C'. x' ∈ Ax + b ± δ": Why do you start
using "±
  δ"? Figure 1, Definition 2.2, and §2.1 just use "+ δ".

* Figure 2 is unnecessary. It does not really show anything.

* "Form a state x ∈ C,"
-> "From a state x ∈ C,"

* "be taken as long as x' ∈ C’ is satisfied."
-> "be taken as long as x' ∈ C' is satisfied."

* "split them in to multiple sets"
-> "split them into multiple sets"

* "PWA transition system which only defines the evolution of states x,"
-> "PWA transition system that only defines the evolution of states x,"

* "trajectory segments form a cell C."
-> "trajectory segments from a cell C."

* "D = {(x, x')|x ∈ C}": Do you not need an extra clause to relate x and x'
  to the model transitions?

* "Figure 5: The data gets split into two sets"
-> "Figure 5: The data is split into two sets"

AN EXAMPLE: VAN DER POL OSCILLATOR

* "The corresponding reachability graph is H(∆ = 0.1).": more notation and
  without a precise definition!

* "which if exists, can indicate the existence of a"
-> "which if it exists, may indicate the existence of a"

* "numerical simulation function SIM ∆~\footnote{...}."
-> "numerical simulation function SIM ∆.\footnote{...}"

EXPERIMENTAL RESULTS

* "Its different parts comprised of the"
-> "Its different parts comprise the"

* "Finally, the The bounded model checking"
-> "Finally, the bounded model checking"

* "due to the differences in which they operate"
-> "due to the differences in how they operate"

* "requires a finer abstraction to begin with in order to minimize the"
-> "modeling errors. On the other hand, a coarser abstraction as the"

* Is table 1 really useful? How is it to be interpreted? What information is
  being conveyed?

* "As apparent form the results,"
-> "As apparent from the results,"

* "but coarse enough to be manageable by the current bmc"
-> "but coarse enough to be manageable by current BMC"

* "during the k-relational modeling process till"
-> "during the k-relational modeling process until"

* "Similarly, small time length trajectory"
-> "Similarly, small time-length trajectory"

REFERENCES

* [6] "Ph.D. Dissertation. Ph. D. thesis," -> "Ph. D. thesis,"

* [22] "(????)": ????

============================================================================
                            REVIEWER 2
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 4
          Presentation/Readability (1-4): 3
                       Originality (1-4): 4
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

This submission presents a model based methodology to find falsifications in
black box dynamical systems. After describing approaches based on data driven
modeling, arguing the pros and cons, the authors put forward a framework that
combines ideas from abstraction based search and relational abstractions and
defines the k-relational modeling to compute PWA relational abstractions. To
estimate the local dynamics of the trajectory segments and compute relational
models, which were then interpreted as infinite state discrete transition
system, the proposed framework utilizes simple regression for an
over-determined case. These infinite state discrete transition system
constructions were model checked for safety violations using an off-the-shelf
bounded model checkers and a naive linear programming based specialized model
checker for comparison. The framework was implemented as a tool chain, namely
the PWA-Rel and its performance was demonstrated on a few benchmarks.

The paper is well organized and written and should be accepted. I list next a
few issues that could improve the readability and the overall manuscript:
- The submission is missing an abstract.
-  The following sentence “For this step, we need to have models that can be
digested by existing symbolic tools, which is harder to do with some of the
aforementioned heavier data-driven modeling methods” requires a more in-depth
discussion. Please enrich the discussion especially because the paper starts on
such a negative tone without providing concrete examples. While I do not
disagree with the authors I feel that things could be made clearer with a few
explanatory sentences.

   ============
   [TODO: Can not use/scale generic neural-networks, non-linear
   kernels for regression,...]
   ============
- In section 2.1. the authors state that when N>n “a single affine function
cannot be found”. Should the problem be stated in finding the best choice of
a and b or should the problem considering determining the optimal number of
piece wise affine approximations? Can the magnitude N-n give a hint of what is
a good number of PWA / affine approximations? I feel that this could lead to a
nice results which the authors can easily cast in optimization (2).

   ============
   In general, finding the `optimal' number of PWA models can be
   framed as a multiobjective optimization problem, where a small
   number of clusters and an accurate affine map is desired. This is a
   hard problem for arbitrary non-linear systems~\cite{[TODO]}.
   Hence, in this paper, we presented a scalable falsification approach
   with a sub-optimal PWA identification approach.

   [TODO: OLS is not the main focus of the paper. We should state that
   and mention that any method for learning affine functions can be
   used.]
   ============

- In section 4.1 the authors state that “in the presence of complex
non-linear behavior, using a single affine map can lead to poor
approximation”. In this context and in relation with recent effort in CPS on
modeling complex dynamics (see “Constructing compact causal mathematical
models for complex dynamics” ICCPS2017) can the authors comment or speculate
on how the complexity of the dynamics might influence the required number of
affine maps? How does the non-smoothness and non-differentiability influence
the computation of R(C,C’)? Is there any relationship between the number of
nondifferentiable points and the number of required affine maps? It will also
be enrich the discussion of data-driven modeling techniques with recent efforts
in modeling non-smooth complex dynamics with fractal properties and hint
towards potential extensions.

   ============
   [TODO: ...]
   ============

- Another critical parameter in the framework seems to be the time steps Delta
used in section 4. Can the authors perform a sensitivity analysis to gauge its
impact on overall quality of the results?

   ============
   [TODO: ...]
   ============

- Since the main goal of this technique is to find violations of safety
properties in the embedded control system, it would prove useful for the reader
if the authors could comment which of the evaluations are considering this
specific goal. Again, this is not a criticism, but I feel that the
investigation in section 6 could be described in a few more sentences to
connect with embedded control.
   ============
   [TODO: ...]
   ============
- Lastly, the paper has a few misspellings and grammar issues. I summarize some
I found: “apparent form the results…”, “We how we can use”,
“Finally, the The bounded…”

============================================================================
                            REVIEWER 3
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 3
                       Originality (1-4): 2
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

The paper presents a method for falsifying safety properties of hybrid systems.
The method first constructs a relational abstraction of the system (which is
considered as a black box) from a set of simulation traces. The relations in
the abstraction are affine with uncertainty. The approach is quite interesting,
however the examples used for demonstration are unfortunately not black boxes,
although these examples are treated as black boxes, since it would be of great
interest to show its usefulness for complex models for which traditional
verification tools cannot be applied. A comparison with S3CAM is interesting,
but is less informative than a comparison with simulation-based validation
tools.
I also recommend using relations more complex than affine, such as polynomial,
for which Z3 can be used (nonlinear relations may reduce the number of
hyper-rectangles).

The method is of heuristic nature and thus convincing experimental results
would be needed. The paper compares with previous work (S3CAM) using a similar
approach. This approach has been tested on more real-life examples from
automative control. I am wondering why there are no comparisons on these
examples.

============================================================================
                            REVIEWER 4
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 2
                       Originality (1-4): 3
               Technical Soundness (1-4): 2
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

The paper presents a model-based methodology to do falsification for hybrid
systems. The technique combines the abstraction based search and relation
abstraction concept to learn a PWA abstract model from the simulations of a
system. The transitions between states of the abstract model characterized by
the affine relations between source and target states. A model checking
technique is then applied to find a counter-example for the abstract model.
This counter-example will be used to guide the search for a counter-example of
the original system. Experimental results of several academic benchmarks are
reported that show the capability of proposed technique in finding false
behavior.

It's clear to the reviewer that the authors may require more time to complete
the paper (e.g., the abstract is missing in the paper. The paper's organization
also needs to be improved. It's not clear for the reviewer to see the
high-level concept and contributions of the paper from the beginning.

Although the author claims that the proposed technique can effectively find a
counter-example for a black-box model, however, the experiment result does not
reflect that statement. Most of the benchmarks are nonlinear continuous systems
that are not black-box. I strongly suggest the author evaluate the proposed
approach on some black-box systems such as industrial Simulink models.

Also, using linear regression to infer affine dynamics and discrete structures
of the abstractions of hybrid systems has been applied in the past. Please
check out the related work on mining hybrid automata from input/output traces

Medhat, Ramy, et al. "A framework for mining hybrid automata from input/output
traces." Embedded Software (EMSOFT), 2015 International Conference on. IEEE,
2015.

Using OLS does not always result in finding the best fit. For example, if some
data points are excessively small or large compared to the rest of data, they
will have disproportionately large effects on the resulting constant. In this
case, using other techniques like ridge regression or lasso regression may
yield a better result.

As the approach is neither sound nor complete and the application perspective
for falsifying black-box system is not adequately demonstrated, it's hard to
assert that the contributions of the proposed approach are novel.

   ============
   [TODO: ...]
   ============

The paper contains some minor typos such as:

Page 2, OLS paragraph: the best choice for a and b, consistency of
capitalization

Page 3, definition 2.2, second bullet: the the valuation function -> delete
one the

Page 3, section 2.2: pwa -> PWA
Page 4, section 3.1, at the first line: a a partitioning of -> delete one
a
Page 7, section 6.1, at the first line: the The bounded model checking ->
delete The
Page 8, section 6.2: bmc -> BMC

============================================================================
                            REVIEWER 5
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 3
                       Originality (1-4): 3
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 3

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

This is sophisticated stuff that, sadly, went over my head. Your motivation is
good and overall, this seems reasonable, but I couldn't easily wrap my head
around the sorts of system for which this technique would work. I suggest
leading with the Van Der Pol oscillator to help ignorant readers such as myself
get a handle on what it is you're actually able to do, then launch into the
formalism. As it stands, I think you need to be fairly expert in hybrid
dynamical systems to even begin to grasp what's going on here.

### 2017-06-12 scam.tex
.

--- 2017-06-09 scam.tex
@@
 
+\begin{example}[Van der Pol Oscillator]
+
+% \inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
+% Pol: continuous trajectories. Red and green boxes indicate unsafe and
+% initial sets.}{vdp-cont}
+%Simulations generated using uniformly random samples are shown in
+%    \figref{vdp-cont}.
+
+\inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
+abstraction $\epsilon=0.2$. Red cells are unsafe cells and green cells are
+initial cells.}{vdp-abs}
+
+We use the Van der Pol oscillator benchmark
+from~\cite{zutshi2014multiple} to demonstrate our falsification
+approach. We want to check the property $P_3$, indicated by
+the red box, given the initial set indicated by the green box. The
+abstraction is defined by $\epsilon=0.2$, which results in an
+evenly gridded state-space, where each cell is of size $0.2 \times
+0.2$ units. $ScatterAndSimulate$ is then used to construct the
+abstraction, using $2$ samples per cell and the time step
+$\Delta = 0.1s$. The result is plotted in ~\figref{vdp-abs}.
+
+% \mypara{Discover}: Using scatter-and-simulate, enumerate a finite number
+%     of cells (vertices) and the relations (edges) of the graph
+%     $\scr{H}(0.1)$. Associate the set of generated trajectory
+%     segments with their respective originating cells using a map $\ds:C
+%         \mapsto \setof{(\vx,\vx') |\vx \in C}$. The discovered
+%     abstraction is show in \figref{vdp-abs}. The red cells
+%     are unsafe and green cells are initial cells.
+\end{example}
+
 \subsection{CEGAR for falsifying $\TSAbstraction{\epsilon}{\System}$}

### 2017-06-12 scam.tex
.

@@
 
-\begin{example}[Van der Pol Oscillator]
+\begin{example}[Van der Pol Oscillator]~\label{ex:vdp}
 
@@
 from~\cite{zutshi2014multiple} to demonstrate our falsification
-approach. We want to check the property $P_3$, indicated by
-the red box, given the initial set indicated by the green box. The
-abstraction is defined by $\epsilon=0.2$, which results in an
-evenly gridded state-space, where each cell is of size $0.2 \times
-0.2$ units. $ScatterAndSimulate$ is then used to construct the
-abstraction, using $2$ samples per cell and the time step
-$\Delta = 0.1s$. The result is plotted in ~\figref{vdp-abs}.
+approach. We want to falsify property $P_3$, $x_{unsafe}\in[-1,
+-6.5]$ and $y_{unsafe}\in[-0.7, -5.6]$, given the initial set
+indicated $x0\in[-0.4, -0.4],\; y_0\in[0.4, 0.4]$. The abstraction
+is defined by $\epsilon=0.2$, which results in an evenly gridded
+state-space, where each cell is of size $0.2 \times 0.2$ units.
+$ScatterAndSimulate$ is then used to construct the abstraction,
+using $2$ samples per cell and the time step $\Delta = 0.1s$. The
+result is plotted in ~\figref{vdp-abs}.
 
@@
 
-\section{An Example: Van der Pol Oscillator}
-\todo{takes four iterations}
+\begin{example}
 \inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
-trajectory segments used by $1$-relational modeling.}{vdp-map}
+trajectory segments used by $1$-relational modeling.}{vdp-abs-paths}
+
+Going back to the example~\exref{vdp}, $ScatterAndSimulate$ is able to
+    find several abstract trajectories as shown
+    in~\figref{vdp-abs-paths}. However, typically, the refinement by
+    increasing the quantization precision requires $4$ refinement
+    iterations and finally finds a concrete counter-example for
+    $\epsilon=0.125$.
+\end{example}
 

### 2017-06-12 learning.tex
.

@@
 %%  complex eigenvalues.
+
+\begin{example}
+    For the~\exref{vdp}, for each cell $C$, regression
+    analysis is performed on the respective set of trajectory segments $D(C)$, and
+    compute a set of affine relations $R_{C,C'}(\x,\x')$ between $\x
+    \in C$ and $\x' \in C'$ s.t., $\text{edge} (C,C') \in
+    \scr{H}(0.1)$. Annotate each edge in the graph $\scr{H}(0.1)$ with
+    the respective relation.
+
+    \figref{vdp-abs-paths} shows the cells and the trajectory segments, which
+    are part of the data sets constructed using the $1$-relational
+    modeling. Against each cell, its unique identifier (integer
+    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
+    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
+    relations. Note that self loops result when an observed trajectory
+    segment has its $start$ and $end$ states in the same cell.
+\end{example}

### 2017-06-12 scam.tex
.

@@
 \begin{example}
-\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
-trajectory segments used by $1$-relational modeling.}{vdp-abs-paths}
+\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{
+Abstract trajectories discovered by $ScatterAndSimulate$. Each
+    rectangle denotes a cell which is uniquely identified  by the
+    tuple against it.
+%Cells and trajectory segments used by $1$-relational modeling.
+}{vdp-abs-paths}
 
-Going back to the example~\exref{vdp}, $ScatterAndSimulate$ is able to
+Going back to the~\exref{vdp}, $ScatterAndSimulate$ is able to
     find several abstract trajectories as shown

### 2017-06-12 learning.tex
cosmetic

--- 2017-06-12 learning.tex
@@
 
-% \begin{equation}
-%     \operatornamewithlimits{argmin}_{A,B}\displaystyle\sum_{i=1}^{N}{\left(\x' - (A\x + B)\right)^2}
-% \label{reg}
-% \end{equation}
+\begin{example}
+\mypara{\emph{Relationalize}}: For each cell $C$, perform regression analysis on
+    the respective set of trajectory segments $D(C)$, and compute a
+    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
+    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(0.1)$. Annotate
+    each edge in the graph $\scr{H}(0.1)$ with the respective
+    relation.
 
-% To ease the presentation, we can rewrite the above as a homogeneous
-% expression by using $\hat{\x} = [\x^T\ 1]^T$ and replacing $A$ and $B$ by the vector
-% $\vec{A} = [A\ B]$. The equation~\ref{reg} now becomes
-% \[
-% \operatornamewithlimits{argmin}_{\vec{A}}\displaystyle\sum_{i=1}^{N}{(\x' - \vec{A}^T \hat{\x})^2}
-% \]
-
-% The solution of OLS can be analytically computed as the Moore-Penrose
-% pseudo-inverse: $\vec{A} = (X^TX)^{-1}X^T X'$, where $X$ is the matrix
-% representing the horizontal stacking of all $\hat{\x}$ and $X'$
-% represents the stacking of all $\x'$ values.  The details can be found
-% in standard texts on learning and statistics
-% \cite{friedman2001elements}. Alternative methods based on gradient
-% descent can also be used to find a suitable $A$ matrix.
-
-% In this fashion, we can we can use OLS to approximate its trajectories
-% of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
-% = \vec{A}\hat{\x}$.
-
-% where
-% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T
-% \end{bmatrix}$ and $\vb = \begin{bmatrix}b_1 \\ \vdots \\ b_n
-% \end{bmatrix}$.
-
-% We note that linear regression tools are often able to provide an
-% $n\times 1$ vector $\vec{\delta}$ of error intervals, where the
-% $i^{th}$ entry of the form $[{\delta_i}_{\min},{\delta_i}_{\max}]$
-% indicates the best and worst possible error in computation of the
-% $i^{th}$ state in $\x'$.
-
-% \mypara{Regression Analysis}
-
-% In statistics, regression is the problem of finding a
-% \textit{predictor}, which can suitably predict the relationship
-% between the given set of observed input $\x$ and output $\y$ vectors.
-% In other words, assuming that $\y$ depends on $\x$, regression
-% strategies find either a parameterized or a non-parameterized
-% prediction function to explain the dependence. We now discuss simple
-% linear regression, which is parametric in nature and searches for an
-% affine predictor. It is also called \textit{ordinary least squares}.
-
-% \mypara{Ordinary Least Squares (OLS)}
-
-% Let the data set be comprised of $N$ input and output pairs $(\x,
-% y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
-% the case in the current context of \textit{finding the best fit}, the
-% problem is over-determined; there are more equations than
-% unknowns. Hence, a single affine function cannot be found which
-% satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
-% need to find the `best' choice for $\vec{a}$ and $b$. This is formally
-% defined using a loss function. For the case of simple linear
-% regression or OLS, the loss function is the sum of squares of the
-% errors in prediction.  The task is then to determine the vector of
-% coefficients $\vec{a}$ and an offset constant $b$, such that the least
-% square error of the affine predictor is minimized for the given data
-% set.
-% \begin{equation}
-%     \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
-% \label{reg}
-% \end{equation}
-% To ease the presentation, we can rewrite the above as a homogeneous
-% expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
-% $\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
-% \[
-%     \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
-% \]
-% The solution of OLS can be analytically computed as
-% \[ Ab = (X^TX)^{-1}X^T\y\]
-% where $X$ is the matrix representing the horizontal stacking of all
-% $\hat{\x}$.
-% The details can be found in several texts on learning and statistics
-% \cite{friedman2001elements}.
-
-% Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
-% where $\x\in\reals^n$ we can use OLS to approximate its trajectories
-% at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
-% $A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
-% \begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
-%     global relational model for the system. The data set for the
-%     regression is a set of pairs $\setof{(\x,\x') | \x' =
-%     \simulate(\x, \Delta)}$ and can be generated on demand. Another
-%     set can be generated to compute the error $\delta$ as an interval
-%     of vectors, where each element is an interval $\delta_i \in
-%     [\delta_{min}, \delta_{max}]$.
-%
-%     Affine maps are poor
-%     approximations for arbitrary non-linear functions. Hence, we use a
-%     collection of affine maps to approximate the local behaviors (in
-%     state-space) of the system $\simulate$. This results in a piece
-%     wise approximation, as described in the next section.
-
-% \subsection{Relational Abstractions}
-% For a hybrid automaton model, they can summarize the continuous
-% dynamics of each mode using a binary relation over continuous states.
-% The resulting relations are timeless (independent) and hence valid for
-% all time as long as the mode invariant is satisfied. The relations
-% take the general form of $R(\x,\x') \bowtie 0$, where $\bowtie$
-% represents one of the relational operators $=, \ge, \le, <, >$. For
-% example, an abstraction that captures the monotonicity with respect to
-% time for the differential equation $\dot{x} = 2$ is $x' > x$. The
-% abstraction capturing the relation between the set of ODEs: $\dot{x} =
-% 2$, $\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.
-
-%%  Similar to timeless relational abstractions, time-aware relational
-%%  abstractions \cite{mover2013time} construct binary relations between
-%%  the current state $\x$ of the system and any future reachable state
-%%  $\x'$. The difference lies with the latter also constructing relations
-%%  between the current time $t$ and any future time $t'$. This is
-%%  achieved by a case by case analysis of the eigen structure of the
-%%  matrix $A$ (of an affine ODE). Separate abstractions are used for the
-%%  case of linear systems with constant rate, real eigen values and
-%%  complex eigenvalues.
+    \figref{vdp-map} shows the cells and the trajectory segments, which
+    are part of the data sets constructed using the $1$-relational
+    modeling. Against each cell, its unique identifier (integer
+    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
+    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
+    relations. Note that self loops result when an observed trajectory
+    segment has its $start$ and $end$ states in the same cell.
+\end{example}

### 2017-06-12 scam.tex
cosmetic

--- 2017-06-12 scam.tex
@@
 $\simmap_\Delta$.  We call the transition system an
-$\epsilon$-precision transition system and denote it by
-$\TSAbstraction{\epsilon}{\System}$.  Let $\epsilon  = 10^{-k}$, for
-some positive integer $k$.  We define a quantization function
-$\quant_\epsilon$ that maps a state $\contState$ in $\ContStates$ to
-$10^{-k}*\lfloor \contState*10^k \rfloor$. This transformation
-essentially truncates the decimal representation of the numeric value
-of the state $\state$ to at most $k$ digits after the decimal point.
-This function induces an equivalence relation $\contState_i
-\equiv_\epsilon \contState_j$ iff $\quant_\epsilon(\contState_i) =
-\quant_\epsilon(\contState_j)$.
+$\epsilon$-precision transition system ($\epsilon$-TS for short) and
+denote it by $\TSAbstraction{\epsilon}{\System}$.  Let $\epsilon  =
+10^{-k}$, for some positive integer $k$.  We define a quantization
+function $\quant_\epsilon$ that maps a state $\contState$ in
+$\ContStates$ to $10^{-k}*\lfloor \contState*10^k \rfloor$. This
+transformation essentially truncates the decimal representation of the
+numeric value of the state $\state$ to at most $k$ digits after the
+decimal point.  This function induces an equivalence relation
+$\contState_i \equiv_\epsilon \contState_j$ iff
+$\quant_\epsilon(\contState_i) = \quant_\epsilon(\contState_j)$.
 
@@
 simulation map $\simmap_\Delta$, and a precision $\epsilon$ creates an
-$\epsilon$-precision transition system abstraction of $\System$.  The
+$\epsilon$-TS abstraction of $\System$.  The
 essence of the procedure is a breadth-first exploration of
@@
 following steps:
-\begin{enumerate}
+\begin{enumerate}[leftmargin=1em,labelsep=1em,label={\arabic*.}]
 \item
@@
 
-\begin{example}[Van der Pol Oscillator]
-
-% \inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
-% Pol: continuous trajectories. Red and green boxes indicate unsafe and
-% initial sets.}{vdp-cont}
-%Simulations generated using uniformly random samples are shown in
-%    \figref{vdp-cont}.
-
+\inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
+Pol: continuous trajectories. Red and green boxes indicate unsafe and
+initial sets.}{vdp-cont}
 \inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
-abstraction $\epsilon=0.2$. Red cells are unsafe cells and green cells are
+$0.2$-TS abstraction. Red cells are unsafe cells and green cells are
 initial cells.}{vdp-abs}
 
-We use the Van der Pol oscillator benchmark
-from~\cite{zutshi2014multiple} to demonstrate our falsification
-approach. We want to check the property $P_3$, indicated by
-the red box, given the initial set indicated by the green box. The
-abstraction is defined by $\epsilon=0.2$, which results in an
-evenly gridded state-space, where each cell is of size $0.2 \times
-0.2$ units. $ScatterAndSimulate$ is then used to construct the
-abstraction, using $2$ samples per cell and the time step
-$\Delta = 0.1s$. The result is plotted in ~\figref{vdp-abs}.
+\begin{example} 
+We now describe $\epsilon$-TS abstraction using the Van der Pol
+oscillator benchmark from~\cite{zutshi2014multiple}. Simulations
+generated using uniformly random samples are shown in
+\figref{vdp-cont}. We want to check the property $P_3$, indicated by
+the red box, given the initial set indicated by the green box. We
+consider the $\epsilon$-TS abstraction defined by the quantization
+function $\quant_{0.2}(\x)$. This results in an evenly gridded
+state-space, where each cell is of size $0.2 \times 0.2$ units.
+$\mathsf{ScatterAndSimulate}$ is then used to construct
+$\TSAbstraction{0.2}{\System}$ using $2$ samples per cell and a time
+step of $\Delta = 0.1s$. The complete process follows.  The
+$\epsilon$-TS abstraction thus discovered is shown in
+\figref{vdp-abs}. The red cells are unsafe and green cells are initial
+cells.
+\end{example}
 
+% \mypara{Abstract}: Assume an implicit abstraction using the quantization
+%         function $\quant_{\epsilon=0.2}(\x)$. The corresponding reachability
+%     graph is $\scr{H}(\Delta=0.1)$.
+% 
 % \mypara{Discover}: Using scatter-and-simulate, enumerate a finite number
@@
 %     segments with their respective originating cells using a map $\ds:C
-%         \mapsto \setof{(\vx,\vx') |\vx \in C}$. The discovered
-%     abstraction is show in \figref{vdp-abs}. The red cells
-%     are unsafe and green cells are initial cells.
-\end{example}
+%         \mapsto \setof{(\vx,\vx') |\vx \in C}$.
 
@@
 nor an underapproximation} of the continuous-time continuous-space
-reachability relation between states. The purpose of the
-$\epsilon$-precision transition system is to discover falsifying
-trajectories. This can be accomplished by
+reachability relation between states. The purpose of the $\epsilon$-TS
+is to discover falsifying trajectories. This can be accomplished by
 Algorithm~\ref{algo:s3camBMC} that uses CounterExample-Guided
@@
 
-\section{An Example: Van der Pol Oscillator}
-\todo{takes four iterations}
-\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Cells and
-trajectory segments used by $1$-relational modeling.}{vdp-map}
-
-% The abstraction of the hybrid dynamical system is a partitioning of
-% the hybrid state space $\HybridStates$ into hyper-rectangles or cells
-% $C_i \in \Cells$. The cells are pairwise disjoint $C_i \cap C_j \neq
-% \emptyset$ iff $i \neq j$ and their union $\bigcup_{C_i\in\Cells}C_i$
-% is the state space $\HybridStates$. The partitions over the continuous
-% state space are implicitly defined using a quantization function
-% $\quant_\epsilon$ parameterized by the precision $\epsilon$.  This further
-% defines an equivalence relations such that $(\vx_i) \equiv (\vx_j)
-% \;\text{iff}\; \quant_\epsilon(\vx_i) = \quant_\epsilon(\vx_j)$.  In other
-% words, $(\vx_i) \in C_i \; \text{iff} \; C_i = \quant_\epsilon(\vx_i)$. The
-% abstraction can be interpreted as a graph and explored using fixed
-% time $(\Delta\in\Reals^+)$ relations obtained using
-% $\simulate(\vx)$~\footnote{$\simulate(\vx)$ is a fixed
-% parameterization of $\mathsc{sim}(\vx,\Delta)$}.
+% \subsection{Example: Van der Pol Oscillator}
+% \todo{Takes 4 iterations}
+% \inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Refinement of
+% $\epsilon$-TS.}{vdp-map}
 %
-% % Such an abstraction is defined by partitioning the
-% % state-space of the system using a grid, and hence each abstract state
-% % is a hyper-rectangle. The relations between abstract states are
-% % obtained by suitably lifting $(\vx,\vx') \in\reach{t,u}$. Considering
-% % the abstract states as vertices of a graph and the relations as edges
-% % between nodes, the problem of finding an abstract counter-example
-% % (CEx) can be cast as finding a path in the graph. Such an abstraction
-% % can be explored on-the-fly by enumerating the abstract relations
-% % between a pair of states.
-%
-% \mypara{A discrete abstraction} over a continuous space uses relations
-% to describe the behavior of the system over time. Such relations
-% abstract the underlying dense-time trajectories of the given dynamical
-% system to a transition system.
-
-% relation between concrete states, and concrete
-% paths can no longer be constructed directly. Instead, search
-% procedures like Counter-Example Guided Refinement (CEGAR) must be used
-% to find concrete paths. In general, such search techniques are based
-% on refinement of the state-space and exponential with respect to
-% computational resources and memory. As an alternative, we introduce
-% data driven `enrichment' of abstractions to approximate the concrete
-% relations underlying the abstract relations.  Specifically, we use
-% linear regression to compute affine maps associated with the relations
-% between two abstract states. We then show how associating the computed
-% maps with the relations transforms a discrete abstraction into a PWA
-% transition system.
-
-% The abstraction of the hybrid dynamical system is a a
-% partitioning of the combined state-input space $\HybridStates \times
-% \Inputs$ into hyper-rectangles or cells $C_i \in \Cells$. The cells are
-% pairwise disjoint $C_i \cap C_j \neq \emptyset$ iff $i \neq j$ and
-% their union $\bigcup_{C_i\in\Cells}C_i$ is the state-input
-% space $\HybridStates \times \Inputs$. The partitions over
-% the continuous input-state space are implicitly defined using a quantization
-% function $\quant_\epsilon$ parameterized by the precision $\epsilon$.
-% This further defines an equivalence relations such that
-% \[
-%     (\vx_i, \vu_i) \equiv (\vx_j, \vu_j) \;\text{iff}\; \quant_\epsilon(\vx_i,
-%     \vu_i) = \quant_\epsilon(\vx_j, \vu_j)
-% \]
-% In other words, $(\vx_i, \vu_i) \in C_i \; \text{iff} \; C_i =
-% \quant_\epsilon(\vx_i, \vu_i)$.
-
-% \begin{definition}[Graph Abstraction]
-%     The quantization function $\quant_\epsilon$ induces a DAG on the discrete
-%     abstraction given by a graph $G(V, E)$, where each vertex
-%     corresponds to a cell $V_i = C$, and an edge exists between two
-%     cells $(C, C') \in E_i$ iff there exist state-input pairs
-%     $(\vx,\vu) \in C$, $(\vx', \vu) \in C'$ such that $\vx' =
-%     \simulate(\vx, \vu )$.
-% \end{definition}
-
-% \begin{definition}[Graph Abstraction]
-%     The quantization function $\quant_\epsilon$ induces a directed acyclic
-%     graph (DAG) on the discrete
-%     abstraction given by a graph $G(V, E)$, where each vertex
-%     corresponds to a cell $V_i = C$, and an edge exists between two
-%     cells $(C, C') \in E_i$ iff there exist state-input pairs
-%     $(\vx) \in C$, $(\vx') \in C'$ such that $\vx' =
-%     \simulate(\vx)$.
-% \end{definition}
-%
-%
-%
-% Given a numerical simulator $\simulate$ and an initial abstraction
-% defined by a quantization function $\quant_\epsilon$ and a time step
-% $\Delta$, we pick samples and simulate them to explore the abstraction
-% on the fly to obtain a graph $G$, which has a finite number of
-% abstract states $C$ (or cells) and edges $(C,C')$ iff $C
-% \areach{\Delta} C'$.
-
-% \begin{definition}[Abstract State Graph]. Let ∆ > 0 be a fixed time
-%     step. The abstract state graph H(∆) for time step ∆ is defined by
-%     the set of vertices C, and edges (Ci, Cj ) whenever Ci ∆ Cj . Let
-%     C0 denote the collection of initial cells in C, i.e., cells Ci
-%     such that Ci ∩ X0 6= ∅. Further, let Cu denote the set of unsafe
-%     cells, or cells Cj such that there is a state xj ∈ Cj that reaches
-%     an unsafe state y ∈ Xf within time 0 ≤ tj < ∆. The abstraction
-% H(∆) is given by D C, ∆ , C0, Cu E .  \end{definition}
-
-% As
-% Instead of using a CEGAR like loop (used in~\cite{zutshi2014multiple}),
-% we use the generated trajectories to learn quantitative models
-% describing the local behavior of the system. These models are defined
-% by a set of relations $R \subseteq \HybridStates \times \HybridStates$
-% for each edge of the reachability graph. In the next section, we
-% summarize this idea.
-
-% \subsection{Enriched Abstraction}
-%
-% \input{figs/segtraj_rel_enriched.tex}
-%
-% Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
-% trajectory between the respective cells. The graph abstraction $G$
-% only states that there exists a state $\x \in C$ from which the system
-% can evolve to a future state $\x' \in C'$. To increase the precision,
-% iterative refinement of the abstraction by state splitting was
-% proposed in~\cite{zutshi2014multiple}. Due to the curse of
-% dimensionality, state splitting is not scalable. Instead, we propose
-% an `enrichment' $G^R$ of $G$ by computing a set of local relations
-% $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$, which
-% non-deterministically describe relations between $\x \in C$ and $\x'
-% \in C'$. This is illustrated in \figref{enriched-edge}.
-% %(compare with \figref{segtraj}).
-%
-% The enriched graph $G^R$ captures the underlying local forward
-% dynamics describing the evolution of the system in each abstract
-% state. We represent the dynamics using an affine model with an
-% interval error. Such a model can either be approximated using learning
-% methods or computed as a sound (over) approximation using reachability
-% set computation methods. Because we assume black box semantics, we
-% only present the former. Using regression analysis on a witnessed
-% trajectory between two cells, we compute an approximate discrete map
-% along with an error estimate.  Moreover, using the simulation function
-% $\simulate$, additional trajectories (or data) can be generated if
-% required. The data can be separated into a training set and testing
-% set to compute the map and the error respectively.
-%
-% %The latter case can be explored if
-% %the symbolic dynamics of the system are known. A tool like
-% %\flowstar~\cite{chen2013flow} can be used to find the reachable set
-% %map.
-%
-% Observe that $G^R$, a directed reachability graph, is rich enough to
-% search for concrete behaviors in the system. We call it a time
-% parameterized PWA relational abstraction. It can be interpreted as an
-% infinite state discrete transition system, and we can use
-% off-the-shelf bounded model checkers to find concrete violations of a
-% given safety property, and even other temporal properties.
-
-% \begin{figure}[!htbp]
-% \begin{center}
-% \tikzstyle{line} = [thick]
-% \tikzstyle{arw} = [->, thick,>=stealth,shorten <=2pt, shorten >=2pt]
-% \tikzstyle{block} = [rectangle, minimum width=0.5cm, minimum height=1.0cm, text centered, draw=black, align=center]%, fill=blue!25]
-% \begin{tikzpicture}[scale=1.0, on grid, auto]
-%
-%     \node (G) at (-4,0) {$G$};
-%     \node (GR) at (0,0) {$G^R$};
-%     \node (CEX) at (4,0) {$CEx$};
-%
-%     \node (enrich) [block] at (-2, 0) {\scriptsize Enrich using \\ \scriptsize affine  maps};
-%     \node (BMC) [block] at (2, 0) {\scriptsize Find CEx \\ \scriptsize using BMC};
-%
-%     %\node[align=center] at (0, 0) {\scriptsize Enrich using \\ \scriptsize Regression \\ \scriptsize(OLS)};
-%
-%     %\draw [line] (-2.0,-0.5) rectangle (-1.0,0.5);
-%     \draw[arw] (G.east) -- (enrich.west);
-%     \draw[arw] (enrich.east) -- (GR.west);
-%     \draw[arw] (GR.east) -- (BMC.west);
-%     \draw[arw] (BMC.east) -- (CEX.west);
-%
-% \end{tikzpicture}
-% \end{center}
-% \vspace*{-.3cm}
-% \caption{Finding counter-examples (CEx) in a graph abstraction using PWA models.}
-% \label{fig:overview}
-% \vspace*{-.3cm}
-% \end{figure}

### 2017-06-12 appendix.tex

@@
-Plan of action:
----------------------------------------------
-
-- What is the paper saying/should be saying?
-- Simplify notations....pwa relational and k-relational
-
-- Modify ordering of sections: Abstraction -> PWA enrichment
-- Simplify notations
-- Fold multiple PWA relational/transitions system into one and use the
-  transitions system as a semantic interpretation for model checking.
-  Can be eliminated if automatons/kripke structures are used instead.
-- Fix the `convex-hull' in the definition 2.2 PWA Relational Model.
-  Showcase clustering using a benchmark. Add it to future work
-  and cite Nimit's  paper. Or, retract it entirely?
-- Clearly state that the approach is not a generic data-drive system
-  identification technique, but focused on the falsification aspect.
-  We only model selective state-space regions: abstract violations.
-- De-emphasize OLS (and the process and the cost functions used for
-  modeling). Add that the idea hinges on selectively identifying the
-  state-space for modeling and not how the models might be created.
-  Emphasize again that the requirement is only on the linearity of
-  the individual models and hence, OLS is one way. Trade-off between
-  efficiency and complexity can be achieved by selecting the `right'
-  modeling template.
-- Add benchmarks: hybrid with multiple modes.
-
-Summary of the reviews
----------------------------------------------
-
-R1: Like the idea but the presentation and results need improvement
-
-- Clarify that the plant is inherently a hybrid dynamical system
-  and in the future (or if time permits) a controller can be added and
-  analyzed together as a infinite state transition system.
-- Clarify the clustering done.
-
-R2: Seems the most positive, but focuses on PWA identification.
-
-- Clarify the purpose of the approach is to use sub-optimal
-  identification techniques but still provide a `reasonable' and
-  scalable falsification technique.
-
-R3: Poor benchmarks: use automotive control examples, try
-polynomial kernels (with Z3), compare with simulation based
-falsification (S-Taliro?).
-
-- Try polynomial kernels on a benchmark?
-- Clarify S3CAM was already compared with S-Taliro.
-
-R4: Probably the most negative review.
-Sloppy presentation; high level concepts are not well presented
-and contributions are not clear. Benchmarks are too simple. Why OLS?
-Contributions are not significant/novel.
-
-- Spell out the contributions clearly.
-- Use Simulink models.
-- Cite papers which learn hybrid automata from i/o traces, and clarify
-  the difference.
-  \cite{medhat2015framework} and many others
-- Clarify that OLS is just one choice out of many.
-
-R5: Probably a borderline.
-
-- Add a motivation section showcasing the contributions.
-
---
 ============================================================================

### 2017-06-12 bmc_section.tex
.

@@
 
-\todo{Tell that we use a DFS in the implementation section}
-
-\section{An Example: Van der Pol Oscillator}
-
-\mypara{\emph{Model Check}}: The graph $\scr{H}(0.1)$ can now be viewed as a
+\inclfig[width=0.50\linewidth]{vdp_graph.pdf}{Enriched
+graph $G^R$. The affine maps $f$ for the transition relations are show
+in \tabref{vdp-pwa}.}{vdp-graph}
+\begin{example}
+    Coming back to~\exref{vdp}, \figref{vdp-graph} %and\tabref{vdp-pwa}
+    shows the abstract reachability graph with its transition
+    relations. The relational maps are omitted for brevity.
+    Note that self loops result when an observed trajectory
+    segment begins and ends in the same cell.
+    The graph $\scr{H}(0.1)$ can now be viewed as a
     transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
@@
     edge(\scr{H}(0.1))}$.
+
 \mypara{\emph{Check Counter-example}}: The infinite state (but finite location) transition system can
@@
         find a counter-example.}.
-
-\inclfig[width=0.50\linewidth]{vdp_graph.pdf}{Enriched
-graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}{vdp-graph}
+\end{example}
 

### 2017-06-12 learning.tex
.

--- 2017-06-12 learning.tex
@@
 where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn an
-affine function described as
-\begin{equation}
-\label{eq:affinemap}
-\x' = A\x + B
-\end{equation}
+affine function described as $\x' = A\x + B$.
 Here, $A$ is a $n\times n$ matrix, and $B$ is a $n\times 1$ matrix.
@@
 section.
+
+\begin{example}
+    For the~\exref{vdp}, for every abstract relation between two cells
+    $C, C'$, linear regression analysis is performed on the respective
+    set of trajectory segments, and the affine relation is computed.
+    \figref{vdp-abs-paths} shows the cells and the trajectory segments, which
+    are part of the data sets constructed using the $1$-relational
+    modeling. Against each cell, its unique identifier (integer
+    co-ordinate) is mentioned.
+    %Finally, \figref{vdp-graph} %and\tabref{vdp-pwa}
+    %show the enriched graph $G^R$ with its transition
+    %relations. Note that self loops result when an observed trajectory
+    %segment has its $start$ and $end$ states in the same cell.
+\end{example}
+    %$R_{C,C'}(\x,\x')$ between $\x
+    %\in C$ and $\x' \in C'$ s.t., $\text{edge} (C,C') \in
+    %\scr{H}(0.1)$. Annotate each edge in the graph $\scr{H}(0.1)$ with
+    %the respective relation.
 

### 2017-06-12 scam.tex
.
fixed the last merge

@@
 
-% Let $\inputsignal$ = $(t_0, u_0)$, $\ldots$, $(t_N,u_N)$ be a
-% timed-sequence of inputs, where $t_0 = 0$, and for all $i > 0$, $t_i -
-% t_{i-1} = \Delta$, and $u_i \in \Inputs$.  We can extend the forward
-% simulation map to a sequence of states to describe the application of
-% an input signal to excite the system as follows:
-% $\simmap_\Delta(\initstate, \mathbf{u}) = \contState_N$, if for all $i
-% \in [0,N-1]$, $\simmap_\Delta(\contState_i, u_i,t_i) =
-% \contState_{i+1}$.
-% 
-% An alternative view of the forward simulation-map is as a
-% parameterized reachability relation $\reach_\Delta{t,u} \subseteq
-% \ContStates \times \ContStates$ defined as below:
-% \[
-% (\contState,\contState') \in \reach_\Delta{t,u}\text{ iff } \contState' = \simmap_\Delta(\contState,t,u).
-% \]
-
-% inputs modeling an embedded control system. 
-% 
-% we assume that the hybrid system is effectively a switched-mode
-% dynamical system, \ie, the mode of the system is completely determined
-% by its state.  An input to $\System$ is assumed to be a time-bounded
-% {\em signal}, \ie, a function from $[0,T]$ to $\Inputs$, with $T$ as
-% the time horizon, and $\Inputs \subseteq \Reals^k$.  In this paper, we
-% view $\System$ as effectively black-box equipped with a forward
-% simulation map, in a simulation time-step of $\Delta_t$ from time $t$,
-% allows computing the unique reachable state $\vx \in \HybridStates$
-% given state-input pair $(\vx,\vu)$.
-
-% time and input parameterized reachability relation $\reach{t,
-% \vu}_\System \subseteq \HybridStates \times \HybridStates$ as follows:
-% $\simulate^{\System}:
-% (\vx, \vu, t) \mapsto \vx$, where $\vx \in \HybridStates$, $\vu \in
-% \Inputs$ and $t \in [0,T]$.
-
-% Given the hybrid state space $\vx\in\HybridStates$ and input space $u
-% \in\Inputs$ and time step $t$, it computes the unique successor state
-% $\vy=\simulate^{\System}(\vx,u,t)$ such that $\vx \reach{t,u} \vy$.
-
-% \begin{definition}[System Evolution]
-%     The behavior of system $\System$ at time $t$ is described as an
-%     evolution in the hybrid state-space $\HybridStates$ under the
-%     effect of the input $\vu \in \Inputs$ at time $t$. The forward
-%     simulation map $\mathsc{sim}^\System : \HybridStates \times \Inputs
-%     \times \Reals^+ \mapsto \HybridStates$, maps a state $\x\in\HybridStates$ at time
-%     $t$ to a new state $\x'\in\HybridStates$ at time $t+\Delta\in\Reals^+$, under a constant
-%     input $\vu$.  Finally, we define the time and input parameterized
-%     reachability relation $\reach{t,u} \subseteq \HybridStates \times
-%     \HybridStates$ such that
-%     \begin{equation}
-%         (\vx,\vx') \in \reach{t,u} \;\text{iff}\; \vx'=\mathsc{sim}^{\System}(\vx,u,t).
-%     \end{equation}
-% \end{definition}
-
-% \mypara{System Assumptions:} 
-% We require that $\System$ can always be
-% simulated forward in time with deterministic results. This in turn
-% requires existence and uniqueness of trajectories over a finite time
-% horizon $[0,T]$, which can be guaranteed by Lipschitz continuity of
-% the vector field in each underlying hybrid mode, ruling away issues
-% such as finite escape times~\cite{Meiss/2007/Differential}, and having
-% well-defined semantics for simulations at switching boundaries. While
-% these assumptions hold for the dynamical system examples which we
-% consider in this paper, such assumptions can be restrictive for
-% general control system models. When analyzing a system that does not
-% obey theoretical conditions that allow deterministic simulation, we
-% assume that the underlying tool used for numerical approximations of
-% system behaviors uses deterministic simulation semantics.  Finally, we
-% assume full observability of the system state, validate our results
-% only against the $\mathsc{sim}^\System$ function (as the analytic closed-form
-% representation of the system dynamics is assumed to be unavailable).
-% From here on, we drop the subscript $\System$ whenever it is clear
-% from the context. Also, to ease the presentation, we conflate the
-% inputs with the state and denote them by $\vx$.
-% 
 \subsection{$\epsilon$-precision Transition System Abstraction}
@@
 
-\inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
-Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}{vdp-cont}
 \inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
@@
 initial cells.}{vdp-abs}
-
-\begin{example} 
+\begin{example}[Van der Pol Oscillator]~\label{ex:vdp}
 We now describe $\epsilon$-TS abstraction using the Van der Pol
@@
     generated using uniformly random samples are shown in
-\figref{vdp-cont}. We want to check the property $P_3$, indicated by
-the red box, given the initial set indicated by the green box. We
-consider the $\epsilon$-TS abstraction defined by the quantization
-function $\quant_{0.2}(\x)$. This results in an evenly gridded
-state-space, where each cell is of size $0.2 \times 0.2$ units.
-$\mathsf{ScatterAndSimulate}$ is then used to construct
-$\TSAbstraction{0.2}{\System}$ using $2$ samples per cell and a time
-step of $\Delta = 0.1s$. The complete process follows.  The
-$\epsilon$-TS abstraction thus discovered is shown in
-\figref{vdp-abs}. The red cells are unsafe and green cells are initial
-cells.
+    \figref{vdp-cont}. We want to check the property $P_3$:
+    $x_{unsafe}\in[-1,-6.5]$ and $y_{unsafe}\in[-0.7, -5.6]$, given
+    the initial set indicated $x0\in[-0.4, -0.4],\; y_0\in[0.4, 0.4]$.
+    We consider the $\epsilon$-TS abstraction defined by the
+    quantization function $\quant_{0.2}(\x)$. This results in an
+    evenly gridded state-space, where each cell is of size $0.2 \times
+    0.2$ units.  $\mathsf{ScatterAndSimulate}$ is then used to
+    construct $\TSAbstraction{0.2}{\System}$ using $2$ samples per
+    cell and a time step of $\Delta = 0.1s$. The complete process
+    follows.  The $\epsilon$-TS abstraction thus discovered is shown
+    in \figref{vdp-abs}. The red cells are unsafe and green cells are
+    initial cells.
 \end{example}
-
-% \mypara{Abstract}: Assume an implicit abstraction using the quantization
-%         function $\quant_{\epsilon=0.2}(\x)$. The corresponding reachability
-%     graph is $\scr{H}(\Delta=0.1)$.
-% 
-% \mypara{Discover}: Using scatter-and-simulate, enumerate a finite number
-%     of cells (vertices) and the relations (edges) of the graph
-%     $\scr{H}(0.1)$. Associate the set of generated trajectory
-%     segments with their respective originating cells using a map $\ds:C
-%         \mapsto \setof{(\vx,\vx') |\vx \in C}$.
 
@@
 %
+\begin{example}
+\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{
+Abstract trajectories discovered by $\mathsf{ScatterAndSimulate}$. Each
+rectangle denotes a cell which is uniquely identified  by the
+tuple against it.
+%Cells and trajectory segments used by $1$-relational modeling.
+}{vdp-abs-paths}
+
+    Going back to the~\exref{vdp}, $\mathsf{ScatterAndSimulate}$ is able to
+    find several abstract trajectories as shown
+    in~\figref{vdp-abs-paths}. However, typically, the refinement by
+    increasing the quantization precision requires $4$ refinement
+    iterations and finally finds a concrete counter-example for
+    $\epsilon=0.125$.
+\end{example}

### 2017-06-12 learning.tex
changes
garbage stuff

--- 2017-06-12 learning.tex
@@
 \begin{example}
-\mypara{\emph{Relationalize}}: For each cell $C$, perform regression analysis on
-    the respective set of trajectory segments $D(C)$, and compute a
-    set of affine relations $R_{C,C'}(\x,\x')$ between $\x \in C$ and $\x'
-    \in C'$ s.t., $\text{edge} (C,C') \in \scr{H}(0.1)$. Annotate
-    each edge in the graph $\scr{H}(0.1)$ with the respective
-    relation.
-
-    \figref{vdp-map} shows the cells and the trajectory segments, which
+    For the~\exref{vdp}, for every abstract relation between two cells
+    $C, C'$, linear regression analysis is performed on the respective
+    set of trajectory segments, and the affine relation is computed.
+    \figref{vdp-abs-paths} shows the cells and the trajectory segments, which
     are part of the data sets constructed using the $1$-relational
     modeling. Against each cell, its unique identifier (integer
-    co-ordinate) is mentioned.  Finally, \figref{vdp-graph} and
-    \tabref{vdp-pwa} show the enriched graph $G^R$ with its transition
-    relations. Note that self loops result when an observed trajectory
-    segment has its $start$ and $end$ states in the same cell.
+    co-ordinate) is mentioned.
 \end{example}

### 2017-06-12 scam.tex
garbage stuff

--- 2017-06-12 scam.tex
@@
 
-\inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
-Pol: continuous trajectories. Red and green boxes indicate unsafe and
-initial sets.}{vdp-cont}
+% \inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
+% Pol: continuous trajectories. Red and green boxes indicate unsafe and
+% initial sets.}{vdp-cont}
 \inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
@@
 
-% \mypara{Abstract}: Assume an implicit abstraction using the quantization
-%         function $\quant_{\epsilon=0.2}(\x)$. The corresponding reachability
-%     graph is $\scr{H}(\Delta=0.1)$.
-% 
-% \mypara{Discover}: Using scatter-and-simulate, enumerate a finite number
-%     of cells (vertices) and the relations (edges) of the graph
-%     $\scr{H}(0.1)$. Associate the set of generated trajectory
-%     segments with their respective originating cells using a map $\ds:C
-%         \mapsto \setof{(\vx,\vx') |\vx \in C}$.
+\inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
+abstraction $\epsilon=0.2$. Red cells are unsafe cells and green cells are
+initial cells.}{vdp-abs}
+
+\begin{example}[Van der Pol Oscillator]~\label{ex:vdp}
+We use the Van der Pol oscillator benchmark
+from~\cite{zutshi2014multiple} to demonstrate our falsification
+approach. We want to falsify property $P_3$, $x_{unsafe}\in[-1,
+-6.5]$ and $y_{unsafe}\in[-0.7, -5.6]$, given the initial set
+indicated $x0\in[-0.4, -0.4],\; y_0\in[0.4, 0.4]$. The abstraction
+is defined by $\epsilon=0.2$, which results in an evenly gridded
+state-space, where each cell is of size $0.2 \times 0.2$ units.
+$ScatterAndSimulate$ is then used to construct the abstraction,
+using $2$ samples per cell and the time step $\Delta = 0.1s$. The
+result is plotted in ~\figref{vdp-abs}.
+\end{example}
 
@@
 
-% \subsection{Example: Van der Pol Oscillator}
-% \todo{Takes 4 iterations}
-% \inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{Refinement of
-% $\epsilon$-TS.}{vdp-map}
-%
+\begin{example}
+\inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{
+Abstract trajectories discovered by $ScatterAndSimulate$. Each
+    rectangle denotes a cell which is uniquely identified  by the
+    tuple against it.
+}{vdp-abs-paths}
+
+Going back to the~\exref{vdp}, $ScatterAndSimulate$ is able to find
+    several abstract trajectories as shown in~\figref{vdp-abs-paths}.
+    However, typically, the refinement by increasing the quantization
+    precision requires $4$ refinement iterations and finally finds a
+    concrete counter-example for $\epsilon=0.125$.
+
+\end{example}

### 2017-06-12 response.txt

Response to the reviwers

============================================================================
EMSOFT 2017 Reviews for Submission 159
============================================================================

Title: Piece-Wise Relational Models for Falsifying Safety Properties of Hybrid Systems

Authors: Aditya Zutshi, Sergio Mover, Sriram Sankaranarayanan and Jyotirmoy Deshmukh
============================================================================
                            REVIEWER 1
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 2
                       Originality (1-4): 4
               Technical Soundness (1-4): 2
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

This paper is about checking safety properties on continuous models of an
embedded controller's environment. The continuous model is provided as a
black-box function capable of simulating Δ seconds from a given state with
constant input values. From what I can understand: (i) the state-space of
the system is explored by running multiple simulations; (ii) the resulting
states are clustered based on a quantization function;

(iii) an affine
relation (the convex hull) is calculated to characterize each cluster;

[TODO: Fix this]

(iv)
linear regression analysis is used to characterize the transitions between
clusters as affine relations, (iv) A finer abstraction is produced by
distinguish trajectories that initially pass through the same (k - 1)
clusters but then end up in different clusters, and (v) safety properties of
the resulting discrete "relational" abstraction are analyzed using SMT-based
bounded model checking. The work does not aim at soundness or completeness,
but rather to find problems in real embedded systems. The technique is a
variation of one proposed at EMSOFT in 2014 ([28]). It is demonstrated on
the academic examples from [28] with mixed results.

   ============
   [TODO: Add an explicit controller+plant benchmark: fuzzy control?]
   ============

I am not an expert in this area, but it seems to me that this paper presents
some novel and interesting ideas. I found, however, that after a great
introduction, the presentation becomes quite frustrating. There are a
succession of different formalisms and notations, which often lack precision
and explanations, and the links between them are not clearly described (I
had to battle to write the above summary and I am not certain that it is
correct). And, even after a careful reading of the text, I am still unsure
about one or two fundamental details:

1. The title states that the paper addresses "hybrid systems" and the system
   model includes discrete states, but thereafter the paper only treats
   continuous systems. Apparently this is reasonable because of an
   assumption that the model is a "switched-mode dynamical system".

   It would be nice to define exactly what you mean by "switched-mode
   dynamical system". Does it mean that only the controller model makes
   discrete transitions and that the environment is a continuous model with
   no discrete events or discontinuities?

   ============
   The systems under test are assumed to be switched-mode to ease the
   presentation, but it is not a limitation of the presented
   technique. Black box systems which have hybrid dynamics due to
   multiple modes, are dealt in the same fashion.
   [TODO: $K$-refinement to differentiate the dynamics of different modes.
   We have added a benchmark to showcase this.
   [TODO: which one?]
   [TODO: clarify the absence of any controllers]
   [TODO: remove or clarify switched-mode systems]
   [TODO: add a hybrid benchmark]
   ============

   I imagine that the idea is to combine the generated discrete abstraction
   with a discrete controller model to verify both together. It would be
   good to state this clearly and to treat the resulting issues throughout
   the paper. I note that the experimental results do not treat this issue.
   Otherwise, why not change the title and focus the text on what you
   actually did?

   ============
   We will explore this in the future, but the current presentation
   only works with the plant, which is itself a hybrid dynamical
   system.
   [TODO: Add this in the conclusion/future work]
   ============

   (Similar remarks hold at a smaller scale. The discussion often describes
    what "can be" done, e.g., in sections 3.2 and 4.1. I would rather read
    about what "has been" done and what the result was!)

   ============
   [TODO]
   ============

2. Why do you need a bounded model checker to check safety invariants when
   you already enumerate and characterize the reachable states? That is, why
   not just test the invariant on the convex hulls as they are generated?

   My guess is that you later want to restrict the state-space by composing
   your abstraction with a controller model. Is that right? This probably
   seems obvious to you, but please put yourself in the place of someone who
   is new to the work and not an expert in the domain! In any case, this
   idea is not clearly stated, not addressed by the text, and not considered
   in the experimental evaluation.

   ============
   [TODO: I don't understand this comment, but it is related to
   point (iii) of the summary]
   Although we enumerate the abstract space, we require the model
   checker to find a concrete counter-example.
   ============

   Is this why the PWA transition system characterizes states by a valuation
   function v and not simply as tuples? If yes, how will this work? I do not
   understand why you have a valuation function when the system only has one
   variable (x).
   ============
   [TODO]
   ============

Other comments and suggestions

* The paper puts a certain emphasis on "learning" a model. Would it not be
  more accurate to talk about "approximating" or "abstracting"?
   ============
   [TODO: Seems reasonable to me. Any suggestions??]
   ============

* It would be great to include an "overview figure" that lists all of the
  different formalisms and explains how they fit together. The enumeration
  in section 5 is very helpful, but it introduces even more notation and
  does not completely describe how all of the other graphs and transition
  systems are related.
   ============
   [TODO]
   ============

INTRODUCTION

* "Automatic approaches for testing the safety and performance of such
closed-loop systems is challenging," ->
  "Developing automatic approaches for testing the safety and performance of
such closed-loop systems is challenging," ?

* "Related approaches using bisimulation relations [20, 25, 27] have
interesting parallels with the work presented here.": What does this sentence
mean exactly?

LEARNING PWA RELATIONAL MODELS

* "The system statespace (denoted X) is a subset of $Q × R^n"$
-> "The system statespace, denoted X, is a subset of $Q × R^n"$

* "An input to S is assumed to be a time-bounded signal, i.e., a function
  from [0,T] to U, with T as the time horizon, and $U \subseteq R^k$.": Why not just
  use functions from [0,T] to $R^k$? What do you gain by introduction U (it is
  not exploited in the rest of the paper)?

* "we view S as effectively black-box"
-> "we view S as a black-box"

* "equipped with a forward simulation map, in a simulation time-step"
-> "equipped with a forward simulation map, which in a simulation time-step"

* "allows computing the unique reachable state x ∈ X"
-> "allows computing the unique reachable state x' ∈ X" (?)

* "the forward simulation map $SIMS^S_Δ : X × U × R^+ ↦ X": why R^+$?

* Why is Δ sometimes subscripted with t? Do you use a constant step-size or
  not? The text could be clearer.

* "Also, to ease the presentation, we conflate the inputs with the state and
  denote them by x.": This sentence could be clearer. I think you mean to
  say that you consider that the input values are included in the continuous
  state x. That said, you keep using u at the start of section 2.1.

* "several “proximal” tuples": "proximate"?

* "In statistics, linear regression is the problem of finding"
-> "Linear regression is the problem of finding"

* "For rest of the section, we elide" -> "For the rest of the paper, we
  elide" ?

* "The equation 2 now becomes"
-> "Equation 2 now becomes"

* "In this fashion, we can we can use OLS to approximate its trajectories of
  system S at fixed time step"
-> "In this fashion, we can use OLS to approximate the trajectories of
system S at a fixed time step"

* The definition of a "PWA transition system" is not a numbered definition
  (while the definition of a PWA relation model is). The "move function"
  seems to define the semantics of the transition systems. This could be
  more clear. You use G here for a set of conjunctions of affine predicates,
  and later it is used for graphs.

* "and use the name a variable as a"
-> "and use the name of the variable as a"

* "We how we can use the formal machinery of a PWA transition"?

* "S defined by a simulator SIMS∆~\footnote{...}."
-> "S defined by a simulator SIMS∆.\footnote{...}"

* Definition 2.2 is not precise enough. Why not systematically state the
  values of $L, x, T, G, U, l_0, Θ$? This would aid the poor reader by
  specifying the initial state and the set of locations. What is the set of
  locations?

* "the the valuation function for variable x"
-> "the valuation function for variable x"

* "A PWA model will be usually non-deterministic,"
-> "A PWA model will usually be non-deterministic,"

* "Abusing the notation,"
-> "Abusing notation,"

* "a single step in ρ by x' = ρ(x)": this abuse of notation is regrettable.
  It would be clearer and more precise, I believe, to avoid it.

* "iff for every state $x ∈ X^n": Why "^n"$?

* "In what follows, we use the terms PWA relational model and PWA
   transition system interchangeably to mean PWA relational model.": Please
   choose one term and use it consistently.

* "As all the guards and transitions are affine, a bounded reachability
  query is equivalent to checking all combinations of discrete locations,
  where each combination can be summarized as a linear program.": I did not
  understand what you are trying to say here.

* "Temporal properties of pwa transitions systems"
  -> "Temporal properties of PWA transitions systems" (and elsewhere).

CONNECTION WITH GRAPH-BASED ABSTRACTION

* I found the transition from the previous section very abrupt and it was
  unclear where all this was going.

* "As an alternative, we introduce data driven ‘enrichment’ of abstractions
  to approximate the concrete relations underlying the abstract relations.
  Informally, we use linear regression to compute affine maps associated
  with the relations between two abstract states.": I would say that the
  first statement is informal and the second is formal. I prefer the second
  as it says precisely what you do. "data driven enrichment of abstractions"
  is a great sounding phrase, but I think it is essentially empty.

* "The abstraction of the hybrid dynamical system is a a partitioning"
-> "The abstraction of a hybrid dynamical system is a partitioning"

* "This further defines an equivalence relations"
-> "This further defines an equivalence relation"

* "We call it a time parameterized PWA relational abstraction": Is there
  a real need to introduce more terminology? This name is never used again.

RELATIONAL MODELING EXTENSIONS

* "such a relation can be rarely represented"
-> "such a relation can rarely be represented"

* "Using a tool like FLOW*": missing citation or URL.

* "we rely on statistical methods like simple linear": which other
  statistical methods did you actually use?

* "with the below relation"
-> "with the following relation"

* "C ~ C' ⟺  ∃x ∈ C. ∃x' ∈ C'. x' ∈ Ax + b ± δ": Why do you start
using "±
  δ"? Figure 1, Definition 2.2, and §2.1 just use "+ δ".

* Figure 2 is unnecessary. It does not really show anything.

* "Form a state x ∈ C,"
-> "From a state x ∈ C,"

* "be taken as long as x' ∈ C’ is satisfied."
-> "be taken as long as x' ∈ C' is satisfied."

* "split them in to multiple sets"
-> "split them into multiple sets"

* "PWA transition system which only defines the evolution of states x,"
-> "PWA transition system that only defines the evolution of states x,"

* "trajectory segments form a cell C."
-> "trajectory segments from a cell C."

* "D = {(x, x')|x ∈ C}": Do you not need an extra clause to relate x and x'
  to the model transitions?

* "Figure 5: The data gets split into two sets"
-> "Figure 5: The data is split into two sets"

AN EXAMPLE: VAN DER POL OSCILLATOR

* "The corresponding reachability graph is H(∆ = 0.1).": more notation and
  without a precise definition!

* "which if exists, can indicate the existence of a"
-> "which if it exists, may indicate the existence of a"

* "numerical simulation function SIM ∆~\footnote{...}."
-> "numerical simulation function SIM ∆.\footnote{...}"

EXPERIMENTAL RESULTS

* "Its different parts comprised of the"
-> "Its different parts comprise the"

* "Finally, the The bounded model checking"
-> "Finally, the bounded model checking"

* "due to the differences in which they operate"
-> "due to the differences in how they operate"

* "requires a finer abstraction to begin with in order to minimize the"
-> "modeling errors. On the other hand, a coarser abstraction as the"

* Is table 1 really useful? How is it to be interpreted? What information is
  being conveyed?

* "As apparent form the results,"
-> "As apparent from the results,"

* "but coarse enough to be manageable by the current bmc"
-> "but coarse enough to be manageable by current BMC"

* "during the k-relational modeling process till"
-> "during the k-relational modeling process until"

* "Similarly, small time length trajectory"
-> "Similarly, small time-length trajectory"

REFERENCES

* [6] "Ph.D. Dissertation. Ph. D. thesis," -> "Ph. D. thesis,"

* [22] "(????)": ????

============================================================================
                            REVIEWER 2
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 4
          Presentation/Readability (1-4): 3
                       Originality (1-4): 4
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

This submission presents a model based methodology to find falsifications in
black box dynamical systems. After describing approaches based on data driven
modeling, arguing the pros and cons, the authors put forward a framework that
combines ideas from abstraction based search and relational abstractions and
defines the k-relational modeling to compute PWA relational abstractions. To
estimate the local dynamics of the trajectory segments and compute relational
models, which were then interpreted as infinite state discrete transition
system, the proposed framework utilizes simple regression for an
over-determined case. These infinite state discrete transition system
constructions were model checked for safety violations using an off-the-shelf
bounded model checkers and a naive linear programming based specialized model
checker for comparison. The framework was implemented as a tool chain, namely
the PWA-Rel and its performance was demonstrated on a few benchmarks.

The paper is well organized and written and should be accepted. I list next a
few issues that could improve the readability and the overall manuscript:
- The submission is missing an abstract.
-  The following sentence “For this step, we need to have models that can be
digested by existing symbolic tools, which is harder to do with some of the
aforementioned heavier data-driven modeling methods” requires a more in-depth
discussion. Please enrich the discussion especially because the paper starts on
such a negative tone without providing concrete examples. While I do not
disagree with the authors I feel that things could be made clearer with a few
explanatory sentences.

   ============
   [TODO: Can not use/scale generic neural-networks, non-linear
   kernels for regression,...]
   ============
- In section 2.1. the authors state that when N>n “a single affine function
cannot be found”. Should the problem be stated in finding the best choice of
a and b or should the problem considering determining the optimal number of
piece wise affine approximations? Can the magnitude N-n give a hint of what is
a good number of PWA / affine approximations? I feel that this could lead to a
nice results which the authors can easily cast in optimization (2).

   ============
   In general, finding the `optimal' number of PWA models can be
   framed as a multiobjective optimization problem, where a small
   number of clusters and an accurate affine map is desired. This is a
   hard problem for arbitrary non-linear systems~\cite{[TODO]}.
   Hence, in this paper, we presented a scalable falsification approach
   with a sub-optimal PWA identification approach.

   [TODO: OLS is not the main focus of the paper. We should state that
   and mention that any method for learning affine functions can be
   used.]
   ============

- In section 4.1 the authors state that “in the presence of complex
non-linear behavior, using a single affine map can lead to poor
approximation”. In this context and in relation with recent effort in CPS on
modeling complex dynamics (see “Constructing compact causal mathematical
models for complex dynamics” ICCPS2017) can the authors comment or speculate
on how the complexity of the dynamics might influence the required number of
affine maps? How does the non-smoothness and non-differentiability influence
the computation of R(C,C’)? Is there any relationship between the number of
nondifferentiable points and the number of required affine maps? It will also
be enrich the discussion of data-driven modeling techniques with recent efforts
in modeling non-smooth complex dynamics with fractal properties and hint
towards potential extensions.

   ============
   [TODO: ...]
   ============

- Another critical parameter in the framework seems to be the time steps Delta
used in section 4. Can the authors perform a sensitivity analysis to gauge its
impact on overall quality of the results?

   ============
   [TODO: ...]
   ============

- Since the main goal of this technique is to find violations of safety
properties in the embedded control system, it would prove useful for the reader
if the authors could comment which of the evaluations are considering this
specific goal. Again, this is not a criticism, but I feel that the
investigation in section 6 could be described in a few more sentences to
connect with embedded control.
   ============
   [TODO: ...]
   ============
- Lastly, the paper has a few misspellings and grammar issues. I summarize some
I found: “apparent form the results…”, “We how we can use”,
“Finally, the The bounded…”

============================================================================
                            REVIEWER 3
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 3
                       Originality (1-4): 2
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

The paper presents a method for falsifying safety properties of hybrid systems.
The method first constructs a relational abstraction of the system (which is
considered as a black box) from a set of simulation traces. The relations in
the abstraction are affine with uncertainty. The approach is quite interesting,
however the examples used for demonstration are unfortunately not black boxes,
although these examples are treated as black boxes, since it would be of great
interest to show its usefulness for complex models for which traditional
verification tools cannot be applied. A comparison with S3CAM is interesting,
but is less informative than a comparison with simulation-based validation
tools.
I also recommend using relations more complex than affine, such as polynomial,
for which Z3 can be used (nonlinear relations may reduce the number of
hyper-rectangles).

The method is of heuristic nature and thus convincing experimental results
would be needed. The paper compares with previous work (S3CAM) using a similar
approach. This approach has been tested on more real-life examples from
automative control. I am wondering why there are no comparisons on these
examples.

============================================================================
                            REVIEWER 4
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 2
                       Originality (1-4): 3
               Technical Soundness (1-4): 2
Account of prior work and references (1-3): 2

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

The paper presents a model-based methodology to do falsification for hybrid
systems. The technique combines the abstraction based search and relation
abstraction concept to learn a PWA abstract model from the simulations of a
system. The transitions between states of the abstract model characterized by
the affine relations between source and target states. A model checking
technique is then applied to find a counter-example for the abstract model.
This counter-example will be used to guide the search for a counter-example of
the original system. Experimental results of several academic benchmarks are
reported that show the capability of proposed technique in finding false
behavior.

It's clear to the reviewer that the authors may require more time to complete
the paper (e.g., the abstract is missing in the paper. The paper's organization
also needs to be improved. It's not clear for the reviewer to see the
high-level concept and contributions of the paper from the beginning.

Although the author claims that the proposed technique can effectively find a
counter-example for a black-box model, however, the experiment result does not
reflect that statement. Most of the benchmarks are nonlinear continuous systems
that are not black-box. I strongly suggest the author evaluate the proposed
approach on some black-box systems such as industrial Simulink models.

Also, using linear regression to infer affine dynamics and discrete structures
of the abstractions of hybrid systems has been applied in the past. Please
check out the related work on mining hybrid automata from input/output traces

Medhat, Ramy, et al. "A framework for mining hybrid automata from input/output
traces." Embedded Software (EMSOFT), 2015 International Conference on. IEEE,
2015.

Using OLS does not always result in finding the best fit. For example, if some
data points are excessively small or large compared to the rest of data, they
will have disproportionately large effects on the resulting constant. In this
case, using other techniques like ridge regression or lasso regression may
yield a better result.

As the approach is neither sound nor complete and the application perspective
for falsifying black-box system is not adequately demonstrated, it's hard to
assert that the contributions of the proposed approach are novel.

   ============
   [TODO: ...]
   ============

The paper contains some minor typos such as:

Page 2, OLS paragraph: the best choice for a and b, consistency of
capitalization

Page 3, definition 2.2, second bullet: the the valuation function -> delete
one the

Page 3, section 2.2: pwa -> PWA
Page 4, section 3.1, at the first line: a a partitioning of -> delete one
a
Page 7, section 6.1, at the first line: the The bounded model checking ->
delete The
Page 8, section 6.2: bmc -> BMC

============================================================================
                            REVIEWER 5
============================================================================

---------------------------------------------------------------------------
Reviewer's Scores
---------------------------------------------------------------------------

               Relevance to EMSOFT (1-4): 3
          Presentation/Readability (1-4): 3
                       Originality (1-4): 3
               Technical Soundness (1-4): 3
Account of prior work and references (1-3): 3

---------------------------------------------------------------------------
Comments
---------------------------------------------------------------------------

This is sophisticated stuff that, sadly, went over my head. Your motivation is
good and overall, this seems reasonable, but I couldn't easily wrap my head
around the sorts of system for which this technique would work. I suggest
leading with the Van Der Pol oscillator to help ignorant readers such as myself
get a handle on what it is you're actually able to do, then launch into the
formalism. As it stands, I think you need to be fairly expert in hybrid
dynamical systems to even begin to grasp what's going on here.

### 2017-06-12 learning.tex
learning

@@
 time as well as storage.  The central idea in this paper is to {\em
-summarize} several tuples as a single map obtained using regression.
-In other words, we can replace several ``proximal'' tuples
-$(\vx,\vx')$ in $\reach{t,u}$ by a map $F$ of the form $\vx' = F(\vx)
-+ \delta$, where $\delta$ is a worst-case error estimate. The form of
-$F$ depends on the kernel used for regression. For the special case of
-linear regression, the kernel is affine and the resulting map can be
-represented by a matrix $A$ as $\vx' = A\vx + \delta$.
+summarize} several tuples as a single map obtained using parametric
+regression.  In other words, we can replace several ``proximal''
+tuples $(\vx,\vx')$ in $\reach{t,u}$ by a map $F$ of the form $\vx' =
+F(\vx, \vu) + \delta$, where $\delta$ is a worst-case error estimate. The
+form of $F$ depends on the kernel used for regression. For the special
+case of linear regression, the kernel is affine and the resulting map
+can be represented by a matrices $A$ and $B$ as $\vx' = A\vx + B\vu +
+\delta$.
 
-In statistics, regression is the problem of finding an affine
-function or \textit{predictor}, which can `best' summarize the
-relationship between the given set of observed input $\x$ and output
-$\x'$ vectors. The notion of `best' is formally captured using a loss
-function which can be non-linear. We use the commonly used loss
-function \textit{ordinary least squares} (OLS) for this presentation.
+In statistics, parametric regression is the problem of finding the
+paremters of a function or \textit{predictor}, which can `best'
+summarize the relationship between the given set of observed input
+$\x$ and output $\x'$ vectors. The model or the kernel for the
+predictor is fixed beforehand and the paramters are searched for.  The
+notion of `best' is formally captured using a loss function which can
+be non-linear. We use the commonly used loss function \textit{ordinary
+least squares} (OLS) to fit affine and polynomial kernels of fixed degree.
+~\footnote{The choice of loss function and regression analysis affects the quality of
+the learnt models. As the procss of selecting an appropriate heuristic
+is not straightforward, we just use the simplest one.}
 
@@
 the best fit}, the problem is over-determined: there are more
-equations than unknowns. Hence, a single affine function cannot be
-found which satisfies the equation \eqref{eq:affinemap}. Instead, we
-need to find the `best' choice for $\vec{a}$ and $b$. This choice is
-formally defined using a loss function. For the case of simple linear
-regression or OLS, the loss function is the sum of squares of the
-errors in prediction.  The task is then to determine $A$ and $B$, such
-that the least square error of the affine predictor is minimized for
-the given data set.
+equations than unknowns. Hence, a single set of parameters cannot be
+found which fits all the data. Instead, we need to find the `best'
+choice. This choice is formally defined using a loss function. For the
+case of simple linear regression or OLS, the loss function is the sum
+of squares of the errors in prediction.  The task is then to determine
+$A$ and $B$, such that the least square error of the affine predictor
+is minimized for the given data set.
 
-\mypara{Ordinary Least Squares (OLS)}
 For rest of the section, we elide the discrete mode from the hybrid
@@
 form $(\x,\x')$ in the reachability relation that we have explored,
-where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn an
-affine function described as $\x' = A\x + B$.
-Here, $A$ is a $n\times n$ matrix, and $B$ is a $n\times 1$ matrix.
-The details can be found in standard texts on learning and
+where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn a
+function described as $\x' = F(\vx, \vu)$.  The details can be found
+in standard texts on learning and
 statistics~\cite{friedman2001elements}.
 
-In this fashion, we can we can use OLS to approximate its trajectories
-of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
-= \vec{A}\hat{\x}$.
+% In this fashion, we can we can use OLS to approximate its trajectories
+% of system $\System$ at fixed time step $\Delta$ by a discrete map $\x'
+% = F(\hat{\x})$.
 

### 2017-06-12 learning.tex
.

@@
 be non-linear. We use the commonly used loss function \textit{ordinary
-least squares} (OLS) to fit affine and polynomial kernels of fixed degree.
-~\footnote{The choice of loss function and regression analysis affects the quality of
-the learnt models. As the procss of selecting an appropriate heuristic
-is not straightforward, we just use the simplest one.}
-
-If $N>n$, which is the case in the current context of \textit{finding
-the best fit}, the problem is over-determined: there are more
-equations than unknowns. Hence, a single set of parameters cannot be
-found which fits all the data. Instead, we need to find the `best'
-choice. This choice is formally defined using a loss function. For the
-case of simple linear regression or OLS, the loss function is the sum
-of squares of the errors in prediction.  The task is then to determine
-$A$ and $B$, such that the least square error of the affine predictor
-is minimized for the given data set.
+least squares} (OLS) to fit affine and polynomial kernels of fixed
+degree.  ~\footnote{The choice of loss function and regression
+analysis affects the quality of the learnt models. As the procss of
+selecting an appropriate heuristic is not straightforward, we just use
+the simplest one.} OLS defines its loss function as the sum of squares
+of the errors in prediction. The task is then to determine the
+parameters, such that the least square error of the predictor is
+minimized for the given data set.
 

### 2017-06-12 implementation.tex
.

@@
 was evluated with both the symbolic BMC SAL (PWA-Rel+SAL) and the
-explicit BMC (PWA-Rel+OPT).  \todo{SatEx was also tried, but ...}. We
+explicit BMC (PWA-Rel+OPT). %\todo{SatEx was also tried, but ...}.
+We
 tabulate both the total time taken by PWA-Rel+SAL and PWA-Rel+OPT to
@@
 refinement procedure involved.~\footnote{In all our experiments we
-used $1$-relational modeling to create the transition system from $G$.
+used $1$-relational modeling to create the transition system.
 This was determined to be a good trade-off between longer simulations

### 2017-06-12 learning.tex

@@
 
-We note that linear regression tools are often able to provide an
+We note that regression tools are often able to provide an
 $n\times 1$ vector $\vec{\delta}$ of error intervals, where the
@@
 
-Affine maps are poor approximations for arbitrary non-linear
-functions. Hence, we use a collection of affine maps to approximate
-the local behaviors (in state-space) of the system $\System$. This
+The choice of kernels for the relational abstraction involves a
+tradeoff between computing precise local models and efficiency of the
+analysis of the resulting relational abstraction. Affine kernels are
+poor approximations for complex dynamics and will result in large
+errors, but they can be very efficiently analyzed in practice.  On the
+other hand, analysis of arbitrary non-linear functions is still not
+tractable. Hence, we restrict ourselves to affine and quadratic
+polynomial kernels.  Moreover, as we use a collection of relations to
+approximate the local behaviors (in state-space) of the system
+$\System$, we can often do away with simple kernels. This however,
 results in a piece wise approximation, as described in the next

### 2017-06-12 appendix.tex

@@
 
-Authors: Aditya Zutshi, Sergio Mover, Sriram Sankaranarayanan and Jyotirmoy Deshmukh
 ============================================================================
@@
 ---------------------------------------------------------------------------
-
-This paper is about checking safety properties on continuous models of an
-embedded controller's environment. The continuous model is provided as a
-black-box function capable of simulating Δ seconds from a given state with
-constant input values. From what I can understand: (i) the state-space of
-the system is explored by running multiple simulations; (ii) the resulting
-states are clustered based on a quantization function;
-
-(iii) an affine
-relation (the convex hull) is calculated to characterize each cluster;
-
-[TODO: Fix this]
-
-(iv)
-linear regression analysis is used to characterize the transitions between
-clusters as affine relations, (iv) A finer abstraction is produced by
-distinguish trajectories that initially pass through the same (k - 1)
-clusters but then end up in different clusters, and (v) safety properties of
-the resulting discrete "relational" abstraction are analyzed using SMT-based
-bounded model checking. The work does not aim at soundness or completeness,
-but rather to find problems in real embedded systems. The technique is a
-variation of one proposed at EMSOFT in 2014 ([28]). It is demonstrated on
-the academic examples from [28] with mixed results.
-
-   ============
-   [TODO: Add an explicit controller+plant benchmark: fuzzy control?]
-   ============
-
-I am not an expert in this area, but it seems to me that this paper presents
-some novel and interesting ideas. I found, however, that after a great
-introduction, the presentation becomes quite frustrating. There are a
-succession of different formalisms and notations, which often lack precision
-and explanations, and the links between them are not clearly described (I
-had to battle to write the above summary and I am not certain that it is
-correct). And, even after a careful reading of the text, I am still unsure
-about one or two fundamental details:
 
@@
    ============
-
-Other comments and suggestions
-
-* The paper puts a certain emphasis on "learning" a model. Would it not be
-  more accurate to talk about "approximating" or "abstracting"?
-   ============
-   [TODO: Seems reasonable to me. Any suggestions??]
-   ============
-
-* It would be great to include an "overview figure" that lists all of the
-  different formalisms and explains how they fit together. The enumeration
-  in section 5 is very helpful, but it introduces even more notation and
-  does not completely describe how all of the other graphs and transition
-  systems are related.
-   ============
-   [TODO]
-   ============
-
-INTRODUCTION
-
-* "Automatic approaches for testing the safety and performance of such
-closed-loop systems is challenging," ->
-  "Developing automatic approaches for testing the safety and performance of
-such closed-loop systems is challenging," ?
-
-* "Related approaches using bisimulation relations [20, 25, 27] have
-interesting parallels with the work presented here.": What does this sentence
-mean exactly?
-
-LEARNING PWA RELATIONAL MODELS
-
-* "The system statespace (denoted X) is a subset of $Q × R^n"$
--> "The system statespace, denoted X, is a subset of $Q × R^n"$
-
-* "An input to S is assumed to be a time-bounded signal, i.e., a function
-  from [0,T] to U, with T as the time horizon, and $U \subseteq R^k$.": Why not just
-  use functions from [0,T] to $R^k$? What do you gain by introduction U (it is
-  not exploited in the rest of the paper)?
-
-* "we view S as effectively black-box"
--> "we view S as a black-box"
-
-* "equipped with a forward simulation map, in a simulation time-step"
--> "equipped with a forward simulation map, which in a simulation time-step"
-
-* "allows computing the unique reachable state x ∈ X"
--> "allows computing the unique reachable state x' ∈ X" (?)
-
-* "the forward simulation map $SIMS^S_Δ : X × U × R^+ ↦ X": why R^+$?
-
-* Why is Δ sometimes subscripted with t? Do you use a constant step-size or
-  not? The text could be clearer.
-
-* "Also, to ease the presentation, we conflate the inputs with the state and
-  denote them by x.": This sentence could be clearer. I think you mean to
-  say that you consider that the input values are included in the continuous
-  state x. That said, you keep using u at the start of section 2.1.
-
-* "several “proximal” tuples": "proximate"?
-
-* "In statistics, linear regression is the problem of finding"
--> "Linear regression is the problem of finding"
-
-* "For rest of the section, we elide" -> "For the rest of the paper, we
-  elide" ?
-
-* "The equation 2 now becomes"
--> "Equation 2 now becomes"
-
-* "In this fashion, we can we can use OLS to approximate its trajectories of
-  system S at fixed time step"
--> "In this fashion, we can use OLS to approximate the trajectories of
-system S at a fixed time step"
-
-* The definition of a "PWA transition system" is not a numbered definition
-  (while the definition of a PWA relation model is). The "move function"
-  seems to define the semantics of the transition systems. This could be
-  more clear. You use G here for a set of conjunctions of affine predicates,
-  and later it is used for graphs.
-
-* "and use the name a variable as a"
--> "and use the name of the variable as a"
-
-* "We how we can use the formal machinery of a PWA transition"?
-
-* "S defined by a simulator SIMS∆~\footnote{...}."
--> "S defined by a simulator SIMS∆.\footnote{...}"
-
-* Definition 2.2 is not precise enough. Why not systematically state the
-  values of $L, x, T, G, U, l_0, Θ$? This would aid the poor reader by
-  specifying the initial state and the set of locations. What is the set of
-  locations?
-
-* "the the valuation function for variable x"
--> "the valuation function for variable x"
-
-* "A PWA model will be usually non-deterministic,"
--> "A PWA model will usually be non-deterministic,"
-
-* "Abusing the notation,"
--> "Abusing notation,"
-
-* "a single step in ρ by x' = ρ(x)": this abuse of notation is regrettable.
-  It would be clearer and more precise, I believe, to avoid it.
-
-* "iff for every state $x ∈ X^n": Why "^n"$?
-
-* "In what follows, we use the terms PWA relational model and PWA
-   transition system interchangeably to mean PWA relational model.": Please
-   choose one term and use it consistently.
-
-* "As all the guards and transitions are affine, a bounded reachability
-  query is equivalent to checking all combinations of discrete locations,
-  where each combination can be summarized as a linear program.": I did not
-  understand what you are trying to say here.
-
-* "Temporal properties of pwa transitions systems"
-  -> "Temporal properties of PWA transitions systems" (and elsewhere).
-
-CONNECTION WITH GRAPH-BASED ABSTRACTION
-
-* I found the transition from the previous section very abrupt and it was
-  unclear where all this was going.
-
-* "As an alternative, we introduce data driven ‘enrichment’ of abstractions
-  to approximate the concrete relations underlying the abstract relations.
-  Informally, we use linear regression to compute affine maps associated
-  with the relations between two abstract states.": I would say that the
-  first statement is informal and the second is formal. I prefer the second
-  as it says precisely what you do. "data driven enrichment of abstractions"
-  is a great sounding phrase, but I think it is essentially empty.
-
-* "The abstraction of the hybrid dynamical system is a a partitioning"
--> "The abstraction of a hybrid dynamical system is a partitioning"
-
-* "This further defines an equivalence relations"
--> "This further defines an equivalence relation"
-
-* "We call it a time parameterized PWA relational abstraction": Is there
-  a real need to introduce more terminology? This name is never used again.
-
-RELATIONAL MODELING EXTENSIONS
-
-* "such a relation can be rarely represented"
--> "such a relation can rarely be represented"
-
-* "Using a tool like FLOW*": missing citation or URL.
-
-* "we rely on statistical methods like simple linear": which other
-  statistical methods did you actually use?
-
-* "with the below relation"
--> "with the following relation"
-
-* "C ~ C' ⟺  ∃x ∈ C. ∃x' ∈ C'. x' ∈ Ax + b ± δ": Why do you start
-using "±
-  δ"? Figure 1, Definition 2.2, and §2.1 just use "+ δ".
-
-* Figure 2 is unnecessary. It does not really show anything.
-
-* "Form a state x ∈ C,"
--> "From a state x ∈ C,"
-
-* "be taken as long as x' ∈ C’ is satisfied."
--> "be taken as long as x' ∈ C' is satisfied."
-
-* "split them in to multiple sets"
--> "split them into multiple sets"
-
-* "PWA transition system which only defines the evolution of states x,"
--> "PWA transition system that only defines the evolution of states x,"
-
-* "trajectory segments form a cell C."
--> "trajectory segments from a cell C."
-
-* "D = {(x, x')|x ∈ C}": Do you not need an extra clause to relate x and x'
-  to the model transitions?
-
-* "Figure 5: The data gets split into two sets"
--> "Figure 5: The data is split into two sets"
-
-AN EXAMPLE: VAN DER POL OSCILLATOR
-
-* "The corresponding reachability graph is H(∆ = 0.1).": more notation and
-  without a precise definition!
-
-* "which if exists, can indicate the existence of a"
--> "which if it exists, may indicate the existence of a"
-
-* "numerical simulation function SIM ∆~\footnote{...}."
--> "numerical simulation function SIM ∆.\footnote{...}"
-
-EXPERIMENTAL RESULTS
-
-* "Its different parts comprised of the"
--> "Its different parts comprise the"
-
-* "Finally, the The bounded model checking"
--> "Finally, the bounded model checking"
-
-* "due to the differences in which they operate"
--> "due to the differences in how they operate"
-
-* "requires a finer abstraction to begin with in order to minimize the"
--> "modeling errors. On the other hand, a coarser abstraction as the"
-
-* Is table 1 really useful? How is it to be interpreted? What information is
-  being conveyed?
-
-* "As apparent form the results,"
--> "As apparent from the results,"
-
-* "but coarse enough to be manageable by the current bmc"
--> "but coarse enough to be manageable by current BMC"
-
-* "during the k-relational modeling process till"
--> "during the k-relational modeling process until"
-
-* "Similarly, small time length trajectory"
--> "Similarly, small time-length trajectory"
-
-REFERENCES
-
-* [6] "Ph.D. Dissertation. Ph. D. thesis," -> "Ph. D. thesis,"
-
-* [22] "(????)": ????
-
 ============================================================================
@@
 
-This submission presents a model based methodology to find falsifications in
-black box dynamical systems. After describing approaches based on data driven
-modeling, arguing the pros and cons, the authors put forward a framework that
-combines ideas from abstraction based search and relational abstractions and
-defines the k-relational modeling to compute PWA relational abstractions. To
-estimate the local dynamics of the trajectory segments and compute relational
-models, which were then interpreted as infinite state discrete transition
-system, the proposed framework utilizes simple regression for an
-over-determined case. These infinite state discrete transition system
-constructions were model checked for safety violations using an off-the-shelf
-bounded model checkers and a naive linear programming based specialized model
-checker for comparison. The framework was implemented as a tool chain, namely
-the PWA-Rel and its performance was demonstrated on a few benchmarks.
-
-The paper is well organized and written and should be accepted. I list next a
-few issues that could improve the readability and the overall manuscript:
-- The submission is missing an abstract.
--  The following sentence “For this step, we need to have models that can be
-digested by existing symbolic tools, which is harder to do with some of the
-aforementioned heavier data-driven modeling methods” requires a more in-depth
-discussion. Please enrich the discussion especially because the paper starts on
-such a negative tone without providing concrete examples. While I do not
-disagree with the authors I feel that things could be made clearer with a few
-explanatory sentences.
-
-   ============
-   [TODO: Can not use/scale generic neural-networks, non-linear
-   kernels for regression,...]
-   ============
 - In section 2.1. the authors state that when N>n “a single affine function
@@
 
-   ============
-   In general, finding the `optimal' number of PWA models can be
-   framed as a multiobjective optimization problem, where a small
-   number of clusters and an accurate affine map is desired. This is a
-   hard problem for arbitrary non-linear systems~\cite{[TODO]}.
-   Hence, in this paper, we presented a scalable falsification approach
-   with a sub-optimal PWA identification approach.
+>> In general, finding the `optimal' number of PWA models can be
+>> framed as a multiobjective optimization problem, where a small
+>> number of clusters and an accurate affine map is desired. This is a
+>> hard problem for arbitrary non-linear systems.
+>> Hence, in this paper, we presented a scalable falsification approach
+>> with a sub-optimal PWA identification approach.
 
-   [TODO: OLS is not the main focus of the paper. We should state that
-   and mention that any method for learning affine functions can be
-   used.]
-   ============
-
-- In section 4.1 the authors state that “in the presence of complex
-non-linear behavior, using a single affine map can lead to poor
-approximation”. In this context and in relation with recent effort in CPS on
-modeling complex dynamics (see “Constructing compact causal mathematical
-models for complex dynamics” ICCPS2017) can the authors comment or speculate
-on how the complexity of the dynamics might influence the required number of
-affine maps? How does the non-smoothness and non-differentiability influence
-the computation of R(C,C’)? Is there any relationship between the number of
-nondifferentiable points and the number of required affine maps? It will also
-be enrich the discussion of data-driven modeling techniques with recent efforts
-in modeling non-smooth complex dynamics with fractal properties and hint
-towards potential extensions.
-
-   ============
-   [TODO: ...]
-   ============
+Can the authors comment or speculate on how the complexity of the
+dynamics might influence the required number of affine maps? How does
+the non-smoothness and non-differentiability influence the computation
+of R(C,C’)? Is there any relationship between the number of
+nondifferentiable points and the number of required affine maps?
 
@@
 
-   ============
-   [TODO: ...]
-   ============
+>> It is a good idea, and it can have a big affect on the quality of
+the results. However, selecting a `good' Delta is not straightforward.
 
@@
 connect with embedded control.
-   ============
-   [TODO: ...]
-   ============
-- Lastly, the paper has a few misspellings and grammar issues. I summarize some
-I found: “apparent form the results…”, “We how we can use”,
-“Finally, the The bounded…”
 
@@
 
-The paper presents a model-based methodology to do falsification for hybrid
-systems. The technique combines the abstraction based search and relation
-abstraction concept to learn a PWA abstract model from the simulations of a
-system. The transitions between states of the abstract model characterized by
-the affine relations between source and target states. A model checking
-technique is then applied to find a counter-example for the abstract model.
-This counter-example will be used to guide the search for a counter-example of
-the original system. Experimental results of several academic benchmarks are
-reported that show the capability of proposed technique in finding false
-behavior.
-
-It's clear to the reviewer that the authors may require more time to complete
-the paper (e.g., the abstract is missing in the paper. The paper's organization
-also needs to be improved. It's not clear for the reviewer to see the
-high-level concept and contributions of the paper from the beginning.
-
 Although the author claims that the proposed technique can effectively find a
@@
 approach on some black-box systems such as industrial Simulink models.
-
-Also, using linear regression to infer affine dynamics and discrete structures
-of the abstractions of hybrid systems has been applied in the past. Please
-check out the related work on mining hybrid automata from input/output traces
-
-Medhat, Ramy, et al. "A framework for mining hybrid automata from input/output
-traces." Embedded Software (EMSOFT), 2015 International Conference on. IEEE,
-2015.
 
@@
 
-As the approach is neither sound nor complete and the application perspective
-for falsifying black-box system is not adequately demonstrated, it's hard to
-assert that the contributions of the proposed approach are novel.
-
-   ============
-   [TODO: ...]
-   ============
-
-The paper contains some minor typos such as:
-
-Page 2, OLS paragraph: the best choice for a and b, consistency of
-capitalization
-
-Page 3, definition 2.2, second bullet: the the valuation function -> delete
-one the
-
-Page 3, section 2.2: pwa -> PWA
-Page 4, section 3.1, at the first line: a a partitioning of -> delete one
-a
-Page 7, section 6.1, at the first line: the The bounded model checking ->
-delete The
-Page 8, section 6.2: bmc -> BMC
+>> As the choice of loss function and type of regression analysis
+>> depends on the problem at hand, we show our methodology using the
+>> simplest one. In practice, sophesticated statistical tests can be used
+>> to find the a good analysis.
 
@@
 
-This is sophisticated stuff that, sadly, went over my head. Your motivation is
-good and overall, this seems reasonable, but I couldn't easily wrap my head
-around the sorts of system for which this technique would work. I suggest
-leading with the Van Der Pol oscillator to help ignorant readers such as myself
-get a handle on what it is you're actually able to do, then launch into the
-formalism. As it stands, I think you need to be fairly expert in hybrid
-dynamical systems to even begin to grasp what's going on here.
+I suggest leading with the Van Der Pol oscillator to help ignorant
+readers such as myself get a handle on what it is you're actually able
+to do, then launch into the formalism.
+
+>> We have made the Van Der Pol oscillator as a running example to improve
+>> the presentation.

### 2017-06-12 learning.tex
pass over learning

--- 2017-06-12 learning.tex
@@
-The approach in \cite{zutshi2014multiple} uses the forward simulation
-map to explore a fixed number of tuples in the reachability relation
-$\reach{t,\vu}$ specified by a user-provided budget.  In general, the
-budget required to adequately explore the transition space requires
-exploring an exponential (in the dimension of $\HybridStates$) number
-of such tuples. This can cause undue burden in terms of computation
+The $\mathsf{ScatterAndSimulate}$ procedure helps in constructing the
+$\epsilon$-TS abstraction using a user-specified budget of trajectory
+fragments (obtained using simulation).  In general, the budget
+required to adequately explore the transition space requires exploring
+a number of states that is exponential in the dimension of
+$\HybridStates$.  This can cause undue burden in terms of computation
 time as well as storage.  The central idea in this paper is to {\em
-summarize} several tuples as a single map obtained using parametric
-regression.  In other words, we can replace several ``proximal''
-tuples $(\vx,\vx')$ in $\reach{t,u}$ by a map $F$ of the form $\vx' =
-F(\vx, \vu) + \delta$, where $\delta$ is a worst-case error estimate. The
-form of $F$ depends on the kernel used for regression. For the special
-case of linear regression, the kernel is affine and the resulting map
-can be represented by a matrices $A$ and $B$ as $\vx' = A\vx + B\vu +
-\delta$.
+summarize} enrich the transitions in the $\epsilon$-TS abstraction
+using relations to obtain an $\epsilon$-RTS.
+
+In other words, suppose there is an edge from cell $C$ to $C'$, and if
+$\vx \in C$ and $\vx' \in C'$, then we can consider all concrete
+trajectories that go from $C$ to $C'$, and use these to replace
+several ``proximal'' tuples $(\vx,\vxp)$ by a map $F$ of the form
+$\vxp = F(\vx, \vu) \pm \delta$, where $\delta$ is a worst-case error
+estimate. The form of $F$ depends on the basis function used for
+regression. We focus on linear regression with different basis
+functions such as affine and polynomial. For the special case of
+linear regression, the resulting update relation can be represented by
+matrices $A$ and $B$, \ie, $r(\vx,\vxp) = \setof{(\vx,\vxp) \mid |\vxp
+- (A\vx + B\vu)| < \delta}$.
 
@@
 summarize the relationship between the given set of observed input
-$\x$ and output $\x'$ vectors. The model or the kernel for the
-predictor is fixed beforehand and the paramters are searched for.  The
-notion of `best' is formally captured using a loss function which can
-be non-linear. We use the commonly used loss function \textit{ordinary
-least squares} (OLS) to fit affine and polynomial kernels of fixed
-degree.  ~\footnote{The choice of loss function and regression
-analysis affects the quality of the learnt models. As the procss of
-selecting an appropriate heuristic is not straightforward, we just use
-the simplest one.} OLS defines its loss function as the sum of squares
-of the errors in prediction. The task is then to determine the
-parameters, such that the least square error of the predictor is
-minimized for the given data set.
+$\x$ and output $\x'$ vectors. The basis functions used for the
+predictor are fixed beforehand and the paramters are searched for.
+The notion of `best' is formally captured using a loss function which
+can be non-linear. We use the commonly used loss function
+\textit{ordinary least squares} (OLS) to fit affine and polynomial
+kernels of fixed degree.  ~\footnote{The choice of loss function and
+regression analysis affects the quality of the learnt models. As the
+procss of selecting an appropriate heuristic is not straightforward,
+we just use the simplest one.} OLS defines its loss function as the
+sum of squares of the errors in prediction. The task is then to
+determine the parameters, such that the least square error of the
+predictor is minimized for the given data set.  The details on
+performing linear regressoin can be found in standard texts on
+learning and statistics~\cite{friedman2001elements}.
 
-For rest of the section, we elide the discrete mode from the hybrid
-state of the system, and only focus on the continuous state. Note that
-we can do this because of our assumption that the system is a
-switched-mode dynamical system.  Suppose we have $N$ tuples of the
-form $(\x,\x')$ in the reachability relation that we have explored,
-where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn a
-function described as $\x' = F(\vx, \vu)$.  The details can be found
-in standard texts on learning and
-statistics~\cite{friedman2001elements}.
+% Suppose we have $N$ tuples of the
+% form $(\vx,\vxp)$ in the reachability relation that we have explored,
+% where $\x\in\reals^n$ and $\x'\in\reals^n$. Then, we wish to learn a
+% function described as $\x' = F(\vx, \vu)$.
 
@@
 indicates the best and worst possible error in computation of the
-$i^{th}$ state in $\x'$.
+in $\vxp_i$.
 
-Affine maps are poor approximations for arbitrary non-linear
-functions. Hence, we use a collection of affine maps to approximate
-the local behaviors (in state-space) of the system $\System$. This
-results in a piece wise approximation, as described in the next
-section.
+The choice of basis functions for constructing an $\epsilon$-RTS
+involves a tradeoff between computing precise local models and
+efficiency of the analysis of the resulting relational abstraction.
+Affine basis functions can be poor approximations for complex dynamics
+and can result in large errors, but they can be very efficiently
+analyzed in practice.  On the other hand, analysis of arbitrary
+non-linear functions is not generally tractable. We restrict ourselves
+to affine and quadratic polynomial basis functions.  Moreover, as we
+use a collection of relations to approximate the local behaviors (in
+state-space) of the system $\System$, we can often do so with simple
+basis functions.
 
@@
     set of trajectory segments, and the affine relation is computed.
-    \figref{vdp-abs-paths} shows the cells and the trajectory segments, which
-    are part of the data sets constructed using the $1$-relational
-    modeling. Against each cell, its unique identifier (integer
-    co-ordinate) is mentioned.
+    \figref{vdp-abs-paths} shows the cells and the trajectory
+    segments, which are part of the data sets constructed using the
+    $1$-relational modeling. Against each cell, its unique identifier
+    (integer co-ordinate) is mentioned.
 \end{example}

### 2017-06-12 appendix.tex
appendix

@@
-============================================================================
+\obeyspaces
+\obeylines
+======================================
 EMSOFT 2017 Reviews for Submission 159
-============================================================================
+======================================
 
@@
 
-============================================================================
+=====================================
                             REVIEWER 1
-============================================================================
+=====================================
 
----------------------------------------------------------------------------
+----------------------------------
 Reviewer's Scores
----------------------------------------------------------------------------
+---------------------------------
 
@@
 
----------------------------------------------------------------------------
+------------------------------------
 Comments
----------------------------------------------------------------------------
+-----------------------------------
 
@@
    ============
-============================================================================
+======================================
                             REVIEWER 2
-============================================================================
+======================================
 
----------------------------------------------------------------------------
+--------------------------------------
 Reviewer's Scores
----------------------------------------------------------------------------
+--------------------------------------
 
@@
 
----------------------------------------------------------------------------
+--------------------------------------
 Comments
----------------------------------------------------------------------------
+--------------------------------------
 
@@
 
-============================================================================
+======================================
                             REVIEWER 3
-============================================================================
+======================================
 
----------------------------------------------------------------------------
+--------------------------------------
 Reviewer's Scores
----------------------------------------------------------------------------
+--------------------------------------
 
@@
 
----------------------------------------------------------------------------
+--------------------------------------
 Comments
----------------------------------------------------------------------------
+--------------------------------------
 
-The paper presents a method for falsifying safety properties of hybrid systems.
-The method first constructs a relational abstraction of the system (which is
-considered as a black box) from a set of simulation traces. The relations in
-the abstraction are affine with uncertainty. The approach is quite interesting,
-however the examples used for demonstration are unfortunately not black boxes,
-although these examples are treated as black boxes, since it would be of great
-interest to show its usefulness for complex models for which traditional
-verification tools cannot be applied. A comparison with S3CAM is interesting,
-but is less informative than a comparison with simulation-based validation
-tools.
-I also recommend using relations more complex than affine, such as polynomial,
-for which Z3 can be used (nonlinear relations may reduce the number of
-hyper-rectangles).
+A comparison with S3CAM is interesting, but is less informative than a
+comparison with simulation-based validation tools.
 
-The method is of heuristic nature and thus convincing experimental results
-would be needed. The paper compares with previous work (S3CAM) using a similar
-approach. This approach has been tested on more real-life examples from
-automative control. I am wondering why there are no comparisons on these
-examples.
+I also recommend using relations more complex than affine, such as
+polynomial, for which Z3 can be used (nonlinear relations may reduce
+the number of hyper-rectangles).
 
-============================================================================
+======================================
                             REVIEWER 4
-============================================================================
+======================================
 
----------------------------------------------------------------------------
+--------------------------------------
 Reviewer's Scores
----------------------------------------------------------------------------
+--------------------------------------
 
@@
 
-============================================================================
+======================================
                             REVIEWER 5
-============================================================================
+======================================
 
----------------------------------------------------------------------------
+--------------------------------------
 Reviewer's Scores
----------------------------------------------------------------------------
+--------------------------------------
 

### 2017-06-12 intro.tex
.

@@
 into a non-deterministic infinite state transition system that is
-required to be an overapproximation of the original system and use
+required to be an over-approximation of the original system and use
 model checking approaches to verify them \cite{asarin2000approximate}.

### 2017-06-12 scam.tex
.

--- 2017-06-12 scam.tex
@@
 
+% \inclfig[width=0.95\linewidth]{vdp_cont_traces.pdf}{Van der
+% Pol: continuous trajectories. Red and green boxes indicate unsafe and
+% initial sets.}{vdp-cont}
 \inclfig[width=0.95\linewidth]{vdp_abs.pdf}{The discovered
@@
     generated using uniformly random samples are shown in
-    \figref{vdp-cont}. We want to check the property $P_3$:
+    \figref{vdp-abs}. We want to check the property $P_3$:
     $x_{unsafe}\in[-1,-6.5]$ and $y_{unsafe}\in[-0.7, -5.6]$, given
@@
 \inclfig[width=0.90\linewidth]{vdp_disc_map.pdf}{
-Abstract trajectories discovered by $\mathsf{ScatterAndSimulate}$. Each
+Abstract trajectories discovered by $ScatterAndSimulate$. Each
     rectangle denotes a cell which is uniquely identified  by the
     tuple against it.
-%Cells and trajectory segments used by $1$-relational modeling.
 }{vdp-abs-paths}
 
-    Going back to the~\exref{vdp}, $\mathsf{ScatterAndSimulate}$ is able to
-    find several abstract trajectories as shown
-    in~\figref{vdp-abs-paths}. However, typically, the refinement by
-    increasing the quantization precision requires $4$ refinement
-    iterations and finally finds a concrete counter-example for
-    $\epsilon=0.125$.
+Going back to the~\exref{vdp}, $ScatterAndSimulate$ is able to find
+    several abstract trajectories as shown in~\figref{vdp-abs-paths}.
+    However, typically, the refinement by increasing the quantization
+    precision requires $4$ refinement iterations and finally finds a
+    concrete counter-example for $\epsilon=0.125$.
+
 \end{example}

### 2017-06-12 bmc_section.tex
.

@@
 
-\inclfig[width=0.50\linewidth]{vdp_graph.pdf}{Enriched
-graph $G^R$. The affine maps $f$ for the transition relations are show
-in \tabref{vdp-pwa}.}{vdp-graph}
+\inclfig[width=0.50\linewidth]{vdp_graph.pdf}{$\epsilon$-RTS for the
+Van der Pol abstraction. Each edge corresponds to an affine map $F$
+and summarizes the transition relation.}{vdp-graph}
 \begin{example}

### 2017-06-12 bmc_section.tex
.

@@
 \begin{example}
-    Coming back to~\exref{vdp}, \figref{vdp-graph} %and\tabref{vdp-pwa}
+    Coming back to~\exref{vdp}, \figref{vdp-graph}
     shows the abstract reachability graph with its transition
@@
     Note that self loops result when an observed trajectory
-    segment begins and ends in the same cell.
-    The graph $\scr{H}(0.1)$ can now be viewed as a
-    transition system $\tupleof{C,\x,\scrT,C_0,\HybridStates_0}$,
-    where $C \in vertices(\scr{H}(0.1))$ and $\x \in
-    \HybridStates$. A transition $\tau \in \scrT$ is of the form
-    $\tupleof{C, C',\rho_{\tau}}$, where $\rho_{\tau}(\x,\x')
-    \subseteq \setof{R_{(C,C')}(\x,\x') |( C,C') \in
-    edge(\scr{H}(0.1))}$.
+    segment begins and ends in the same cell. This graph can now be
+    model checked for counter-examples. Note that each path in the
+    graph corresponds to a conunction of affine constraints (when the
+    maps are affine). A counter-example exists if any of the
+    linear programs (paths) are feasible. The over all BMC problem
+    involves reasoning over disjunctions due to the presence of
+    multiple paths.
 
-\mypara{\emph{Check Counter-example}}: The infinite state (but finite location) transition system can
-    be model checked to find a concrete counter-example, which if
-    exists, can indicate the existence of a similar trace in the
-    original black-box system $\System$. The latter check is carried
-    out as before, using the numerical simulation function
-        $\simulate$~\footnote{For the given example, the model checker is unable to
-        find a counter-example.}.
+% \mypara{\emph{Check Counter-example}}: The infinite state (but finite location) transition system can
+%     be model checked to find a concrete counter-example, which if
+%     exists, can indicate the existence of a similar trace in the
+%     original black-box system $\System$. The latter check is carried
+%     out as before, using the numerical simulation function
+%         $\simulate$~\footnote{For the given example, the model checker is unable to
+%         find a counter-example.}.
 \end{example}

### 2017-06-12 implementation.tex
.

@@
-The presented technique was prototyped as a tool: PWA-Rel.  We
+The presented technique was prototyped as a tool: PW-Rel.  We
 compared it against S3CAM~\cite{zutshi2014multiple} on several
@@
 
-PWA-Rel is implemented in Python. It uses S3CAM to generate the graph
+PW-Rel is implemented in Python. It uses S3CAM to generate the graph
 abstraction and regression routines from
@@
 
-To check the transition system, PWA-Rel can either invoke symbolic
+To check the transition system, PW-Rel can either invoke symbolic
 bounded model checkers or explicit model checkers. For symbolic BMC,
@@
 and Ipopt. If feasible, the returned concrete counter-example is
-simulated and checked for existance in $\System$. Also, PWA-Rel
+simulated and checked for existance in $\System$. Also, PW-Rel
 extracts the bounds on the initial parameters by formulating multiple
@@
 
-% PWA-Rel uses S3CAM to discover the abstraction graph $G$, which is
+% PW-Rel uses S3CAM to discover the abstraction graph $G$, which is
 % then filtered to remove the nodes from which error nodes are not
@@
 
-\mypara{The search parameters} for PWA-Rel include the parameters of
+\mypara{The search parameters} for PW-Rel include the parameters of
 S3CAM: $N$, $\epsilon$ and $\Delta$. Additionally, they also include
@@
 Van der Pol oscillator, Brusselator, Lorenz attractor and the
-Navigation benchmark. We also tried PWA-Rel on the insulin
+Navigation benchmark. We also tried PW-Rel on the insulin
 benchmark~\cite{chenformal} and lane change maneuver
-benchmark~\cite{althoff2012reachability}. PWA-Rel was compared with
+benchmark~\cite{althoff2012reachability}. PW-Rel was compared with
 S3CAM by running both $10$ times with different seeds (as both use
-randomized algorithms), and their running times were averaged. PWA-Rel
-was evluated with both the symbolic BMC SAL (PWA-Rel+SAL) and the
-explicit BMC (PWA-Rel+OPT). %\todo{SatEx was also tried, but ...}.
+randomized algorithms), and their running times were averaged. PW-Rel
+was evluated with both the symbolic BMC SAL (PW-Rel+SAL) and the
+explicit BMC (PW-Rel+OPT). %\todo{SatEx was also tried, but ...}.
 We
-tabulate both the total time taken by PWA-Rel+SAL and PWA-Rel+OPT to
+tabulate both the total time taken by PW-Rel+SAL and PW-Rel+OPT to
 compute the counter-example and compare against an implementation of
 S3CAM.  Different abstraction parameters are used for S3CAM and
-PWA-Rel, due to the differences in which they operate. PWA-Rel
+PW-Rel, due to the differences in which they operate. PW-Rel
 requires a finer abstraction to begin with in order to minimize the
@@
 time taken by the BMC engine. The total time in minutes is noted under
-    \textbf{PWA-Rel+BMC solver} and \textbf{S3CAM}.
+    \textbf{PW-Rel+BMC solver} and \textbf{S3CAM}.
 $TO$ signifies time $>5hr$, after which the search was killed.}
@@
 \toprule
-    Benchmark & PWA Rel + SAL & PWA Rel + LP & S3CAM\\
+    Benchmark & PW-Rel + SAL & PW-Rel + LP & S3CAM\\
 \midrule
@@
 apparent form the results, the performance is comparable in all
-benchmarks, but the last benchmark: bouncing ball. PWA-Rel determines
-it to be `safe' as it can not find a counter-example in the PWA
-relational model. Also, apparent from the results, PWA-Rel is more
-scalable with the LP solver and PWA-Rel+SAL times out on most of the
+benchmarks, but the last benchmark: bouncing ball. PW-Rel determines
+it to be `safe' as it can not find a counter-example in the piecewise
+relational model. Also, apparent from the results, PW-Rel is more
+scalable with the LP solver and PW-Rel+SAL times out on most of the
 benchmarks.  The efficiency of the naive LP solver can be attributed
@@
 However, we need to explore more benchmarks to be conclusive on the
-tool's performance on real world examples. Although, PWA-Rel does not
+tool's performance on real world examples. Although, PW-Rel does not
 need to refine like S3CAM, and should be more scalable, it is still
-constrained by the accuracy of the PWA models. Its performance can be
+constrained by the accuracy of the piecewise models. Its performance can be
 explained by the difficulty in finding an abstraction

### 2017-06-12 appendix.tex
.

@@
 
-   ============
-   The systems under test are assumed to be switched-mode to ease the
-   presentation, but it is not a limitation of the presented
-   technique. Black box systems which have hybrid dynamics due to
-   multiple modes, are dealt in the same fashion.
-   [TODO: $K$-refinement to differentiate the dynamics of different modes.
-   We have added a benchmark to showcase this.
-   [TODO: which one?]
-   [TODO: clarify the absence of any controllers]
-   [TODO: remove or clarify switched-mode systems]
-   [TODO: add a hybrid benchmark]
-   ============
-
-   I imagine that the idea is to combine the generated discrete abstraction
-   with a discrete controller model to verify both together. It would be
-   good to state this clearly and to treat the resulting issues throughout
-   the paper. I note that the experimental results do not treat this issue.
-   Otherwise, why not change the title and focus the text on what you
-   actually did?
-
-   ============
-   We will explore this in the future, but the current presentation
-   only works with the plant, which is itself a hybrid dynamical
-   system.
-   [TODO: Add this in the conclusion/future work]
-   ============
-
-   (Similar remarks hold at a smaller scale. The discussion often describes
-    what "can be" done, e.g., in sections 3.2 and 4.1. I would rather read
-    about what "has been" done and what the result was!)
-
-   ============
-   [TODO]
-   ============
+>> The systems under test are assumed to be switched-mode to ease the
+>> presentation, but it is not a limitation of the presented
+>> technique. Black box systems which have hybrid dynamics due to
+>> multiple modes, are dealt in the same fashion.
 
@@
 
+>> BMC is required to find a concrete counter-example. The invariants
+>> found through clustering (which has been since removed due to the
+>> impending confusion) can not be used to conclude the presence of a
+>> concrete coutner-examples by themselves as they are just an
+>> approximation.
+
    My guess is that you later want to restrict the state-space by composing
-   your abstraction with a controller model. Is that right? This probably
-   seems obvious to you, but please put yourself in the place of someone who
-   is new to the work and not an expert in the domain! In any case, this
-   idea is not clearly stated, not addressed by the text, and not considered
-   in the experimental evaluation.
+   your abstraction with a controller model. Is that right?
 
-   ============
-   [TODO: I don't understand this comment, but it is related to
-   point (iii) of the summary]
-   Although we enumerate the abstract space, we require the model
-   checker to find a concrete counter-example.
-   ============
+>> This is a very important idea which we plan to explore in future
+>> work. But it is not being presented in the paper as of now.
 
-   Is this why the PWA transition system characterizes states by a valuation
-   function v and not simply as tuples? If yes, how will this work? I do not
-   understand why you have a valuation function when the system only has one
-   variable (x).
-   ============
-   [TODO]
-   ============
 ======================================
@@
 >> In general, finding the `optimal' number of PWA models can be
->> framed as a multiobjective optimization problem, where a small
+>> framed as a multi-objective optimization problem, where a small
 >> number of clusters and an accurate affine map is desired. This is a
@@
 
+>> Current examples include generic dynamical systems. But the same
+>> procedure can be used to find falsifications in closed loop control
+>> systems. We plan to do extensive case studies in the future.
+
 ======================================
@@
 
+>> In earlier works, S3CAM was shown to be better than S-Taliro.
+>> Hence, for now only S3CAM was compared. In the future we plan to have
+>> extensive case-studies and comparison with other tools.
+
 I also recommend using relations more complex than affine, such as
@@
 the number of hyper-rectangles).
+
+>> We tried polynomial kernels but Z3 can not scale even on small
+>> number of polynomial constraints. Instead we used non-linear
+>> optimization solvers. But, they were very sensitive to the initial
+>> guess.
 
@@
 >> depends on the problem at hand, we show our methodology using the
->> simplest one. In practice, sophesticated statistical tests can be used
+>> simplest one. In practice, sophisticated statistical tests can be used
 >> to find the a good analysis.

### 2017-06-12 implementation.tex
.

@@
 confirms if the counter-example is valid for $\System$, in which case
-a successfull falsification run is concluded.
+a successful falsification run is concluded.
 
@@
 bounded model checkers or explicit model checkers. For symbolic BMC,
-the off-the-shelf SAL modelchecker~\cite{SAL-SRI} is used. It is
+the off-the-shelf SAL model checker~\cite{SAL-SRI} is used. It is
 configured to encode the problem as a single query for a given depth
@@
 Explicit model checking was implemented out using a naive DFS
-(depth-first-exploration) stratergy, wherein, each abstract path is
+(depth-first-exploration) strategy, wherein, each abstract path is
 enumerated and checked for feasibility using a an optimization
-routine. Intefaces to various solvers were implemented, including GLPK
+routine. Interfaces to various solvers were implemented, including GLPK
 and Ipopt. If feasible, the returned concrete counter-example is
-simulated and checked for existance in $\System$. Also, PW-Rel
+simulated and checked for existence in $\System$. Also, PW-Rel
 extracts the bounds on the initial parameters by formulating multiple
@@
 
-% If the existance check fails, multiple
+% If the existence check fails, multiple
 % parameter values can be sampled within the determined bounds to search
-% for counter-examples in the neighbhourhood of the returned concrete
+% for counter-examples in the neighborhood of the returned concrete
 % counter-example.
@@
 Van der Pol oscillator, Brusselator, Lorenz attractor and the
-Navigation benchmark. We also tried PW-Rel on the insulin
-benchmark~\cite{chenformal} and lane change maneuver
-benchmark~\cite{althoff2012reachability}. PW-Rel was compared with
+Navigation benchmark.
+%We also tried PW-Rel on the insulin
+%benchmark~\cite{chenformal} and lane change maneuver
+%benchmark~\cite{althoff2012reachability}.
+PW-Rel was compared with
 S3CAM by running both $10$ times with different seeds (as both use
 randomized algorithms), and their running times were averaged. PW-Rel
-was evluated with both the symbolic BMC SAL (PW-Rel+SAL) and the
+was evaluated with both the symbolic BMC SAL (PW-Rel+SAL) and the
 explicit BMC (PW-Rel+OPT). %\todo{SatEx was also tried, but ...}.
-We
-tabulate both the total time taken by PW-Rel+SAL and PW-Rel+OPT to
+An affine kernel was used for modeling and the linear programming
+solver GLPK was used for checking feasibility of the constraints.
+We also tried polynomial kernels, and used Z3 SMT
+solver~\cite{DeMoura+Bjorner/08/Z3} to check the symbolic BMC.
+Unfortunately, Z3 does not scale well on polynomial constraints and
+consistently timed out. For explicit BMC, several non-linear
+optimization engines were tried, but the results were inconsistent due
+to sensitivity towards the initial guess.
+
+We tabulate both the total time taken by PW-Rel+SAL and PW-Rel+OPT to
 compute the counter-example and compare against an implementation of
@@
 \toprule
-    Benchmark & PW-Rel + SAL & PW-Rel + LP & S3CAM\\
+    Benchmark & PW-Rel + SAL & PW-Rel + OPT & S3CAM\\
 \midrule
@@
 
-    insulin-st-1             &$Safe$ & $Safe$  & $0.0$\\
-    insulin-st-2             &$Safe$ & $Safe$  & $0.0$\\
-    lane change              &$Safe$ & $Safe$  & $0.0$\\
-    HVAC                     &$Safe$ & $Safe$  & $0.0$\\
+    %insulin-st-1             &$Safe$ & $Safe$  & $0.0$\\
+    %insulin-st-2             &$Safe$ & $Safe$  & $0.0$\\
+    %lane change              &$Safe$ & $Safe$  & $0.0$\\
+    %HVAC                     &$Safe$ & $Safe$  & $0.0$\\
 
@@
 prediction models but coarse enough to be manageable by the current
-bmc tools.  To obtain a reproducible counter-example from the model
+BMC tools.  To obtain a reproducible counter-example from the model
 checker (and avoid false positives), the transition system should be

### 2017-06-12 learning.tex
.

@@
 In statistics, parametric regression is the problem of finding the
-paremters of a function or \textit{predictor}, which can `best'
+parameters of a function or \textit{predictor}, which can `best'
 summarize the relationship between the given set of observed input
 $\x$ and output $\x'$ vectors. The basis functions used for the
-predictor are fixed beforehand and the paramters are searched for.
+predictor are fixed beforehand and the parameters are searched for.
 The notion of `best' is formally captured using a loss function which
@@
 regression analysis affects the quality of the learnt models. As the
-procss of selecting an appropriate heuristic is not straightforward,
+process of selecting an appropriate heuristic is not straightforward,
 we just use the simplest one.} OLS defines its loss function as the
@@
 predictor is minimized for the given data set.  The details on
-performing linear regressoin can be found in standard texts on
+performing linear regression can be found in standard texts on
 learning and statistics~\cite{friedman2001elements}.
@@
 in $\vxp_i$.
+
+\begin{algorithm}[t]
+\DontPrintSemicolon
+\caption{RelationalAbstraction\label{algo:relabs}}
+\KwIn{$\simmap_\Delta$, $\TSAbstraction{\epsilon}{\System}$,
+    regression kernel $F_k$}
+\KwOut{$\RTSAbstraction{\epsilon}{\System}$}
+    \ForAll{$(\Cells,\Cells') \in \CellEdges$}{
+    $F, \delta = \mathsf{Regression(F_k, \simmap_\Delta, (\Cells,\Cells'))}$ \;
+    $r(\vx,\vx^+) = F, \delta$
+    %$\RTSAbstraction{\epsilon}{\System}$ $\assign$
+    %$\mathsf{k-RelationalAbstraction}$($\TSAbstraction{\epsilon}{\System}, F$) \nllabel{algoline:relabs} \;
+    %$\Pi$ $\assign$ $\mathsf{BoundedModelCheck}$($\RTSAbstraction{\epsilon}{\System}$) \nllabel{algoline:bmc} \;
+    %\lIf {$\Pi = \emptyset$}{ return FAIL \nllabel{algoline:nocex} }
+    %\lIf {$\mathsf{checkIfConcretizable}(\Pi)$}{ return VIOLATION
+    %\nllabel{algoline:checkspurious} }
+}
+\end{algorithm}
 

### 2017-06-12 prelims2.tex
.

@@
 function (which can be non-linear), resulting in different kinds of
-regression analysis including ridge, lasso, huber, generelized and
+regression analysis including ridge, lasso, huber, generalized and
 many other regressions. For simplicity, we present the commonly used

### 2017-06-12 scam.tex
.

@@
 system which has discrete modes and a continuous state-space, and which
-evolvution is either (1) continuous in time according to the dynamics
+evolution is either (1) continuous in time according to the dynamics
 associated with its current discrete mode (typically described using
@@
 an initial state $\initstate \in \InitStates$ at time $t=0$.  We
-assume that the system has $m$ exogenuous inputs which take values
+assume that the system has $m$ exogenous inputs which take values
 from the set $\Inputs$ which is a compact subset of $\Reals^m$. We
@@
 
-We remark that our abstraction is {\em neither an overapproximation
-nor an underapproximation} of the continuous-time continuous-space
+We remark that our abstraction is {\em neither an over-approximation
+nor an under-approximation} of the continuous-time continuous-space
 reachability relation between states. The purpose of the $\epsilon$-TS
